### 2013-04-26 paper.tex
paper skeleton

\documentclass[a4paper]{article}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage[round]{natbib}

% working title, please change
\title{Python as a machine learning language:
       experiences from the scikit-learn project}

\begin{document}

\maketitle

\bibliographystyle{plainnat}
\bibliography{paper}

\end{document}

### 2013-04-26 api.tex
draft section about estimator API

\section{Application programming interface (API)}

The bulk of the scikit-learn application programming interface
consists of so-called \textit{estimator} objects.
All learning algorithms for classification, regression and clustering
are offered as such objects,
but so are feature extraction, feature selection and dimensionality reduction.
What all estimators have in common is a method called \texttt{fit},
which causes the estimator to learn from data supplied as an argument.
Such a method may seem superfluous for tasks such as vectorizing text documents
for subsequent learning,
but even the classes that handle this task can be said
to learn their vocabulary from the training corpus.
(Only in a very small number of cases,
specifically two estimators that implement the ``hashing trick'' of
\citealt{weinberger2009} for feature extraction,
is \texttt{fit} a no-op offered only for consistency and composibility.)

The \texttt{fit} method's task is to learn model-specific parameters
from a training set and set these as attributes on the estimator object,
which can then be used to transform data or make predictions.
As in other object-oriented machine learning libraries
(e.g.\ Weka's Java API, \citealp{hall2009weka})
no distinction is made in the type hierarchy
between what is variously called an estimator or a \textit{learner}
in the machine learning literature,
and the \textit{model} that results from applying an estimator to data.
When the \texttt{fit} method has been called,
an estimator object serves dual purpose as a model
(and is referred to as such in the documentation).

Importantly, the parameters learned by an estimator
are exposed as \textit{public} attributes on the trained object.\footnote{
  The Python language does not actually distinguish between public and private
  attributes and methods,
  but by convention, any identifier prefixed with an underscore (\texttt{\_})
  is considered ``a non-public part of the API\ldots
  an implementation detail and subject to change without notice.''}
This facilitates model inspection
and makes it possible to train a model using scikit-learn,
export it to a (standardized or custom) external data format
and use the learned parameters in a different piece of code,
perhaps a prediction algorithm implemented in a different language
or one that combines the model parameters with other information
to build a custom model.
Of course, it also forces the developers to think carefully
about what might in a more ``black box''-style toolkit
be considered the internals of objects,
since authors are client code are allowed to rely
on the names and formats of model parameters.

\subsection{Pipelines and model selection}

A distinguishing feature of scikit-learn estimators
is that they can be composed into \texttt{Pipeline} objects.
Not only do such objects combine typical machine learning workflows
(extract features, center/scale/normalize, select features, 
make predictions) in a single object that is itself an estimator
whose methods cause data to flow through the pipeline,
but more importantly they can be used for
systematic, automated model selection.
This is best explained by an example.

By composing, say, a \texttt{CountVectorizer} object
(which extracts term frequency features from text documents
according to a bag-words model),
a \textsf{tf--idf} transformer, a $\chi^2$ feature selector and a linear SVM,
one obtains a pipeline object with various hyperparameters:
whether to do stopword filtering,
linear vs.\ logarithmic \textsf{tf}, number of features to keep,
SVM regularization ($L_1$ or $L_2$ norm, strength $C$) a.o.
Such a composite \texttt{Pipeline} object
exposes all of these parameters using a special syntax;
each estimators must be given a name,
and if the SVM's name is, e.g., \texttt{linearsvm},
then its $C$ parameter is exposed as \texttt{linearsvm\_\_C}.

The aforementioned \texttt{Pipeline} object
can be passed as an argument to a model selection algorithm.
Available algorithms at the time of writing include random search
\citep{bergstra2012} and ``grid search''.
The latter, which we shall use in this example,
performs an exhaustive search through a grid formed by the Cartesian product
of a set of possible values for each parameter, specified by the user.
In our example, this grid might be given as

\begin{align*}
         & \textsf{stopwords} \in \{0, 1\}                      \\
  \times & \; \textsf{tf} \in \{\textsf{linear}, \textsf{log}\} \\
  \times & \; n_\textsf{features} \in \{1000, 2000, 5000\}      \\
  \times & \; \textsf{norm} \in \{L_1, L_2\}                    \\
  \times & \; C \in \{10, 100, 1000\}
\end{align*}

At each point is this grid, $k$-fold cross validation is run
to estimate the performance of the estimator according to some measure
(e.g.\ accuracy, $F_\beta$-score or a host of clustering quality metrics)
and the best performing set of hyperparameters is stored
on the \texttt{GridSearchCV} object; again, as a public attribute.

### 2013-04-26 api.tex
describe/defend use of NumPy arrays

@@
 
+% subsection here? also discuss scipy.sparse?
+Input to an estimator is typically given as NumPy arrays \citep{vanderwalt2011};
+typically, a clustering or dimensionality reduction algorithm
+takes a single NumPy array, conventionally called $X$ of size $n \times m$
+to represent $n$ samples of $m$ features each.
+A regression or classification estimator takes, at training time,
+an additional array of size $n$, by convention called $y$,
+that represents the target variable.
+While the NumPy array may seem like a ``bare'' representation of data
+when compared to more object-oriented representations
+(again, Weka is an example of this style of API),
+it brings the prime advantages of allowing \textit{vector operations}
+on entire arrays,
+which are often orders of magnitude faster than the corresponding Python loops
+while at the same time keeping the code short and readable.
+Although conversion to the NumPy format may seem like a burden on the user,
+we argue that this is not the case.
+Python's lists (its dynamic array structure)
+are easily transformed into NumPy arrays
+and many scientific users of Python will already be handling such arrays,
+since they pervade in other science-oriented Python packages.
+For tasks where input is likely to consists of text files
+or semi-structured objects, ``vectorizers'' are provided
+that efficiently convert from such data formats to the NumPy format.
+Finally, offering a custom datatype for samples
+would require a conversion \textit{in any case}.
+
 \subsection{Pipelines and model selection}

### 2013-04-26 api.tex
typo

@@
 
-At each point is this grid, $k$-fold cross validation is run
+At each point in this grid, $k$-fold cross validation is run
 to estimate the performance of the estimator according to some measure

### 2013-04-26 api.tex
expand/restructure API section

@@
 \section{Application programming interface (API)}
+
+\subsection{Estimators}
+
+% TODO describe transformer vs. classifiers/regressors
 
@@
 but so are feature extraction, feature selection and dimensionality reduction.
-What all estimators have in common is a method called \texttt{fit},
-which causes the estimator to learn from data supplied as an argument.
-Such a method may seem superfluous for tasks such as vectorizing text documents
-for subsequent learning,
-but even the classes that handle this task can be said
-to learn their vocabulary from the training corpus.
-(Only in a very small number of cases,
-specifically two estimators that implement the ``hashing trick'' of
-\citealt{weinberger2009} for feature extraction,
-is \texttt{fit} a no-op offered only for consistency and composibility.)
 
-The \texttt{fit} method's task is to learn model-specific parameters
+Construction and application of an estimator object
+proceeds in a way that is somewhat reminiscent of partial function application.
+The constructor, \texttt{\_\_init\_\_}\footnote{
+  Strictly speaking, \texttt{\_\_init\_\_} is Python's
+  class \textit{initialization} method,
+  while \texttt{\_\_new\_\_} is the constructor.
+  The distinction is not relevant to the present description of scikit-learn.}
+takes a set of named (hyper-)parameters
+and attaches these to the estimator as public attributes.\footnote{
+  The Python language does not actually distinguish between public and private
+  attributes and methods,
+  but by convention, any identifier prefixed with an underscore (\texttt{\_})
+  is considered ``a non-public part of the API\ldots
+  an implementation detail and subject to change without notice.''
+  % TODO cite Python tutorial here
+  }
+It never performs actual learning, nor does it even see a training set.
+
+The workhorse of most estimators, and one of the few methods
+that all estimators have in common, is called \texttt{fit}.
+Calling this method causes the estimator to learn from a training set,
+supplied as one or more arguments.
+Its task, from an API point of view,
+is to determine (via a learning algorithm) model-specific parameters
 from a training set and set these as attributes on the estimator object,
@@
 Importantly, the parameters learned by an estimator
-are exposed as \textit{public} attributes on the trained object.\footnote{
-  The Python language does not actually distinguish between public and private
-  attributes and methods,
-  but by convention, any identifier prefixed with an underscore (\texttt{\_})
-  is considered ``a non-public part of the API\ldots
-  an implementation detail and subject to change without notice.''}
+are exposed as \textit{public} attributes on the trained object.
 This facilitates model inspection
@@
 
-% subsection here? also discuss scipy.sparse?
+A \texttt{fit} method is present even on objects
+that perform seemingly mundane tasks such as vectorizing text documents
+for subsequent learning.
+Such classes, however, can be said to learn their vocabulary
+from the training corpus,
+and follow the conventions set out above.
+Even those estimators that are entirely stateless and therefore
+do not require a \texttt{fit} method to perform useful work,
+nevertheless have such a method for consistency and composibility,
+and to perform parameter validation.
+Examples of this latter kind of objects are the normalizing transformer,
+kernel approximation transformers,
+and two feature extraction classes that implement the ``hashing trick''
+\citep{weinberger2009}.
+
+\subsection{Data formats}
+
+% TODO also discuss scipy.sparse
 Input to an estimator is typically given as NumPy arrays \citep{vanderwalt2011};
 typically, a clustering or dimensionality reduction algorithm
-takes a single NumPy array, conventionally called $X$ of size $n \times m$
-to represent $n$ samples of $m$ features each.
+takes a single NumPy array, conventionally called $X$ of size $n \times d$
+to represent $n$ samples of $d$ features each.
 A regression or classification estimator takes, at training time,
@@
 Not only do such objects combine typical machine learning workflows
-(extract features, center/scale/normalize, select features, 
-make predictions) in a single object that is itself an estimator
-whose methods cause data to flow through the pipeline,
-but more importantly they can be used for
+(extracting features, centering/scaling/normalization,
+feature selection, making predictions)
+in a single object that is itself an estimator,
+and whose methods cause data to flow through the pipeline.
+More importantly, though, they can be used for
 systematic, automated model selection.
-This is best explained by an example.
+This is best explained by example.
 

### 2013-04-27 api.tex
citations for kernel approximations
Thanks @amueller!

@@
 Examples of this latter kind of objects are the normalizing transformer,
-kernel approximation transformers,
+kernel approximation transformers
+\citep{rahimi2007random, li2010random, vedaldi2010efficient},
 and two feature extraction classes that implement the ``hashing trick''

### 2013-04-27 api.tex
example code for partial application/method chaining style

@@
 (and is referred to as such in the documentation).
+
+An example to clarify the partial application style
+is the following code snippet.
+If $X$ is a dataset in the format described below and $k$ is some integer,
+then the Python code to train a $k$-means clustering model on it
+can be written as
+% XXX we could use Pygments for syntax highlighting
+\begin{verbatim}
+km = KMeans(n_clusters=k).fit(X)
+\end{verbatim}
+This snippet relies on the fact that the \texttt{fit} method
+always returns the object it was called on (its ``\texttt{self}''),
+making a method chaining style possible.
 

### 2013-04-27 api.tex
typo

@@
 and many scientific users of Python will already be handling such arrays,
-since they pervade in other science-oriented Python packages.
+since they are pervasive in other science-oriented Python packages.
 For tasks where input is likely to consists of text files

### 2013-04-27 api.tex

@@
   an implementation detail and subject to change without notice.''
-  % TODO cite Python tutorial here
+  \Citep{pythontut}
   }
@@
 % TODO also discuss scipy.sparse
-Input to an estimator is typically given as NumPy arrays \citep{vanderwalt2011};
+Input to an estimator is typically given as NumPy arrays
+\Citep{vanderwalt2011};
 typically, a clustering or dimensionality reduction algorithm

### 2013-04-28 api.tex
Rework "Data formats" subsection.

@@
 
-\subsection{Data formats}
-
-% TODO also discuss scipy.sparse
-Input to an estimator is typically given as NumPy arrays
-\Citep{vanderwalt2011};
-typically, a clustering or dimensionality reduction algorithm
-takes a single NumPy array, conventionally called $X$ of size $n \times d$
-to represent $n$ samples of $d$ features each.
-A regression or classification estimator takes, at training time,
-an additional array of size $n$, by convention called $y$,
-that represents the target variable.
-While the NumPy array may seem like a ``bare'' representation of data
-when compared to more object-oriented representations
-(again, Weka is an example of this style of API),
-it brings the prime advantages of allowing \textit{vector operations}
-on entire arrays,
-which are often orders of magnitude faster than the corresponding Python loops
-while at the same time keeping the code short and readable.
-Although conversion to the NumPy format may seem like a burden on the user,
-we argue that this is not the case.
-Python's lists (its dynamic array structure)
-are easily transformed into NumPy arrays
-and many scientific users of Python will already be handling such arrays,
-since they are pervasive in other science-oriented Python packages.
-For tasks where input is likely to consists of text files
-or semi-structured objects, ``vectorizers'' are provided
-that efficiently convert from such data formats to the NumPy format.
-Finally, offering a custom datatype for samples
-would require a conversion \textit{in any case}.
-
 \subsection{Pipelines and model selection}

### 2013-04-28 vision.tex
Rework "Data formats" subsection.

\newcommand{\Xtr}{\texttt{X\_tr}}
\newcommand{\Xte}{\texttt{X\_te}}
\newcommand{\ytr}{\texttt{y\_tr}}
\newcommand{\yte}{\texttt{y\_te}}

\section{Projection vision}

\subsection{General principles}

ToDo.

\subsection{Data formats}

As a machine learning library, one of the most important decisions we had to
make is how to represent data.  Rather than reinventing the wheel, we opted for
NumPy multidimensional arrays \Citep{vanderwalt2011} for dense data and SciPy
sparse matrices for sparse data.  While those may seem like a ``bare''
representation of data when compared to more object-oriented representations
(Weka is an example of this style of representation), it brings the prime
advantage of allowing us to rely on NumPy and SciPy's \textit{vector
operations} which are often orders of magnitude faster than the corresponding
Python loops, while at the same time keeping the code short and readable.
Conversion to NumPy and SciPy formats is usually easy and many scientific users
of Python will already be used to such formats, since they are pervasive in
other scientific Python packages.  For tasks where input is likely to consist
of text files or semi-structured objects, we provide ``vectorizers'' -- objects
that efficiently convert such data to the NumPy or SciPy formats.
% FIXME: elaborate what is meant below
%Finally, offering a custom datatype for samples would require a conversion
%\textit{in any case}.

In the remaing of the paper, we use the following notation. Training data are
denoted by \Xtr ~and test data (i.e., unseen data to which the learning
algorithm need be able to generalize) by \Xte.  Following the discussion above,
we represent \Xtr ~and \Xte ~by 2-dimensional NumPy arrays or SciPy sparse
matrices. Using NumPy/SciPy's notation, \Xtr\texttt{[i, j]} denotes the
$i^{th}$ training sample's $j^{th}$ feature. For supervised learning tasks,
such as classification or regression, we denote the training and test labels by
\ytr ~and \yte, respectively. We typically store \ytr ~and \yte ~as
1-dimensional NumPy arrays (\ytr\texttt{[i]} denotes the $i^{th}$ training
sample's label value).

### 2013-04-28 paper.tex
draft section about estimator API
solve Dutch name sorting problem
Rework "Data formats" subsection.
Merge everything into one file.

@@
 
+% Dutch name sorting hack as per http://tex.stackexchange.com/a/40750/2806
+\DeclareRobustCommand{\VAN}[3]{#2}
+
+\newcommand{\Xtr}{\texttt{X\_tr}}
+\newcommand{\Xte}{\texttt{X\_te}}
+\newcommand{\ytr}{\texttt{y\_tr}}
+\newcommand{\yte}{\texttt{y\_te}}
+
 \begin{document}
@@
 
+\section{Project vision}
+
+\subsection{General principles}
+
+ToDo.
+
+\subsection{Data formats}
+
+As a machine learning library, one of the most important decisions we had to
+make is how to represent data.  Rather than reinventing the wheel, we opted for
+NumPy multidimensional arrays \Citep{vanderwalt2011} for dense data and SciPy
+sparse matrices for sparse data.  While those may seem like a ``bare''
+representation of data when compared to more object-oriented representations
+(Weka is an example of this style of representation), it brings the prime
+advantage of allowing us to rely on NumPy and SciPy's \textit{vector
+operations} which are often orders of magnitude faster than the corresponding
+Python loops, while at the same time keeping the code short and readable.
+Conversion to NumPy and SciPy formats is usually easy and many scientific users
+of Python will already be used to such formats, since they are pervasive in
+other scientific Python packages.  For tasks where input is likely to consist
+of text files or semi-structured objects, we provide ``vectorizers'' -- objects
+that efficiently convert such data to the NumPy or SciPy formats.
+% FIXME: elaborate what is meant below
+%Finally, offering a custom datatype for samples would require a conversion
+%\textit{in any case}.
+
+In the remaing of the paper, we use the following notation. Training data are
+denoted by \Xtr ~and test data (i.e., unseen data to which the learning
+algorithm need be able to generalize) by \Xte.  Following the discussion above,
+we represent \Xtr ~and \Xte ~by 2-dimensional NumPy arrays or SciPy sparse
+matrices. Using NumPy/SciPy's notation, \Xtr\texttt{[i, j]} denotes the
+$i^{th}$ training sample's $j^{th}$ feature. For supervised learning tasks,
+such as classification or regression, we denote the training and test labels by
+\ytr ~and \yte, respectively. We typically store \ytr ~and \yte ~as
+1-dimensional NumPy arrays (\ytr\texttt{[i]} denotes the $i^{th}$ training
+sample's label value).
+
+\section{Application programming interface (API)}
+
+\subsection{Estimators}
+
+% TODO describe transformer vs. classifiers/regressors
+
+The bulk of the scikit-learn application programming interface
+consists of so-called \textit{estimator} objects.
+All learning algorithms for classification, regression and clustering
+are offered as such objects,
+but so are feature extraction, feature selection and dimensionality reduction.
+
+Construction and application of an estimator object
+proceeds in a way that is somewhat reminiscent of partial function application.
+The constructor, \texttt{\_\_init\_\_}\footnote{
+  Strictly speaking, \texttt{\_\_init\_\_} is Python's
+  class \textit{initialization} method,
+  while \texttt{\_\_new\_\_} is the constructor.
+  The distinction is not relevant to the present description of scikit-learn.}
+takes a set of named (hyper-)parameters
+and attaches these to the estimator as public attributes.\footnote{
+  The Python language does not actually distinguish between public and private
+  attributes and methods,
+  but by convention, any identifier prefixed with an underscore (\texttt{\_})
+  is considered ``a non-public part of the API\ldots
+  an implementation detail and subject to change without notice.''
+  \Citep{pythontut}
+  }
+It never performs actual learning, nor does it even see a training set.
+
+The workhorse of most estimators, and one of the few methods
+that all estimators have in common, is called \texttt{fit}.
+Calling this method causes the estimator to learn from a training set,
+supplied as one or more arguments.
+Its task, from an API point of view,
+is to determine (via a learning algorithm) model-specific parameters
+from a training set and set these as attributes on the estimator object,
+which can then be used to transform data or make predictions.
+As in other object-oriented machine learning libraries
+(e.g.\ Weka's Java API, \citealp{hall2009weka})
+no distinction is made in the type hierarchy
+between what is variously called an estimator or a \textit{learner}
+in the machine learning literature,
+and the \textit{model} that results from applying an estimator to data.
+When the \texttt{fit} method has been called,
+an estimator object serves dual purpose as a model
+(and is referred to as such in the documentation).
+
+An example to clarify the partial application style
+is the following code snippet.
+If $X$ is a dataset in the format described below and $k$ is some integer,
+then the Python code to train a $k$-means clustering model on it
+can be written as
+% XXX we could use Pygments for syntax highlighting
+\begin{verbatim}
+km = KMeans(n_clusters=k).fit(X)
+\end{verbatim}
+This snippet relies on the fact that the \texttt{fit} method
+always returns the object it was called on (its ``\texttt{self}''),
+making a method chaining style possible.
+
+Importantly, the parameters learned by an estimator
+are exposed as \textit{public} attributes on the trained object.
+This facilitates model inspection
+and makes it possible to train a model using scikit-learn,
+export it to a (standardized or custom) external data format
+and use the learned parameters in a different piece of code,
+perhaps a prediction algorithm implemented in a different language
+or one that combines the model parameters with other information
+to build a custom model.
+Of course, it also forces the developers to think carefully
+about what might in a more ``black box''-style toolkit
+be considered the internals of objects,
+since authors are client code are allowed to rely
+on the names and formats of model parameters.
+
+A \texttt{fit} method is present even on objects
+that perform seemingly mundane tasks such as vectorizing text documents
+for subsequent learning.
+Such classes, however, can be said to learn their vocabulary
+from the training corpus,
+and follow the conventions set out above.
+Even those estimators that are entirely stateless and therefore
+do not require a \texttt{fit} method to perform useful work,
+nevertheless have such a method for consistency and composibility,
+and to perform parameter validation.
+Examples of this latter kind of objects are the normalizing transformer,
+kernel approximation transformers
+\citep{rahimi2007random, li2010random, vedaldi2010efficient},
+and two feature extraction classes that implement the ``hashing trick''
+\citep{weinberger2009}.
+
+\subsection{Pipelines and model selection}
+
+A distinguishing feature of scikit-learn estimators
+is that they can be composed into \texttt{Pipeline} objects.
+Not only do such objects combine typical machine learning workflows
+(extracting features, centering/scaling/normalization,
+feature selection, making predictions)
+in a single object that is itself an estimator,
+and whose methods cause data to flow through the pipeline.
+More importantly, though, they can be used for
+systematic, automated model selection.
+This is best explained by example.
+
+By composing, say, a \texttt{CountVectorizer} object
+(which extracts term frequency features from text documents
+according to a bag-words model),
+a \textsf{tf--idf} transformer, a $\chi^2$ feature selector and a linear SVM,
+one obtains a pipeline object with various hyperparameters:
+whether to do stopword filtering,
+linear vs.\ logarithmic \textsf{tf}, number of features to keep,
+SVM regularization ($L_1$ or $L_2$ norm, strength $C$) a.o.
+Such a composite \texttt{Pipeline} object
+exposes all of these parameters using a special syntax;
+each estimators must be given a name,
+and if the SVM's name is, e.g., \texttt{linearsvm},
+then its $C$ parameter is exposed as \texttt{linearsvm\_\_C}.
+
+The aforementioned \texttt{Pipeline} object
+can be passed as an argument to a model selection algorithm.
+Available algorithms at the time of writing include random search
+\citep{bergstra2012} and ``grid search''.
+The latter, which we shall use in this example,
+performs an exhaustive search through a grid formed by the Cartesian product
+of a set of possible values for each parameter, specified by the user.
+In our example, this grid might be given as
+
+\begin{align*}
+         & \textsf{stopwords} \in \{0, 1\}                      \\
+  \times & \; \textsf{tf} \in \{\textsf{linear}, \textsf{log}\} \\
+  \times & \; n_\textsf{features} \in \{1000, 2000, 5000\}      \\
+  \times & \; \textsf{norm} \in \{L_1, L_2\}                    \\
+  \times & \; C \in \{10, 100, 1000\}
+\end{align*}
+
+At each point in this grid, $k$-fold cross validation is run
+to estimate the performance of the estimator according to some measure
+(e.g.\ accuracy, $F_\beta$-score or a host of clustering quality metrics)
+and the best performing set of hyperparameters is stored
+on the \texttt{GridSearchCV} object; again, as a public attribute.
+
 \bibliographystyle{plainnat}
+\DeclareRobustCommand{\VAN}[3]{#3}
 \bibliography{paper}

### 2013-04-28 paper.tex
Add new sections (to be written).
typos

@@
 
-As a machine learning library, one of the most important decisions we had to
+For a machine learning library, one of the most important decisions we had to
 make is how to represent data.  Rather than reinventing the wheel, we opted for
@@
 
-In the remaing of the paper, we use the following notation. Training data are
+In the remainder of the paper, we use the following notation. Training data are
 denoted by \Xtr ~and test data (i.e., unseen data to which the learning
@@
 
-\subsection{Pipelines and model selection}
+\section{Pipelines and model selection}
 
@@
 
+\section{Comparison with other packages}
+
+ToDo.
+
+\section{Future work}
+
+ToDo.
+
 \bibliographystyle{plainnat}

### 2013-04-28 paper.tex
project vision

@@
 
-ToDo.
+scikit-learn \citep{pedregosa2011} is a machine learning library
+for the Python programming language,
+intended to tie in with the set of numeric and scientific packages
+centered around the NumPy and SciPy libraries
+\Citep{vanderwalt2011, varoquaux2013scipy}.
+It is a library, rather than a novel (domain-specific) programming language,
+intended to be called from Python programs;
+at present, it offers no command line interface or GUI.
+
+The main goal of the scikit-learn project
+is to provide high-quality implementations
+of ``classic'' machine learning algorithms
+with a consistent API.
 
@@
 representation of data when compared to more object-oriented representations
-(Weka is an example of this style of representation), it brings the prime
-advantage of allowing us to rely on NumPy and SciPy's \textit{vector
-operations} which are often orders of magnitude faster than the corresponding
-Python loops, while at the same time keeping the code short and readable.
+(Weka is an example of this style of representation, \citealp{hall2009weka}),
+it brings the prime advantage of allowing us to rely on NumPy and SciPy's
+\textit{vector operations} which are often orders of magnitude faster
+than the corresponding Python loops,
+while at the same time keeping the code short and readable.
 Conversion to NumPy and SciPy formats is usually easy and many scientific users
@@
 As in other object-oriented machine learning libraries
-(e.g.\ Weka's Java API, \citealp{hall2009weka})
+(e.g.\ Weka's Java API)
 no distinction is made in the type hierarchy

### 2013-04-28 paper.tex
elaborated on object conversion (still commented out)
start section on implementation

@@
 % FIXME: elaborate what is meant below
+% Lars: elaborated, but I'm not sure how important this is.
+% It mostly reflects experiences with OO libraries that require conversions
+% between types just to pass information to a method.
 %Finally, offering a custom datatype for samples would require a conversion
-%\textit{in any case}.
+%\textit{in any case}, since such a datatype would be toolkit-specific
+%and no program would store its data using such a type
+%if it weren't from the outset designed to use scikit-learn.
 
@@
 
+\section{Implementation}
+
+scikit-learn is primarily implemented in Python and Cython
+\citep{behnel2011cython},
+a language that extends Python with static typing
+and a compiler that produces C extension modules
+for the Python runtime system.
+In addition, it includes (modified versions of)
+the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
+used for training support vector machines
+and logistic regression models \citep{chang2011libsvm, fan2008}.
+These libraries, both written in C{}\verb!++!,
+and wrapped using Cython modules.
+
+As much as possible is implemented in ``pure'' Python,
+to avoid the compile-edit-debug cycle needed for Cython
+and to keep the codebase readable for as large an audience as possible.
+However, since the speed of the current Python implementation
+is not sufficient for numeric programming
+and NumPy's vector operations are not appropriate for all use cases
+(among other factors, because they often have linear space requirements),
+some of the core algorithms have to be implemented in Cython.\footnote{
+  The only Python implementation currently supported by scikit-learn
+  is the reference implementation, CPython.
+  The alternative Python implementation PyPy \citep{bolz2009tracing}
+  promises to solve this problem by JIT compilation,
+  but does not yet interoperate with NumPy and SciPy.
+}
+Examples include the aforementioned *k*-means implementation,
+stochastic gradient descent for linear models
+and some graph-based clustering algorithms.
+
 \section{Comparison with other packages}

### 2013-04-28 paper.tex
minor stuff (LaTeX != rST)

@@
 matrices. Using NumPy/SciPy's notation, \Xtr\texttt{[i, j]} denotes the
-$i^{th}$ training sample's $j^{th}$ feature. For supervised learning tasks,
+$i^\textrm{th}$ training sample's $j^\textrm{th}$ feature. For supervised learning tasks,
 such as classification or regression, we denote the training and test labels by
 \ytr ~and \yte, respectively. We typically store \ytr ~and \yte ~as
-1-dimensional NumPy arrays (\ytr\texttt{[i]} denotes the $i^{th}$ training
+1-dimensional NumPy arrays (\ytr\texttt{[i]} denotes the $i^\textrm{th}$ training
 sample's label value).
@@
 }
-Examples include the aforementioned *k*-means implementation,
+Examples include the aforementioned $k$-means implementation,
 stochastic gradient descent for linear models

### 2013-04-29 paper.tex
and it's open source, too

@@
 
-scikit-learn \citep{pedregosa2011} is a machine learning library
+scikit-learn \citep{pedregosa2011} is an open source machine learning library
 for the Python programming language,

### 2013-04-29 paper.tex
duck typing

@@
 
+While object-oriented techniques and design patterns
+are used throughout scikit-learn,
+they mostly serve to facilitate code reuse and are not considered a project goal.
+Where possible, the Python principle of ``duck typing'' is exploited
+to simplify the implementation and skip the introduction of superfluous classes.
+This allows for extensibility and flexibility at the same time:
+user-defined estimators that follow scikit-learn conventions
+should be usable in pipelines and other composite objects
+without them actually inheriting from scikit-learn base classes.
+
 \section{Comparison with other packages}

### 2013-04-29 paper.tex
popularity of scikit-learn

@@
 
+% XXX the following is more of a general introduction than a project vision
 scikit-learn \citep{pedregosa2011} is an open source machine learning library
@@
 at present, it offers no command line interface or GUI.
+
+The scikit-learn project started in 2009 and has since been developed
+by a team largely consisting of scientists
+(both from computer science and other fields).
+Its popularity can be gauged from various sources:
+% TODO: list volume on the mailing list, number of contributors,
+% number of issues opened and closed, number of surveys filled in.
+at the time of writing,
+the popular programmer's Q~\&{}A site StackOverflow\footnote{
+  \url{http://stackoverflow.com/questions/tagged/scikit-learn}}
+% update the following figure when paper is finalized
+lists, at the time of writing, 322 scikit-learn-related questions.
 

### 2013-05-01 paper.tex
Typo
counted number of contributors
explain double underscore syntax on pipelines

@@
 % TODO: list volume on the mailing list, number of contributors,
-% number of issues opened and closed, number of surveys filled in.
+% number of issues opened and closed, number of forks, number of surveys filled in.
 at the time of writing,
@@
 % update the following figure when paper is finalized
-lists, at the time of writing, 322 scikit-learn-related questions.
+lists 322 scikit-learn-related questions.
+At the time of writing,  % ddaa494c116e3c16bf032003c5cccbed851733d2
+version control logs for the development version of scikit-learn
+show 183 contributors to the project's source code and documentation
+(after manual deduplication).
 
@@
 then its $C$ parameter is exposed as \texttt{linearsvm\_\_C}.
+The pipeline's \texttt{set\_params} splits the parameter names
+on double underscores, uses the first part of the result
+to look up the estimator,
+and passes the second part to the estimator's own \texttt{set\_params}
+as a keyword argument; due to Python's dynamic parameter passing,
+the result is the same as if the user had called \texttt{set\_params}
+on the SVM themselves.
+(The double underscore syntax is not special to pipelines;
+it is also employed in ensemble estimators such as random forests.)
 

### 2013-05-01 paper.tex
name/affilition for current authors (per git shortlog :)
@mblondel, @jakevdp, please check and correct.

I've sorted author names alphabetically; let's argue about the order
later.
some problems with SciPy and Cython
Hope I hedged this carefully enough.

@@
        experiences from the scikit-learn project}
+
+% Add your name and affiliation here.
+% Don't worry too much about the formatting,
+% as we'll have to redo that for the LNCS format.
+\author{Mathieu Blondel \\
+        Graduate School of System Informatics \\
+        Kobe University \\
+        \small{\texttt{mblondel@ai.cs.kobe-u.ac.jp}}
+      \and
+        Lars Buitinck \\
+        Informatics Institute \\
+        University of Amsterdam \\
+        \small{\texttt{L.J.Buitinck@uva.nl}}
+      \and
+        Jake Vanderplas \\
+        Astronomy Department \\
+        University of Washington \\
+        \small{\texttt{vanderplas@astro.washington.edu}}
+}
 
@@
 
+\section{Weaknesses of the NumPy/SciPy environment}
+
+% This needs an introductory paragraph, but the best I could come up with was:
+As can be expected, the NumPy/SciPy environment is not completely flawless.
+
+% The following hedge needs to be in here; we don't want to diss our peers
+It should be noted that this section is meant
+as a summary of challenges faced by scikit-learn development
+and an attempt at constructive criticism.
+There is considerable overlap between the communities
+that develop or contribute to NumPy, SciPy, scikit-learn, and Cython,
+so the issues raised here may eventually be resolved
+by the scikit-learn developers themselves.
+
+\subsection{Sparse matrix support}
+
+SciPy's handling of sparse matrices is one source of confusion
+for scikit-learn developers and users alike.
+The \texttt{scipy.sparse} package offers quite full-featured support
+for this kind of data structure,
+which is crucial to achieving performance and scalability
+in the face of high-dimensional inputs
+(including, but not limited to, NLP applications).
+However, the interface exposed by its classes
+deviates from the NumPy array interface, sometimes subtly so,
+sometimes entirely.  % XXX is this grammatically correct?
+For example, while NumPy arrays have a variable number of dimensions,
+so that vectors can be considered 1-d arrays
+(without a distinction between row and column vectors),
+matrices are 2-d, etc., \texttt{scipy.sparse} matrices are always 2-dimensional,
+meaning that special care has to be taken when handling sparse vectors.
+
+The development team has built up sufficient familiarity
+with the sparse/array distinction to factor out common patterns
+for supporting both sparse and dense input data structures in many estimators.
+In other cases, though, sparse matrix support has to implemented
+on a per-estimator basis.
+This situation is aggravated by the lack of a C API for sparse matrices,
+so that Cython modules have to implement their own version
+of operations like sparse dot products
+that ideally would have been handled by a library.
+
+However, users of scikit-learn cannot always be expected
+to have similar familiarity, and in some cases
+(e.g., when users want to leverage feature extraction code
+outside the context of a \texttt{Pipeline})
+they will have to face the \texttt{scipy.sparse} interface.
+This is, of course, as much an artifact of the design choice
+to not implement separate classes and interfaces for samples/instances
+as it a consequence of SciPy's implementation of sparse matrices.
+
+\subsection{Python/Cython mismatch}
+
+The Python/Cython divide is another source of potential difficulties
+both for users and developers/contributors.
+While Cython makes writing performance-critical code possible
+in a language that is very much like Python,
+doing so properly still requires some knowledge of C.
+One source of trouble has been the various fixed-size integer types
+that are used by the Python and NumPy C runtimes,
+whose sizes are practically all platform-dependent.
+Choosing the appropriate integer type from among
+\texttt{int}, \texttt{size\_t}, \texttt{Py\_ssize\_t} and \texttt{npy\_intp}
+can require knowledge of the Python extension API,
+the NumPy C API and the C standard.
+Neither Cython nor C compilers provide much in the way of static checking
+for the right type,
+while picking the wrong type may cause segmentation faults
+when users try to use arrays that are too large for the type used
+and using a dynamic type often leads to unacceptable slowdowns.
+
+Another problem concerns the mismatch between on the one hand Python functions,
+which are first-class runtime objects endowed with methods and metadata
+that identify them as ``callable'',
+that dynamically check their number and types of arguments,
+and that can return any Python type of object,
+and compiled C or Cython functions on the other hand,
+which exhibit static typing.
+Where it may sometimes be convenient for estimators
+to take functions as arguments
+(e.g., in the case of kernel machine learning,
+where one might like to use a custom kernel),
+but such usage is not at present possible and may never be.
+Passing C or Cython-compiled functions would require an API for the kernel method
+to be implemented at the C or Cython level,
+only to be used from compiled code.
+Passing arbitrary Python functions would be possible,
+but performance can be expected to degrade significantly
+because of the runtime checks needed for each invocation of the kernel function.
+(The current solution in scikit-learn is to let the user pass not a kernel function,
+but a precomputed Gram matrix for an entire set of objects.
+The linear, polynomial and RBF kernels are treated as special cases in
+the \textsf{LIBSVM}-based SVM learner, where they are implemented in C{}\verb!++!.)
+
 \section{Comparison with other packages}

### 2013-05-01 paper.tex
criticize Python's lack of proper multithreading

@@
 
+\subsection{Parallelism}
+
+At this point in time, support for multi-core programming in Python,
+at least in the CPython reference implementation targeted by scikit-learn,
+is downright weak.
+The interpreter supports multithreading, but due to the ``Global Interpreter Lock'' (GIL),
+no two threads in a process may at the same time execute Python code.
+It follows that multithreading is not a viable solution
+even two embarrassingly parallel problems,
+such as fitting and evaluation multiple models in a grid search procedure.
+The workaround is to use Python's multiprocessing module,
+which emulates threads using multiple operating system-level processes,
+but this in turn leads to high memory use and overhead
+because of the copying of training and test sets
+into the various processes' address spaces.\footnote{
+  The fact that copying occurs at all is an artifact of NumPy's implementation;
+  no attempt at implementing a multiprocessing module
+  that shares NumPy arrays using copy-on-write semantics or memory mapping
+  has so far been completed.}
+
 \subsection{Python/Cython mismatch}

### 2013-05-03 paper.tex
weaken PyPy "promise"

@@
   The alternative Python implementation PyPy \citep{bolz2009tracing}
-  promises to solve this problem by JIT compilation,
+  promises to alleviate this problem by JIT compilation,
   but does not yet interoperate with NumPy and SciPy.

### 2013-05-03 paper.tex
tentative abstract

@@
 \maketitle
+
+\begin{abstract}
+The scikit-learn project aims to create an open source
+machine learning library for the Python programming language,
+offering performant implementations of well-established learning algorithms,
+and tying in with the ecosystem of scientific software
+that has evolved around the NumPy and SciPy packages.
+Since the start of the project in 2009,
+it has gathered a sizeable user base.
+
+This paper discusses design choices made
+in both the programmers' interface (API) and the implementation
+of scikit-learn.
+It also highlights obstacles faced by users and implementers
+of machine learning algorithms in the Python language.
+\end{abstract}
 

### 2013-05-03 paper.tex
repetitive "at the time of writing"

@@
 lists 322 scikit-learn-related questions.
-At the time of writing,  % ddaa494c116e3c16bf032003c5cccbed851733d2
-version control logs for the development version of scikit-learn
+The version control logs for the development version of scikit-learn
+% ddaa494c116e3c16bf032003c5cccbed851733d2
 show 183 contributors to the project's source code and documentation

### 2013-05-03 paper.tex
confusing repetition of the verb "to call"
[The method] is called "fit". Calling this method...
"workhorse of an estimator" looks strange

@@
 
-The workhorse of most estimators, and one of the few methods
+The algorithmic workhorse of most estimators, and one of the few methods
 that all estimators have in common, is called \texttt{fit}.
-Calling this method causes the estimator to learn from a training set,
+Invoking this method causes the estimator to learn from a training set,
 supplied as one or more arguments.

### 2013-05-09 paper.tex
Small typo and style changes

@@
 denoted by \Xtr ~and test data (i.e., unseen data to which the learning
-algorithm need be able to generalize) by \Xte.  Following the discussion above,
+algorithm needs to be able to generalize) by \Xte.  Following the discussion above,
 we represent \Xtr ~and \Xte ~by 2-dimensional NumPy arrays or SciPy sparse
@@
 are offered as such objects,
-but so are feature extraction, feature selection and dimensionality reduction.
+but so are feature extraction, feature selection and dimensionality reduction
+algorithms.
 
@@
 to build a custom model.
+% XXX Hard to follow phrasing below
 Of course, it also forces the developers to think carefully
-about what might in a more ``black box''-style toolkit
+about what might, in a more ``black box''-style toolkit,
 be considered the internals of objects,
-since authors are client code are allowed to rely
+since authors of client code are allowed to rely
 on the names and formats of model parameters.
@@
 \citep{rahimi2007random, li2010random, vedaldi2010efficient},
+% XXX is it relevant that they are exactly two? Propose s/two/the
 and two feature extraction classes that implement the ``hashing trick''
@@
 is that they can be composed into \texttt{Pipeline} objects.
-Not only do such objects combine typical machine learning workflows
+Such objects combine typical machine learning workflows
 (extracting features, centering/scaling/normalization,
@@
 (The double underscore syntax is not special to pipelines;
-it is also employed in ensemble estimators such as random forests.)
+it is also employed in ensemble estimators such as random forests
+% Is the following clear? Suggestions to replace "direct attributes" with"?
+It is chosen as a convention to avoid ambiguity, as direct
+attributes and methods of scikit-learn objects never contain
+double underscores.)
 
@@
 of a set of possible values for each parameter, specified by the user.
-In our example, this grid might be given as
+In our example, this grid might be given as:
 
@@
 and the best performing set of hyperparameters is stored
-on the \texttt{GridSearchCV} object; again, as a public attribute.
+on the \texttt{GridSearchCV} object, again, as a public attribute.
 
@@
 and logistic regression models \citep{chang2011libsvm, fan2008}.
-These libraries, both written in C{}\verb!++!,
+These libraries are both written in C{}\verb!++!
 and wrapped using Cython modules.

### 2013-05-12 paper.tex
focus on hyperparameter optimization

--- 2013-05-03 paper.tex
@@
 \documentclass[a4paper]{article}
+
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{hyperref}
@@
 % working title, please change
-\title{Python as a machine learning language:
-       experiences from the scikit-learn project}
+\title{APIs for hyperparameter optimization in scikit-learn}
 
@@
 in both the programmers' interface (API) and the implementation
-of scikit-learn.
-It also highlights obstacles faced by users and implementers
-of machine learning algorithms in the Python language.
+of scikit-learn,
+with a focus on the problem of hyperparameter optimization.
+%It also highlights obstacles faced by users and implementers
+%of machine learning algorithms in the Python language.
 \end{abstract}
 
-\section{Project vision}
+\section{Introduction}
 
-\subsection{General principles}
+\subsection{Project vision}
 
@@
 with a consistent API.
+This paper explores the role that such an API plays
+for handling the problem of hyperparameter optimization,
+or \textit{model selection} as it is called in the statistics literature.
+
+The rest of this section explains some fundamental design choices
+made throughout scikit-learn.
+% TODO and the rest of the paper...
 
@@
 
+\section{The hyperparameter optimization problem}
+
+The focus of many machine learning toolkits is the \textit{learning} task,
+which, at least in the supervised case,
+can be formalized as the problem of finding a function
+$\mathsf{model} : \mathcal{X} \to \mathcal{Y}$,
+where $\mathcal{X}$ is some set of inputs
+and $\mathcal{Y}$ some set of targets
+($\mathbb{R}^k$ in regression, $\{1,\dots,K\}$ in classification).
+An algorithm that finds such a function is called a \textit{learner}
+% TODO: cite Mitchell 1998
+or alternatively an \textit{estimator}
+(in the scikit-learn documentation, and therefore in this paper).
+Using function currying, its function signature is
+
+\[
+  L : \mathcal{H} \to \mathcal{X} \to \mathcal{Y}
+\]
+
+where $\mathcal{H}$ is the set of all possible \textit{hyperparameter} settings
+for the estimator in question.\footnote{
+  We disregard the fact that different estimators
+  have widely varying sets of hyperparameters for notational simplicity.}
+Hyperparameters include, for example,
+regularization strength of linear models or SVMs,
+choice of loss function for generic linear model learners,
+or the number of features to keep in a feature selection method.
+They are often treated as user-settable parameters
+and not part of the learning problem proper,
+either because they would affect the learning problem
+in a way that would make it intractable \citep{bergstra2012}
+or because their effects cannot properly be measured
+in terms of a loss function on a training set.
+
+However, in recent years, algorithms have been published
+that systematically explore spaces of possible hyperparameter settings.
+The problem of hyperparameter tuning
+is crucial to achieving good generalization performance in machine learning.
+The present paper does not go into hyperparameter optimization algorithms
+in depth, focusing instead on API support for using them.
+
 \section{Application programming interface (API)}
@@
 The bulk of the scikit-learn application programming interface
-consists of so-called \textit{estimator} objects.
+consists of estimator classes.
 All learning algorithms for classification, regression and clustering
-are offered as such objects,
+are offered as such classes,
 but so are feature extraction, feature selection and dimensionality reduction.
+The three latter types of estimator are examples of a subtype
+called \textit{transformers}:
+in this case, both input and output are NumPy arrays representing samples,
+but commonly with different numbers of columns.
+Transformers that perform feature extraction on e.g.\ free text
+are an exception in that their input is not usually a NumPy array.
 
@@
 no distinction is made in the type hierarchy
-between what is variously called an estimator or a \textit{learner}
-in the machine learning literature,
-and the \textit{model} that results from applying an estimator to data.
+between an estimator and the model that results from applying it to data.
 When the \texttt{fit} method has been called,
-an estimator object serves dual purpose as a model
-(and is referred to as such in the documentation).
+an estimator object serves dual purpose as a model.
 
-An example to clarify the partial application style
-is the following code snippet.
-If $X$ is a dataset in the format described below and $k$ is some integer,
+As an example to clarify the partial application style,
+consider a linear support vector machine (SVM).
+% Rewrite example using LinearSVC
+If $X$ is a dataset in the format described above and $k$ is some integer,
 then the Python code to train a $k$-means clustering model on it
@@
 
-\section{Pipelines and model selection}
+\section{Pipelines}
 
-A distinguishing feature of scikit-learn estimators
-is that they can be composed into \texttt{Pipeline} objects.
+A feature of scikit-learn estimators
+that makes effective hyperparameter optimization possible
+is the fact that they can be composed into \texttt{Pipeline} objects.
 Not only do such objects combine typical machine learning workflows

### 2013-05-12 paper.tex
start section about nested pipelines

@@
 
+\subsection{Recursive pipelines}
+
+In some cases, it is useful to build recursive pipelines,
+and scikit-learn has support for these:
+pipelines are themselves estimators,
+and can thus be nested inside larger pipelines.
+
+While direct use of this idiom would not have any benefit
+over using the individual stages of the sub-pipeline directly,
+the recursion support shines when in more complex situations
+that involve the construction of estimators using \textit{meta-estimators}.
+These are objects in the scikit-learn API
+that wrap an estimator to give it additional functionality.
+An example of meta-estimators are object of the class
+\texttt{OneVsRestClassifier},
+which can be used to construct a multi-label classifier
+\citep{tsoumakas2007multi}
+from an ordinary binary classifier object
+by constructing one binary classifier per class
+(a strategy variously called one-vs.-the-rest,
+one-vs.-all or binary relevance).
+
+% TODO insert example from http://stackoverflow.com/q/16437022/166749 here
+
 \section{Implementation}

### 2013-05-13 paper.tex
two-column layout for more realistic view of final paper
copyedit

@@
-\documentclass[a4paper]{article}
+\documentclass[a4paper,twocolumn]{article}
 
@@
 for handling the problem of hyperparameter optimization,
-or \textit{model selection} as it is called in the statistics literature.
+also known as \textit{model selection} (\citealp{elemstatlearn}:219--257).
 
@@
 
-While direct use of this idiom would not have any benefit
+Direct use of this idiom would not have any benefit
 over using the individual stages of the sub-pipeline directly,
-the recursion support shines when in more complex situations
+as both approaches would simply produce a linear pipeline.
+However, the recursion support shines in more complex situations
 that involve the construction of estimators using \textit{meta-estimators}.

### 2013-05-13 paper.tex
add example from StackOverflow

@@
 (a strategy variously called one-vs.-the-rest,
-one-vs.-all or binary relevance).
+one-vs.-all or binary relevance):
 
-% TODO insert example from http://stackoverflow.com/q/16437022/166749 here
+\begin{verbatim}
+ovr_svm = OneVsRestClassifier(LinearSVC())
+\end{verbatim}
+
+The \texttt{ovr\_svm} object can be passed to a hyperparameter search
+producedure directly to determine parameters such as $C$,
+but the syntax to do that has changed.
+The object under consideration is no longer a \texttt{LinearSVC}
+but a \texttt{OneVsRestClassifier},
+which exposes the former object as its public attribute \texttt{estimator}.
+Double underscore syntax must be used to address $C$,
+i.e.\ \texttt{estimator\_\_C}.
+
+To make this example more interesting,
+feature selection using $\chi^2$ tests can be added to the SVM,
+such that each of the separate binary classification problems
+gets its own set of features.
+This can be done by wrapping a pipeline in an OvR object:
+
+\begin{verbatim}
+svm_chi2 = Pipeline(["chi2", SelectKBest(chi2)],
+                    ["svm", LinearSVC()])
+ovr_svm = OneVsRestClassifier(svm_chi2)
+\end{verbatim}
+
+Now, the parameter $k$ specifying number of features to be kept
+in each $\chi^2$ feature selection step
+is exposed by \texttt{ovr\_svm} as \texttt{estimator\_\_chi2\_\_k},
+mirroring the nesting of estimators in meta-estimators,
+and can be optimized by grid search or random search.
+
+Taking the example one step further to arrive at recursive pipelines,
+the \texttt{ovr\_svm} object can be combined
+with a \textsf{tf--idf} document vectorizer
+to obtain a full document classification pipeline
+that takes a stream of text files, paired with class labels, as input:
+
+\begin{verbatim}
+doc_clf = Pipeline(["tfidf", TfidfVectorizer()],
+                   ["clf", ovr_svm])
+\end{verbatim}
+
+The double underscore syntax again serves to address attributes
+in this triply nested structure,
+so the linear SVM's $C$ parameter
+is now called \texttt{clf\_\_estimator\_\_svm\_\_C}.
+
+Following the Python mantra that ``explicit is better than implicit,''
+the implementers of scikit-learn have resisted the temptation
+to implement shortcuts in the notation,
+for example to allow addressing of attributes
+of the underlying estimator of the OvR classifier
+without the intermediate \texttt{estimator}.
+Doing so might require conflict resolution
+when an attribute of the meta-estimator object itself
+has the same name as an attribute on the wrapped estimator.
+The rules for such conflict resolution would need to be
+designed, implemented, tested, and, equally important, documented
+such as to be understandable for users.
+So, the cost of ``syntactic sugar'' was avoided
+at the expense of some typing by the user and some loss of elegance.
 

### 2013-05-13 paper.tex
tone down criticism of CPython

@@
 at least in the CPython reference implementation targeted by scikit-learn,
-is downright weak.
-The interpreter supports multithreading, but due to the ``Global Interpreter Lock'' (GIL),
+is one of the weak spots in the language and its libraries.
+The interpreter supports multithreading,
+but due to the ``Global Interpreter Lock'' (GIL),
 no two threads in a process may at the same time execute Python code.
@@
 even two embarrassingly parallel problems,
-such as fitting and evaluation multiple models in a grid search procedure.
+such as fitting and evaluating multiple models in a grid search procedure.
 The workaround is to use Python's multiprocessing module,

### 2013-05-13 paper.tex
cleanup and ruthless propaganda

@@
 of ``classic'' machine learning algorithms
-with a consistent API.
+with a consistent and clear API.
 This paper explores the role that such an API plays
@@
 but such usage is not at present possible and may never be.
-Passing C or Cython-compiled functions would require an API for the kernel method
-to be implemented at the C or Cython level,
+Passing C or Cython-compiled functions would require an API
+for the kernel method to be implemented at the C or Cython level,
 only to be used from compiled code.
@@
 because of the runtime checks needed for each invocation of the kernel function.
-(The current solution in scikit-learn is to let the user pass not a kernel function,
-but a precomputed Gram matrix for an entire set of objects.
+(The current solution in scikit-learn is to let the user pass
+not a kernel function, but a precomputed Gram matrix
+for an entire set of objects.
 The linear, polynomial and RBF kernels are treated as special cases in
-the \textsf{LIBSVM}-based SVM learner, where they are implemented in C{}\verb!++!.)
+the \textsf{LIBSVM}-based SVM learner,
+where they are implemented in C{}\verb!++!.)
 

### 2013-05-13 paper.tex
elaborate on the H -> X -> Y notation

@@
 \section{The hyperparameter optimization problem}
+\label{hyperparams}
 
@@
 Construction and application of an estimator object
-proceeds in a way that is somewhat reminiscent of partial function application.
+proceeds in a way that is similar to partial function application,
+mirroring the $L : \mathcal{H} \to \mathcal{X} \to \mathcal{Y}$ notation
+given for the supervised learning problem in Section~\ref{hyperparams}.
 The constructor, \texttt{\_\_init\_\_}\footnote{

### 2013-05-13 paper.tex
Xtr etc. notation not actually used yet, comment out

@@
 
-In the remainder of the paper, we use the following notation. Training data are
-denoted by \Xtr ~and test data (i.e., unseen data to which the learning
-algorithm need be able to generalize) by \Xte.  Following the discussion above,
-we represent \Xtr ~and \Xte ~by 2-dimensional NumPy arrays or SciPy sparse
-matrices. Using NumPy/SciPy's notation, \Xtr\texttt{[i, j]} denotes the
-$i^\textrm{th}$ training sample's $j^\textrm{th}$ feature. For supervised learning tasks,
-such as classification or regression, we denote the training and test labels by
-\ytr ~and \yte, respectively. We typically store \ytr ~and \yte ~as
-1-dimensional NumPy arrays (\ytr\texttt{[i]} denotes the $i^\textrm{th}$ training
-sample's label value).
+%In the remainder of the paper, we use the following notation. Training data are
+%denoted by \Xtr ~and test data (i.e., unseen data to which the learning
+%algorithm need be able to generalize) by \Xte.  Following the discussion above,
+%we represent \Xtr ~and \Xte ~by 2-dimensional NumPy arrays or SciPy sparse
+%matrices. Using NumPy/SciPy's notation, \Xtr\texttt{[i, j]} denotes the
+%$i^\textrm{th}$ training sample's $j^\textrm{th}$ feature. For supervised learning tasks,
+%such as classification or regression, we denote the training and test labels by
+%\ytr ~and \yte, respectively. We typically store \ytr ~and \yte ~as
+%1-dimensional NumPy arrays (\ytr\texttt{[i]} denotes the $i^\textrm{th}$ training
+%sample's label value).
 

### 2013-05-13 paper.tex
elaborate on public attribute status of learned parameters

@@
 
-The algorithmic workhorse of most estimators, and one of the few methods
-that all estimators have in common, is called \texttt{fit}.
+Instead, the actual learning is performed by a method called \texttt{fit}.
+This is one of the very few methods that all estimators have in common.
 Invoking this method causes the estimator to learn from a training set,
@@
 is to determine (via a learning algorithm) model-specific parameters
-from a training set and set these as attributes on the estimator object,
-which can then be used to transform data or make predictions.
-As in other object-oriented machine learning libraries
+from a training set and set these as attributes on the estimator object.
+To distinguish learned parameters from hyperparameters,
+the former always have a trailing underscore appended to their name;
+e.g., the weights of a linear model are stored in attributes
+\texttt{coef\_} and \texttt{intercept\_}
+(the underscore is suggestive of a hat,
+commonly used to denote estimates in mathematical notation: $\hat{w}$).
+
+After learning, the estimator object serves dual purpose as a model
+and can then be used to transform data or make predictions;
+as in other object-oriented machine learning libraries
 (e.g.\ Weka's Java API)
@@
 between an estimator and the model that results from applying it to data.
-When the \texttt{fit} method has been called,
-an estimator object serves dual purpose as a model.
+While it would be possible to decouple models and estimators
+at the class hierarchy level,
+doing so would complicate APIs for online learning.
 

### 2013-05-13 paper.tex
mention of ensemble methods no longer needed
formalize model selection and copyedit

@@
 
-where $\mathcal{H}$ is the set of all possible \textit{hyperparameter} settings
+Here, $\mathcal{H}$ is the set of all possible \textit{hyperparameter} settings
 for the estimator in question.\footnote{
@@
 However, in recent years, algorithms have been published
-that systematically explore spaces of possible hyperparameter settings.
+that systematically explore spaces of possible hyperparameter settings,
+a problem that can be formalized as learning and optimization
+within the set of subset of hyperparameter space:
+
+\[
+  L' : 2^\mathcal{H} \to \mathcal{X} \to \mathcal{Y}
+\]
+
 The problem of hyperparameter tuning
@@
 The present paper does not go into hyperparameter optimization algorithms
-in depth, focusing instead on API support for using them.
+in depth, but instead focuses on the design of an API
+that facilitates the use of such algorithms.
 
@@
 on the SVM themselves.
-(The double underscore syntax is not special to pipelines;
-it is also employed in ensemble estimators such as random forests.)
 

### 2013-06-04 paper.tex
Use LLNCS documentclass

--- 2013-05-13 paper.tex
@@
-\documentclass[a4paper,twocolumn]{article}
+\documentclass{llncs}
 
+\usepackage[utf8]{inputenc}
 \usepackage{amsmath}
-\usepackage{amssymb}
 \usepackage{hyperref}
@@
 % working title, please change
-\title{APIs for hyperparameter optimization in scikit-learn}
+\title{Python as a machine learning language:
+       experiences from the scikit-learn project}
 
-% Add your name and affiliation here.
-% Don't worry too much about the formatting,
-% as we'll have to redo that for the LNCS format.
-\author{Mathieu Blondel \\
-        Graduate School of System Informatics \\
-        Kobe University \\
-        \small{\texttt{mblondel@ai.cs.kobe-u.ac.jp}}
-      \and
-        Lars Buitinck \\
-        Informatics Institute \\
-        University of Amsterdam \\
-        \small{\texttt{L.J.Buitinck@uva.nl}}
-      \and
-        Jake Vanderplas \\
-        Astronomy Department \\
-        University of Washington \\
-        \small{\texttt{vanderplas@astro.washington.edu}}
-}
+\author{Mathieu Blondel\inst{1} \and
+        Lars Buitinck\inst{2} \and
+        Gilles Louppe\inst{3} \and
+        Jake Vanderplas\inst{4}}
+
+\institute{Graduate School of System Informatics, Kobe University \and
+           Informatics Institute, University of Amsterdam \and
+           Dept. of EE \& CS, University of Liège \and
+           Astronomy Department, University of Washington}
 
@@
 in both the programmers' interface (API) and the implementation
-of scikit-learn,
-with a focus on the problem of hyperparameter optimization.
-%It also highlights obstacles faced by users and implementers
-%of machine learning algorithms in the Python language.
+of scikit-learn.
+It also highlights obstacles faced by users and implementers
+of machine learning algorithms in the Python language.
 \end{abstract}
 
-\section{Introduction}
+\section{Project vision}
 
-\subsection{Project vision}
+\subsection{General principles}
 
@@
 of ``classic'' machine learning algorithms
-with a consistent and clear API.
-This paper explores the role that such an API plays
-for handling the problem of hyperparameter optimization,
-also known as \textit{model selection} (\citealp{elemstatlearn}:219--257).
-
-The rest of this section explains some fundamental design choices
-made throughout scikit-learn.
-% TODO and the rest of the paper...
+with a consistent API.
 
@@
 
-%In the remainder of the paper, we use the following notation. Training data are
-%denoted by \Xtr ~and test data (i.e., unseen data to which the learning
-%algorithm need be able to generalize) by \Xte.  Following the discussion above,
-%we represent \Xtr ~and \Xte ~by 2-dimensional NumPy arrays or SciPy sparse
-%matrices. Using NumPy/SciPy's notation, \Xtr\texttt{[i, j]} denotes the
-%$i^\textrm{th}$ training sample's $j^\textrm{th}$ feature. For supervised learning tasks,
-%such as classification or regression, we denote the training and test labels by
-%\ytr ~and \yte, respectively. We typically store \ytr ~and \yte ~as
-%1-dimensional NumPy arrays (\ytr\texttt{[i]} denotes the $i^\textrm{th}$ training
-%sample's label value).
-
-\section{The hyperparameter optimization problem}
-\label{hyperparams}
-
-The focus of many machine learning toolkits is the \textit{learning} task,
-which, at least in the supervised case,
-can be formalized as the problem of finding a function
-$\mathsf{model} : \mathcal{X} \to \mathcal{Y}$,
-where $\mathcal{X}$ is some set of inputs
-and $\mathcal{Y}$ some set of targets
-($\mathbb{R}^k$ in regression, $\{1,\dots,K\}$ in classification).
-An algorithm that finds such a function is called a \textit{learner}
-% TODO: cite Mitchell 1998
-or alternatively an \textit{estimator}
-(in the scikit-learn documentation, and therefore in this paper).
-Using function currying, its function signature is
-
-\[
-  L : \mathcal{H} \to \mathcal{X} \to \mathcal{Y}
-\]
-
-where $\mathcal{H}$ is the set of all possible \textit{hyperparameter} settings
-for the estimator in question.\footnote{
-  We disregard the fact that different estimators
-  have widely varying sets of hyperparameters for notational simplicity.}
-Hyperparameters include, for example,
-regularization strength of linear models or SVMs,
-choice of loss function for generic linear model learners,
-or the number of features to keep in a feature selection method.
-They are often treated as user-settable parameters
-and not part of the learning problem proper,
-either because they would affect the learning problem
-in a way that would make it intractable \citep{bergstra2012}
-or because their effects cannot properly be measured
-in terms of a loss function on a training set.
-
-However, in recent years, algorithms have been published
-that systematically explore spaces of possible hyperparameter settings.
-The problem of hyperparameter tuning
-is crucial to achieving good generalization performance in machine learning.
-The present paper does not go into hyperparameter optimization algorithms
-in depth, focusing instead on API support for using them.
+In the remainder of the paper, we use the following notation. Training data are
+denoted by \Xtr ~and test data (i.e., unseen data to which the learning
+algorithm needs to be able to generalize) by \Xte.  Following the discussion above,
+we represent \Xtr ~and \Xte ~by 2-dimensional NumPy arrays or SciPy sparse
+matrices. Using NumPy/SciPy's notation, \Xtr\texttt{[i, j]} denotes the
+$i^\textrm{th}$ training sample's $j^\textrm{th}$ feature. For supervised learning tasks,
+such as classification or regression, we denote the training and test labels by
+\ytr ~and \yte, respectively. We typically store \ytr ~and \yte ~as
+1-dimensional NumPy arrays (\ytr\texttt{[i]} denotes the $i^\textrm{th}$ training
+sample's label value).
 
@@
 The bulk of the scikit-learn application programming interface
-consists of estimator classes.
+consists of so-called \textit{estimator} objects.
 All learning algorithms for classification, regression and clustering
-are offered as such classes,
-but so are feature extraction, feature selection and dimensionality reduction.
-The three latter types of estimator are examples of a subtype
-called \textit{transformers}:
-in this case, both input and output are NumPy arrays representing samples,
-but commonly with different numbers of columns.
-Transformers that perform feature extraction on e.g.\ free text
-are an exception in that their input is not usually a NumPy array.
+are offered as such objects,
+but so are feature extraction, feature selection and dimensionality reduction
+algorithms.
 
 Construction and application of an estimator object
-proceeds in a way that is similar to partial function application,
-mirroring the $L : \mathcal{H} \to \mathcal{X} \to \mathcal{Y}$ notation
-given for the supervised learning problem in Section~\ref{hyperparams}.
+proceeds in a way that is somewhat reminiscent of partial function application.
 The constructor, \texttt{\_\_init\_\_}\footnote{
@@
 no distinction is made in the type hierarchy
-between an estimator and the model that results from applying it to data.
+between what is variously called an estimator or a \textit{learner}
+in the machine learning literature,
+and the \textit{model} that results from applying an estimator to data.
 When the \texttt{fit} method has been called,
-an estimator object serves dual purpose as a model.
+an estimator object serves dual purpose as a model
+(and is referred to as such in the documentation).
 
-As an example to clarify the partial application style,
-consider a linear support vector machine (SVM).
-% Rewrite example using LinearSVC
-If $X$ is a dataset in the format described above and $k$ is some integer,
+An example to clarify the partial application style
+is the following code snippet.
+If $X$ is a dataset in the format described below and $k$ is some integer,
 then the Python code to train a $k$-means clustering model on it
@@
 
-\section{Pipelines}
+\section{Pipelines and model selection}
 
-A feature of scikit-learn estimators
-that makes effective hyperparameter optimization possible
-is the fact that they can be composed into \texttt{Pipeline} objects.
-Not only do such objects combine typical machine learning workflows
+A distinguishing feature of scikit-learn estimators
+is that they can be composed into \texttt{Pipeline} objects.
+Such objects combine typical machine learning workflows
 (extracting features, centering/scaling/normalization,
@@
 and the best performing set of hyperparameters is stored
-on the \texttt{GridSearchCV} object; again, as a public attribute.
-
-\subsection{Recursive pipelines}
-
-In some cases, it is useful to build recursive pipelines,
-and scikit-learn has support for these:
-pipelines are themselves estimators,
-and can thus be nested inside larger pipelines.
-
-Direct use of this idiom would not have any benefit
-over using the individual stages of the sub-pipeline directly,
-as both approaches would simply produce a linear pipeline.
-However, the recursion support shines in more complex situations
-that involve the construction of estimators using \textit{meta-estimators}.
-These are objects in the scikit-learn API
-that wrap an estimator to give it additional functionality.
-An example of meta-estimators are object of the class
-\texttt{OneVsRestClassifier},
-which can be used to construct a multi-label classifier
-\citep{tsoumakas2007multi}
-from an ordinary binary classifier object
-by constructing one binary classifier per class
-(a strategy variously called one-vs.-the-rest,
-one-vs.-all or binary relevance):
-
-\begin{verbatim}
-ovr_svm = OneVsRestClassifier(LinearSVC())
-\end{verbatim}
-
-The \texttt{ovr\_svm} object can be passed to a hyperparameter search
-producedure directly to determine parameters such as $C$,
-but the syntax to do that has changed.
-The object under consideration is no longer a \texttt{LinearSVC}
-but a \texttt{OneVsRestClassifier},
-which exposes the former object as its public attribute \texttt{estimator}.
-Double underscore syntax must be used to address $C$,
-i.e.\ \texttt{estimator\_\_C}.
-
-To make this example more interesting,
-feature selection using $\chi^2$ tests can be added to the SVM,
-such that each of the separate binary classification problems
-gets its own set of features.
-This can be done by wrapping a pipeline in an OvR object:
-
-\begin{verbatim}
-svm_chi2 = Pipeline(["chi2", SelectKBest(chi2)],
-                    ["svm", LinearSVC()])
-ovr_svm = OneVsRestClassifier(svm_chi2)
-\end{verbatim}
-
-Now, the parameter $k$ specifying number of features to be kept
-in each $\chi^2$ feature selection step
-is exposed by \texttt{ovr\_svm} as \texttt{estimator\_\_chi2\_\_k},
-mirroring the nesting of estimators in meta-estimators,
-and can be optimized by grid search or random search.
-
-Taking the example one step further to arrive at recursive pipelines,
-the \texttt{ovr\_svm} object can be combined
-with a \textsf{tf--idf} document vectorizer
-to obtain a full document classification pipeline
-that takes a stream of text files, paired with class labels, as input:
-
-\begin{verbatim}
-doc_clf = Pipeline(["tfidf", TfidfVectorizer()],
-                   ["clf", ovr_svm])
-\end{verbatim}
-
-The double underscore syntax again serves to address attributes
-in this triply nested structure,
-so the linear SVM's $C$ parameter
-is now called \texttt{clf\_\_estimator\_\_svm\_\_C}.
-
-Following the Python mantra that ``explicit is better than implicit,''
-the implementers of scikit-learn have resisted the temptation
-to implement shortcuts in the notation,
-for example to allow addressing of attributes
-of the underlying estimator of the OvR classifier
-without the intermediate \texttt{estimator}.
-Doing so might require conflict resolution
-when an attribute of the meta-estimator object itself
-has the same name as an attribute on the wrapped estimator.
-The rules for such conflict resolution would need to be
-designed, implemented, tested, and, equally important, documented
-such as to be understandable for users.
-So, the cost of ``syntactic sugar'' was avoided
-at the expense of some typing by the user and some loss of elegance.
+on the \texttt{GridSearchCV} object, again, as a public attribute.
 
@@
 at least in the CPython reference implementation targeted by scikit-learn,
-is one of the weak spots in the language and its libraries.
-The interpreter supports multithreading,
-but due to the ``Global Interpreter Lock'' (GIL),
+is downright weak.
+The interpreter supports multithreading, but due to the ``Global Interpreter Lock'' (GIL),
 no two threads in a process may at the same time execute Python code.
@@
 even two embarrassingly parallel problems,
-such as fitting and evaluating multiple models in a grid search procedure.
+such as fitting and evaluation multiple models in a grid search procedure.
 The workaround is to use Python's multiprocessing module,
@@
 but such usage is not at present possible and may never be.
-Passing C or Cython-compiled functions would require an API
-for the kernel method to be implemented at the C or Cython level,
+Passing C or Cython-compiled functions would require an API for the kernel method
+to be implemented at the C or Cython level,
 only to be used from compiled code.
@@
 because of the runtime checks needed for each invocation of the kernel function.
-(The current solution in scikit-learn is to let the user pass
-not a kernel function, but a precomputed Gram matrix
-for an entire set of objects.
+(The current solution in scikit-learn is to let the user pass not a kernel function,
+but a precomputed Gram matrix for an entire set of objects.
 The linear, polynomial and RBF kernels are treated as special cases in
-the \textsf{LIBSVM}-based SVM learner,
-where they are implemented in C{}\verb!++!.)
+the \textsf{LIBSVM}-based SVM learner, where they are implemented in C{}\verb!++!.)
 

### 2013-06-05 paper.tex
Abstract + Introduction

@@
 \usepackage[round]{natbib}
+\usepackage{xcolor}
 
@@
 \title{Python as a machine learning language:
-       experiences from the scikit-learn project}
+       experiences from the \textit{scikit-learn} project}
 
@@
            Informatics Institute, University of Amsterdam \and
-           Dept. of EE \& CS, University of Liège \and
+           Department of EE \& CS, University of Liège \and
            Astronomy Department, University of Washington}
@@
 \begin{abstract}
-The scikit-learn project aims to create an open source
-machine learning library for the Python programming language,
-offering performant implementations of well-established learning algorithms,
-and tying in with the ecosystem of scientific software
-that has evolved around the NumPy and SciPy packages.
-Since the start of the project in 2009,
-it has gathered a sizeable user base.
-
-This paper discusses design choices made
-in both the programmers' interface (API) and the implementation
-of scikit-learn.
-It also highlights obstacles faced by users and implementers
-of machine learning algorithms in the Python language.
+The \textit{scikit-learn} project is an increasingly popular open source machine
+learning library designed to be simple and efficient, accessible to non-experts,
+and reusable in various contexts. In this paper, we present and discuss our
+design choices for the application programming interface (API) of the project.
+In particular, we describe the \textit{estimator} interface shared by all
+learning and processing units in the library and then discuss its advantages in
+terms of composition and reusability. The paper also comments on implementation
+details specific to the Python ecosystem and highlights obstacles faced by users
+and developers of machine learning algorithms.
 \end{abstract}
 
-\section{Project vision}
+\section{The \textit{scikit-learn} project}
 
-\subsection{General principles}
+The \textit{scikit-learn} project~\citep{pedregosa2011} is an open source
+machine learning library for the Python programming language. The ambition of
+the project is to provide  effective and long-established machine learning tools
+within a programming environment that is easy-to-use, efficient, accessible to
+non-machine learning experts and reusable in various scientific areas. In
+itself, the project is a library of machine learning tools including, among
+others, learning algorithms, model selection tools and preprocessing procedures
+rather than a novel domain-specific programming language. From its core, the
+library has been designed to tie in with the set of numeric and scientific
+packages centered around the Numpy and SciPy~\citep{vanderwalt2011,
+varoquaux2013scipy} libraries. As such, \textit {scikit-learn} is intended to be
+used within Python scripts and requires some Python programming knowledge. At
+the time, it offers no command line or graphical user interface.
 
-% XXX the following is more of a general introduction than a project vision
-scikit-learn \citep{pedregosa2011} is an open source machine learning library
-for the Python programming language,
-intended to tie in with the set of numeric and scientific packages
-centered around the NumPy and SciPy libraries
-\Citep{vanderwalt2011, varoquaux2013scipy}.
-It is a library, rather than a novel (domain-specific) programming language,
-intended to be called from Python programs;
-at present, it offers no command line interface or GUI.
+Started in 2009, \textit{scikit-learn} counts a large team of volunteer
+international developers mainly consisting of scientists and researchers from
+various areas (e.g., machine learning, natural language processing or
+neuroscience). The project also greatly benefits from many occasional
+contributors proposing small code and/or documentation fixes or improvements.
+Because of this large base of developers, important efforts are demanded to make
+the project maintainable. In particular, we emphasize on high-quality and well-documented
+code  in order to ease future developments by external contributors.
 
-The scikit-learn project started in 2009 and has since been developed
-by a team largely consisting of scientists
-(both from computer science and other fields).
-Its popularity can be gauged from various sources:
-% TODO: list volume on the mailing list, number of contributors,
-% number of issues opened and closed, number of forks, number of surveys filled in.
-at the time of writing,
-the popular programmer's Q~\&{}A site StackOverflow\footnote{
-  \url{http://stackoverflow.com/questions/tagged/scikit-learn}}
-% update the following figure when paper is finalized
-lists 322 scikit-learn-related questions.
-The version control logs for the development version of scikit-learn
-% ddaa494c116e3c16bf032003c5cccbed851733d2
-show 183 contributors to the project's source code and documentation
-(after manual deduplication).
+Within the past few years, the project has proven to be valuable to an  ever-
+increasing user base. Its popularity can be gauged from various indicators,
+including hundreds of citations in scientific papers, successes in various
+machine learning challenges, or statistics from our repositories and mailing
+lists.
 
-The main goal of the scikit-learn project
-is to provide high-quality implementations
-of ``classic'' machine learning algorithms
-with a consistent API.
+\textcolor{red}{The rest of this text is organized as follows: in Section 2, ...}
+
+% % XXX the following is more of a general introduction than a project vision
+% scikit-learn \citep{pedregosa2011} is an open source machine learning library
+% for the Python programming language,
+% intended to tie in with the set of numeric and scientific packages
+% centered around the NumPy and SciPy libraries
+% \Citep{vanderwalt2011, varoquaux2013scipy}.
+% It is a library, rather than a novel (domain-specific) programming language,
+% intended to be called from Python programs;
+% at present, it offers no command line interface or GUI.
+
+% The scikit-learn project started in 2009 and has since been developed
+% by a team largely consisting of scientists
+% (both from computer science and other fields).
+% Its popularity can be gauged from various sources:
+% % TODO: list volume on the mailing list, number of contributors,
+% % number of issues opened and closed, number of forks, number of surveys filled in.
+% at the time of writing,
+% the popular programmer's Q~\&{}A site StackOverflow\footnote{
+%   \url{http://stackoverflow.com/questions/tagged/scikit-learn}}
+% % update the following figure when paper is finalized
+% lists 322 scikit-learn-related questions.
+% The version control logs for the development version of scikit-learn
+% % ddaa494c116e3c16bf032003c5cccbed851733d2
+% show 183 contributors to the project's source code and documentation
+% (after manual deduplication).
+
+% The main goal of the scikit-learn project
+% is to provide high-quality implementations
+% of ``classic'' machine learning algorithms
+% with a consistent API.
 
@@
 
-In the remainder of the paper, we use the following notation. Training data are
-denoted by \Xtr ~and test data (i.e., unseen data to which the learning
-algorithm needs to be able to generalize) by \Xte.  Following the discussion above,
-we represent \Xtr ~and \Xte ~by 2-dimensional NumPy arrays or SciPy sparse
-matrices. Using NumPy/SciPy's notation, \Xtr\texttt{[i, j]} denotes the
-$i^\textrm{th}$ training sample's $j^\textrm{th}$ feature. For supervised learning tasks,
-such as classification or regression, we denote the training and test labels by
-\ytr ~and \yte, respectively. We typically store \ytr ~and \yte ~as
-1-dimensional NumPy arrays (\ytr\texttt{[i]} denotes the $i^\textrm{th}$ training
-sample's label value).
+% In the remainder of the paper, we use the following notation. Training data are
+% denoted by \Xtr ~and test data (i.e., unseen data to which the learning
+% algorithm needs to be able to generalize) by \Xte.  Following the discussion above,
+% we represent \Xtr ~and \Xte ~by 2-dimensional NumPy arrays or SciPy sparse
+% matrices. Using NumPy/SciPy's notation, \Xtr\texttt{[i, j]} denotes the
+% $i^\textrm{th}$ training sample's $j^\textrm{th}$ feature. For supervised learning tasks,
+% such as classification or regression, we denote the training and test labels by
+% \ytr ~and \yte, respectively. We typically store \ytr ~and \yte ~as
+% 1-dimensional NumPy arrays (\ytr\texttt{[i]} denotes the $i^\textrm{th}$ training
+% sample's label value).
 

### 2013-06-05 paper.tex
Added some todos

@@
 \usepackage{hyperref}
-\usepackage[round]{natbib}
+\usepackage[authoryear]{natbib}
 \usepackage{xcolor}
@@
 code  in order to ease future developments by external contributors.
+\textcolor{red}{TODO: Mention some good coding practices (automatic testing, code
+reviews, etc)}
 
@@
 including hundreds of citations in scientific papers, successes in various
-machine learning challenges, or statistics from our repositories and mailing
-lists.
+machine learning challenges, or statistics\footnote{\textcolor{red}{Cite some
+statistics here.}} from our repositories and mailing lists.
 
-\textcolor{red}{The rest of this text is organized as follows: in Section 2, ...}
+\textcolor{red}{TODO: The rest of this text is organized as follows: in Section 2, ...}
 
@@
 % with a consistent API.
+
+\section{API}
 
@@
 % sample's label value).
-
-\section{Application programming interface (API)}
 
@@
 
-\section{Pipelines and model selection}
+\subsection{Composition}
 
@@
 double underscores.)
+
+\subsection{Model selection}
 

### 2013-06-05 paper.tex
Some rewordings

@@
 design choices for the application programming interface (API) of the project.
-In particular, we describe the \textit{estimator} interface shared by all
+In particular, we describe our core \textit{estimator} interface shared by all
 learning and processing units in the library and then discuss its advantages in
@@
 details specific to the Python ecosystem and highlights obstacles faced by users
-and developers of machine learning algorithms.
+and developers of the library.
 \end{abstract}
@@
 others, learning algorithms, model selection tools and preprocessing procedures
-rather than a novel domain-specific programming language. From its core, the
-library has been designed to tie in with the set of numeric and scientific
+rather than being a novel domain-specific programming language. From its core,
+the library has been designed to tie in with the set of numeric and scientific
 packages centered around the Numpy and SciPy~\citep{vanderwalt2011,
@@
 international developers mainly consisting of scientists and researchers from
-various areas (e.g., machine learning, natural language processing or
+various fields (e.g., machine learning, natural language processing or
 neuroscience). The project also greatly benefits from many occasional
 contributors proposing small code and/or documentation fixes or improvements.
-Because of this large base of developers, important efforts are demanded to make
+Because of this large base of developers, important efforts are deployed to make
 the project maintainable. In particular, we emphasize on high-quality and well-documented
@@
 
-Within the past few years, the project has proven to be valuable to an  ever-
+Within the past few years, the project has proven to be valuable to an  ever
 increasing user base. Its popularity can be gauged from various indicators,
 including hundreds of citations in scientific papers, successes in various
-machine learning challenges, or statistics\footnote{\textcolor{red}{Cite some
-statistics here.}} from our repositories and mailing lists.
+machine learning challenges, or significant statistics\footnote{\textcolor{red}{Cite some
+statistics here.}} derived from our repositories and mailing lists.
 

### 2013-06-05 paper.tex
Intro of section 2

@@
 design choices for the application programming interface (API) of the project.
-In particular, we describe our core \textit{estimator} interface shared by all
+In particular, we describe the core \textit{estimator} interface shared by all
 learning and processing units in the library and then discuss its advantages in
@@
 \section{API}
+
+In this section, we discuss the general principles of the application programming
+interface (API) of \textit{scikit-learn}. We first specify the representation
+of data and then proceed by describing the interface shared by all learning
+processing units of the library. We then discuss the composition and reusability
+advantages of that interface.
 

### 2013-06-05 paper.tex
Data representation

@@
 the project is to provide  effective and long-established machine learning tools
-within a programming environment that is easy-to-use, efficient, accessible to
-non-machine learning experts and reusable in various scientific areas. In
-itself, the project is a library of machine learning tools including, among
-others, learning algorithms, model selection tools and preprocessing procedures
-rather than being a novel domain-specific programming language. From its core,
-the library has been designed to tie in with the set of numeric and scientific
-packages centered around the Numpy and SciPy~\citep{vanderwalt2011,
-varoquaux2013scipy} libraries. As such, \textit {scikit-learn} is intended to be
-used within Python scripts and requires some Python programming knowledge. At
-the time, it offers no command line or graphical user interface.
+within a programming environment that is accessible to non-machine learning
+experts, efficient, and reusable in various scientific areas. In itself, the
+project is a library of machine learning tools  rather than  a novel domain-
+specific programming language. Among other things, the library includes
+classical learning algorithms, model selection tools and preprocessing
+procedures. From its core, the library has been designed to tie in with the set
+of numeric and scientific packages centered around the Numpy and
+SciPy~\citep{vanderwalt2011, varoquaux2013scipy} libraries. As such, \textit
+{scikit-learn} is intended to be used within Python scripts and requires some
+Python programming knowledge. At the time, it offers no command line or
+graphical user interface.
 
@@
 
-\subsection{Data formats}
+\subsection{Data representation}
 
-For a machine learning library, one of the most important decisions we had to
-make is how to represent data.  Rather than reinventing the wheel, we opted for
-NumPy multidimensional arrays \Citep{vanderwalt2011} for dense data and SciPy
-sparse matrices for sparse data.  While those may seem like a ``bare''
-representation of data when compared to more object-oriented representations
-(Weka is an example of this style of representation, \citealp{hall2009weka}),
-it brings the prime advantage of allowing us to rely on NumPy and SciPy's
-\textit{vector operations} which are often orders of magnitude faster
-than the corresponding Python loops,
-while at the same time keeping the code short and readable.
-Conversion to NumPy and SciPy formats is usually easy and many scientific users
-of Python will already be used to such formats, since they are pervasive in
-other scientific Python packages.  For tasks where input is likely to consist
-of text files or semi-structured objects, we provide ``vectorizers'' -- objects
-that efficiently convert such data to the NumPy or SciPy formats.
+In most machine learning tasks, data is modeled as a set of variables.  For
+example, in a classification or regression task, the goal is to find a mapping
+between input variables $X_1, ..., X_p$ -- called attributes or features -- and
+some output variable $Y$. A sample is then defined as a tuple of values $(x_1,
+..., x_p, y)$ of these variables and a dataset as a collection of such samples.
+As such, matrices of numerical values\footnote{In the case of categorical
+variables, numerical values can be obtained using various encoding schemes, such
+as one-hot encoding.} constitute a natural representation used almost
+unanimously within all areas of machine learning. In particular, each row in the
+matrix then corresponds to a sample of the dataset and each column to one of the
+variables of the problem.
+
+In \textit{scikit-learn}, we chose a representation of data that is as close as
+possible to the matrix representation: datasets are encoded as NumPy
+multidimensional arrays for dense data and as SciPy sparse matrices for sparse
+data. While those may seem like bare representations of data when compared to
+more object-oriented representations\footnote{Weka~\citep{hall2009weka} is an
+example of a toolkit-specific object-oriented data representation.}, they bring
+the prime advantage of allowing us to rely on all NumPy and SciPy vector
+operations -- often orders of magnitude faster that the corresponding Python
+loops -- while allowing us at the same time to keep the code short and readable.
+From a practical point of view, these formats also provide a bunch of data
+loading and conversion tools which make them very easy to use in many
+contexts\footnote{For tasks where the inputs are text files or semi-structured
+objects, we provide \textit{vectorizer} objects that efficiently convert such
+data to the NumPy or SciPy formats.}. Finally, this design choice has also been
+motivated by the fact that, given their pervasiveness in many other scientific
+Python packages, many scientific users of Python are likely to be already used
+to those formats, hence helping them to familiarize with the project.
+
+% For a machine learning library, one of the most important decisions we had to
+% make is how to represent data.  Rather than reinventing the wheel, we opted for
+% NumPy multidimensional arrays \Citep{vanderwalt2011} for dense data and SciPy
+% sparse matrices for sparse data.  While those may seem like a ``bare''
+% representation of data when compared to more object-oriented representations
+% (Weka is an example of this style of representation, \citealp{hall2009weka}),
+% it brings the prime advantage of allowing us to rely on NumPy and SciPy's
+% \textit{vector operations} which are often orders of magnitude faster
+% than the corresponding Python loops,
+% while at the same time keeping the code short and readable.
+% Conversion to NumPy and SciPy formats is usually easy and many scientific users
+% of Python will already be used to such formats, since they are pervasive in
+% other scientific Python packages.  For tasks where input is likely to consist
+% of text files or semi-structured objects, we provide ``vectorizers'' -- objects
+% that efficiently convert such data to the NumPy or SciPy formats.
+
 % FIXME: elaborate what is meant below

### 2013-06-05 paper.tex
Wordsmithing

@@
 multidimensional arrays for dense data and as SciPy sparse matrices for sparse
-data. While those may seem like bare representations of data when compared to
-more object-oriented representations\footnote{Weka~\citep{hall2009weka} is an
-example of a toolkit-specific object-oriented data representation.}, they bring
-the prime advantage of allowing us to rely on all NumPy and SciPy vector
-operations -- often orders of magnitude faster that the corresponding Python
-loops -- while allowing us at the same time to keep the code short and readable.
-From a practical point of view, these formats also provide a bunch of data
-loading and conversion tools which make them very easy to use in many
-contexts\footnote{For tasks where the inputs are text files or semi-structured
-objects, we provide \textit{vectorizer} objects that efficiently convert such
-data to the NumPy or SciPy formats.}. Finally, this design choice has also been
-motivated by the fact that, given their pervasiveness in many other scientific
-Python packages, many scientific users of Python are likely to be already used
-to those formats, hence helping them to familiarize with the project.
+data. While those may seem like bare and minimalistic data representations
+when compared to more object-oriented
+representations\footnote{Weka~\citep{hall2009weka} is an example of a toolkit-specific
+object-oriented data representation.}, they bring the prime advantage
+of allowing us to rely on all NumPy and SciPy vector operations -- often orders
+of magnitude faster that the corresponding Python loops -- while allowing us at
+the same time to keep the code short and readable. From a practical point of
+view, these formats also provide a bunch of data loading and conversion tools
+which make them very easy to use in many contexts\footnote{For tasks where the
+inputs are text files or semi-structured objects, we provide \textit{vectorizer}
+objects that efficiently convert such data to the NumPy or SciPy formats.}.
+Finally, this design choice has also been motivated by the fact that, given
+their pervasiveness in many other scientific Python packages, many scientific
+users of Python are likely to be already familiar with those formats, hence helping
+them to familiarize with the project.
 

### 2013-06-06 paper.tex
Section 2.2 (not finished)

@@
 design choices for the application programming interface (API) of the project.
-In particular, we describe the core \textit{estimator} interface shared by all
+In particular, we describe the simple and limited interface shared by all
 learning and processing units in the library and then discuss its advantages in
@@
 
-In this section, we discuss the general principles of the application programming
-interface (API) of \textit{scikit-learn}. We first specify the representation
-of data and then proceed by describing the interface shared by all learning
-processing units of the library. We then discuss the composition and reusability
+In this section, we discuss the general principles of the application
+programming interface (API) of \textit{scikit-learn}. We first specify the
+representation of data and then proceed by describing the interface shared by
+all objects of the library. We then discuss the composition and reusability
 advantages of that interface.
@@
 
-\subsection{Estimators}
+\subsection{Estimators, predictors and transformers}
 
-% TODO describe transformer vs. classifiers/regressors
+All objects within \textit{scikit-learn} share a uniform common basic API
+consisting of three complementary interfaces: an \textit{estimator} interface
+for building and fitting objects, a \textit{predictor} interface for making
+predictions and a \textit{transformer} interface for converting data.
+As much as possible, our design choices have been guided so as to avoid
+the proliferation of framework code. We try to adopt simple conventions and
+to limit to a minimum the number of methods an object must implement.
 
-The bulk of the scikit-learn application programming interface
-consists of so-called \textit{estimator} objects.
-All learning algorithms for classification, regression and clustering
-are offered as such objects,
-but so are feature extraction, feature selection and dimensionality reduction
-algorithms.
+\textbf{Estimators.} The \textit{estimator} interface is at the core of the
+library. It defines instantiation mechanisms of objects and exposes a
+\texttt{fit} method for learning a model from training data.  All supervised and
+unsupervised learning algorithms (e.g., for classification, regression or
+clustering) are offered as objects implementing that interface. Machine
+learning tasks like feature extraction, feature selection or dimensionality
+reduction are also provided as estimators.
 
-Construction and application of an estimator object
-proceeds in a way that is somewhat reminiscent of partial function application.
-The constructor, \texttt{\_\_init\_\_}\footnote{
-  Strictly speaking, \texttt{\_\_init\_\_} is Python's
-  class \textit{initialization} method,
-  while \texttt{\_\_new\_\_} is the constructor.
-  The distinction is not relevant to the present description of scikit-learn.}
-takes a set of named (hyper-)parameters
-and attaches these to the estimator as public attributes.\footnote{
-  The Python language does not actually distinguish between public and private
-  attributes and methods,
-  but by convention, any identifier prefixed with an underscore (\texttt{\_})
-  is considered ``a non-public part of the API\ldots
-  an implementation detail and subject to change without notice.''
-  \Citep{pythontut}
-  }
-It never performs actual learning, nor does it even see a training set.
+The constructor of an estimator object takes as input a set of named constant
+hyper-parameters. At initialization, all given parameters are attached to the
+estimator instance as public attributes\footnote{Python does not actually
+distinguish between public and private  attributes or methods. By convention
+however, any identifier prefixed with an underscore is considered a non-public
+part of the API~\citep{pythontut}.} and serve to determine the estimator
+behavior (e.g., the $C$ constant in SVMs). Parameters set as public attributes
+are also used for introspection purposes, e.g., in model selection routines. To
+ease their use, appropriate default hyper-parameter values are provided for all
+built-in estimators. These default values are set to be relevant in most common
+situations in order to make estimators as efficient as possible out-of-box for
+non-experts. Finally, the constructor of an estimator never performs any actual
+learning, nor does it even see any training dataset. Its only goal
+is to set the parameter values and freeze the behavior of the estimator.
+
+\vspace{3cm}
+
+% The bulk of the scikit-learn application programming interface
+% consists of so-called \textit{estimator} objects.
+% All learning algorithms for classification, regression and clustering
+% are offered as such objects,
+% but so are feature extraction, feature selection and dimensionality reduction
+% algorithms.
+
+% Construction and application of an estimator object
+% proceeds in a way that is somewhat reminiscent of partial function application.
+% The constructor, \texttt{\_\_init\_\_}\footnote{
+%   Strictly speaking, \texttt{\_\_init\_\_} is Python's
+%   class \textit{initialization} method,
+%   while \texttt{\_\_new\_\_} is the constructor.
+%   The distinction is not relevant to the present description of scikit-learn.}
+% takes a set of named (hyper-)parameters
+% and attaches these to the estimator as public attributes.\footnote{
+%   The Python language does not actually distinguish between public and private
+%   attributes and methods,
+%   but by convention, any identifier prefixed with an underscore (\texttt{\_})
+%   is considered ``a non-public part of the API\ldots
+%   an implementation detail and subject to change without notice.''
+%   \Citep{pythontut}
+%   }
+% It never performs actual learning, nor does it even see a training set.
 

### 2013-06-06 paper.tex
Estimators

@@
 built-in estimators. These default values are set to be relevant in most common
-situations in order to make estimators as efficient as possible out-of-box for
-non-experts. Finally, the constructor of an estimator never performs any actual
-learning, nor does it even see any training dataset. Its only goal
-is to set the parameter values and freeze the behavior of the estimator.
+situations in order to make estimators as efficient as possible \textit{out-of-box}
+for non-experts. Finally, the constructor of an estimator never performs
+any actual learning, nor does it even see any training dataset. Its only goal is
+to set the parameter values and freeze the behavior of the estimator.
+
+The estimator interface also defines a \texttt{fit} method. Invoking this method
+causes the estimator to learn from a training set (e.g., supplied as two arrays
+\texttt{X} and \texttt{y} in supervised learning estimators). Its task, from an
+API point of view, is to run a learning algorithm and to determine model-
+specific parameters from the training data and set these as attributes on the
+estimator object.
+
+As an example, let us consider a supervised learning task using linear regression.
+Given the API defined above, solving this problem is as simple as:
+\begin{center}
+\texttt{clf = LinearRegression(penalty="l1").fit(X, y)}
+\end{center}
+In this snippet, a \texttt{LinearRegression} estimator is first initialized by
+setting the \texttt{penalty} hyper-parameter to \texttt{l1}, hence enabling $L1$
+regularization, and using default values for all other hyper-parameters. Upon
+\texttt{fit}, the model is then learned from the arrays \texttt{X} and
+\texttt{y}. Also, since all estimators share the same interface, using a
+different learning algorithm is as simple as replacing the constructor. For
+example, to build a Random Forest on the same training data, one would simply
+have to replace \texttt{LinearRegression(penalty="l1")} in the snippet above by
+\texttt{RandomForestClassifier()}. Let us also note the snippet relies on the
+fact that the \texttt{fit} method always returns the object it was called
+on\footnote{I.e., the \texttt{self} argument it was called on.}, hence making
+method chaining  possible.
+
+In \textit{scikit-learn}, classical learning algorithms are not the only objects
+to be implemented as estimators. For example, preprocessing routines (e.g.,
+vectorization of text documents, scaling of attributes, etc) or feature
+extraction techniques also implement the same interface. Even those estimators
+that are entirely stateless and therefore do not require the \texttt{fit} method
+to perform any useful work implement the estimator interface. As we will
+illustrate in the next section, this design pattern is indeed of prime
+importance for consistency, composition and parameter validation reasons.
+
+\textbf{Predictors.}
+
+\textbf{Transformers.}
 
@@
 
-The algorithmic workhorse of most estimators, and one of the few methods
-that all estimators have in common, is called \texttt{fit}.
-Invoking this method causes the estimator to learn from a training set,
-supplied as one or more arguments.
-Its task, from an API point of view,
-is to determine (via a learning algorithm) model-specific parameters
-from a training set and set these as attributes on the estimator object,
-which can then be used to transform data or make predictions.
-As in other object-oriented machine learning libraries
-(e.g.\ Weka's Java API)
-no distinction is made in the type hierarchy
-between what is variously called an estimator or a \textit{learner}
-in the machine learning literature,
-and the \textit{model} that results from applying an estimator to data.
-When the \texttt{fit} method has been called,
-an estimator object serves dual purpose as a model
-(and is referred to as such in the documentation).
+% The algorithmic workhorse of most estimators, and one of the few methods
+% that all estimators have in common, is called \texttt{fit}.
+% Invoking this method causes the estimator to learn from a training set,
+% supplied as one or more arguments.
+% Its task, from an API point of view,
+% is to determine (via a learning algorithm) model-specific parameters
+% from a training set and set these as attributes on the estimator object,
+% which can then be used to transform data or make predictions.
+% As in other object-oriented machine learning libraries
+% (e.g.\ Weka's Java API)
+% no distinction is made in the type hierarchy
+% between what is variously called an estimator or a \textit{learner}
+% in the machine learning literature,
+% and the \textit{model} that results from applying an estimator to data.
+% When the \texttt{fit} method has been called,
+% an estimator object serves dual purpose as a model
+% (and is referred to as such in the documentation).
 
-An example to clarify the partial application style
-is the following code snippet.
-If $X$ is a dataset in the format described below and $k$ is some integer,
-then the Python code to train a $k$-means clustering model on it
-can be written as
-% XXX we could use Pygments for syntax highlighting
-\begin{verbatim}
-km = KMeans(n_clusters=k).fit(X)
-\end{verbatim}
-This snippet relies on the fact that the \texttt{fit} method
-always returns the object it was called on (its ``\texttt{self}''),
-making a method chaining style possible.
+% An example to clarify the partial application style
+% is the following code snippet.
+% If $X$ is a dataset in the format described below and $k$ is some integer,
+% then the Python code to train a $k$-means clustering model on it
+% can be written as
+% % XXX we could use Pygments for syntax highlighting
+% \begin{verbatim}
+% km = KMeans(n_clusters=k).fit(X)
+% \end{verbatim}
+% This snippet relies on the fact that the \texttt{fit} method
+% always returns the object it was called on (its ``\texttt{self}''),
+% making a method chaining style possible.
+
+% A \texttt{fit} method is present even on objects
+% that perform seemingly mundane tasks such as vectorizing text documents
+% for subsequent learning.
+% Such classes, however, can be said to learn their vocabulary
+% from the training corpus,
+% and follow the conventions set out above.
+% Even those estimators that are entirely stateless and therefore
+% do not require a \texttt{fit} method to perform useful work,
+% nevertheless have such a method for consistency and composibility,
+% and to perform parameter validation.
+% Examples of this latter kind of objects are the normalizing transformer,
+% kernel approximation transformers
+% \citep{rahimi2007random, li2010random, vedaldi2010efficient},
+% % XXX is it relevant that they are exactly two? Propose s/two/the
+% and two feature extraction classes that implement the ``hashing trick''
+% \citep{weinberger2009}.
 
@@
 on the names and formats of model parameters.
-
-A \texttt{fit} method is present even on objects
-that perform seemingly mundane tasks such as vectorizing text documents
-for subsequent learning.
-Such classes, however, can be said to learn their vocabulary
-from the training corpus,
-and follow the conventions set out above.
-Even those estimators that are entirely stateless and therefore
-do not require a \texttt{fit} method to perform useful work,
-nevertheless have such a method for consistency and composibility,
-and to perform parameter validation.
-Examples of this latter kind of objects are the normalizing transformer,
-kernel approximation transformers
-\citep{rahimi2007random, li2010random, vedaldi2010efficient},
-% XXX is it relevant that they are exactly two? Propose s/two/the
-and two feature extraction classes that implement the ``hashing trick''
-\citep{weinberger2009}.
 

### 2013-06-06 paper.tex
Predictors and transformers

@@
 specific  parameters from the training data and set these as attributes on the
-estimator object.
+estimator object. As a convention, the parameters learned by an estimator are
+exposed as public attributes\footnote{Note that the learned parameters are
+different from the hyper-parameters given to the constructor. To distinguish
+between the two, learned parameters are all suffixed with a trailing underscore
+(e.g., \texttt{clf.coef\_} in the case of linear models is the vector of
+coefficients learned from the data).} on the trained object.
+
+% Importantly, the parameters learned by an estimator
+% are exposed as \textit{public} attributes on the trained object.
+% This facilitates model inspection
+% and makes it possible to train a model using scikit-learn,
+% export it to a (standardized or custom) external data format
+% and use the learned parameters in a different piece of code,
+% perhaps a prediction algorithm implemented in a different language
+% or one that combines the model parameters with other information
+% to build a custom model.
+% % XXX Hard to follow phrasing below
+% Of course, it also forces the developers to think carefully
+% about what might, in a more ``black box''-style toolkit,
+% be considered the internals of objects,
+% since authors of client code are allowed to rely
+% on the names and formats of model parameters.
 
@@
 have to replace \texttt{LinearRegression(penalty="l1")} in the snippet above by
-\texttt{RandomForestClassifier()}. Let us also note the snippet relies on the
+\texttt{RandomForestClassifier()}. Let us also note that the snippet relies on the
 fact that the \texttt{fit} method always returns the object it was called
@@
 to perform any useful work implement the estimator interface. As we will
-illustrate in the next section, this design pattern is indeed of prime
+illustrate in the next sections, this design pattern is indeed of prime
 importance for consistency, composition and parameter validation reasons.
 
-\textbf{Predictors.}
+\textbf{Predictors.} The \textit{predictor} interface defines a \texttt{predict}
+method taking as input some dataset \texttt{X} and producing as output
+predictions for \texttt{X} based on the learned parameters of the estimator. In
+the case of supervised learning estimators, this method typically returns the
+predicted labels or values computed by the model. Some unsupervised learning
+estimators may however also implement that interface: for example, in clustering
+estimators, \texttt{predict} returns the identifiers of the closest clusters for
+the samples in \texttt{X}.
 
-\textbf{Transformers.}
+Along with the \texttt{predict} method, predictors may also implement methods
+for quantifying the certainty of a prediction. In the case of estimators
+implementing linear models, the \texttt{decision\_function} method can be used
+to measure the distance of the samples from the discriminating hyperplane. Some
+predictors also provide a \texttt{predict\_proba} method which returns
+class probabilities for the input samples \texttt{X}.
 
-\vspace{3cm}
+Finally, predictors should also provide a \texttt{score} function to assess the
+goodness of fit on some input data. In supervised learning estimators, this
+method takes as inputs a dataset \texttt{X} and \texttt{y} and typically
+computes the coefficient of determination between \texttt{y} and
+\texttt{predict(X)} in the case of classification or the mean squared error in
+the case of regression. Not however that, from an API point of view, the only
+guideline is that the \texttt{score} method returns a numerical value such that
+the higher the better. As such, some unsupervised learning estimators also
+exposes a \texttt{score} function, computing for instance the likelihood of the
+given data.
+
+Continuing with the previous example, predictions on unseen data
+\texttt{X\_test}, \texttt{y\_test} can therefore be obtained and evaluated using
+the following snippet:
+\begin{center}
+\texttt{y\_hat = clf.predict(X\_test)}\\
+\texttt{score = clf.score(X\_test, y\_test)}
+\end{center}
+
+\textbf{Transformers.} Since it is common to modify or filter data before
+feeding it to some learning algorithm, some estimators in the library implement
+a \textit{transformer} interface which defines a \texttt{transform} method
+taking as input some dataset \texttt{X} and yielding as output a transformed
+version of \texttt{X}\footnote{From an implementation point of view, it may
+sometimes be more efficient to implement \texttt{fit} and \texttt{transform} in
+a single method. In that case, transformers may expose an additional
+\texttt{fit\_transform} method.}. In particular, all preprocessing algorithms
+and feature selection routines are offered as transformers within the library.
+
+In our example, if one wanted to transform the inputs \texttt{X} before
+fitting the linear regression estimator, e.g., by standardizing the feature values,
+one could thus simply write:
+\begin{center}
+\texttt{X = StandardScaler().fit(X).transform(X)}
+\end{center}
 
@@
 % \citep{weinberger2009}.
-
-Importantly, the parameters learned by an estimator
-are exposed as \textit{public} attributes on the trained object.
-This facilitates model inspection
-and makes it possible to train a model using scikit-learn,
-export it to a (standardized or custom) external data format
-and use the learned parameters in a different piece of code,
-perhaps a prediction algorithm implemented in a different language
-or one that combines the model parameters with other information
-to build a custom model.
-% XXX Hard to follow phrasing below
-Of course, it also forces the developers to think carefully
-about what might, in a more ``black box''-style toolkit,
-be considered the internals of objects,
-since authors of client code are allowed to rely
-on the names and formats of model parameters.
 

### 2013-06-06 paper.tex
Composition (not finished)

@@
 \textbf{Transformers.} Since it is common to modify or filter data before
-feeding it to some learning algorithm, some estimators in the library implement
+feeding it to a learning algorithm, some estimators in the library implement
 a \textit{transformer} interface which defines a \texttt{transform} method
@@
 
-A distinguishing feature of scikit-learn estimators
+A distinguishing feature of the \textit{scikit-learn} API is its ability to
+compose new estimators from other estimators. Such objects combine typical
+machine learning  workflows (e.g., feature extraction,
+centering/scaling/normalization, feature selection, learning) into  a single
+object which is itself an estimator. As such, composed estimators can be used
+anywhere usual estimators are used. In particular, they can be used within model
+selection routines, allowing one to optimize at once every step of a complex
+workflow. Composition of estimators can be done in two ways: either sequentially
+through \texttt{Pipeline} objects or in a parallel fashion through
+\texttt{FeatureUnion} objects.
+
+\vspace{3cm}
+
+A distinguishing feature of \textit{scikit-learn} estimators
 is that they can be composed into \texttt{Pipeline} objects.

### 2013-06-06 paper.tex
Small changes

--- 2013-06-06 paper.tex
@@
 \end{abstract}
+
+\textcolor{red}{TODO: focus on the experience we accumulated (general remark).}
 
@@
 \textbf{Transformers.} Since it is common to modify or filter data before
-feeding it to some learning algorithm, some estimators in the library implement
-a \textit{transformer} interface which defines a \texttt{transform} method
-taking as input some dataset \texttt{X} and yielding as output a transformed
+feeding it to a learning algorithm, some estimators in the library implement a
+\textit{transformer} interface which defines a \texttt{transform} method. It
+takes as input some dataset \texttt{X} and yields as output a transformed
 version of \texttt{X}\footnote{From an implementation point of view, it may
@@
 
-A distinguishing feature of scikit-learn estimators
+A distinguishing feature of the \textit{scikit-learn} API is its ability to
+compose new estimators from several base estimators. Composition mechanisms can
+be used to combine typical machine learning workflows (e.g., feature extraction,
+centering/scaling/normalization, feature selection, learning) into a single
+object which is itself an estimator. Relying on duck typing, composed estimators
+can be used wherever usual estimators can be used. In particular, they integrate
+within model selection routines, allowing one to optimize at once every step of
+a complex workflow. Composition of estimators can be done in two ways: either
+sequentially through \texttt{Pipeline} objects or in a parallel fashion through
+\texttt{FeatureUnion} objects.
+
+\vspace{3cm}
+
+A distinguishing feature of \textit{scikit-learn} estimators
 is that they can be composed into \texttt{Pipeline} objects.

### 2013-06-07 paper.tex
Pipelines

@@
 \DeclareRobustCommand{\VAN}[3]{#2}
-
-\newcommand{\Xtr}{\texttt{X\_tr}}
-\newcommand{\Xte}{\texttt{X\_te}}
-\newcommand{\ytr}{\texttt{y\_tr}}
-\newcommand{\yte}{\texttt{y\_te}}
 
@@
 compose new estimators from several base estimators. Composition mechanisms can
-be used to combine typical machine learning workflows (e.g., feature extraction,
-centering/scaling/normalization, feature selection, learning) into a single
-object which is itself an estimator. Relying on duck typing, composed estimators
-can be used wherever usual estimators can be used. In particular, they integrate
-within model selection routines, allowing one to optimize at once every step of
-a complex workflow. Composition of estimators can be done in two ways: either
-sequentially through \texttt{Pipeline} objects or in a parallel fashion through
-\texttt{FeatureUnion} objects.
+be used to combine typical machine learning workflows into a single object which
+is itself an estimator. Through duck typing, composed
+estimators can be used wherever usual estimators can be used. In particular,
+they integrate within model selection routines, allowing one to optimize at once
+every step of a complex workflow. Composition of estimators can be done in two
+ways: either sequentially through \texttt{Pipeline} objects or in a parallel
+fashion through \texttt{FeatureUnion} objects.
+
+\texttt{Pipeline} objects are used for chaining multiple estimators into a
+single one. This is useful since a machine learning workflow typically involves
+a fixed sequence of steps in processing data (e.g., feature extraction,
+centering / scaling / normalization, feature selection, learning and making
+predictions). For convenience reasons, those steps can be combined into a
+pipeline such that \texttt{fit} and \texttt{predict} need to be called only once
+to fit the whole sequence. A pipeline also allows for the hyper-parameters of
+all the base estimators to be grid-searched all at once. From an API point of
+view, a \texttt{Pipeline} object takes as input a sequence of $N$ estimators,
+where all $N-1$ first are transformers and the last is either a predictor or a
+transformer. Calling \texttt{fit} on the pipeline is the same as calling
+\texttt{fit} on each estimator in turn, \texttt{transform} the input and then
+pass it on to the next step. The pipeline exposes all the methods the last
+estimator in the pipe exposes. That is, if the last estimator is a predictor,
+the pipeline can be used itself as a predictor. If the last estimator is a
+transformer, then the pipeline is itself a transformer.
+
+\texttt{FeatureUnion} objects
 
@@
 \subsection{Model selection}
+
+\textcolor{red}{TODO: rewrite + describe Scorer API}
 

### 2013-06-07 paper.tex
Feature unions + example

@@
 
-\texttt{FeatureUnion} objects
+A \texttt{FeatureUnion} object is a transformer combining multiple transformers
+into a single one that joins their outputs. From an API point of view, a
+\texttt{FeatureUnion} takes as input a list of transformers. Calling
+\texttt{fit} on the union is the same as calling \texttt{fit} independently on
+each of the transformers. For transforming data, the base transformers are
+applied in parallel are their outputs are concatenated end-to-end into a larger
+output.
 
-\vspace{3cm}
+As an example, the following snippet illustrates how to create a complex worflow
+computing both PCA and Kernel PCA components on \texttt{X\_train} (through a
+\texttt{FeatureUninon}), selecting the best of those features and then feeding
+them to a linear model:
+\begin{verbatim}
+Pipeline([("pca", FeatureUnion([('linear_pca', PCA()),
+                                ('kernel_pca', KernelPCA())])),
+          ("feature_selection", SelectKBest()),
+          ("linear_model", LinearRegression())
+]).fit(X_train, y_train).predict(X_test)
+\end{verbatim}
 
-A distinguishing feature of \textit{scikit-learn} estimators
-is that they can be composed into \texttt{Pipeline} objects.
-Such objects combine typical machine learning workflows
-(extracting features, centering/scaling/normalization,
-feature selection, making predictions)
-in a single object that is itself an estimator,
-and whose methods cause data to flow through the pipeline.
-More importantly, though, they can be used for
-systematic, automated model selection.
-This is best explained by example.
+As the above snippet shows, \texttt{Pipeline} and \texttt{FeatureUninon}
+can be combined to create complex objects. In particular, since both
+are themselves estimators, they can be used to form nested workflows.
 
-By composing, say, a \texttt{CountVectorizer} object
-(which extracts term frequency features from text documents
-according to a bag-words model),
-a \textsf{tf--idf} transformer, a $\chi^2$ feature selector and a linear SVM,
-one obtains a pipeline object with various hyperparameters:
-whether to do stopword filtering,
-linear vs.\ logarithmic \textsf{tf}, number of features to keep,
-SVM regularization ($L_1$ or $L_2$ norm, strength $C$) a.o.
-Such a composite \texttt{Pipeline} object
-exposes all of these parameters using a special syntax;
-each estimators must be given a name,
-and if the SVM's name is, e.g., \texttt{linearsvm},
-then its $C$ parameter is exposed as \texttt{linearsvm\_\_C}.
-The pipeline's \texttt{set\_params} splits the parameter names
-on double underscores, uses the first part of the result
-to look up the estimator,
-and passes the second part to the estimator's own \texttt{set\_params}
-as a keyword argument; due to Python's dynamic parameter passing,
-the result is the same as if the user had called \texttt{set\_params}
-on the SVM themselves.
-(The double underscore syntax is not special to pipelines;
-it is also employed in ensemble estimators such as random forests
-% Is the following clear? Suggestions to replace "direct attributes" with"?
-It is chosen as a convention to avoid ambiguity, as direct
-attributes and methods of scikit-learn objects never contain
-double underscores.)
+% A distinguishing feature of \textit{scikit-learn} estimators
+% is that they can be composed into \texttt{Pipeline} objects.
+% Such objects combine typical machine learning workflows
+% (extracting features, centering/scaling/normalization,
+% feature selection, making predictions)
+% in a single object that is itself an estimator,
+% and whose methods cause data to flow through the pipeline.
+% More importantly, though, they can be used for
+% systematic, automated model selection.
+% This is best explained by example.
+
+% By composing, say, a \texttt{CountVectorizer} object
+% (which extracts term frequency features from text documents
+% according to a bag-words model),
+% a \textsf{tf--idf} transformer, a $\chi^2$ feature selector and a linear SVM,
+% one obtains a pipeline object with various hyperparameters:
+% whether to do stopword filtering,
+% linear vs.\ logarithmic \textsf{tf}, number of features to keep,
+% SVM regularization ($L_1$ or $L_2$ norm, strength $C$) a.o.
+% Such a composite \texttt{Pipeline} object
+% exposes all of these parameters using a special syntax;
+% each estimators must be given a name,
+% and if the SVM's name is, e.g., \texttt{linearsvm},
+% then its $C$ parameter is exposed as \texttt{linearsvm\_\_C}.
+% The pipeline's \texttt{set\_params} splits the parameter names
+% on double underscores, uses the first part of the result
+% to look up the estimator,
+% and passes the second part to the estimator's own \texttt{set\_params}
+% as a keyword argument; due to Python's dynamic parameter passing,
+% the result is the same as if the user had called \texttt{set\_params}
+% on the SVM themselves.
+% (The double underscore syntax is not special to pipelines;
+% it is also employed in ensemble estimators such as random forests
+% % Is the following clear? Suggestions to replace "direct attributes" with"?
+% It is chosen as a convention to avoid ambiguity, as direct
+% attributes and methods of scikit-learn objects never contain
+% double underscores.)
 

### 2013-06-07 paper.tex
todo
Wording

@@
 the pipeline can be used itself as a predictor. If the last estimator is a
-transformer, then the pipeline is itself a transformer.
+transformer, then the pipeline is itself a transformer. \textcolor{red}{TODO:
+Talk about nested attribute syntax.}
 
@@
 As an example, the following snippet illustrates how to create a complex worflow
-computing both PCA and Kernel PCA components on \texttt{X\_train} (through a
+computing both linear PCA and kernel PCA components on \texttt{X\_train} (through a
 \texttt{FeatureUninon}), selecting the best of those features and then feeding

### 2013-06-07 paper.tex
todos

@@
 
-\textcolor{red}{TODO: focus on the experience we accumulated (general remark).}
+\textcolor{red}{TODO: Focus on the experience we accumulated (general remark); motivate our design choices.}
 
@@
 
-\textcolor{red}{TODO: rewrite + describe Scorer API}
+\textcolor{red}{TODO: rewrite and briefly describe the API (see user guide)}\\
+\textcolor{red}{TODO: describe Scorer API?}\\
+\textcolor{red}{TODO: mention that GridSearchCV objects are themselves estimators}
 
@@
 The latter, which we shall use in this example,
-performs an exhaustive search through a grid formed by the Cartesian product
+performs an exhaustive search through a grid formed by the cartesian product
 of a set of possible values for each parameter, specified by the user.

### 2013-06-08 paper.tex
todos
TODOs

@@
 
-\textcolor{red}{TODO: Focus on the experience we accumulated (general remark); motivate our design choices.}
+\noindent \textcolor{red}{TODO: Focus on the experience we accumulated (general remark)}\\
+\textcolor{red}{TODO: Motivate our design choices.}
 
@@
 
-As the above snippet shows, \texttt{Pipeline} and \texttt{FeatureUninon}
-can be combined to create complex objects. In particular, since both
-are themselves estimators, they can be used to form nested workflows.
+As the above snippet shows, \texttt{Pipeline} and \texttt{FeatureUninon} can be
+combined to create complex objects. In particular, since both are themselves
+estimators, they can be used to form nested workflows. \textcolor{red}{TODO:
+elaborate a bit more on nested constructions.}
 
@@
 
+\noindent \textcolor{red}{TODO: Mention small set of dependencies.}\\
+\textcolor{red}{TODO: elaborate on use of mixins and duck typing}
+
 scikit-learn is primarily implemented in Python and Cython
@@
 \section{Weaknesses of the NumPy/SciPy environment}
+
+\textcolor{red}{TODO: move this into the Implementation section? These are mostly implementation considerations.}
 
@@
 
-ToDo.
+\textcolor{red}{TODO: compare with weka, R, vowpal rabbit, ... ?}\\
+\textcolor{red}{TODO: emphasize scikit-learn's place in the overall scipy ecosystem, and its role in spurring the development of more basic tools.}\\
+\textcolor{red}{TODO: some packages start to adopt our own API.}
 

### 2013-06-08 paper.tex
LinearRegression -> LogisticRegression

@@
 
-As an example, let us consider a supervised learning task using linear regression.
+As an example, let us consider a supervised learning task using logistic regression.
 Given the API defined above, solving this problem is as simple as:
 \begin{center}
-\texttt{clf = LinearRegression(penalty="l1").fit(X, y)}
+\texttt{clf = LogisticRegression(penalty="l1").fit(X\_train, y\_train)}
 \end{center}
-In this snippet, a \texttt{LinearRegression} estimator is first initialized by
+In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
 setting the \texttt{penalty} hyper-parameter to \texttt{l1}, hence enabling $L1$
 regularization, and using default values for all other hyper-parameters. Upon
-\texttt{fit}, the model is then learned from the arrays \texttt{X} and
-\texttt{y}. Also, since all estimators share the same interface, using a
+\texttt{fit}, the model is then learned from the training arrays \texttt{X\_train} and
+\texttt{y\_train}. Also, since all estimators share the same interface, using a
 different learning algorithm is as simple as replacing the constructor. For
 example, to build a Random Forest on the same training data, one would simply
-have to replace \texttt{LinearRegression(penalty="l1")} in the snippet above by
+have to replace \texttt{LogisticRegression(penalty="l1")} in the snippet above by
 \texttt{RandomForestClassifier()}. Let us also note that the snippet relies on the
@@
 In our example, if one wanted to transform the inputs \texttt{X} before
-fitting the linear regression estimator, e.g., by standardizing the feature values,
+fitting the logistic regression estimator, e.g., by standardizing the feature values,
 one could thus simply write:

### 2013-06-08 paper.tex
Update snippets

@@
 
-In our example, if one wanted to transform the inputs \texttt{X} before
+In our example, if one wanted to transform the inputs \texttt{X\_train} before
 fitting the logistic regression estimator, e.g., by standardizing the feature values,
@@
 \begin{center}
-\texttt{X = StandardScaler().fit(X).transform(X)}
+\texttt{X\_train = StandardScaler().fit(X\_train).transform(X\_train)}
 \end{center}
@@
 \texttt{FeatureUninon}), selecting the best of those features and then feeding
-them to a linear model:
+them to a logistic regression model:
 \begin{verbatim}
@@
           ("feature_selection", SelectKBest()),
-          ("linear_model", LinearRegression())
+          ("logistic_model", LogisticRegression())
 ]).fit(X_train, y_train).predict(X_test)

### 2013-06-08 paper.tex
Minor rewording and a few remarks

@@
 graphical user interface.
+% XXX Gael: maybe we should say that it augments a general purpose
+% language with machine learning concepts, like numpy augments it with
+% array-computing concepts.
 
@@
 multidimensional arrays for dense data and as SciPy sparse matrices for sparse
-data. While those may seem like bare and minimalistic data representations
+data. While those may seem rather unsophisticated data representations
 when compared to more object-oriented
-representations\footnote{Weka~\citep{hall2009weka} is an example of a toolkit-specific
-object-oriented data representation.}, they bring the prime advantage
+constructs, such as those used by Weka~\citep{hall2009weka}. However,
+they bring the prime advantage
 of allowing us to rely on all NumPy and SciPy vector operations -- often orders
@@
 Given the API defined above, solving this problem is as simple as:
+% XXX: Gael: I think that we should break the following lines in 2: the
+% 'return self' hidden in the call to fit can be confusing to the casual
+% reader
 \begin{center}

### 2013-06-08 paper.tex
L1 regularization is written \ell_1

@@
 In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
-setting the \texttt{penalty} hyper-parameter to \texttt{l1}, hence enabling $L1$
+setting the \texttt{penalty} hyper-parameter to \texttt{l1}, hence
+enabling $\ell_1$
 regularization, and using default values for all other hyper-parameters. Upon
@@
 % linear vs.\ logarithmic \textsf{tf}, number of features to keep,
-% SVM regularization ($L_1$ or $L_2$ norm, strength $C$) a.o.
+% SVM regularization ($\ell_1$ or $\ell_2$ norm, strength $C$) a.o.
 % Such a composite \texttt{Pipeline} object

### 2013-06-08 paper.tex
wording

@@
 machine learning library for the Python programming language. The ambition of
-the project is to provide  effective and long-established machine learning tools
+the project is to provide efficient and well-established machine learning tools
 within a programming environment that is accessible to non-machine learning
-experts, efficient, and reusable in various scientific areas. In itself, the
+experts and reusable in various scientific areas. In itself, the
 project is a library of machine learning tools  rather than  a novel domain-

### 2013-06-08 paper.tex
wording

@@
 within a programming environment that is accessible to non-machine learning
-experts and reusable in various scientific areas. In itself, the
-project is a library of machine learning tools  rather than  a novel domain-
-specific programming language. Among other things, the library includes
-classical learning algorithms, model selection tools and preprocessing
+experts and reusable in various scientific areas. In itself the project isn't 
+a novel domain-specific language, but library that provides machine
+learning idioms to a general-purpose high-level language. Among other things,
+it includes classical learning algorithms, model selection tools and
+preprocessing
 procedures. From its core, the library has been designed to tie in with the set

### 2013-06-08 paper.tex
wording

@@
 
-Started in 2009, \textit{scikit-learn} counts a large team of volunteer
-international developers mainly consisting of scientists and researchers from
-various fields (e.g., machine learning, natural language processing or
+Started in 2009, \textit{scikit-learn} is developed by a large and
+international team of volunteers, mostly researchers from
+various fields (e.g., machine learning, natural language processing, or
 neuroscience). The project also greatly benefits from many occasional
-contributors proposing small code and/or documentation fixes or improvements.
-Because of this large base of developers, important efforts are deployed to make
-the project maintainable. In particular, we emphasize on high-quality and well-documented
-code  in order to ease future developments by external contributors.
-\textcolor{red}{TODO: Mention some good coding practices (automatic testing, code
-reviews, etc)}
+contributors proposing small code or documentation fixes or improvements.
+Because of this large base of developers, emphasis is put on keeping
+the project maintainable. In particular, code must follow specific
+quality guidelines, such as style consistency, unit-test coverage. In
+addition, complete documentation and examples are required for all
+features. Finally, any addition to the code base goes through thorough
+code review.
 

### 2013-06-08 paper.tex
iter

@@
 
-Within the past few years, the project has proven to be valuable to an  ever
-increasing user base. Its popularity can be gauged from various indicators,
-including hundreds of citations in scientific papers, successes in various
-machine learning challenges, or significant statistics\footnote{\textcolor{red}{Cite some
-statistics here.}} derived from our repositories and mailing lists.
+Within the past few years, the project has proved valuable to an ever
+increasing user base. Its popularity can be gaged from various indicators
+such as hundreds of citations in scientific papers, successes in various
+machine learning challenges (\emph{e.g.} kaggle), or significant
+statistics\footnote{As of writing, 1309 stars and 667 forks of the
+project on github, 300 mails per months on the user mailing list, 385
+questions tagged \emph{scikit-learn} on {\tt stackoverflow}} derived from
+our repositories and mailing lists.
 

### 2013-06-08 paper.tex
introduce layout

@@
 
-\textcolor{red}{TODO: The rest of this text is organized as follows: in Section 2, ...}
+In this paper, we describe and motivate the design choices used in
+building the library. First, in section \ref{sec:api}, we give the
+central application programming interface (API). Then, in section
+\ref{sec:implementation}, we describe the implementation trade-offs.
+In section \ref{sec:weaknesses}, we analyze the limitations of the
+technologies we rely on; and finally, in section \ref{sec:comparison}, we
+position \emph{scikit-learn} with respect to other major projects.
 
@@
 \section{API}
+
+\label{sec:api}
 
@@
 
+\label{sec:implementation}
+
 \noindent \textcolor{red}{TODO: Mention small set of dependencies.}\\
@@
 \section{Weaknesses of the NumPy/SciPy environment}
+
+\label{sec:weaknesses}
 
@@
 
+\label{sec:comparison}
+
 \textcolor{red}{TODO: compare with weka, R, vowpal rabbit, ... ?}\\

### 2013-06-08 paper.tex
typo

@@
 project on github, 300 mails per months on the user mailing list, 385
-questions tagged \emph{scikit-learn} on {\tt stackoverflow}} derived from
+questions tagged \emph{scikit-learn} on {\tt stackoverflow}.} derived from
 our repositories and mailing lists.

### 2013-06-08 paper.tex
wording

@@
 are also used for introspection purposes, e.g., in model selection routines. To
-ease their use, appropriate default hyper-parameter values are provided for all
+ease their use, default hyper-parameter values are provided for all
 built-in estimators. These default values are set to be relevant in most common
@@
 
-The estimator interface also defines a \texttt{fit} method. Invoking this method
-causes the estimator to learn from a training set (e.g., supplied as two arrays
-\texttt{X} and \texttt{y} in supervised learning estimators). Its task, from an
-API point of view, is to run a learning algorithm and to determine model-
+The estimator interface defines a \texttt{fit} method. This method is
+called with training data (e.g., supplied as two arrays
+\texttt{X} and \texttt{y} in supervised learning estimators). Its task
+is to run a learning algorithm and to determine model-
 specific parameters from the training data and set these as attributes on the

### 2013-06-09 paper.tex
Wording

@@
 when compared to more object-oriented
-constructs, such as those used by Weka~\citep{hall2009weka}. However,
+constructs, such as those used by Weka~\citep{hall2009weka},
 they bring the prime advantage

### 2013-06-09 paper.tex
typo

@@
 experts and reusable in various scientific areas. In itself the project isn't
-a novel domain-specific language, but library that provides machine
+a novel domain-specific language, but a library that provides machine
 learning idioms to a general-purpose high-level language. Among other things,

### 2013-06-09 paper.tex
Break down the call to fit.

--- 2013-06-09 paper.tex
@@
 % reader
-\begin{center}
-\texttt{clf = LogisticRegression(penalty="l1").fit(X\_train, y\_train)}
-\end{center}
+\begin{verbatim}
+clf = LogisticRegression(penalty="l1")
+clf.fit(X_train, y_train)
+\end{verbatim}
 In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
@@
 have to replace \texttt{LogisticRegression(penalty="l1")} in the snippet above by
-\texttt{RandomForestClassifier()}. Let us also note that the snippet relies on the
-fact that the \texttt{fit} method always returns the object it was called
-on\footnote{I.e., the \texttt{self} argument it was called on.}, hence making
-method chaining  possible.
+\texttt{RandomForestClassifier()}.
 
@@
 the following snippet:
-\begin{center}
-\texttt{y\_hat = clf.predict(X\_test)}\\
-\texttt{score = clf.score(X\_test, y\_test)}
-\end{center}
+\begin{verbatim}
+y_pred = clf.predict(X_test)
+score = clf.score(X_test, y_test)
+\end{verbatim}
 
@@
 one could thus simply write:
-\begin{center}
-\texttt{X\_train = StandardScaler().fit(X\_train).transform(X\_train)}
-\end{center}
+\begin{verbatim}
+X_train = StandardScaler().fit(X_train).transform(X_train)
+\end{verbatim}
+
+Note that the snippet relies on the fact that the \texttt{fit} method always
+returns the object it was called on , hence making method chaining  possible.
 

### 2013-06-09 paper.tex
typo
Cosmit.

@@
 the project maintainable. In particular, code must follow specific
-quality guidelines, such as style consistency, unit-test coverage. In
+quality guidelines, such as style consistency and unit-test coverage. In
 addition, complete documentation and examples are required for all
@@
 each of the transformers. For transforming data, the base transformers are
-applied in parallel are their outputs are concatenated end-to-end into a larger
+applied in parallel and their outputs are concatenated end-to-end into a larger
 output.

### 2013-06-09 paper.tex
Cosmit.
Avoid footnote.

@@
 addition, complete documentation and examples are required for all
-features. Finally, any addition to the code base goes through thorough
+features. Finally, any addition to the code base goes through rigorous
 code review.
 
-Within the past few years, the project has proved valuable to an ever
-increasing user base. Its popularity can be gaged from various indicators
-such as hundreds of citations in scientific papers, successes in various
-machine learning challenges (\emph{e.g.} kaggle), or significant
-statistics\footnote{As of writing, 1309 stars and 667 forks of the
-project on github, 300 mails per months on the user mailing list, 385
-questions tagged \emph{scikit-learn} on {\tt stackoverflow}.} derived from
-our repositories and mailing lists.
+Within the past few years, the project has proved valuable to an ever increasing
+user base. Its popularity can be gaged from various indicators such as hundreds
+of citations in scientific publications of a previous paper
+\citep{pedregosa2011}, successes in various machine learning challenges
+(\emph{e.g.} kaggle), or significant statistics derived from our repositories
+and mailing lists.  At the time of writing, the project is watched by 1309
+people and forked 667 time on github, and the mailing-list receives 300 mails
+per month.
 

### 2013-06-09 paper.tex
Cosmit and avoid footnote.
Let us not drown the reader into obvious details.

@@
 between input variables $X_1, ..., X_p$ -- called attributes or features -- and
-some output variable $Y$. A sample is then defined as a tuple of values $(x_1,
-..., x_p, y)$ of these variables and a dataset as a collection of such samples.
-As such, matrices of numerical values\footnote{In the case of categorical
-variables, numerical values can be obtained using various encoding schemes, such
-as one-hot encoding.} constitute a natural representation used almost
-unanimously within all areas of machine learning. In particular, each row in the
+some output variable $Y$. A sample is then defined as a tuple of values $([x_1,
+..., x_p]^\mathrm{T}, y)$ of these variables and a dataset as a collection of
+such samples.  As such, matrices of numerical values constitute a natural and
+widely used representation in machine learning. In this view, each row in the
 matrix then corresponds to a sample of the dataset and each column to one of the

### 2013-06-09 paper.tex
Avoid footnote.

@@
 multidimensional arrays for dense data and as SciPy sparse matrices for sparse
-data. While those may seem rather unsophisticated data representations
-when compared to more object-oriented
-constructs, such as those used by Weka~\citep{hall2009weka},
-they bring the prime advantage
-of allowing us to rely on all NumPy and SciPy vector operations -- often orders
-of magnitude faster that the corresponding Python loops -- while allowing us at
-the same time to keep the code short and readable. From a practical point of
-view, these formats also provide a bunch of data loading and conversion tools
-which make them very easy to use in many contexts\footnote{For tasks where the
-inputs are text files or semi-structured objects, we provide \textit{vectorizer}
-objects that efficiently convert such data to the NumPy or SciPy formats.}.
-Finally, this design choice has also been motivated by the fact that, given
-their pervasiveness in many other scientific Python packages, many scientific
-users of Python are likely to be already familiar with those formats, hence helping
-them to familiarize with the project.
+data. While those may seem rather unsophisticated data representations when
+compared to more object-oriented constructs, such as those used by
+Weka~\citep{hall2009weka}, they bring the prime advantage of allowing us to rely
+on all NumPy and SciPy vector operations -- often orders of magnitude faster
+that the corresponding Python loops -- while allowing us at the same time to
+keep the code short and readable.  This design choice has also been motivated by
+the fact that, given their pervasiveness in many other scientific Python
+packages, many scientific users of Python are already familiar with NumPy dense
+arrays and SciPy sparse matrices, hence helping them to familiarize with the
+project.  From a practical point of view, these formats also provide a bunch of
+data loading and conversion tools which make them very easy to use in many
+contexts. Moreover, for tasks where the inputs are text files or semi-structured
+objects, we provide \textit{vectorizer} objects that efficiently convert such
+data to the NumPy or SciPy formats.
 

### 2013-06-09 paper.tex
Remove footnote.

@@
 hyper-parameters. At initialization, all given parameters are attached to the
-estimator instance as public attributes\footnote{Python does not actually
-distinguish between public and private  attributes or methods. By convention
-however, any identifier prefixed with an underscore is considered a non-public
-part of the API~\citep{pythontut}.} and serve to determine the estimator
+estimator instance as public attributes and serve to determine the estimator
 behavior (e.g., the $C$ constant in SVMs). Parameters set as public attributes
 are also used for introspection purposes, e.g., in model selection routines. To
-ease their use, default hyper-parameter values are provided for all
-built-in estimators. These default values are set to be relevant in most common
-situations in order to make estimators as efficient as possible \textit{out-of-box}
-for non-experts. Finally, the constructor of an estimator never performs
-any actual learning, nor does it even see any training dataset. Its only goal is
-to set the parameter values and freeze the behavior of the estimator.
+ease their use, default hyper-parameter values are provided for all built-in
+estimators. These default values are set to be relevant in most common
+situations in order to make estimators as efficient as possible
+\textit{out-of-box} for non-experts. Finally, the constructor of an estimator
+never performs any actual learning, nor does it even see any training dataset.
+Its only goal is to set the parameter values and freeze the behavior of the
+estimator.
 

### 2013-06-09 paper.tex
X_train, y_train

@@
 
-The estimator interface defines a \texttt{fit} method. This method is
-called with training data (e.g., supplied as two arrays
-\texttt{X} and \texttt{y} in supervised learning estimators). Its task
-is to run a learning algorithm and to determine model-
-specific parameters from the training data and set these as attributes on the
-estimator object. As a convention, the parameters learned by an estimator are
-exposed as public attributes\footnote{Note that the learned parameters are
-different from the hyper-parameters given to the constructor. To distinguish
-between the two, learned parameters are all suffixed with a trailing underscore
-(e.g., \texttt{clf.coef\_} in the case of linear models is the vector of
+The estimator interface defines a \texttt{fit} method. This method is called
+with training data (e.g., supplied as two arrays \texttt{X\_train} and
+\texttt{y\_train} in supervised learning estimators). Its task is to run a
+learning algorithm and to determine model- specific parameters from the training
+data and set these as attributes on the estimator object. As a convention, the
+parameters learned by an estimator are exposed as public
+attributes\footnote{Note that the learned parameters are different from the
+hyper-parameters given to the constructor. To distinguish between the two,
+learned parameters are all suffixed with a trailing underscore (e.g.,
+\texttt{clf.coef\_} in the case of linear models is the vector of
 coefficients learned from the data).} on the trained object.

### 2013-06-09 paper.tex
Avoid footnote.

@@
 \textit{out-of-box} for non-experts. Finally, the constructor of an estimator
-never performs any actual learning, nor does it even see any training dataset.
-Its only goal is to set the parameter values and freeze the behavior of the
-estimator.
+never performs any actual learning (this is the role of the \texttt{fit} method,
+see below).  Its only goal is to set the parameter values and freeze the
+behavior of the estimator.
 
@@
 data and set these as attributes on the estimator object. As a convention, the
-parameters learned by an estimator are exposed as public
-attributes\footnote{Note that the learned parameters are different from the
-hyper-parameters given to the constructor. To distinguish between the two,
-learned parameters are all suffixed with a trailing underscore (e.g.,
-\texttt{clf.coef\_} in the case of linear models is the vector of
-coefficients learned from the data).} on the trained object.
+parameters learned by an estimator are exposed as public attributes whose name
+is suffixed with a trailing underscore (e.g., \texttt{coef\_} for the
+learned coefficients of a linear model).
+
+on the trained object.
 

### 2013-06-09 paper.tex
Import.

@@
 
-on the trained object.
-
 % Importantly, the parameters learned by an estimator
@@
 \begin{verbatim}
+from sklearn.linear_model import LogisticRegression
 clf = LogisticRegression(penalty="l1")
@@
 In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
-setting the \texttt{penalty} hyper-parameter to \texttt{l1}, hence
-enabling $\ell_1$
-regularization, and using default values for all other hyper-parameters. Upon
-\texttt{fit}, the model is then learned from the training arrays \texttt{X\_train} and
-\texttt{y\_train}. Also, since all estimators share the same interface, using a
-different learning algorithm is as simple as replacing the constructor. For
-example, to build a Random Forest on the same training data, one would simply
-have to replace \texttt{LogisticRegression(penalty="l1")} in the snippet above by
+setting the \texttt{penalty} hyper-parameter to \texttt{"l1"}, hence enabling
+$\ell_1$ regularization. Other hyper-parameters (such as \texttt{C}, which
+defines the strength of the regularization) are not explicitly defined and are
+thus set to the default values. Upon calling \texttt{fit}, the model is then
+learned from the training arrays \texttt{X\_train} and \texttt{y\_train}. Since
+all estimators share the same interface, using a different learning algorithm is
+as simple as replacing the constructor. For example, to build a Random Forest on
+the same training data, one would simply have to replace
+\texttt{LogisticRegression(penalty="l1")} in the snippet above by \\
 \texttt{RandomForestClassifier()}.

### 2013-06-09 paper.tex
KMeans example.

@@
 \textbf{Predictors.} The \textit{predictor} interface defines a \texttt{predict}
-method taking as input some dataset \texttt{X} and producing as output
-predictions for \texttt{X} based on the learned parameters of the estimator. In
-the case of supervised learning estimators, this method typically returns the
-predicted labels or values computed by the model. Some unsupervised learning
-estimators may however also implement that interface: for example, in clustering
-estimators, \texttt{predict} returns the identifiers of the closest clusters for
-the samples in \texttt{X}.
+method taking as input some new data \texttt{X\_test} and producing as output
+predictions for \texttt{X\_test} based on the learned parameters of the
+estimator. In the case of supervised learning estimators, this method typically
+returns the predicted labels or values computed by the model.  Continuing with
+the previous example, predicted labels for \texttt{X\_test} can be obtained
+using the following snippet:
+\begin{verbatim}
+y_pred = clf.predict(X_test)
+\end{verbatim}
+Some unsupervised learning
+estimators may however also implement that interface, as illustrated below:
+\begin{verbatim}
+from sklearn.cluster import KMeans
+km = KMeans(n_clusters=10)
+km.fit(X_train)
+clust_pred = km.predict(X_test)
+\end{verbatim}
+In the above example, a \texttt{KMeans} estimator is first fit on the training
+data \texttt{X\_train}. Then the \texttt{predict} method is called on the test
+data \texttt{X\_test} in order to retrieve predicted cluster labels.
+The \texttt{predict} method in \textit{scikit-learn} thus offers a unified way
+to compute predictions with respect to unseen data.
 
@@
 given data.
-
-Continuing with the previous example, predictions on unseen data
-\texttt{X\_test}, \texttt{y\_test} can therefore be obtained and evaluated using
-the following snippet:
-\begin{verbatim}
-y_pred = clf.predict(X_test)
-score = clf.score(X_test, y_test)
-\end{verbatim}
 

### 2013-06-09 paper.tex
Method chaining.

@@
 method taking as input some new data \texttt{X\_test} and producing as output
-predictions for \texttt{X\_test} based on the learned parameters of the
-estimator. In the case of supervised learning estimators, this method typically
-returns the predicted labels or values computed by the model.  Continuing with
-the previous example, predicted labels for \texttt{X\_test} can be obtained
-using the following snippet:
+predictions for \texttt{X\_test}, based on the learned parameters of the
+estimator (we call the input to \texttt{predict} ``\texttt{X\_test}'' in order
+to emphasize that \texttt{predict} generalizes to new data; however,
+\texttt{predict} can also be called on \texttt{X\_train}). In the case of
+supervised learning estimators, this method typically returns the predicted
+labels or values computed by the model.  Continuing with the previous example,
+predicted labels for \texttt{X\_test} can be obtained using the following
+snippet:
 \begin{verbatim}
@@
 predictors also provide a \texttt{predict\_proba} method which returns
-class probabilities for the input samples \texttt{X}.
+class probabilities for the input samples \texttt{X\_test}.
 
 Finally, predictors should also provide a \texttt{score} function to assess the
-goodness of fit on some input data. In supervised learning estimators, this
-method takes as inputs a dataset \texttt{X} and \texttt{y} and typically
-computes the coefficient of determination between \texttt{y} and
-\texttt{predict(X)} in the case of classification or the mean squared error in
-the case of regression. Not however that, from an API point of view, the only
-guideline is that the \texttt{score} method returns a numerical value such that
-the higher the better. As such, some unsupervised learning estimators also
-exposes a \texttt{score} function, computing for instance the likelihood of the
-given data.
+performance on some input data. In supervised learning estimators, this method
+takes as inputs a dataset \texttt{X\_test} and \texttt{y\_test} and typically
+computes the coefficient of determination between \texttt{y\_test} and
+\texttt{predict(X\_test)} in the case of classification or the mean squared
+error in the case of regression. Not however that, from an API point of view,
+the only guideline is that the \texttt{score} method returns a numerical value
+such that the higher the better. As such, some unsupervised learning estimators
+also exposes a \texttt{score} function, computing for instance the likelihood of
+the given data.
 
@@
 \textit{transformer} interface which defines a \texttt{transform} method. It
-takes as input some dataset \texttt{X} and yields as output a transformed
-version of \texttt{X}\footnote{From an implementation point of view, it may
-sometimes be more efficient to implement \texttt{fit} and \texttt{transform} in
-a single method. In that case, transformers may expose an additional
-\texttt{fit\_transform} method.}. In particular, all preprocessing algorithms
-and feature selection routines are offered as transformers within the library.
+takes as input some new data \texttt{X\_test} and yields as output a transformed
+version of \texttt{X\_test} Preprocessing, feature selection and dimensionality
+reduction algorithms are all provided as transformers within the library.  In
+our example, if one wanted to preprocess the inputs \texttt{X\_train} before
+fitting the logistic regression estimator, e.g., by standardizing the feature
+values, one could thus simply write:
+\begin{verbatim}
+scaler = StandardScaler()
+scaler.fit(X_train)
+X_train = scaler.transform(X_train)
+\end{verbatim}
+Of course, in practice, it is important to apply the same preprocessing to the
+test data \texttt{X\_test}. This can easily be done as follows:
+\begin{verbatim}
+X_test = scaler.transform(X_test)
+\end{verbatim}
 
-In our example, if one wanted to transform the inputs \texttt{X\_train} before
-fitting the logistic regression estimator, e.g., by standardizing the feature values,
-one could thus simply write:
+\subsection{Method chaining}
+
+In \textit{scikit-learn}, the \texttt{fit} method always returns the estimator
+it was called on, hence making method chaining possible.  Using this technique,
+we can thus rewrite the \texttt{StandardScaler} example above in a single line:
 \begin{verbatim} 
@@
 \end{verbatim} 
-
-Note that the snippet relies on the fact that the \texttt{fit} method always
-returns the object it was called on , hence making method chaining  possible.
+In \textit{scikit-learn}, \texttt{fit(X\_train).transform(X\_train)} can further
+be rewritten as \texttt{fit\_transform(X\_train)}.  Combining the \texttt{fit}
+and \texttt{transform} methods into a single method \texttt{fit\_transform}
+allows to share computations, which often makes a huge difference in practice.
+In the same spirit, \textit{scikit-learn} sometimes provides the
+\texttt{fit\_predict} method for combining the \texttt{fit} and \texttt{predict}
+methods. This is especially useful for retrieving cluster predictions with
+respect to the training data \texttt{X\_train} in clustering algorithms.
 
@@
 
-\subsection{Composition}
+\subsection{Estimator composition}
 

### 2013-06-09 paper.tex
Import.
iter

@@
 \begin{verbatim}
+from sklearn.preprocessing import StandardScaler
 scaler = StandardScaler()
@@
 \end{verbatim}
+Transformers also include a variety of learning algorithms, such as
+dimension reduction (PCA, manifold learning techniques), Kernel
+approximation techniques, and any other approach that can be used to
+derive a modified representation of the data.
 
@@
 \end{verbatim} 
-In \textit{scikit-learn}, \texttt{fit(X\_train).transform(X\_train)} can further
+In addition, \texttt{fit(X\_train).transform(X\_train)} can further
 be rewritten as \texttt{fit\_transform(X\_train)}.  Combining the \texttt{fit}
@@
 for scikit-learn developers and users alike.
-The \texttt{scipy.sparse} package offers quite full-featured support
+The \texttt{scipy.sparse} package offers support
 for this kind of data structure,

### 2013-06-10 paper.tex
refactor \textit{scikit-learn} to a command

@@
 \usepackage{xcolor}
+\usepackage{xspace}
+
+\newcommand{\sklearn}{\textit{scikit-learn}\xspace}
 
@@
 \title{Python as a machine learning language:
-       experiences from the \textit{scikit-learn} project}
+       experiences from the \sklearn project}
 
@@
 \begin{abstract}
-The \textit{scikit-learn} project is an increasingly popular open source machine
+The \sklearn project is an increasingly popular open source machine
 learning library designed to be simple and efficient, accessible to non-experts,
@@
 
-\section{The \textit{scikit-learn} project}
+\section{The \sklearn project}
 
-The \textit{scikit-learn} project~\citep{pedregosa2011} is an open source
+The \sklearn project~\citep{pedregosa2011} is an open source
 machine learning library for the Python programming language. The ambition of
@@
 of numeric and scientific packages centered around the Numpy and
-SciPy~\citep{vanderwalt2011, varoquaux2013scipy} libraries. As such, \textit
-{scikit-learn} is intended to be used within Python scripts and requires some
-Python programming knowledge. At the time, it offers no command line or
-graphical user interface.
+SciPy~\citep{vanderwalt2011, varoquaux2013scipy} libraries.
+As such, \sklearn is intended to be used within Python scripts
+and requires some Python programming knowledge.
+At the time, it offers no command line or graphical user interface.
 % XXX Gael: maybe we should say that it augments a general purpose
@@
 
-Started in 2009, \textit{scikit-learn} is developed by a large and
+Started in 2009, \sklearn is developed by a large and
 international team of volunteers, mostly researchers from
@@
 In this section, we discuss the general principles of the application
-programming interface (API) of \textit{scikit-learn}. We first specify the
+programming interface (API) of \sklearn. We first specify the
 representation of data and then proceed by describing the interface shared by
@@
 
-In \textit{scikit-learn}, we chose a representation of data that is as close as
+In \sklearn, we chose a representation of data that is as close as
 possible to the matrix representation: datasets are encoded as NumPy
@@
 
-All objects within \textit{scikit-learn} share a uniform common basic API
+All objects within \sklearn share a uniform common basic API
 consisting of three complementary interfaces: an \textit{estimator} interface
@@
 
-In \textit{scikit-learn}, classical learning algorithms are not the only objects
+In \sklearn, classical learning algorithms are not the only objects
 to be implemented as estimators. For example, preprocessing routines (e.g.,
@@
 data \texttt{X\_test} in order to retrieve predicted cluster labels.
-The \texttt{predict} method in \textit{scikit-learn} thus offers a unified way
+The \texttt{predict} method in \sklearn thus offers a unified way
 to compute predictions with respect to unseen data.
@@
 
-In \textit{scikit-learn}, the \texttt{fit} method always returns the estimator
+In \sklearn, the \texttt{fit} method always returns the estimator
 it was called on, hence making method chaining possible.  Using this technique,
@@
 allows to share computations, which often makes a huge difference in practice.
-In the same spirit, \textit{scikit-learn} sometimes provides the
+In the same spirit, \sklearn sometimes provides the
 \texttt{fit\_predict} method for combining the \texttt{fit} and \texttt{predict}
@@
 
-A distinguishing feature of the \textit{scikit-learn} API is its ability to
+A distinguishing feature of the \sklearn API is its ability to
 compose new estimators from several base estimators. Composition mechanisms can
@@
 
-% A distinguishing feature of \textit{scikit-learn} estimators
+% A distinguishing feature of \sklearn estimators
 % is that they can be composed into \texttt{Pipeline} objects.

### 2013-06-10 paper.tex
no need for nbsp before citations

@@
 \usepackage{hyperref}
-\usepackage[authoryear]{natbib}
+\usepackage[authoryear,round]{natbib}
 \usepackage{xcolor}
@@
 
-The \sklearn project~\citep{pedregosa2011} is an open source
+The \sklearn project \citep{pedregosa2011} is an open source
 machine learning library for the Python programming language. The ambition of
@@
 of numeric and scientific packages centered around the Numpy and
-SciPy~\citep{vanderwalt2011, varoquaux2013scipy} libraries.
+SciPy \citep{vanderwalt2011, varoquaux2013scipy} libraries.
 As such, \sklearn is intended to be used within Python scripts
 and requires some Python programming knowledge.
-At the time, it offers no command line or graphical user interface.
+At present, it offers no command line or graphical user interface.
 % XXX Gael: maybe we should say that it augments a general purpose
@@
 compared to more object-oriented constructs, such as those used by
-Weka~\citep{hall2009weka}, they bring the prime advantage of allowing us to rely
+Weka \citep{hall2009weka}, they bring the prime advantage of allowing us to rely
 on all NumPy and SciPy vector operations -- often orders of magnitude faster

### 2013-06-10 paper.tex
quantify dev team and code review

@@
 
-Started in 2009, \sklearn is developed by a large and
-international team of volunteers, mostly researchers from
-various fields (e.g., machine learning, natural language processing, or
-neuroscience). The project also greatly benefits from many occasional
+Started in 2009, \sklearn is developed by an international team
+of several dozen volunteers, mostly researchers from various fields
+(e.g., machine learning, natural language processing, or neuroscience).
+The project also greatly benefits from many occasional
 contributors proposing small code or documentation fixes or improvements.
@@
 addition, complete documentation and examples are required for all
-features. Finally, any addition to the code base goes through rigorous
-code review.
+features.
+Finally, additions to the code base must pass code review
+by at least two developers not involved in the implementation
+of the proposed addition.
 
@@
 and mailing lists.  At the time of writing, the project is watched by 1309
-people and forked 667 time on github, and the mailing-list receives 300 mails
+people and forked 667 time on GitHub, and the mailing-list receives 300 mails
 per month.
+The version control logs for the development version of scikit-learn
+% ddaa494c116e3c16bf032003c5cccbed851733d2
+show 183 contributors to the project's source code and documentation
+(after manual deduplication).
 
@@
 % lists 322 scikit-learn-related questions.
-% The version control logs for the development version of scikit-learn
-% % ddaa494c116e3c16bf032003c5cccbed851733d2
-% show 183 contributors to the project's source code and documentation
-% (after manual deduplication).
 

### 2013-06-10 paper.tex
page numbers
capitalization

@@
 \usepackage{xspace}
+
+\pagestyle{headings}
 
@@
 Transformers also include a variety of learning algorithms, such as
-dimension reduction (PCA, manifold learning techniques), Kernel
+dimension reduction (PCA, manifold learning techniques), kernel
 approximation techniques, and any other approach that can be used to

### 2013-06-10 paper.tex
library vs. CLI/GUI toolkit (suggest IPython)
Also address @GaelVaroquaux's comment.

@@
 it includes classical learning algorithms, model selection tools and
-preprocessing
-procedures. From its core, the library has been designed to tie in with the set
-of numeric and scientific packages centered around the Numpy and
+preprocessing procedures.
+
+\sklearn is explicitly a library,
+i.e. a collection of classes (in the object-oriented sense)
+that users ``import'' into Python scripts/programs.
+Use of the package requires basic Python programming knowledge,
+and no command-line interface, let alone a GUI,
+is offered for non-programmer users.
+Instead, interactive use is made possible
+by the Python interactive interpreter,
+and its enhanced replacement IPython \citep{perez2007ipython},
+which offers a \textsc{matlab}-like working environment
+specifically designed for scientific use.
+
+From its core, the library has been designed to tie in with the set
+of numeric and scientific packages centered around the NumPy and
 SciPy \citep{vanderwalt2011, varoquaux2013scipy} libraries.
-As such, \sklearn is intended to be used within Python scripts
-and requires some Python programming knowledge.
-At present, it offers no command line or graphical user interface.
-% XXX Gael: maybe we should say that it augments a general purpose
-% language with machine learning concepts, like numpy augments it with
-% array-computing concepts.
+Of these, NumPy augments Python with a numeric array datatype
+and associated array computing primitives \citep{vanderwalt2011},
+while SciPy extends it further
+by making common numeric operations available,
+either by implementing those in Python/NumPy,
+or by wrapping existing C/C{}\verb!++!/Fortran implementations.
+In a similar vein, \sklearn extends the general-purpose language Python
+with a set of composable machine learning operations.
 

### 2013-06-10 paper.tex
minor stuff

@@
 computing both linear PCA and kernel PCA components on \texttt{X\_train} (through a
-\texttt{FeatureUninon}), selecting the best of those features and then feeding
+\texttt{FeatureUnion}), selecting the best of those features and then feeding
 them to a logistic regression model:
@@
 
-As the above snippet shows, \texttt{Pipeline} and \texttt{FeatureUninon} can be
+As the above snippet shows, \texttt{Pipeline} and \texttt{FeatureUnion} can be
 combined to create complex objects. In particular, since both are themselves

### 2013-06-10 paper.tex
be consistent in spelling \ell_1
Also turned off hyperref as it consistently breaks stuff.
duplicate citations

@@
 \usepackage{amsmath}
-\usepackage{hyperref}
+%\usepackage{hyperref}
 \usepackage[authoryear,round]{natbib}
@@
 From its core, the library has been designed to tie in with the set
-of numeric and scientific packages centered around the NumPy and
-SciPy \citep{vanderwalt2011, varoquaux2013scipy} libraries.
-Of these, NumPy augments Python with a numeric array datatype
-and associated array computing primitives \citep{vanderwalt2011},
-while SciPy extends it further
+of numeric and scientific packages centered around the NumPy and SciPy libraries.
+Of these, NumPy \citep{vanderwalt2011} augments Python
+with a numeric array datatype and associated array computing primitives,
+while SciPy \citep{varoquaux2013scipy} extends it further
 by making common numeric operations available,
@@
   \times & \; n_\textsf{features} \in \{1000, 2000, 5000\}      \\
-  \times & \; \textsf{norm} \in \{L_1, L_2\}                    \\
+  \times & \; \textsf{norm} \in \{\ell_1, \ell_2\}              \\
   \times & \; C \in \{10, 100, 1000\}

### 2013-06-10 paper.tex
praise GitHub for making collaboration easy

@@
 contributors proposing small code or documentation fixes or improvements.
-Because of this large base of developers, emphasis is put on keeping
-the project maintainable. In particular, code must follow specific
+Development proceeds on GitHub (\url{https://github.com/scikit-learn}),
+a platform which greatly facilitates this kind of collaboration.
+Because of the large number of developers,
+particular emphasis is put on keeping the project maintainable.
+In particular, code must follow specific
 quality guidelines, such as style consistency and unit-test coverage. In

### 2013-06-10 paper.tex
capitalize Kaggle

@@
 \citep{pedregosa2011}, successes in various machine learning challenges
-(\emph{e.g.} kaggle), or significant statistics derived from our repositories
+(\emph{e.g.} Kaggle), or significant statistics derived from our repositories
 and mailing lists.  At the time of writing, the project is watched by 1309

### 2013-06-10 paper.tex
attributes -> features per scikit-learn convention
Attributes is ambiguous in Python context.

@@
 example, in a classification or regression task, the goal is to find a mapping
-between input variables $X_1, ..., X_p$ -- called attributes or features -- and
+between input variables $X_1, ..., X_p$, called features, and
 some output variable $Y$. A sample is then defined as a tuple of values $([x_1,
@@
 to be implemented as estimators. For example, preprocessing routines (e.g.,
-vectorization of text documents, scaling of attributes, etc) or feature
+vectorization of text documents, scaling of features, etc) or feature
 extraction techniques also implement the same interface. Even those estimators

### 2013-06-10 paper.tex
clarify pipeline fitting
Also, the concept of "grid-searching" is not yet introduced.

@@
 estimator (we call the input to \texttt{predict} ``\texttt{X\_test}'' in order
-to emphasize that \texttt{predict} generalizes to new data; however,
-\texttt{predict} can also be called on \texttt{X\_train}). In the case of
+to emphasize that \texttt{predict} generalizes to new data). In the case of
 supervised learning estimators, this method typically returns the predicted
@@
 a fixed sequence of steps in processing data (e.g., feature extraction,
-centering / scaling / normalization, feature selection, learning and making
-predictions). For convenience reasons, those steps can be combined into a
-pipeline such that \texttt{fit} and \texttt{predict} need to be called only once
-to fit the whole sequence. A pipeline also allows for the hyper-parameters of
-all the base estimators to be grid-searched all at once. From an API point of
-view, a \texttt{Pipeline} object takes as input a sequence of $N$ estimators,
-where all $N-1$ first are transformers and the last is either a predictor or a
-transformer. Calling \texttt{fit} on the pipeline is the same as calling
-\texttt{fit} on each estimator in turn, \texttt{transform} the input and then
-pass it on to the next step. The pipeline exposes all the methods the last
+normalization, feature selection, learning and making predictions).
+A sequence of $N$ such steps can be combined into a composite object
+called a \texttt{Pipeline}.
+The first $N-1$ steps must be transformers
+and the last can be either a predictor or a transformer.
+Fitting a pipeline to a training set amounts, conceptually,
+to the following recursive procedure:
+\begin{enumerate}
+  \item When only one step remains, call its \texttt{fit} method.
+  \item Otherwise, fit the first step,
+        use it to \texttt{transform} the training set
+        and fit the rest of the pipeline with the transformed data.
+\end{enumerate}
+The pipeline exposes all the methods the last
 estimator in the pipe exposes. That is, if the last estimator is a predictor,
-the pipeline can be used itself as a predictor. If the last estimator is a
+the pipeline can itself be used as a predictor. If the last estimator is a
 transformer, then the pipeline is itself a transformer. \textcolor{red}{TODO:

### 2013-06-11 paper.tex
missing full stop

@@
 takes as input some new data \texttt{X\_test} and yields as output a transformed
-version of \texttt{X\_test} Preprocessing, feature selection and dimensionality
+version of \texttt{X\_test}.
+Preprocessing, feature selection and dimensionality
 reduction algorithms are all provided as transformers within the library.  In

### 2013-06-11 paper.tex
missing full stop

@@
 \title{Python as a machine learning language:
-       experiences from the \sklearn project}
+       experiences from the scikit-learn project}
 
@@
 design choices for the application programming interface (API) of the project.
-In particular, we describe the simple and limited interface shared by all
+In particular, we describe the simple and elegant interface shared by all
 learning and processing units in the library and then discuss its advantages in
 terms of composition and reusability. The paper also comments on implementation
-details specific to the Python ecosystem and highlights obstacles faced by users
+details specific to the Python ecosystem and analyzes obstacles faced by users
 and developers of the library.
@@
 
-\section{The \sklearn project}
+\section{The scikit-learn project}
 
@@
 Started in 2009, \sklearn is developed by an international team
-of several dozen volunteers, mostly researchers from various fields
+of a dozen of core developers, mostly researchers from various fields
 (e.g., machine learning, natural language processing, or neuroscience).
@@
 
-In this paper, we describe and motivate the design choices used in
-building the library. First, in section \ref{sec:api}, we give the
+In a previous paper \citep{pedregosa2011}, \sklearn was briefly presented and
+benchmarked against several competitors. In this paper, we instead focus on an
+in-depth discussion and analysis of the design choices used in
+building the library. First, in section \ref{sec:api}, we present the
 central application programming interface (API). Then, in section
@@
 technologies we rely on; and finally, in section \ref{sec:comparison}, we
-position \emph{scikit-learn} with respect to other major projects.
+compare \emph{scikit-learn} to other major projects in terms of API.
 
@@
 be used to combine typical machine learning workflows into a single object which
-is itself an estimator. Through duck typing, composed
+is itself an estimator. Thanks to duck typing, composed
 estimators can be used wherever usual estimators can be used. In particular,
@@
 As an example, the following snippet illustrates how to create a complex worflow
-computing both linear PCA and kernel PCA components on \texttt{X\_train} (through a
-\texttt{FeatureUnion}), selecting the best of those features and then feeding
-them to a logistic regression model:
+computing both linear PCA and kernel PCA components on \texttt{X\_train}
+(through a \texttt{FeatureUnion}), selecting the 100 best of those features and
+then feeding them to an $\ell_2$-regularized logistic regression model:
 \begin{verbatim}
-Pipeline([("pca", FeatureUnion([('linear_pca', PCA()),
-                                ('kernel_pca', KernelPCA())])),
-          ("feature_selection", SelectKBest()),
-          ("logistic_model", LogisticRegression())
+from sklearn.pipeline import Pipeline
+from sklearn.decomposition import PCA, KernelPCA
+from sklearn.feature_selection import SelectKBest
+
+Pipeline([("pca", FeatureUnion([('pca', PCA()),
+                                ('kpca', KernelPCA(kernel="rbf"))])),
+          ("feature_selection", SelectKBest(k=100)),
+          ("logistic_model", LogisticRegression(penalty="l2"))
 ]).fit(X_train, y_train).predict(X_test)

### 2013-06-11 paper.tex
some general principles for the API

--- 2013-06-11 paper.tex
@@
 \title{Python as a machine learning language:
-       experiences from the scikit-learn project}
+       experiences from the \sklearn project}
 
@@
 design choices for the application programming interface (API) of the project.
-In particular, we describe the simple and elegant interface shared by all
+In particular, we describe the simple and limited interface shared by all
 learning and processing units in the library and then discuss its advantages in
 terms of composition and reusability. The paper also comments on implementation
-details specific to the Python ecosystem and analyzes obstacles faced by users
+details specific to the Python ecosystem and highlights obstacles faced by users
 and developers of the library.
@@
 
-\section{The scikit-learn project}
+\section{The \sklearn project}
 
@@
 Started in 2009, \sklearn is developed by an international team
-of a dozen of core developers, mostly researchers from various fields
+of several dozen volunteers, mostly researchers from various fields
 (e.g., machine learning, natural language processing, or neuroscience).
@@
 advantages of that interface.
+
+The API is designed to adhere to the following broad principles:
+\begin{description}
+  \item[Inspectability.]
+       \sklearn objects store their constructor parameters
+       as well as parameter values determined by learning algorithms
+       as public attributes in a documented format.
+       (The Python language does not actually distinguish
+       between public and private attributes of objects,
+       but by convention, any identifier
+       prefixed with an underscore (\texttt{\_})
+       is considered ``a non-public part of the API''; \citealp{pythontut}.)
+  \item[No proliferation of classes.]
+       Learning algorithms are objects of custom classes,
+       but datasets and hyperparameter names and values,
+       are represented as standard Python, NumPy and SciPy classes
+       rather than custom ones whenever feasible.
+       Custom classes adhere to Python conventions and protocols
+       for constructs such as iteration over the ``folds''
+       in a $k$-fold cross validation.
+       This should keep \sklearn easy to use
+       and easy to combine with other libraries.
+       % XXX do we need to elaborate more on this, or do we expect everyone
+       % to be an experienced Java/adapter pattern hater?
+  \item[Composability.]
+       Many machine learning tasks are expressible
+       as \textit{pipelines} of transformations to data.
+       Other tasks are naturally viewed
+       as modifications to other tasks;
+       e.g., multi-label classification
+       can be reduced to multiple binary classification tasks.
+       Accordingly, multi-label classification
+       is implemented by a class that takes an existing classifier object
+       as input and modifies its function.
+  \item[Sensible defaults.]
+       Whenever an operation requires a user-defined parameter,
+       a default value for that parameter must be defined by the library
+       if at all possible,
+       and the default should cause the operation to be performed
+       in a sensible way (giving a baseline solution for the task at hand).
+\end{description}
 
@@
 % takes a set of named (hyper-)parameters
-% and attaches these to the estimator as public attributes.\footnote{
-%   The Python language does not actually distinguish between public and private
-%   attributes and methods,
-%   but by convention, any identifier prefixed with an underscore (\texttt{\_})
-%   is considered ``a non-public part of the API\ldots
-%   an implementation detail and subject to change without notice.''
-%   \Citep{pythontut}
-%   }
+% and attaches these to the estimator as public attributes.
 % It never performs actual learning, nor does it even see a training set.

### 2013-06-11 paper.tex
remove subsection about method chaining (too detailed)

@@
 advantages of that interface.
-
-The API is designed to adhere to the following broad principles:
-\begin{description}
-  \item[Inspectability.]
-       \sklearn objects store their constructor parameters
-       as well as parameter values determined by learning algorithms
-       as public attributes in a documented format.
-       (The Python language does not actually distinguish
-       between public and private attributes of objects,
-       but by convention, any identifier
-       prefixed with an underscore (\texttt{\_})
-       is considered ``a non-public part of the API''; \citealp{pythontut}.)
-  \item[No proliferation of classes.]
-       Learning algorithms are objects of custom classes,
-       but datasets and hyperparameter names and values,
-       are represented as standard Python, NumPy and SciPy classes
-       rather than custom ones whenever feasible.
-       Custom classes adhere to Python conventions and protocols
-       for constructs such as iteration over the ``folds''
-       in a $k$-fold cross validation.
-       This should keep \sklearn easy to use
-       and easy to combine with other libraries.
-       % XXX do we need to elaborate more on this, or do we expect everyone
-       % to be an experienced Java/adapter pattern hater?
-  \item[Composability.]
-       Many machine learning tasks are expressible
-       as \textit{pipelines} of transformations to data.
-       Other tasks are naturally viewed
-       as modifications to other tasks;
-       e.g., multi-label classification
-       can be reduced to multiple binary classification tasks.
-       Accordingly, multi-label classification
-       is implemented by a class that takes an existing classifier object
-       as input and modifies its function.
-  \item[Sensible defaults.]
-       Whenever an operation requires a user-defined parameter,
-       a default value for that parameter must be defined by the library
-       if at all possible,
-       and the default should cause the operation to be performed
-       in a sensible way (giving a baseline solution for the task at hand).
-\end{description}
 
@@
 
-\subsection{Method chaining}
-
-In \sklearn, the \texttt{fit} method always returns the estimator
-it was called on, hence making method chaining possible.  Using this technique,
-we can thus rewrite the \texttt{StandardScaler} example above in a single line:
-\begin{verbatim} 
-X_train = StandardScaler().fit(X_train).transform(X_train)
-\end{verbatim} 
-In addition, \texttt{fit(X\_train).transform(X\_train)} can further
-be rewritten as \texttt{fit\_transform(X\_train)}.  Combining the \texttt{fit}
-and \texttt{transform} methods into a single method \texttt{fit\_transform}
-allows to share computations, which often makes a huge difference in practice.
-In the same spirit, \sklearn sometimes provides the
-\texttt{fit\_predict} method for combining the \texttt{fit} and \texttt{predict}
-methods. This is especially useful for retrieving cluster predictions with
-respect to the training data \texttt{X\_train} in clustering algorithms.
-
 % The bulk of the scikit-learn application programming interface
@@
 % takes a set of named (hyper-)parameters
-% and attaches these to the estimator as public attributes.
+% and attaches these to the estimator as public attributes.\footnote{
+%   The Python language does not actually distinguish between public and private
+%   attributes and methods,
+%   but by convention, any identifier prefixed with an underscore (\texttt{\_})
+%   is considered ``a non-public part of the API\ldots
+%   an implementation detail and subject to change without notice.''
+%   \Citep{pythontut}
+%   }
 % It never performs actual learning, nor does it even see a training set.

### 2013-06-11 paper.tex
comment: implem. detail in API section?

@@
 Weka \citep{hall2009weka}, they bring the prime advantage of allowing us to rely
+% Lars: "orders of magnitude faster" is an implementation detail
 on all NumPy and SciPy vector operations -- often orders of magnitude faster

### 2013-06-11 paper.tex
principles: address @mblondel's comments

@@
        is considered ``a non-public part of the API''; \citealp{pythontut}.)
-  \item[No proliferation of classes.]
+  \item[Non-proliferation of classes.]
        Learning algorithms are objects of custom classes,
@@
        as \textit{pipelines} of transformations to data.
-       Other tasks are naturally viewed
-       as modifications to other tasks;
-       e.g., multi-label classification
-       can be reduced to multiple binary classification tasks.
-       Accordingly, multi-label classification
-       is implemented by a class that takes an existing classifier object
-       as input and modifies its function.
+       Also, some learning algorithms are naturally viewed
+       as meta-algorithms parameterized on other algorithms.
+       To be able to implement and use these algorithms in a simple way,
+       all learners are usable with a consistent set of methods
+       and a simple syntax is implemented
+       for setting/retrieving parameters of ``sub-estimators''.
+       % XXX need to describe this!
   \item[Sensible defaults.]

### 2013-06-11 paper.tex
elaborate on effect of array implementation to API
Also, at least mention the online learning algorithms that we have.

@@
 % Lars: "orders of magnitude faster" is an implementation detail
-on all NumPy and SciPy vector operations -- often orders of magnitude faster
-that the corresponding Python loops -- while allowing us at the same time to
+on all NumPy and SciPy vector operations while allowing us at the same time to
 keep the code short and readable.  This design choice has also been motivated by
@@
 data to the NumPy or SciPy formats.
+
+The public API is oriented toward batch processing
+to enable efficient implementation.
+While classification and regression algorithms
+can make predictions for single samples,
+and some algorithms in \sklearn can perform online learning
+(e.g. stochastic gradient descent for linear SVMs and logistic regression,
+an online variant of $k$-means based on \citealt{sculley2010web}),
+these algorithm are really intended to be used on small batches of samples.
+Doing so makes optimal use of NumPy and SciPy
+because they prevent the overhead inherent in Python function calls
+and the language's per-element dynamic type checking.
 

### 2013-06-11 paper.tex
trim the weaknesses section a bit
Needs further trimming and merge with implementation section.

--- 2013-06-11 paper.tex
@@
 on a per-estimator basis.
-This situation is aggravated by the lack of a C API for sparse matrices,
-so that Cython modules have to implement their own version
-of operations like sparse dot products
-that ideally would have been handled by a library.
-
-However, users of scikit-learn cannot always be expected
-to have similar familiarity, and in some cases
-(e.g., when users want to leverage feature extraction code
-outside the context of a \texttt{Pipeline})
-they will have to face the \texttt{scipy.sparse} interface.
-This is, of course, as much an artifact of the design choice
-to not implement separate classes and interfaces for samples/instances
-as it a consequence of SciPy's implementation of sparse matrices.
 
@@
 because of the copying of training and test sets
-into the various processes' address spaces.\footnote{
-  The fact that copying occurs at all is an artifact of NumPy's implementation;
-  no attempt at implementing a multiprocessing module
-  that shares NumPy arrays using copy-on-write semantics or memory mapping
-  has so far been completed.}
+into the various processes' address spaces.
 
-\subsection{Python/Cython mismatch}
-
-The Python/Cython divide is another source of potential difficulties
-both for users and developers/contributors.
 While Cython makes writing performance-critical code possible
@@
 whose sizes are practically all platform-dependent.
-Choosing the appropriate integer type from among
-\texttt{int}, \texttt{size\_t}, \texttt{Py\_ssize\_t} and \texttt{npy\_intp}
-can require knowledge of the Python extension API,
-the NumPy C API and the C standard.
-Neither Cython nor C compilers provide much in the way of static checking
-for the right type,
-while picking the wrong type may cause segmentation faults
-when users try to use arrays that are too large for the type used
-and using a dynamic type often leads to unacceptable slowdowns.
 

### 2013-06-11 paper.tex
start comparison section

@@
 
+Recent years have witnessed a rising interest in machine learning and data mining
+with applications in many fields.
+With this rise comes a host of machine learning packages
+(both open source and proprietary) with which \sklearn competes.
+
+% XXX Is the Julia comparison fair?
+% The goal of that language seems to be a bit different:
+% to provide a compiler to implement numeric algorithms without vectorized idioms,
+% maybe more like Cython than Python...
+In comparison to specialized languages for numeric and statistical programming
+such as \textsc{matlab}, R \citep{trancon2012r} and Julia \citep{bezanson2012julia},
+the use of Python as a host language has the distinct advantage
+that it is a mature, \textit{general purpose} language:
+it has strong language and standard library support for such tasks as
+string handling/text processing, interprocess communication, networking
+and many of the other tasks that many machine learning programs
+(whether academic or commercial)
+will have to perform ``on the side''.
+In addition, third-party libraries are available for virtually any task
+that modern computer programs routinely perform.
+
 \section{Future work}

### 2013-06-12 paper.tex
wording

@@
 
-\sklearn is explicitly a library,
-i.e. a collection of classes (in the object-oriented sense)
-that users ``import'' into Python scripts/programs.
-Use of the package requires basic Python programming knowledge,
-and no command-line interface, let alone a GUI,
-is offered for non-programmer users.
-Instead, interactive use is made possible
-by the Python interactive interpreter,
-and its enhanced replacement IPython \citep{perez2007ipython},
-which offers a \textsc{matlab}-like working environment
-specifically designed for scientific use.
+The project is a programming library, i.e. a collection of classes (in the
+object-oriented sense) that users import into Python programs. Its use
+therefore requires basic Python programming knowledge. No command-line
+interface, let alone a GUI, is offered for non-programmer users. Instead,
+interactive use is made possible by the Python interactive interpreter, and its
+enhanced replacement IPython \citep{perez2007ipython}, which offers a
+\textsc{matlab}-like working environment specifically designed for scientific
+use.
 
-From its core, the library has been designed to tie in with the set
-of numeric and scientific packages centered around the NumPy and SciPy libraries.
-Of these, NumPy \citep{vanderwalt2011} augments Python
-with a numeric array datatype and associated array computing primitives,
-while SciPy \citep{varoquaux2013scipy} extends it further
-by making common numeric operations available,
-either by implementing those in Python/NumPy,
-or by wrapping existing C/C{}\verb!++!/Fortran implementations.
-In a similar vein, \sklearn extends the general-purpose language Python
-with a set of composable machine learning operations.
+From its core, the library has been designed to tie in with the set of numeric
+and scientific packages centered around the NumPy and SciPy libraries. NumPy
+\citep{vanderwalt2011} augments Python with a numeric array datatype and
+associated array computing primitives while SciPy \citep{varoquaux2013scipy}
+extends it further by making common numeric operations available, either by
+implementing those in Python/NumPy or by wrapping existing
+C/C{}\verb!++!/Fortran implementations. In a similar vein, \sklearn extends
+Python with a set of composable machine learning operations.
 
-Started in 2009, \sklearn is developed by an international team
-of a dozen of core developers, mostly researchers from various fields
-(e.g., machine learning, natural language processing, or neuroscience).
-The project also greatly benefits from many occasional
-contributors proposing small code or documentation fixes or improvements.
-Development proceeds on GitHub (\url{https://github.com/scikit-learn}),
-a platform which greatly facilitates this kind of collaboration.
-Because of the large number of developers,
-particular emphasis is put on keeping the project maintainable.
-In particular, code must follow specific
-quality guidelines, such as style consistency and unit-test coverage. In
-addition, complete documentation and examples are required for all
-features.
-Finally, additions to the code base must pass code review
-by at least two developers not involved in the implementation
-of the proposed addition.
+Started in 2009, \sklearn is developed by an international team of a dozen of
+core developers, mostly researchers from various fields (e.g., machine learning,
+natural language processing, or neuroscience). The project also greatly benefits
+from many occasional contributors proposing small code or documentation fixes or
+improvements. Development proceeds on GitHub\footnote{\url{https://github.com/scikit-learn}},
+a platform which greatly facilitates this kind of
+collaboration. Because of the large number of developers, emphasis is
+put on keeping the project maintainable. In particular, code must follow
+specific quality guidelines, such as style consistency and unit-test coverage.
+In addition, complete documentation and examples are required for all features.
+Finally, additions to the code base must pass code review by at least two
+developers not involved in the implementation of the proposed addition.
 
@@
 user base. Its popularity can be gaged from various indicators such as hundreds
-of citations in scientific publications of a previous paper
-\citep{pedregosa2011}, successes in various machine learning challenges
+of citations in scientific publications, successes in various machine learning challenges
 (\emph{e.g.} Kaggle), or significant statistics derived from our repositories

### 2013-06-12 paper.tex
remark
wording

@@
   \item[Inspectability.]
-       \sklearn objects store their constructor parameters
-       as well as parameter values determined by learning algorithms
-       as public attributes in a documented format.
-       (The Python language does not actually distinguish
-       between public and private attributes of objects,
-       but by convention, any identifier
-       prefixed with an underscore (\texttt{\_})
-       is considered ``a non-public part of the API''; \citealp{pythontut}.)
+       Constructor parameters
+       and parameter values determined by learning algorithms
+       are stored as public attributes. All of them are consistently documented.
+       % Gilles: Do we really need to bother the reader with such details?
+       % (The Python language does not actually distinguish
+       % between public and private attributes of objects,
+       % but by convention, any identifier
+       % prefixed with an underscore (\texttt{\_})
+       % is considered ``a non-public part of the API''; \citealp{pythontut}.)
   \item[Non-proliferation of classes.]
        Learning algorithms are objects of custom classes,
-       but datasets and hyperparameter names and values,
+       but datasets and hyperparameter names and values
        are represented as standard Python, NumPy and SciPy classes
-       rather than custom ones whenever feasible.
+       rather than new ones whenever feasible.
+       % Gilles: I don't understand the phrase below
        Custom classes adhere to Python conventions and protocols
-       for constructs such as iteration over the ``folds''
+       for constructs such as iteration over the folds
        in a $k$-fold cross validation.
@@
        % to be an experienced Java/adapter pattern hater?
-  \item[Composability.]
+  \item[Composition.]
        Many machine learning tasks are expressible
@@
        and a simple syntax is implemented
-       for setting/retrieving parameters of ``sub-estimators''.
+       for setting/retrieving parameters of sub-estimators.
        % XXX need to describe this!
@@
        Whenever an operation requires a user-defined parameter,
-       a default value for that parameter must be defined by the library
-       if at all possible,
-       and the default should cause the operation to be performed
+       a default and appropriate value is defined by the library.
+       The default value should cause the operation to be performed
        in a sensible way (giving a baseline solution for the task at hand).

### 2013-06-12 paper.tex
batch processing

@@
 
-The public API is oriented toward batch processing
-to enable efficient implementation.
-While classification and regression algorithms
-can make predictions for single samples,
-and some algorithms in \sklearn can perform online learning
-(e.g. stochastic gradient descent for linear SVMs and logistic regression,
-an online variant of $k$-means based on \citealt{sculley2010web}),
-these algorithm are really intended to be used on small batches of samples.
-Doing so makes optimal use of NumPy and SciPy
-because they prevent the overhead inherent in Python function calls
-and the language's per-element dynamic type checking.
+The public API is oriented towards batch processing. While classification and
+regression algorithms can make predictions for single samples, most \sklearn
+algorithms are intended to be used on batches of data. For efficiency reasons,
+this indeed makes the use of NumPy and SciPy optimal by preventing the overhead
+inherent to Python function calls and to per-element dynamic type checking.
+% Gilles: I don't think it is worth mentionning online learning. They are exceptions in scikit-learn.
+
+% The public API is oriented toward batch processing
+% to enable efficient implementation.
+% While classification and regression algorithms
+% can make predictions for single samples,
+% and some algorithms in \sklearn can perform online learning
+% (e.g. stochastic gradient descent for linear SVMs and logistic regression,
+% an online variant of $k$-means based on \citealt{sculley2010web}),
+% these algorithm are really intended to be used on small batches of samples.
+% Doing so makes optimal use of NumPy and SciPy
+% because they prevent the overhead inherent in Python function calls
+% and the language's per-element dynamic type checking.
 

### 2013-06-12 paper.tex
remove old comments
indent snippets
api principles

@@
 
-% % XXX the following is more of a general introduction than a project vision
-% scikit-learn \citep{pedregosa2011} is an open source machine learning library
-% for the Python programming language,
-% intended to tie in with the set of numeric and scientific packages
-% centered around the NumPy and SciPy libraries
-% \Citep{vanderwalt2011, varoquaux2013scipy}.
-% It is a library, rather than a novel (domain-specific) programming language,
-% intended to be called from Python programs;
-% at present, it offers no command line interface or GUI.
-
-% The scikit-learn project started in 2009 and has since been developed
-% by a team largely consisting of scientists
-% (both from computer science and other fields).
-% Its popularity can be gauged from various sources:
-% % TODO: list volume on the mailing list, number of contributors,
-% % number of issues opened and closed, number of forks, number of surveys filled in.
-% at the time of writing,
-% the popular programmer's Q~\&{}A site StackOverflow\footnote{
-%   \url{http://stackoverflow.com/questions/tagged/scikit-learn}}
-% % update the following figure when paper is finalized
-% lists 322 scikit-learn-related questions.
-
-% The main goal of the scikit-learn project
-% is to provide high-quality implementations
-% of ``classic'' machine learning algorithms
-% with a consistent API.
-
 \section{API}
@@
        and parameter values determined by learning algorithms
-       are stored as public attributes. All of them are consistently documented.
+       are stored as public attributes. All of them are documented in a consistent manner.
        % Gilles: Do we really need to bother the reader with such details?
@@
        but datasets and hyperparameter names and values
-       are represented as standard Python, NumPy and SciPy classes
+       are represented as standard Python, NumPy or SciPy classes
        rather than new ones whenever feasible.
        % Gilles: I don't understand the phrase below
-       Custom classes adhere to Python conventions and protocols
-       for constructs such as iteration over the folds
-       in a $k$-fold cross validation.
+       % Custom classes adhere to Python conventions and protocols
+       % for constructs such as iteration over the folds
+       % in a $k$-fold cross validation.
        This should keep \sklearn easy to use
@@
 inherent to Python function calls and to per-element dynamic type checking.
-% Gilles: I don't think it is worth mentionning online learning. They are exceptions in scikit-learn.
+% Gilles: I don't think it is worth elaborating too much (if at all) on online learning. They are exceptions in scikit-learn.
 
@@
 % and the language's per-element dynamic type checking.
-
-% For a machine learning library, one of the most important decisions we had to
-% make is how to represent data.  Rather than reinventing the wheel, we opted for
-% NumPy multidimensional arrays \Citep{vanderwalt2011} for dense data and SciPy
-% sparse matrices for sparse data.  While those may seem like a ``bare''
-% representation of data when compared to more object-oriented representations
-% (Weka is an example of this style of representation, \citealp{hall2009weka}),
-% it brings the prime advantage of allowing us to rely on NumPy and SciPy's
-% \textit{vector operations} which are often orders of magnitude faster
-% than the corresponding Python loops,
-% while at the same time keeping the code short and readable.
-% Conversion to NumPy and SciPy formats is usually easy and many scientific users
-% of Python will already be used to such formats, since they are pervasive in
-% other scientific Python packages.  For tasks where input is likely to consist
-% of text files or semi-structured objects, we provide ``vectorizers'' -- objects
-% that efficiently convert such data to the NumPy or SciPy formats.
-
-% FIXME: elaborate what is meant below
-% Lars: elaborated, but I'm not sure how important this is.
-% It mostly reflects experiences with OO libraries that require conversions
-% between types just to pass information to a method.
-%Finally, offering a custom datatype for samples would require a conversion
-%\textit{in any case}, since such a datatype would be toolkit-specific
-%and no program would store its data using such a type
-%if it weren't from the outset designed to use scikit-learn.
-
-% In the remainder of the paper, we use the following notation. Training data are
-% denoted by \Xtr ~and test data (i.e., unseen data to which the learning
-% algorithm needs to be able to generalize) by \Xte.  Following the discussion above,
-% we represent \Xtr ~and \Xte ~by 2-dimensional NumPy arrays or SciPy sparse
-% matrices. Using NumPy/SciPy's notation, \Xtr\texttt{[i, j]} denotes the
-% $i^\textrm{th}$ training sample's $j^\textrm{th}$ feature. For supervised learning tasks,
-% such as classification or regression, we denote the training and test labels by
-% \ytr ~and \yte, respectively. We typically store \ytr ~and \yte ~as
-% 1-dimensional NumPy arrays (\ytr\texttt{[i]} denotes the $i^\textrm{th}$ training
-% sample's label value).
 
@@
 learned coefficients of a linear model).
-
-% Importantly, the parameters learned by an estimator
-% are exposed as \textit{public} attributes on the trained object.
-% This facilitates model inspection
-% and makes it possible to train a model using scikit-learn,
-% export it to a (standardized or custom) external data format
-% and use the learned parameters in a different piece of code,
-% perhaps a prediction algorithm implemented in a different language
-% or one that combines the model parameters with other information
-% to build a custom model.
-% % XXX Hard to follow phrasing below
-% Of course, it also forces the developers to think carefully
-% about what might, in a more ``black box''-style toolkit,
-% be considered the internals of objects,
-% since authors of client code are allowed to rely
-% on the names and formats of model parameters.
 
@@
 
-% The bulk of the scikit-learn application programming interface
-% consists of so-called \textit{estimator} objects.
-% All learning algorithms for classification, regression and clustering
-% are offered as such objects,
-% but so are feature extraction, feature selection and dimensionality reduction
-% algorithms.
-
-% Construction and application of an estimator object
-% proceeds in a way that is somewhat reminiscent of partial function application.
-% The constructor, \texttt{\_\_init\_\_}\footnote{
-%   Strictly speaking, \texttt{\_\_init\_\_} is Python's
-%   class \textit{initialization} method,
-%   while \texttt{\_\_new\_\_} is the constructor.
-%   The distinction is not relevant to the present description of scikit-learn.}
-% takes a set of named (hyper-)parameters
-% and attaches these to the estimator as public attributes.
-% It never performs actual learning, nor does it even see a training set.
-
-% The algorithmic workhorse of most estimators, and one of the few methods
-% that all estimators have in common, is called \texttt{fit}.
-% Invoking this method causes the estimator to learn from a training set,
-% supplied as one or more arguments.
-% Its task, from an API point of view,
-% is to determine (via a learning algorithm) model-specific parameters
-% from a training set and set these as attributes on the estimator object,
-% which can then be used to transform data or make predictions.
-% As in other object-oriented machine learning libraries
-% (e.g.\ Weka's Java API)
-% no distinction is made in the type hierarchy
-% between what is variously called an estimator or a \textit{learner}
-% in the machine learning literature,
-% and the \textit{model} that results from applying an estimator to data.
-% When the \texttt{fit} method has been called,
-% an estimator object serves dual purpose as a model
-% (and is referred to as such in the documentation).
-
-% An example to clarify the partial application style
-% is the following code snippet.
-% If $X$ is a dataset in the format described below and $k$ is some integer,
-% then the Python code to train a $k$-means clustering model on it
-% can be written as
-% % XXX we could use Pygments for syntax highlighting
-% \begin{verbatim}
-% km = KMeans(n_clusters=k).fit(X)
-% \end{verbatim}
-% This snippet relies on the fact that the \texttt{fit} method
-% always returns the object it was called on (its ``\texttt{self}''),
-% making a method chaining style possible.
-
-% A \texttt{fit} method is present even on objects
-% that perform seemingly mundane tasks such as vectorizing text documents
-% for subsequent learning.
-% Such classes, however, can be said to learn their vocabulary
-% from the training corpus,
-% and follow the conventions set out above.
-% Even those estimators that are entirely stateless and therefore
-% do not require a \texttt{fit} method to perform useful work,
-% nevertheless have such a method for consistency and composibility,
-% and to perform parameter validation.
-% Examples of this latter kind of objects are the normalizing transformer,
-% kernel approximation transformers
-% \citep{rahimi2007random, li2010random, vedaldi2010efficient},
-% % XXX is it relevant that they are exactly two? Propose s/two/the
-% and two feature extraction classes that implement the ``hashing trick''
-% \citep{weinberger2009}.
-
 \subsection{Estimator composition}
@@
 elaborate a bit more on nested constructions.}
-
-% A distinguishing feature of \sklearn estimators
-% is that they can be composed into \texttt{Pipeline} objects.
-% Such objects combine typical machine learning workflows
-% (extracting features, centering/scaling/normalization,
-% feature selection, making predictions)
-% in a single object that is itself an estimator,
-% and whose methods cause data to flow through the pipeline.
-% More importantly, though, they can be used for
-% systematic, automated model selection.
-% This is best explained by example.
-
-% By composing, say, a \texttt{CountVectorizer} object
-% (which extracts term frequency features from text documents
-% according to a bag-words model),
-% a \textsf{tf--idf} transformer, a $\chi^2$ feature selector and a linear SVM,
-% one obtains a pipeline object with various hyperparameters:
-% whether to do stopword filtering,
-% linear vs.\ logarithmic \textsf{tf}, number of features to keep,
-% SVM regularization ($\ell_1$ or $\ell_2$ norm, strength $C$) a.o.
-% Such a composite \texttt{Pipeline} object
-% exposes all of these parameters using a special syntax;
-% each estimators must be given a name,
-% and if the SVM's name is, e.g., \texttt{linearsvm},
-% then its $C$ parameter is exposed as \texttt{linearsvm\_\_C}.
-% The pipeline's \texttt{set\_params} splits the parameter names
-% on double underscores, uses the first part of the result
-% to look up the estimator,
-% and passes the second part to the estimator's own \texttt{set\_params}
-% as a keyword argument; due to Python's dynamic parameter passing,
-% the result is the same as if the user had called \texttt{set\_params}
-% on the SVM themselves.
-% (The double underscore syntax is not special to pipelines;
-% it is also employed in ensemble estimators such as random forests
-% % Is the following clear? Suggestions to replace "direct attributes" with"?
-% It is chosen as a convention to avoid ambiguity, as direct
-% attributes and methods of scikit-learn objects never contain
-% double underscores.)
 

### 2013-06-12 paper.tex
avoid enumerate where not necessary

@@
 a fixed sequence of steps in processing data (e.g., feature extraction,
-normalization, feature selection, learning and making predictions).
-A sequence of $N$ such steps can be combined into a composite object
-called a \texttt{Pipeline}.
-The first $N-1$ steps must be transformers
-and the last can be either a predictor or a transformer.
-Fitting a pipeline to a training set amounts, conceptually,
-to the following recursive procedure:
-\begin{enumerate}
-  \item When only one step remains, call its \texttt{fit} method.
-  \item Otherwise, fit the first step,
-        use it to \texttt{transform} the training set
-        and fit the rest of the pipeline with the transformed data.
-\end{enumerate}
+normalization, feature selection, learning and making predictions). A sequence
+of $N$ such steps can be combined into a composite object called a
+\texttt{Pipeline}. The first $N-1$ steps must be transformers and the last can
+be either a predictor or a transformer. Conceptually, fitting a pipeline to a
+training set amounts to the following recursive procedure: i) when only one step
+remains, call its \texttt{fit} method; ii) otherwise, fit the first step, use it
+to \texttt{transform} the training set and fit the rest of the pipeline with the
+transformed data.
+
 The pipeline exposes all the methods the last

### 2013-06-12 paper.tex
wording

@@
        as \textit{pipelines} of transformations to data.
-       Also, some learning algorithms are naturally viewed
+       Some learning algorithms are also naturally viewed
        as meta-algorithms parameterized on other algorithms.
@@
        and a simple syntax is implemented
-       for setting/retrieving parameters of sub-estimators.
+       for setting and retrieving parameters of sub-estimators.
        % XXX need to describe this!
@@
 The public API is oriented towards batch processing. While classification and
-regression algorithms can make predictions for single samples, most \sklearn
-algorithms are intended to be used on batches of data. For efficiency reasons,
+regression algorithms can indeed make predictions for single samples, most \sklearn
+objects are however intended to be used on batches of data. For efficiency reasons,
 this indeed makes the use of NumPy and SciPy optimal by preventing the overhead
-inherent to Python function calls and to per-element dynamic type checking.
+inherent to Python function calls and per-element dynamic type checking.
 % Gilles: I don't think it is worth elaborating too much (if at all) on online learning. They are exceptions in scikit-learn.
@@
 transformed data.
-
 The pipeline exposes all the methods the last

### 2013-06-12 paper.tex
typo

@@
 of citations in scientific publications, successes in various machine learning challenges
-(\emph{e.g.} Kaggle), or significant statistics derived from our repositories
+(e.g., Kaggle), or significant statistics derived from our repositories
 and mailing lists.  At the time of writing, the project is watched by 1309
@@
 to estimate the performance of the estimator according to some measure
-(e.g.\ accuracy, $F_\beta$-score or a host of clustering quality metrics)
+(e.g., accuracy, $F_\beta$-score or a host of clustering quality metrics)
 and the best performing set of hyperparameters is stored

### 2013-06-12 paper.tex
some more wording

@@
 
-The project is a programming library, i.e. a collection of classes (in the
-object-oriented sense) that users import into Python programs. Its use
+The project is a programming library, i.e. a collection of classes -- in the
+object-oriented sense -- that users import into Python programs. Using \sklearn
 therefore requires basic Python programming knowledge. No command-line
-interface, let alone a GUI, is offered for non-programmer users. Instead,
-interactive use is made possible by the Python interactive interpreter, and its
-enhanced replacement IPython \citep{perez2007ipython}, which offers a
-\textsc{matlab}-like working environment specifically designed for scientific
-use.
+interface, let alone a graphical user interface, is offered for non-programmer
+users. Instead, interactive use is made possible by the Python interactive
+interpreter, and its enhanced replacement IPython \citep{perez2007ipython},
+which offer a \textsc{matlab}-like working environment specifically designed for
+scientific use.
 
@@
 user base. Its popularity can be gaged from various indicators such as hundreds
-of citations in scientific publications, successes in various machine learning challenges
-(e.g., Kaggle), or significant statistics derived from our repositories
-and mailing lists.  At the time of writing, the project is watched by 1309
-people and forked 667 time on GitHub, and the mailing-list receives 300 mails
-per month.
-The version control logs for the development version of scikit-learn
+of citations in scientific publications, successes in various machine learning
+challenges (e.g., Kaggle), or significant statistics derived from our
+repositories and mailing lists.  At the time of writing, the project is watched
+by 1309 people and forked 667 times on GitHub, and the mailing-list receives
+more than 300 mails per month. The version control logs for the development
+version
 % ddaa494c116e3c16bf032003c5cccbed851733d2
-show 183 contributors to the project's source code and documentation
-(after manual deduplication).
+also count 183 contributors to the project (after manual deduplication).
 
@@
 \begin{description}
-  \item[Inspectability.]
+  \item[Inspection.]
        Constructor parameters
@@
        Many machine learning tasks are expressible
-       as \textit{pipelines} of transformations to data.
+       as sequences or combinations of transformations to data.
        Some learning algorithms are also naturally viewed

### 2013-06-12 paper.tex
remove old comments
wording

@@
 
-The public API is oriented towards batch processing. While classification and
-regression algorithms can indeed make predictions for single samples, most \sklearn
-objects are however intended to be used on batches of data. For efficiency reasons,
-this indeed makes the use of NumPy and SciPy optimal by preventing the overhead
-inherent to Python function calls and per-element dynamic type checking.
+The public API is oriented towards batch processing rather than online
+processing. While classification and regression algorithms can indeed make
+predictions for single samples, most \sklearn objects are however intended to be
+used on batches of data. For efficiency reasons, this makes the use of NumPy and
+SciPy optimal by preventing the overhead inherent to Python function calls or
+due to per-element dynamic type checking.
 % Gilles: I don't think it is worth elaborating too much (if at all) on online learning. They are exceptions in scikit-learn.
@@
 Given the API defined above, solving this problem is as simple as:
-% XXX: Gael: I think that we should break the following lines in 2: the
-% 'return self' hidden in the call to fit can be confusing to the casual
-% reader
 \begin{verbatim}

### 2013-06-12 paper.tex
wording

@@
        To be able to implement and use these algorithms in a simple way,
-       all learners are usable with a consistent set of methods
-       and a simple syntax is implemented
-       for setting and retrieving parameters of sub-estimators.
+       all learners (basic or composed) are usable with a consistent set of methods.
+       % Gilles: let elaborate later on the syntax
+       % and a simple syntax is implemented
+       % for setting and retrieving parameters of sub-estimators.
        % XXX need to describe this!

### 2013-06-12 paper.tex
project actually started in 2007

@@
 
-Started in 2009, \sklearn is developed by an international team of a dozen of
+Started in 2007, \sklearn is developed by an international team of a dozen of
 core developers, mostly researchers from various fields (e.g., machine learning,

### 2013-06-12 paper.tex
motivate batch processing API further
@glouppe: I snuck the online learning algos back in so as not to give
the false impression that we're tied to batch learning algorithms.
The remark is shorter now.

@@
 
-The public API is oriented towards batch processing rather than online
-processing. While classification and regression algorithms can indeed make
-predictions for single samples, most \sklearn objects are however intended to be
-used on batches of data. For efficiency reasons, this makes the use of NumPy and
-SciPy optimal by preventing the overhead inherent to Python function calls or
-due to per-element dynamic type checking.
+The public API is oriented towards batch processing
+rather than processing single samples per API call.
+While classification and regression algorithms can indeed make predictions
+for single samples,
+\sklearn objects are not optimized for this use case.
+(The few online learning algorithms implemented
+are intended to take minibatches.)
+Batch processing makes optimal use of NumPy and SciPy
+by preventing the overhead inherent to Python function calls
+and per-element dynamic type checking.
+Although this might seem to be an artifact of the Python language,
+and therefore an implementation detail that leaks into the API,
+we argue that APIs should be designed
+so as to not tie a library to a suboptimal implementation strategy,
+and batch processing can enable fast implementations
+in lower-level languages as well
+(where locality of reference effects
+and the possibility of internal parallelization come into play).
 % Gilles: I don't think it is worth elaborating too much (if at all) on online learning. They are exceptions in scikit-learn.
-
-% The public API is oriented toward batch processing
-% to enable efficient implementation.
-% While classification and regression algorithms
-% can make predictions for single samples,
-% and some algorithms in \sklearn can perform online learning
-% (e.g. stochastic gradient descent for linear SVMs and logistic regression,
-% an online variant of $k$-means based on \citealt{sculley2010web}),
-% these algorithm are really intended to be used on small batches of samples.
-% Doing so makes optimal use of NumPy and SciPy
-% because they prevent the overhead inherent in Python function calls
-% and the language's per-element dynamic type checking.
 

### 2013-06-12 paper.tex
more an array processing: gensim comparison

@@
 \subsection{Data representation}
+\label{sec:arrays}
 
@@
 in lower-level languages as well
-(where locality of reference effects
+(where memory hierarchy effects
 and the possibility of internal parallelization come into play).
@@
 
+Another package that deserves mention is the Gensim topic modeling toolkit
+\citep{rehurek2010gensim}.
+Although its stated goal is not machine learning,
+it does exemplify a style of API design for scalable processing of ``big data''
+that takes a different approach than the comparable models in \sklearn.
+Gensim's way of dealing with large datasets is to use algorithms
+that have $O(1)$ space complexity and can be updated online,
+called \textit{document streaming} by \citeauthor{rehurek2010gensim}.
+The API is designed around the Python concept of an \textit{iterable}
+(supported in the language by a restricted form of co-routines called
+\textit{generators}).
+While the text vectorization part of \sklearn
+also uses irerables to some extent,
+the emphasis is still largely on (mini-)batch processing,
+even in the truly stateless, O(1) memory vectorizers
+that implement the hashing trick of \citet{weinberger2009}.
+This way of processing, as argued earlier in Section~\ref{sec:arrays},
+reduces various forms of overhead
+and allows users to make use of the abundance of memory in modern machines.
+No attempt is made to hide this batch-oriented processing from the user,
+allowing control over the amount of memory actually dedicated
+to \sklearn algorithms.
+
 \section{Future work}

### 2013-06-12 paper.tex
Use variable for FeatureUnion.

@@
 
-    Pipeline([("pca", FeatureUnion([('pca', PCA()),
-                                    ('kpca', KernelPCA(kernel="rbf"))])),
+    feat_union = FeatureUnion([('pca', PCA()),
+                               ('kpca', KernelPCA(kernel="rbf"))])
+
+    Pipeline([("pca", feat_union,
               ("feature_selection", SelectKBest(k=100)),

### 2013-06-12 paper.tex
New structure

@@
 benchmarked against several competitors. In this paper, we instead focus on an
-in-depth discussion and analysis of the design choices used in
-building the library. First, in section \ref{sec:api}, we present the
-central application programming interface (API). Then, in section
-\ref{sec:implementation}, we describe the implementation trade-offs.
-In section \ref{sec:weaknesses}, we analyze the limitations of the
-technologies we rely on; and finally, in section \ref{sec:comparison}, we
-compare \emph{scikit-learn} to other major projects in terms of API.
+in-depth discussion and analysis of the design choices used in building the
+library. First, in section \ref{sec:core-api}, we present the central application
+programming interface (API) and then describe, in section \ref{sec:advanced-api},
+advanced API mechanisms built on this core API. Section
+\ref{sec:implementation} describes the implementation trade-offs and analyzes
+the limitations of the technologies we rely on while section
+\ref{sec:comparison} then compares \emph{scikit-learn} to other major projects in
+terms of API. Finally, conclusions and directions of future works are included
+in section \ref{sec:conclusions}.
 
-\section{API}
+\section{Core API}
 
-\label{sec:api}
+\label{sec:core-api}
 
 In this section, we discuss the general principles of the application
-programming interface (API) of \sklearn. We first specify the
-representation of data and then proceed by describing the interface shared by
-all objects of the library. We then discuss the composition and reusability
-advantages of that interface.
+programming interface (API) of \sklearn. We first specify the representation of
+data and then proceed by describing the core interface of the library.
 
-The API is designed to adhere to the following broad principles:
+All objects within \sklearn share a uniform common basic API consisting of three
+complementary interfaces: an \textit{estimator} interface for building and
+fitting objects, a \textit{predictor} interface for making predictions and a
+\textit{transformer} interface for converting data. As much as possible, our
+design choices have been guided so as to avoid the proliferation of framework
+code. We try to adopt simple conventions and to limit to a minimum the number of
+methods an object must implement. The API is designed to adhere to the following
+broad principles:
+
 \begin{description}
@@
 
-The public API is oriented towards batch processing
-rather than processing single samples per API call.
-While classification and regression algorithms can indeed make predictions
-for single samples,
-\sklearn objects are not optimized for this use case.
-(The few online learning algorithms implemented
-are intended to take minibatches.)
-Batch processing makes optimal use of NumPy and SciPy
-by preventing the overhead inherent to Python function calls
-and per-element dynamic type checking.
-Although this might seem to be an artifact of the Python language,
-and therefore an implementation detail that leaks into the API,
-we argue that APIs should be designed
-so as to not tie a library to a suboptimal implementation strategy,
-and batch processing can enable fast implementations
-in lower-level languages as well
-(where memory hierarchy effects
-and the possibility of internal parallelization come into play).
-% Gilles: I don't think it is worth elaborating too much (if at all) on online learning. They are exceptions in scikit-learn.
+The public API is oriented towards batch processing rather than processing
+single samples per API call. While classification and regression algorithms can
+indeed make predictions for single samples, \sklearn objects are not optimized
+for this use case. (The few online learning algorithms implemented are intended
+to take minibatches.) Batch processing makes optimal use of NumPy and SciPy by
+preventing the overhead inherent to Python function calls or due to per-element
+dynamic type checking. Although this might seem to be an artifact of the Python
+language, and therefore an implementation detail that leaks into the API, we
+argue that APIs should be designed so as not to tie a library to a suboptimal
+implementation strategy. As such, batch processing enable fast implementations
+in lower-level languages (where memory hierarchy effects and the possibility of
+internal parallelization come into play).
 
-\subsection{Estimators, predictors and transformers}
+\subsection{Estimators}
 
-All objects within \sklearn share a uniform common basic API
-consisting of three complementary interfaces: an \textit{estimator} interface
-for building and fitting objects, a \textit{predictor} interface for making
-predictions and a \textit{transformer} interface for converting data.
-As much as possible, our design choices have been guided so as to avoid
-the proliferation of framework code. We try to adopt simple conventions and
-to limit to a minimum the number of methods an object must implement.
-
-\textbf{Estimators.} The \textit{estimator} interface is at the core of the
+The \textit{estimator} interface is at the core of the
 library. It defines instantiation mechanisms of objects and exposes a
@@
 
-\textbf{Predictors.} The \textit{predictor} interface defines a \texttt{predict}
+\subsection{Predictors}
+
+The \textit{predictor} interface defines a \texttt{predict}
 method taking as input some new data \texttt{X\_test} and producing as output
@@
 
-\textbf{Transformers.} Since it is common to modify or filter data before
-feeding it to a learning algorithm, some estimators in the library implement a
-\textit{transformer} interface which defines a \texttt{transform} method. It
-takes as input some new data \texttt{X\_test} and yields as output a transformed
-version of \texttt{X\_test}.
-Preprocessing, feature selection and dimensionality
-reduction algorithms are all provided as transformers within the library.  In
-our example, if one wanted to preprocess the inputs \texttt{X\_train} before
-fitting the logistic regression estimator, e.g., by standardizing the feature
-values, one could thus simply write:
+\subsection{Transformers}
+
+Since it is common to modify or filter data before feeding it to a learning
+algorithm, some estimators in the library implement a \textit{transformer}
+interface which defines a \texttt{transform} method. It takes as input some new
+data \texttt{X\_test} and yields as output a transformed version of
+\texttt{X\_test}. Preprocessing, feature selection and dimensionality reduction
+algorithms are all provided as transformers within the library.  In our example,
+if one wanted to preprocess the inputs \texttt{X\_train} before fitting the
+logistic regression estimator, e.g., by standardizing the feature values, one
+could thus simply write:
 \begin{verbatim}
@@
 
-\subsection{Method chaining}
-
-In \sklearn, the \texttt{fit} method always returns the estimator
+Finally, let us also note that, the \texttt{fit} method always returns the estimator
 it was called on, hence making method chaining possible.  Using this technique,
@@
 
-\subsection{Estimator composition}
+\section{Advanced API}
+
+\label{sec:advanced-api}
+
+Building on the core interface introduced in the previous section, we now
+present in this section advanced API mechanisms for i) building meta-estimators,
+ii) composing complex estimators and iii) selecting models.
+
+\subsection{Meta-estimators}
+
+\textcolor{red}{TODO}
+
+\subsection{Pipelines and feature unions}
 
@@
 
-\section{Weaknesses of the NumPy/SciPy environment}
+%\section{Weaknesses of the NumPy/SciPy environment}
 
-\label{sec:weaknesses}
-
-\textcolor{red}{TODO: move this into the Implementation section? These are mostly implementation considerations.}
+\textcolor{red}{TODO: The following corresponds to the old weaknesses section. It should be properly rewritten.}
 
@@
 
-\section{Future work}
+\section{Conclusions and future works}
+\label{sec:conclusions}
 

### 2013-06-12 paper.tex
wordings

@@
 advanced API mechanisms built on this core API. Section
-\ref{sec:implementation} describes the implementation trade-offs and analyzes
-the limitations of the technologies we rely on while section
+\ref{sec:implementation} describes implementation trade-offs and analyzes
+limitations of the technologies we rely on, while section
 \ref{sec:comparison} then compares \emph{scikit-learn} to other major projects in
@@
 \begin{description}
+  \item[Consistency.]
+       All objects (basic or composed) share a consistent interface composed of
+       a limited set of methods.
   \item[Inspection.]
-       Constructor parameters
-       and parameter values determined by learning algorithms
-       are stored as public attributes. All of them are documented in a consistent manner.
+       Constructor parameters and parameter values determined by learning
+       algorithms are stored as public attributes. All of them are documented in
+       a consistent manner.
        % Gilles: Do we really need to bother the reader with such details?
@@
   \item[Non-proliferation of classes.]
-       Learning algorithms are objects of custom classes,
-       but datasets and hyperparameter names and values
+       Only learning algorithms are objects of custom classes,
+       Datasets and hyperparameter names and values
        are represented as standard Python, NumPy or SciPy classes
@@
        % to be an experienced Java/adapter pattern hater?
-  \item[Composition.]
-       Many machine learning tasks are expressible
-       as sequences or combinations of transformations to data.
-       Some learning algorithms are also naturally viewed
-       as meta-algorithms parameterized on other algorithms.
-       To be able to implement and use these algorithms in a simple way,
-       all learners (basic or composed) are usable with a consistent set of methods.
-       % Gilles: let elaborate later on the syntax
-       % and a simple syntax is implemented
-       % for setting and retrieving parameters of sub-estimators.
-       % XXX need to describe this!
+  % Gilles: this is actually not a design principle
+  % \item[Composition.]
+  %      Many machine learning tasks are expressible
+  %      as sequences or combinations of transformations to data.
+  %      Some learning algorithms are also naturally viewed
+  %      as meta-algorithms parameterized on other algorithms.
+  %      To be able to implement and use these algorithms in a simple way,
+  %      % Gilles: let elaborate later on the syntax
+  %      % and a simple syntax is implemented
+  %      % for setting and retrieving parameters of sub-estimators.
+  %      % XXX need to describe this!
   \item[Sensible defaults.]

### 2013-06-12 paper.tex
wording

@@
   \item[Non-proliferation of classes.]
-       Only learning algorithms are objects of custom classes,
+       Only learning algorithms are objects of custom classes.
        Datasets and hyperparameter names and values
@@
 Building on the core interface introduced in the previous section, we now
-present in this section advanced API mechanisms for i) building meta-estimators,
-ii) composing complex estimators and iii) selecting models.
+present in this section advanced API mechanisms for building meta-estimators,
+composing complex estimators and then selecting models.
 

### 2013-06-12 paper.tex
api principles

@@
        All objects (basic or composed) share a consistent interface composed of
-       a limited set of methods.
+       a limited set of methods. This interface is documented in a consistent
+       manner for all objects.
   \item[Inspection.]
        Constructor parameters and parameter values determined by learning
-       algorithms are stored as public attributes. All of them are documented in
-       a consistent manner.
-       % Gilles: Do we really need to bother the reader with such details?
-       % (The Python language does not actually distinguish
-       % between public and private attributes of objects,
-       % but by convention, any identifier
-       % prefixed with an underscore (\texttt{\_})
-       % is considered ``a non-public part of the API''; \citealp{pythontut}.)
+       algorithms are stored and exposed as public attributes.
   \item[Non-proliferation of classes.]
-       Only learning algorithms are objects of custom classes.
-       Datasets and hyperparameter names and values
-       are represented as standard Python, NumPy or SciPy classes
-       rather than new ones whenever feasible.
-       % Gilles: I don't understand the phrase below
-       % Custom classes adhere to Python conventions and protocols
-       % for constructs such as iteration over the folds
-       % in a $k$-fold cross validation.
-       This should keep \sklearn easy to use
-       and easy to combine with other libraries.
+       Learning algorithms are the only objects to be of custom classes.
+       Datasets and hyperparameter names and values are represented as standard
+       Python, NumPy or SciPy classes rather than new ones whenever feasible.
+       This keeps \sklearn easy to use and easy to combine with other libraries.
        % XXX do we need to elaborate more on this, or do we expect everyone
@@
   % Gilles: this is actually not a design principle
-  % \item[Composition.]
-  %      Many machine learning tasks are expressible
-  %      as sequences or combinations of transformations to data.
-  %      Some learning algorithms are also naturally viewed
-  %      as meta-algorithms parameterized on other algorithms.
-  %      To be able to implement and use these algorithms in a simple way,
-  %      % Gilles: let elaborate later on the syntax
-  %      % and a simple syntax is implemented
-  %      % for setting and retrieving parameters of sub-estimators.
-  %      % XXX need to describe this!
+  \item[Composition.]
+       Many machine learning tasks are expressible
+       as sequences or combinations of transformations to data.
+       Some learning algorithms are also naturally viewed
+       as meta-algorithms parameterized on other algorithms.
+       Whenever possible, those algorithms are composed from
+       existing building blocks.
+       % Gilles: let elaborate later on the syntax
+       % and a simple syntax is implemented
+       % for setting and retrieving parameters of sub-estimators.
+       % XXX need to describe this!
   \item[Sensible defaults.]

### 2013-06-12 paper.tex
remove some comments
wording

@@
 programming interface (API) and then describe, in section \ref{sec:advanced-api},
-advanced API mechanisms built on this core API. Section
+advanced API mechanisms built from the core interface. Section
 \ref{sec:implementation} describes implementation trade-offs and analyzes
@@
        % to be an experienced Java/adapter pattern hater?
-  % Gilles: this is actually not a design principle
   \item[Composition.]
@@
        as meta-algorithms parameterized on other algorithms.
-       Whenever possible, those algorithms are composed from
+       Whenever possible, those algorithms are implemented and composed from
        existing building blocks.
-       % Gilles: let elaborate later on the syntax
-       % and a simple syntax is implemented
-       % for setting and retrieving parameters of sub-estimators.
-       % XXX need to describe this!
   \item[Sensible defaults.]

### 2013-06-12 paper.tex
typo
trim/rewrite the old weaknesses section to fit in an implementation section

@@
 
-Finally, let us also note that, the \texttt{fit} method always returns the estimator
+Finally, let us also note that the \texttt{fit} method always returns the estimator
 it was called on, hence making method chaining possible.  Using this technique,
@@
 
-% This needs an introductory paragraph, but the best I could come up with was:
-As can be expected, the NumPy/SciPy environment is not completely flawless.
-
-% The following hedge needs to be in here; we don't want to diss our peers
-It should be noted that this section is meant
-as a summary of challenges faced by scikit-learn development
-and an attempt at constructive criticism.
-There is considerable overlap between the communities
-that develop or contribute to NumPy, SciPy, scikit-learn, and Cython,
-so the issues raised here may eventually be resolved
-by the scikit-learn developers themselves.
-
-\subsection{Sparse matrix support}
-
-SciPy's handling of sparse matrices is one source of confusion
-for scikit-learn developers and users alike.
-The \texttt{scipy.sparse} package offers support
-for this kind of data structure,
-which is crucial to achieving performance and scalability
-in the face of high-dimensional inputs
-(including, but not limited to, NLP applications).
-However, the interface exposed by its classes
-deviates from the NumPy array interface, sometimes subtly so,
-sometimes entirely.  % XXX is this grammatically correct?
-For example, while NumPy arrays have a variable number of dimensions,
-so that vectors can be considered 1-d arrays
-(without a distinction between row and column vectors),
-matrices are 2-d, etc., \texttt{scipy.sparse} matrices are always 2-dimensional,
-meaning that special care has to be taken when handling sparse vectors.
-
-The development team has built up sufficient familiarity
-with the sparse/array distinction to factor out common patterns
-for supporting both sparse and dense input data structures in many estimators.
-In other cases, though, sparse matrix support has to implemented
-on a per-estimator basis.
-
 \subsection{Parallelism}
@@
 at least in the CPython reference implementation targeted by scikit-learn,
-is downright weak.
-The interpreter supports multithreading, but due to the ``Global Interpreter Lock'' (GIL),
+is suboptimal.
+The interpreter supports multithreading,
+but due to the ``Global Interpreter Lock'' (GIL),
 no two threads in a process may at the same time execute Python code.
 It follows that multithreading is not a viable solution
-even two embarrassingly parallel problems,
+even for embarrassingly parallel problems,
 such as fitting and evaluation multiple models in a grid search procedure.
@@
 
-While Cython makes writing performance-critical code possible
-in a language that is very much like Python,
-doing so properly still requires some knowledge of C.
-One source of trouble has been the various fixed-size integer types
-that are used by the Python and NumPy C runtimes,
-whose sizes are practically all platform-dependent.
+\subsection{Cython for inner loops}
 
-Another problem concerns the mismatch between on the one hand Python functions,
+A problem arising from the use of Cython
+concerns the mismatch between on the one hand Python functions,
 which are first-class runtime objects endowed with methods and metadata
@@
 that dynamically check their number and types of arguments,
-and that can return any Python type of object,
+and that can return any Python type of object;
 and compiled C or Cython functions on the other hand,
@@
 but such usage is not at present possible and may never be.
-Passing C or Cython-compiled functions would require an API for the kernel method
+Passing C or Cython-compiled functions
+would require an API for the kernel method
 to be implemented at the C or Cython level,
@@
 because of the runtime checks needed for each invocation of the kernel function.
-(The current solution in scikit-learn is to let the user pass not a kernel function,
+(The current solution in scikit-learn is to let the user pass
+not a kernel function,
 but a precomputed Gram matrix for an entire set of objects.
 The linear, polynomial and RBF kernels are treated as special cases in
-the \textsf{LIBSVM}-based SVM learner, where they are implemented in C{}\verb!++!.)
+the \textsf{LIBSVM}-based SVM learner,
+where they are implemented in C{}\verb!++!.)
 

### 2013-06-12 paper.tex
start a future work section

@@
 \subsection{Parallelism}
+\label{sec:parallel}
 
@@
 
+\subsection{Future work}
+
+The \sklearn toolkit is by no means finished,
+although it is coming close to a 1.0 release.
+
+\sklearn does not support all tasks in the broad field of machine learning,
+and does not intend to either;
+in particular, structured prediction and bandit-style online learning algorithms
+are considered ``off-topic'' for the library,
+and submissions of implementations of these for inclusion
+will (at least at present) be rejected.
+Collaborative filtering algorithms are not implemented
+and it is remains to be seen
+whether an API for them can be designed that matches the \sklearn API.
+
+At a lower-level, a potential point of improvement
+is parallel processing, which, as described in Section~\ref{sec:parallel},
+is only implemented at a rather coarse-grained level.
+Cython's built-in support for OpenMP \citep{dagum1998openmp}
+remains a viable candidate technology
+for more fine-grained multicore support in \sklearn.
+
 \bibliographystyle{plainnat}

### 2013-06-12 paper.tex
trim section on Cython

@@
 
-As much as possible is implemented in ``pure'' Python,
+As much as possible is implemented in ``pure'' Python
+with NumPy for numerical work,
 to avoid the compile-edit-debug cycle needed for Cython
 and to keep the codebase readable for as large an audience as possible.
-However, since the speed of the current Python implementation
-is not sufficient for numeric programming
-and NumPy's vector operations are not appropriate for all use cases
-(among other factors, because they often have linear space requirements),
-some of the core algorithms have to be implemented in Cython.\footnote{
-  The only Python implementation currently supported by scikit-learn
-  is the reference implementation, CPython.
-  The alternative Python implementation PyPy \citep{bolz2009tracing}
-  promises to alleviate this problem by JIT compilation,
-  but does not yet interoperate with NumPy and SciPy.
-}
-Examples include the aforementioned $k$-means implementation,
-stochastic gradient descent for linear models
+However, not all algorithms are easily and efficiently expressed
+as NumPy vector operations
+(among other factors, because they often have linear space requirements).
+Algorithms that fall into this category have to be implemented in Cython
+to obtain decent performance and scalability.
+Examples include $k$-means, stochastic gradient descent for linear models
 and some graph-based clustering algorithms.

### 2013-06-13 paper.tex
wordings in section 1

@@
 
-\noindent \textcolor{red}{TODO: Focus on the experience we accumulated (general remark)}\\
-\textcolor{red}{TODO: Motivate our design choices.}
-
 \section{The scikit-learn project}
 
-The \sklearn project \citep{pedregosa2011} is an open source
-machine learning library for the Python programming language. The ambition of
-the project is to provide efficient and well-established machine learning tools
-within a programming environment that is accessible to non-machine learning
-experts and reusable in various scientific areas. In itself the project isn't
-a novel domain-specific language, but a library that provides machine
-learning idioms to a general-purpose high-level language. Among other things,
-it includes classical learning algorithms, model selection tools and
-preprocessing procedures.
+The \sklearn project \citep{pedregosa2011} is an open source machine learning
+library for the Python programming language. The ambition of the project is to
+provide efficient and well-established machine learning tools within a
+programming environment that is accessible to non-machine learning experts and
+reusable in various scientific areas. The project is not a novel domain-specific
+language, but a library that provides machine learning idioms to a general-
+purpose high-level language. Among other things, it includes classical learning
+algorithms, model selection tools and preprocessing procedures.
 
-The project is a programming library, i.e. a collection of classes -- in the
-object-oriented sense -- that users import into Python programs. Using \sklearn
-therefore requires basic Python programming knowledge. No command-line
-interface, let alone a graphical user interface, is offered for non-programmer
-users. Instead, interactive use is made possible by the Python interactive
-interpreter, and its enhanced replacement IPython \citep{perez2007ipython},
-which offer a \textsc{matlab}-like working environment specifically designed for
-scientific use.
+In itself, the project is a programming library, i.e. a collection of classes
+and functions that users import into Python programs. Using \sklearn therefore
+requires basic Python programming knowledge. No command-line interface, let
+alone a graphical user interface, is offered for non-programmer users. Instead,
+interactive use is made possible by the Python interactive interpreter, and its
+enhanced replacement IPython \citep{perez2007ipython}, which offer a
+\textsc{matlab}-like working environment specifically designed for scientific
+use.
 
@@
 and scientific packages centered around the NumPy and SciPy libraries. NumPy
-\citep{vanderwalt2011} augments Python with a numeric array datatype and
-associated array computing primitives while SciPy \citep{varoquaux2013scipy}
-extends it further by making common numeric operations available, either by
-implementing those in Python/NumPy or by wrapping existing
-C/C{}\verb!++!/Fortran implementations. In a similar vein, \sklearn extends
-Python with a set of composable machine learning operations.
+\citep{vanderwalt2011} augments Python with a numeric array datatype and proper
+array computing primitives while SciPy \citep{varoquaux2013scipy} extends it
+further by making common numerical operations available, either by implementing
+those in Python/NumPy or by wrapping existing C/C{}\verb!++!/Fortran
+implementations. In a similar vein, \sklearn extends Python with a set of
+composable machine learning operations.
 
@@
 in-depth discussion and analysis of the design choices used in building the
-library. First, in section \ref{sec:core-api}, we present the central application
+library. We first present in section \ref{sec:core-api} the central application
 programming interface (API) and then describe, in section \ref{sec:advanced-api},
 advanced API mechanisms built from the core interface. Section
-\ref{sec:implementation} describes implementation trade-offs and analyzes
-limitations of the technologies we rely on, while section
-\ref{sec:comparison} then compares \emph{scikit-learn} to other major projects in
-terms of API. Finally, conclusions and directions of future works are included
-in section \ref{sec:conclusions}.
+\ref{sec:implementation} describes implementation trade-offs and discusses
+limitations of the technologies we rely on. Section \ref{sec:comparison}
+then compares \emph{scikit-learn} to other major projects in terms of API.
+Finally, conclusions and directions of future works are included in section
+\ref{sec:conclusions}.
 

### 2013-06-13 paper.tex
wordings in section 2

@@
 packages, many scientific users of Python are already familiar with NumPy dense
-arrays and SciPy sparse matrices, hence helping them to familiarize with the
-project.  From a practical point of view, these formats also provide a bunch of
+arrays and SciPy sparse matrices.
+% Gilles: the phrase was too long
+% hence helping them to familiarize with the project.
+From a practical point of view, these formats also provide a bunch of
 data loading and conversion tools which make them very easy to use in many
@@
 
-The public API is oriented towards batch processing rather than processing
-single samples per API call. While classification and regression algorithms can
-indeed make predictions for single samples, \sklearn objects are not optimized
-for this use case. (The few online learning algorithms implemented are intended
-to take minibatches.) Batch processing makes optimal use of NumPy and SciPy by
-preventing the overhead inherent to Python function calls or due to per-element
-dynamic type checking. Although this might seem to be an artifact of the Python
-language, and therefore an implementation detail that leaks into the API, we
-argue that APIs should be designed so as not to tie a library to a suboptimal
-implementation strategy. As such, batch processing enable fast implementations
-in lower-level languages (where memory hierarchy effects and the possibility of
-internal parallelization come into play).
+For efficiency reasons, the public interface is oriented towards processing
+batches of samples rather than single samples per API call. While classification
+and regression algorithms can indeed make predictions for single samples,
+\sklearn objects are not optimized for this use case. (The few online learning
+algorithms implemented are intended to take mini-batches.) Batch processing makes
+optimal use of NumPy and SciPy by preventing the overhead inherent to Python
+function calls or due to per-element dynamic type checking. Although this might
+seem to be an artifact of the Python language, and therefore an implementation
+detail that leaks into the API, we argue that APIs should be designed so as not
+to tie a library to a suboptimal implementation strategy. As such, batch
+processing enable fast implementations in lower-level languages (where memory
+hierarchy effects and the possibility of internal parallelization come into
+play).
 
@@
 performance on some input data. In supervised learning estimators, this method
-takes as inputs a dataset \texttt{X\_test} and \texttt{y\_test} and typically
+takes as inputs arrays \texttt{X\_test} and \texttt{y\_test} and typically
 computes the coefficient of determination between \texttt{y\_test} and
-\texttt{predict(X\_test)} in the case of classification or the mean squared
+\texttt{predict(X\_test)} in the case of classification, or the mean squared
 error in the case of regression. Not however that, from an API point of view,

### 2013-06-13 paper.tex
wordings in section 3

@@
 training set amounts to the following recursive procedure: i) when only one step
-remains, call its \texttt{fit} method; ii) otherwise, fit the first step, use it
-to \texttt{transform} the training set and fit the rest of the pipeline with the
-transformed data.
-The pipeline exposes all the methods the last
-estimator in the pipe exposes. That is, if the last estimator is a predictor,
-the pipeline can itself be used as a predictor. If the last estimator is a
-transformer, then the pipeline is itself a transformer. \textcolor{red}{TODO:
-Talk about nested attribute syntax.}
+remains, call its \texttt{fit} method; ii) otherwise, \texttt{fit} the first
+step, use it to \texttt{transform} the training set and \texttt{fit} the rest of
+the pipeline with the transformed data. The pipeline exposes all the methods the
+last estimator in the pipe exposes. That is, if the last estimator is a
+predictor, the pipeline can itself be used as a predictor. If the last estimator
+is a transformer, then the pipeline is itself a transformer.
 
@@
 combined to create complex objects. In particular, since both are themselves
-estimators, they can be used to form nested workflows. \textcolor{red}{TODO:
-elaborate a bit more on nested constructions.}
+estimators, they can be used to form nested workflows.
 
@@
 \textcolor{red}{TODO: mention that GridSearchCV objects are themselves estimators}
+\textcolor{red}{TODO: Talk about nested attribute syntax.}
 

### 2013-06-14 paper.tex
model selection

@@
 \subsection{Estimators}
+\label{sec:estimators}
 
@@
 
-\textcolor{red}{TODO: rewrite and briefly describe the API (see user guide)}\\
-\textcolor{red}{TODO: describe Scorer API?}\\
-\textcolor{red}{TODO: mention that GridSearchCV objects are themselves estimators}
-\textcolor{red}{TODO: Talk about nested attribute syntax.}
+As introduced in Section~\ref{sec:estimators}, hyper-parameter values set in the
+constructor of an estimator determine its behavior. In particular, depending on
+the problem, some values produce estimators that are better than others. In this
+setting, the problem of \textit{model selection} consists in finding, within
+some hyper-parameter space, the combination of hyper-parameter values yielding
+an estimator whose performance optimizes some score function on unseen data. For
+example, in the case of a decision tree, too small values for the
+\texttt{max\_depth} parameter will make the tree to underfit while too large
+values will make the tree to overfit the training data. The problem is therefore
+to find the best trade-off value in order to maximize the performance of the
+tree on unseen data.
 
-The aforementioned \texttt{Pipeline} object
-can be passed as an argument to a model selection algorithm.
-Available algorithms at the time of writing include random search
-\citep{bergstra2012} and ``grid search''.
-The latter, which we shall use in this example,
-performs an exhaustive search through a grid formed by the cartesian product
-of a set of possible values for each parameter, specified by the user.
-In our example, this grid might be given as:
+In \sklearn, both \texttt{GridSearchCV} and \texttt{RandomizedSearchCV} meta-estimators
+implement model selection. They take as inputs an estimator (basic or
+complex), a grid of hyper-parameter values, a cross-validation scheme and a
+score function. At \texttt{fit}, a \texttt{GridSearchCV} object generates
+exhaustively all hyper-parameter combinations from the grid, while a
+\texttt{RandomizedSearchCV} instance generates a limited number of combinations
+randomly sampled from the grid. Then, for each combination, the meta-estimator
+fits the given estimator with the corresponding hyper-parameter values and
+evaluates its performance through the given cross-validation scheme. Finally,
+the best performing model is retained and exposed as a public attribute.
 
-\begin{align*}
-         & \textsf{stopwords} \in \{0, 1\}                      \\
-  \times & \; \textsf{tf} \in \{\textsf{linear}, \textsf{log}\} \\
-  \times & \; n_\textsf{features} \in \{1000, 2000, 5000\}      \\
-  \times & \; \textsf{norm} \in \{\ell_1, \ell_2\}              \\
-  \times & \; C \in \{10, 100, 1000\}
-\end{align*}
+From an API point of view, cross-validation schemes are given as iterator
+objects over train and test folds on the training data. The library provides
+various such iterators, including K-Fold, Leave-One-Out or shuffled
+cross-validation schemes. The score function used by default is the function
+provided by the estimator itself. However, the library provides a variety of
+score functions that can be used instead, including -- among many others --
+accuracy, AUC or $F_\beta$-score for classifiers; $R^2$ score or mean squared
+error for regressors; or the Rand Index for clustering estimators. Also, let us
+note that \texttt{GridSearchCV} and \texttt{RandomizedSearchCV} instances are
+themselves estimators. In particular, they provide \texttt{predict},
+\texttt{predict\_proba}, \texttt{transform} and \texttt{score} methods that are
+all transparently delegated to the best estimator found during \texttt{fit}.
 
-At each point in this grid, $k$-fold cross validation is run
-to estimate the performance of the estimator according to some measure
-(e.g., accuracy, $F_\beta$-score or a host of clustering quality metrics)
-and the best performing set of hyperparameters is stored
-on the \texttt{GridSearchCV} object, again, as a public attribute.
+As an example, the snippet below illustrates how to optimize the hyper-parameter
+values of an SVM classifier. It considers an exhaustive exploration of all
+hyper-parameter combinations from two distinct grids of parameter values and
+maximizes the AUC of the classifier through a 10-fold cross-validation:
+\begin{verbatim}
+    from sklearn.svm import SVC
+    from sklearn.grid_search import GridSearchCV
+
+    param_grid = [
+      {'C': [1, 10, 100, 1000], 'kernel': ['linear']},
+      {'C': [1, 10, 100, 1000], 'gamma': [0.001, 0.0001], 'kernel': ['rbf']},
+    ]
+
+    clf = GridSearchCV(SVC(), param_grid, scoring="roc_auc", cv=10)
+    clf.fit(X_train, y_train)
+\end{verbatim}
+
+\textcolor{red}{TODO: Mention that pipelines can also be grid-searched.}
 

### 2013-06-14 paper.tex
cite Bergstra paper
some work on the conclusions

@@
 \texttt{RandomizedSearchCV} instance generates a limited number of combinations
-randomly sampled from the grid. Then, for each combination, the meta-estimator
+randomly sampled from the grid \citep{bergstra2012}. Then, for each combination, the meta-estimator
 fits the given estimator with the corresponding hyper-parameter values and
@@
 
-ToDo.
+\textcolor{red}{TODO: include some overall conclusions (e.g., reasons of the success sklearn)}
 
-\subsection{Future work}
+Although it is coming close to a 1.0 release, the \sklearn toolkit is by no
+means finished. At the time, it does not support all tasks in the broad field of
+machine learning, but does not intend to either. Among others, missing features
+include neural networks, ensemble meta-estimators for bagging or subsampling
+strategies, or missing value completion algorithms provided as transformers.
+However, topics like structured prediction or reinforcement learning are
+considered off-topic for the project. The library also does not provide any
+collaborative filtering algorithms -- it remains to be seen whether a compatible
+and matching API can be written for them.
 
-The \sklearn toolkit is by no means finished,
-although it is coming close to a 1.0 release.
-
-\sklearn does not support all tasks in the broad field of machine learning,
-and does not intend to either;
-in particular, structured prediction and bandit-style online learning algorithms
-are considered ``off-topic'' for the library,
-and submissions of implementations of these for inclusion
-will (at least at present) be rejected.
-Collaborative filtering algorithms are not implemented
-and it is remains to be seen
-whether an API for them can be designed that matches the \sklearn API.
-
-At a lower-level, a potential point of improvement
-is parallel processing, which, as described in Section~\ref{sec:parallel},
-is only implemented at a rather coarse-grained level.
-Cython's built-in support for OpenMP \citep{dagum1998openmp}
-remains a viable candidate technology
-for more fine-grained multicore support in \sklearn.
+At a lower-level, a potential point of improvement is parallel processing,
+which, as described in Section~\ref{sec:parallel}, is only implemented at a
+rather coarse-grained level. Cython built-in support for OpenMP
+\citep{dagum1998openmp} remains a viable candidate technology for more fine-
+grained multicore support in \sklearn.
 

### 2013-06-14 paper.tex
move remark on gridsearchcv after the snippet

@@
 accuracy, AUC or $F_\beta$-score for classifiers; $R^2$ score or mean squared
-error for regressors; or the Rand Index for clustering estimators. Also, let us
-note that \texttt{GridSearchCV} and \texttt{RandomizedSearchCV} instances are
-themselves estimators. In particular, they provide \texttt{predict},
-\texttt{predict\_proba}, \texttt{transform} and \texttt{score} methods that are
-all transparently delegated to the best estimator found during \texttt{fit}.
+error for regressors; or the Rand Index for clustering estimators.
 
@@
     clf.fit(X_train, y_train)
+    y_pred = clf.predict(X_test)
 \end{verbatim}
+Let us finally note \texttt{GridSearchCV} and \texttt{RandomizedSearchCV}
+instances are themselves estimators. In particular, they provide
+\texttt{predict}, \texttt{predict\_proba}, \texttt{transform} and \texttt{score}
+methods that are all transparently delegated to the best estimator found during
+\texttt{fit}, as the example shows.
 

### 2013-06-14 paper.tex
author list

@@
 
-\author{Mathieu Blondel\inst{1} \and
-        Lars Buitinck\inst{2} \and
-        Gilles Louppe\inst{3} \and
-        Jake Vanderplas\inst{4}}
+\author{Lars Buitinck~\inst{1,2} \and
+        Gilles Louppe~\inst{1,2} \and
+        Mathieu Blondel~\inst{1,2} \and
+        Fabian Pedregosa~\inst{2} \and
+        Andreas Mueller~\inst{2} \and
+        Gaël Varoquaux~\inst{2} \and
+        Olivier Grisel~\inst{2} \and
+        Vlad Niculae~\inst{2} \and
+        Peter Prettenhofer~\inst{2} \and
+        Alexandre Gramfort~\inst{2} \and
+        Jacques Grobler~\inst{2} \and
+        Robert Layton~\inst{2} \and
+        Jake Vanderplas~\inst{2} \and
+        Arnaud Joly~\inst{2} \and
+        Bernard Thirion}
 
-\institute{Graduate School of System Informatics, Kobe University \and
-           Informatics Institute, University of Amsterdam \and
-           Department of EE \& CS, University of Liège \and
-           Astronomy Department, University of Washington}
+\institute{Main authors of the present paper \and
+           Core contributors to the project}
 

### 2013-06-14 paper.tex
author list (2)

--- 2013-06-14 paper.tex
@@
 
-\author{Mathieu Blondel\inst{1} \and
-        Lars Buitinck\inst{2} \and
-        Gilles Louppe\inst{3} \and
-        Jake Vanderplas\inst{4}}
+\author{Lars Buitinck~\inst{1,2} \and
+        Gilles Louppe~\inst{1,2} \and
+        Mathieu Blondel~\inst{1,2} \and
+        Fabian Pedregosa~\inst{2} \and
+        Andreas Mueller~\inst{2} \and
+        Gaël Varoquaux~\inst{2} \and
+        Olivier Grisel~\inst{2} \and
+        Vlad Niculae~\inst{2} \and
+        Peter Prettenhofer~\inst{2} \and
+        Alexandre Gramfort~\inst{2} \and
+        Jacques Grobler~\inst{2} \and
+        Robert Layton~\inst{2} \and
+        Jake Vanderplas~\inst{2} \and
+        Arnaud Joly~\inst{2} \and
+        Bernard Thirion~\inst{2}}
 
-\institute{Graduate School of System Informatics, Kobe University \and
-           Informatics Institute, University of Amsterdam \and
-           Department of EE \& CS, University of Liège \and
-           Astronomy Department, University of Washington}
+\institute{Main authors of the present paper \and
+           Core contributors to the project}
 

### 2013-06-14 paper.tex
do not break names

--- 2013-06-14 paper.tex
@@
 
-\author{Mathieu Blondel\inst{1} \and
-        Lars Buitinck\inst{2} \and
-        Gilles Louppe\inst{3} \and
-        Jake Vanderplas\inst{4}}
+\author{Lars~Buitinck~\inst{1,2} \and
+        Gilles~Louppe~\inst{1,2} \and
+        Mathieu~Blondel~\inst{1,2} \and
+        Fabian~Pedregosa~\inst{2} \and
+        Andreas~Mueller~\inst{2} \and
+        Gaël~Varoquaux~\inst{2} \and
+        Olivier~Grisel~\inst{2} \and
+        Vlad~Niculae~\inst{2} \and
+        Peter~Prettenhofer~\inst{2} \and
+        Alexandre~Gramfort~\inst{2} \and
+        Jacques~Grobler~\inst{2} \and
+        Robert~Layton~\inst{2} \and
+        Jake~Vanderplas~\inst{2} \and
+        Arnaud~Joly~\inst{2} \and
+        Bernard~Thirion~\inst{2}}
 
-\institute{Graduate School of System Informatics, Kobe University \and
-           Informatics Institute, University of Amsterdam \and
-           Department of EE \& CS, University of Liège \and
-           Astronomy Department, University of Washington}
+\institute{Main authors of the present paper \and
+           Core contributors to the project}
 

### 2013-06-14 paper.tex
remove Bertrand, move Gael as last author

--- 2013-06-14 paper.tex
@@
 
-\author{Lars Buitinck~\inst{1,2} \and
-        Gilles Louppe~\inst{1,2} \and
-        Mathieu Blondel~\inst{1,2} \and
-        Fabian Pedregosa~\inst{2} \and
-        Andreas Mueller~\inst{2} \and
-        Gaël Varoquaux~\inst{2} \and
-        Olivier Grisel~\inst{2} \and
-        Vlad Niculae~\inst{2} \and
-        Peter Prettenhofer~\inst{2} \and
-        Alexandre Gramfort~\inst{2} \and
-        Jacques Grobler~\inst{2} \and
-        Robert Layton~\inst{2} \and
-        Jake Vanderplas~\inst{2} \and
-        Arnaud Joly~\inst{2} \and
-        Bernard Thirion}
+\author{Lars~Buitinck~\inst{1,2} \and
+        Gilles~Louppe~\inst{1,2} \and
+        Mathieu~Blondel~\inst{1,2} \and
+        Fabian~Pedregosa~\inst{2} \and
+        Andreas~Mueller~\inst{2} \and
+        Olivier~Grisel~\inst{2} \and
+        Vlad~Niculae~\inst{2} \and
+        Peter~Prettenhofer~\inst{2} \and
+        Alexandre~Gramfort~\inst{2} \and
+        Jacques~Grobler~\inst{2} \and
+        Robert~Layton~\inst{2} \and
+        Jake~Vanderplas~\inst{2} \and
+        Arnaud~Joly~\inst{2} \and
+        Gaël~Varoquaux~\inst{2}}
 

### 2013-06-15 paper.tex
Add Brian.
This way that makes it 15.

@@
         Arnaud~Joly~\inst{2} \and
+        Brian Holt~\inst{2} \and
         Gaël~Varoquaux~\inst{2}}

### 2013-06-15 paper.tex
Add blank line after imports.

@@
     from sklearn.linear_model import LogisticRegression
+
     clf = LogisticRegression(penalty="l1")
@@
     from sklearn.cluster import KMeans
+
     km = KMeans(n_clusters=10)
@@
     from sklearn.preprocessing import StandardScaler
+
     scaler = StandardScaler()

### 2013-06-15 paper.tex
Rework model selection.

@@
 
-As introduced in Section~\ref{sec:estimators}, hyper-parameter values set in the
+As introduced in Section~\ref{sec:estimators}, hyper-parameters set in the
 constructor of an estimator determine its behavior. In particular, depending on
-the problem, some values produce estimators that are better than others. In this
-setting, the problem of \textit{model selection} consists in finding, within
-some hyper-parameter space, the combination of hyper-parameter values yielding
-an estimator whose performance optimizes some score function on unseen data. For
-example, in the case of a decision tree, too small values for the
-\texttt{max\_depth} parameter will make the tree to underfit while too large
-values will make the tree to overfit the training data. The problem is therefore
-to find the best trade-off value in order to maximize the performance of the
-tree on unseen data.
+the problem, some hyper-parameters produce estimators that are better than
+others. The problem of \textit{model selection} is therefore to find, within
+some hyper-parameter space, the best combination of hyper-parameters, with
+respect to some user-chosen criterion. For example, in the case of a decision
+tree, too small values of the \texttt{max\_depth} parameter will make the tree
+underfit while too large values will make the tree overfit the training data.
+The problem is therefore to find the best trade-off in order to maximize
+the performance of the tree on unseen data.
 
-In \sklearn, both \texttt{GridSearchCV} and \texttt{RandomizedSearchCV} meta-estimators
-implement model selection. They take as inputs an estimator (basic or
-complex), a grid of hyper-parameter values, a cross-validation scheme and a
-score function. At \texttt{fit}, a \texttt{GridSearchCV} object generates
-exhaustively all hyper-parameter combinations from the grid, while a
-\texttt{RandomizedSearchCV} instance generates a limited number of combinations
-randomly sampled from the grid \citep{bergstra2012}. Then, for each combination, the meta-estimator
-fits the given estimator with the corresponding hyper-parameter values and
-evaluates its performance through the given cross-validation scheme. Finally,
-the best performing model is retained and exposed as a public attribute.
+In \sklearn, model selection is supported in two distinct meta-estimators,
+\texttt{GridSearchCV} and \texttt{RandomizedSearchCV}.  They take as inputs an
+estimator (basic or composite), whose hyper-parameters must be optimized, and a
+hyper-parameter grid, which defines the values to search. Optionally, they can
+also take a cross-validation scheme and a score function.  The library provides
+various such cross-validation schemes, including K-Fold (default), Leave-One-Out
+and Bootstrap. The score function used by default is the \texttt{score} method
+provided by the estimator. However, the library provides a variety of
+score functions that the user can choose instead, including -- among many others
+-- accuracy, AUC, $F_1$-score for classification; $R^2$ score or mean squared
+error for regression and the Rand Index for clustering.
 
-From an API point of view, cross-validation schemes are given as iterator
-objects over train and test folds on the training data. The library provides
-various such iterators, including K-Fold, Leave-One-Out or shuffled
-cross-validation schemes. The score function used by default is the function
-provided by the estimator itself. However, the library provides a variety of
-score functions that can be used instead, including -- among many others --
-accuracy, AUC or $F_\beta$-score for classifiers; $R^2$ score or mean squared
-error for regressors; or the Rand Index for clustering estimators.
+\texttt{GridSearchCV} and \texttt{RandomizedSearchCV} differ in the way they
+search the hyper-parameter space.  \texttt{GridSearchCV} exhaustively searches
+for all hyper-parameter combinations from the grid. This scheme can be expensive
+when there are many possible hyper-parameter combinations.
+\texttt{RandomizedSearchCV}, on the other hand, only searches for a limited
+number of combinations, randomly sampled from the grid. Such scheme has been
+shown to work well in practice while considerably reducing the computational
+cost \citep{bergstra2012}. For each hyper-parameter combination and each train /
+validation split generated by the cross-validation scheme, \texttt{GridSearchCV}
+and \texttt{RandomizedSearchCV} fit the base estimator on the training set and
+evaluate its performance on the validation set.  In the end, the best performing
+model on average is retained and exposed as the public attribute
+\texttt{best\_estimator\_}.
 
-As an example, the snippet below illustrates how to optimize the hyper-parameter
-values of an SVM classifier. It considers an exhaustive exploration of all
-hyper-parameter combinations from two distinct grids of parameter values and
-maximizes the AUC of the classifier through a 10-fold cross-validation:
+As an example, the snippet below illustrates how to optimize the
+hyper-parameters of an SVM classifier. It maximizes the $F_1$ score of the
+classifier through a 10-fold cross-validation scheme.
 \begin{verbatim}
@@
 
-    clf = GridSearchCV(SVC(), param_grid, scoring="roc_auc", cv=10)
+    clf = GridSearchCV(SVC(), param_grid, scoring="f1", cv=10)
     clf.fit(X_train, y_train)
@@
 \end{verbatim}
-Let us finally note \texttt{GridSearchCV} and \texttt{RandomizedSearchCV}
-instances are themselves estimators. In particular, they provide
-\texttt{predict}, \texttt{predict\_proba}, \texttt{transform} and \texttt{score}
-methods that are all transparently delegated to the best estimator found during
-\texttt{fit}, as the example shows.
+Note that, in the above example, two distinct hyper-parameter grids are
+considered. This is because it only makes sense to search for \texttt{gamma}
+when \texttt{kernel} is set to \texttt{"rbf"}.  Additionally, we see that
+\texttt{GridSearchCV} has a \texttt{predict} method, like any other classifier.
+This is because \texttt{GridSearchCV} and \texttt{RandomizedSearchCV} in fact
+delegate the \texttt{predict}, \texttt{predict\_proba}, \texttt{transform} and
+\texttt{score} methods to the best estimator, thus mimicking its behavior.
 

### 2013-06-15 paper.tex
Cosmit.
Remove remark on collaborative filtering.
Nicolas will work on matrix completion during his GSOC, which can be
used to do for collaborative filtering.
Cosmit conclusion.

@@
 % ddaa494c116e3c16bf032003c5cccbed851733d2
-also count 183 contributors to the project (after manual deduplication).
+also count 183 unique contributors to the project.
 
@@
 
-Although it is coming close to a 1.0 release, the \sklearn toolkit is by no
-means finished. At the time, it does not support all tasks in the broad field of
-machine learning, but does not intend to either. Among others, missing features
-include neural networks, ensemble meta-estimators for bagging or subsampling
-strategies, or missing value completion algorithms provided as transformers.
-However, topics like structured prediction or reinforcement learning are
-considered off-topic for the project. The library also does not provide any
-collaborative filtering algorithms -- it remains to be seen whether a compatible
-and matching API can be written for them.
+Although it is coming close to a 1.0 release, \sklearn is by no means finished.
+At the time, it does not support many classical machine learning tasks. Among
+others, missing features include neural networks, ensemble meta-estimators for
+bagging or subsampling strategies, or missing value completion algorithms
+provided as transformers.  However, topics like structured prediction or
+reinforcement learning are considered off-topic for the project, since they
+would require quite different data representations and API.
 

### 2013-06-15 paper.tex
Syntax highlighting.

@@
 \usepackage{xspace}
+\usepackage{listings}
+\usepackage{color}
+
+\definecolor{htmltitlecolor}{rgb}{0.23, 0.04, 0}
+\definecolor{darkred}{rgb}{0.7, 0, 0}
+
+\definecolor{SkoleBrown}{RGB}{150,100,50}
+\definecolor{SkoleYellow}{RGB}{232,183,3}
+\definecolor{SkoleOrange}{RGB}{252,135,36}
+\colorlet{SkoleLightBrown}{SkoleBrown!50!SkoleYellow}
+%\colorlet{titlecolor}{SkoleBrown!85!black}
+\definecolor{darkgreen}{cmyk}{0.7, 0, 1, 0.5}
+\definecolor{darkblue}{cmyk}{1, 0.8, 0, 0}
+\definecolor{lightblue}{cmyk}{0.05,0,0,0.05}
+\definecolor{grey}{cmyk}{0.1,0.1,0.1,1}
+\definecolor{lightgrey}{cmyk}{0,0,0,0.5}
+\definecolor{purple}{cmyk}{0.8,1,0,0}
+
+\colorlet{titlecolor}{SkoleOrange!50!black}
+\colorlet{codeblockcolorbase}{SkoleLightBrown!80!black}
+\colorlet{codeblockcolor}{codeblockcolorbase!08}
+
+\lstset{language=python,
+extendedchars=true,
+backgroundcolor=\color{codeblockcolor},
+frame=single,
+xleftmargin = 0pt,
+numbersep = 0pt,
+framexleftmargin = 0pt,
+resetmargins = true,
+framesep = 0pt,
+rulesep = 0pt,
+basicstyle=\ttfamily\footnotesize,
+fillcolor=\color{codeblockcolor},
+rulecolor=\color{lightgrey},
+aboveskip = 0.5ex,
+belowskip = 0.6ex,
+escapebegin={\color{darkgreen}},
+keywordstyle=\sffamily\color{lightgrey},
+identifierstyle=\ttfamily,
+commentstyle=\slshape\color{darkgreen},
+stringstyle=\ttfamily\color{blue},
+showstringspaces=false,
+tabsize=2,
+%breaklines=true,
+morekeywords={{,},=,:}, 
+classoffset=0,
+}
+
+% tt font with bold support
+%\renewcommand{\ttdefault}{pcr}
 
@@
 Given the API defined above, solving this problem is as simple as:
-\begin{verbatim}
+\begin{lstlisting}
     from sklearn.linear_model import LogisticRegression
@@
     clf.fit(X_train, y_train)
-\end{verbatim}
+\end{lstlisting}
 In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
@@
 snippet:
-\begin{verbatim}
+\begin{lstlisting}
     y_pred = clf.predict(X_test)
-\end{verbatim}
+\end{lstlisting}
 Some unsupervised learning
 estimators may however also implement that interface, as illustrated below:
-\begin{verbatim}
+\begin{lstlisting}
     from sklearn.cluster import KMeans
@@
     clust_pred = km.predict(X_test)
-\end{verbatim}
+\end{lstlisting}
 In the above example, a \texttt{KMeans} estimator is first fit on the training
@@
 could thus simply write:
-\begin{verbatim}
+\begin{lstlisting}
     from sklearn.preprocessing import StandardScaler
@@
     X_train = scaler.transform(X_train)
-\end{verbatim}
+\end{lstlisting}
 Of course, in practice, it is important to apply the same preprocessing to the
 test data \texttt{X\_test}. This can easily be done as follows:
-\begin{verbatim}
+\begin{lstlisting}
     X_test = scaler.transform(X_test)
-\end{verbatim}
+\end{lstlisting}
 Transformers also include a variety of learning algorithms, such as
@@
 we can thus rewrite the \texttt{StandardScaler} example above in a single line:
-\begin{verbatim}
+\begin{lstlisting}
     X_train = StandardScaler().fit(X_train).transform(X_train)
-\end{verbatim}
+\end{lstlisting}
 In addition, \texttt{fit(X\_train).transform(X\_train)} can further
@@
 then feeding them to an $\ell_2$-regularized logistic regression model:
-\begin{verbatim}
+\begin{lstlisting}
     from sklearn.pipeline import Pipeline
@@
 
-    feat_union = FeatureUnion([('pca', PCA()),
+    fu = FeatureUnion([('pca', PCA()),
                        ('kpca', KernelPCA(kernel="rbf"))])
 
-    Pipeline([("pca", feat_union,
-              ("feature_selection", SelectKBest(k=100)),
-              ("logistic_model", LogisticRegression(penalty="l2"))
+    Pipeline([("pca", fu,
+              ("feat_sel", SelectKBest(k=100)),
+              ("log_reg", LogisticRegression(penalty="l2"))
     ]).fit(X_train, y_train).predict(X_test)
-\end{verbatim}
+\end{lstlisting}
 
@@
 classifier through a 10-fold cross-validation scheme.
-\begin{verbatim}
+\begin{lstlisting}
     from sklearn.svm import SVC
@@
     param_grid = [
-      {'C': [1, 10, 100, 1000], 'kernel': ['linear']},
-      {'C': [1, 10, 100, 1000], 'gamma': [0.001, 0.0001], 'kernel': ['rbf']},
+      {"C": [1, 10, 100, 1000], "kernel": ["linear"]},
+      {"C": [1, 10, 100, 1000], "gamma": [0.001, 0.0001], 
+       "kernel": ["rbf"]},
     ]
@@
     y_pred = clf.predict(X_test)
-\end{verbatim}
+\end{lstlisting}
 Note that, in the above example, two distinct hyper-parameter grids are

### 2013-06-15 paper.tex
snippet style

@@
 
+\definecolor{rulecolor}{RGB}{210,210,210}
 \colorlet{titlecolor}{SkoleOrange!50!black}
@@
 
-\lstset{language=python,
+\lstset{language=Python,
 extendedchars=true,
@@
 fillcolor=\color{codeblockcolor},
-rulecolor=\color{lightgrey},
+rulecolor=\color{rulecolor},
 aboveskip = 0.5ex,
@@
 escapebegin={\color{darkgreen}},
-keywordstyle=\sffamily\color{lightgrey},
+keywordstyle=\sffamily\color{darkblue},
 identifierstyle=\ttfamily,
 commentstyle=\slshape\color{darkgreen},
-stringstyle=\ttfamily\color{blue},
+stringstyle=\ttfamily\color{orange},
 showstringspaces=false,

### 2013-06-15 paper.tex
Switch to minted.

--- 2013-06-15 paper.tex
@@
 \usepackage{xspace}
+\usepackage{minted}
+
+\newminted{python}{frame=single}
+
+% tt font with bold support
+%\renewcommand{\ttdefault}{pcr}
 
@@
 Given the API defined above, solving this problem is as simple as:
-\begin{verbatim}
+\begin{pythoncode}
     from sklearn.linear_model import LogisticRegression
@@
     clf.fit(X_train, y_train)
-\end{verbatim}
+\end{pythoncode}
 In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
@@
 snippet:
-\begin{verbatim}
+\begin{pythoncode}
     y_pred = clf.predict(X_test)
-\end{verbatim}
+\end{pythoncode}
 Some unsupervised learning
 estimators may however also implement that interface, as illustrated below:
-\begin{verbatim}
+\begin{pythoncode}
     from sklearn.cluster import KMeans
@@
     clust_pred = km.predict(X_test)
-\end{verbatim}
+\end{pythoncode}
 In the above example, a \texttt{KMeans} estimator is first fit on the training
@@
 could thus simply write:
-\begin{verbatim}
+\begin{pythoncode}
     from sklearn.preprocessing import StandardScaler
@@
     X_train = scaler.transform(X_train)
-\end{verbatim}
+\end{pythoncode}
 Of course, in practice, it is important to apply the same preprocessing to the
 test data \texttt{X\_test}. This can easily be done as follows:
-\begin{verbatim}
+\begin{pythoncode}
     X_test = scaler.transform(X_test)
-\end{verbatim}
+\end{pythoncode}
 Transformers also include a variety of learning algorithms, such as
@@
 we can thus rewrite the \texttt{StandardScaler} example above in a single line:
-\begin{verbatim}
+\begin{pythoncode}
     X_train = StandardScaler().fit(X_train).transform(X_train)
-\end{verbatim}
+\end{pythoncode}
 In addition, \texttt{fit(X\_train).transform(X\_train)} can further
@@
 then feeding them to an $\ell_2$-regularized logistic regression model:
-\begin{verbatim}
+\begin{pythoncode}
     from sklearn.pipeline import Pipeline
@@
 
-    feat_union = FeatureUnion([('pca', PCA()),
+    fu = FeatureUnion([('pca', PCA()),
                        ('kpca', KernelPCA(kernel="rbf"))])
 
-    Pipeline([("pca", feat_union,
-              ("feature_selection", SelectKBest(k=100)),
-              ("logistic_model", LogisticRegression(penalty="l2"))
+    Pipeline([("pca", fu,
+              ("feat_sel", SelectKBest(k=100)),
+              ("log_reg", LogisticRegression(penalty="l2"))
     ]).fit(X_train, y_train).predict(X_test)
-\end{verbatim}
+\end{pythoncode}
 
@@
 classifier through a 10-fold cross-validation scheme.
-\begin{verbatim}
+\begin{pythoncode}
     from sklearn.svm import SVC
@@
     y_pred = clf.predict(X_test)
-\end{verbatim}
+\end{pythoncode}
 Note that, in the above example, two distinct hyper-parameter grids are

### 2013-06-15 paper.tex
Indentation.
wordings in the conclusions

@@
 others, missing features include neural networks, ensemble meta-estimators for
-bagging or subsampling strategies, or missing value completion algorithms
-provided as transformers.  However, topics like structured prediction or
-reinforcement learning are considered off-topic for the project, since they
-would require quite different data representations and API.
+bagging or subsampling strategies, or missing value completion algorithms.
+Topics like structured prediction or reinforcement learning are however
+considered off-topic for the project, since they would require quite different
+data representations and API.
 

### 2013-06-15 paper.tex
Double quote consistency.
Cosmit.

@@
 
-In \sklearn, classical learning algorithms are not the only objects
-to be implemented as estimators. For example, preprocessing routines (e.g.,
-vectorization of text documents, scaling of features, etc) or feature
-extraction techniques also implement the same interface. Even those estimators
-that are entirely stateless and therefore do not require the \texttt{fit} method
-to perform any useful work implement the estimator interface. As we will
-illustrate in the next sections, this design pattern is indeed of prime
-importance for consistency, composition and parameter validation reasons.
+In \sklearn, classical learning algorithms are not the only objects to be
+implemented as estimators. For example, preprocessing routines (e.g., scaling of
+features) or feature extraction techniques (e.g., vectorization of text
+documents) also implement the same interface. Even those estimators that are
+entirely stateless and therefore do not require the \texttt{fit} method to
+perform any useful work implement the estimator interface. As we will illustrate
+in the next sections, this design pattern is indeed of prime importance for
+consistency, composition and model selection reasons.
 
@@
 
-    fu = FeatureUnion([('pca', PCA()),
-                       ('kpca', KernelPCA(kernel="rbf"))])
+fu = FeatureUnion([("pca", PCA()),
+                   ("kpca", KernelPCA(kernel="rbf"))])
 
@@
 \texttt{GridSearchCV} and \texttt{RandomizedSearchCV} differ in the way they
-search the hyper-parameter space.  \texttt{GridSearchCV} exhaustively searches
+search the hyper-parameter grid.  \texttt{GridSearchCV} exhaustively searches
 for all hyper-parameter combinations from the grid. This scheme can be expensive

### 2013-06-15 paper.tex
rulecolor in snippets
Syntax error.
Shorten section beginning.

@@
 
-\newminted{python}{frame=single}
+\definecolor{rulecolor}{rgb}{0.80,0.80,0.80}
+\newminted{python}{frame=single,rulecolor=\color{rulecolor}}
 
@@
 
-In this section, we discuss the general principles of the application
-programming interface (API) of \sklearn. We first specify the representation of
-data and then proceed by describing the core interface of the library.
-
 All objects within \sklearn share a uniform common basic API consisting of three
 complementary interfaces: an \textit{estimator} interface for building and
-fitting objects, a \textit{predictor} interface for making predictions and a
+fitting models, a \textit{predictor} interface for making predictions and a
 \textit{transformer} interface for converting data. As much as possible, our
@@
 
-Pipeline([("pca", fu,
+Pipeline([("feat_union", fu),
           ("feat_sel", SelectKBest(k=100)),

### 2013-06-15 paper.tex
remove remark
implementation

@@
 
-\textcolor{red}{TODO: Mention that pipelines can also be grid-searched.}
-
 \section{Implementation}
@@
 
-\noindent \textcolor{red}{TODO: Mention small set of dependencies.}\\
-\textcolor{red}{TODO: elaborate on use of mixins and duck typing}
+Our implementation guidelines emphasize on writing short and readable code,
+easily maintainable by external contributors, while being as efficient as
+possible.  Whenever practicable, algorithms implemented in \sklearn are thus
+directly written in pure Python. For intensive numerical computations, the
+library mostly relies on NumPy vector operations, which allows for the code to
+remain concise and readable while being efficient. For all algorithms that
+cannot be easily and efficiently expressed as NumPy operations, we rely on
+Cython \citep{behnel2011cython}  for writing the critical parts and achieving
+decent performance and scalability. Cython is a compiled programming language
+that extends Python with static typing. It produces efficient C extension
+modules that are directly importable from the Python runtime system. Examples of
+algorithms written in Cython include stochastic gradient descent for linear
+models, some graph-based clustering algorithms or decision trees. To facilitate
+the adoption of the library, we also limit to a bare minimum the set of external
+depencies. As such, NumPy and SciPy are actually the only two required
+dependencies for the library to fully work. Some modules may rely on other
+external libraries, but then we tend to prefer to directly ship them within the
+codebase rather than adding new dependencies. In particular, the codebase ships
+modified versions of the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
+\citep{chang2011libsvm,fan2008}, both written in C{}\verb!++! and wrapped using
+Cython modules.
 
-scikit-learn is primarily implemented in Python and Cython
-\citep{behnel2011cython},
-a language that extends Python with static typing
-and a compiler that produces C extension modules
-for the Python runtime system.
-In addition, it includes (modified versions of)
-the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
-used for training support vector machines
-and logistic regression models \citep{chang2011libsvm, fan2008}.
-These libraries are both written in C{}\verb!++!
-and wrapped using Cython modules.
+% scikit-learn is primarily implemented in Python and Cython
+% ,
+% a language that extends Python with static typing
+% and a compiler that produces C extension modules
+% for the Python runtime system.
+% In addition, it includes (modified versions of)
+% the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
+% used for training support vector machines
+% and logistic regression models \citep{chang2011libsvm, fan2008}.
+% These libraries are both written in C{}\verb!++!
+% and wrapped using Cython modules.
 
-As much as possible is implemented in ``pure'' Python
-with NumPy for numerical work,
-to avoid the compile-edit-debug cycle needed for Cython
-and to keep the codebase readable for as large an audience as possible.
-However, not all algorithms are easily and efficiently expressed
-as NumPy vector operations
-(among other factors, because they often have linear space requirements).
-Algorithms that fall into this category have to be implemented in Cython
-to obtain decent performance and scalability.
-Examples include $k$-means, stochastic gradient descent for linear models
-and some graph-based clustering algorithms.
+% As much as possible is implemented in ``pure'' Python
+% with NumPy for numerical work,
+% to avoid the compile-edit-debug cycle needed for Cython
+% and to keep the codebase readable for as large an audience as possible.
+% However, not all algorithms are easily and efficiently expressed
+% as NumPy vector operations
+% (among other factors, because they often have linear space requirements).
+% Algorithms that fall into this category have to be implemented in Cython
+% to obtain decent performance and scalability.
+% Examples include $k$-means, stochastic gradient descent for linear models
+% and some graph-based clustering algorithms.
 
-While object-oriented techniques and design patterns
-are used throughout scikit-learn,
-they mostly serve to facilitate code reuse and are not considered a project goal.
-Where possible, the Python principle of ``duck typing'' is exploited
-to simplify the implementation and skip the introduction of superfluous classes.
-This allows for extensibility and flexibility at the same time:
-user-defined estimators that follow scikit-learn conventions
-should be usable in pipelines and other composite objects
-without them actually inheriting from scikit-learn base classes.
-
-%\section{Weaknesses of the NumPy/SciPy environment}
-
-\textcolor{red}{TODO: The following corresponds to the old weaknesses section. It should be properly rewritten.}
+% While object-oriented techniques and design patterns
+% are used throughout scikit-learn,
+% they mostly serve to facilitate code reuse and are not considered a project goal.
+% Where possible, the Python principle of ``duck typing'' is exploited
+% to simplify the implementation and skip the introduction of superfluous classes.
+% This allows for extensibility and flexibility at the same time:
+% user-defined estimators that follow scikit-learn conventions
+% should be usable in pipelines and other composite objects
+% without them actually inheriting from scikit-learn base classes.
 

### 2013-06-15 paper.tex
duck typing

@@
 the adoption of the library, we also limit to a bare minimum the set of external
-depencies. As such, NumPy and SciPy are actually the only two required
-dependencies for the library to fully work. Some modules may rely on other
-external libraries, but then we tend to prefer to directly ship them within the
-codebase rather than adding new dependencies. In particular, the codebase ships
-modified versions of the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
+depencies. As such, NumPy and SciPy -- the two most common dependencies in the
+Python scientific ecosystem -- are actually the only two required dependencies
+for the library to fully work. Some modules may rely on other external
+libraries, but then we tend to prefer to directly ship them within the codebase
+rather than adding new dependencies. In particular, the codebase ships modified
+versions of the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
 \citep{chang2011libsvm,fan2008}, both written in C{}\verb!++! and wrapped using
@@
 % and some graph-based clustering algorithms.
+
+To ease code reuse, object-oriented techniques and design patterns are used all
+throughout the codebase. In particular, the \textit{duck typing} design pattern
+is widely exploited to simplify implementation and skip the introduction of
+superfluous classes. This technique indeed allows both for extensibility and
+flexibility: user-defined estimators that follow our API and conventions should
+be reusable (e.g., in a pipeline) as any built-in estimators, without them
+actually inheriting from our base classes.
 

### 2013-06-15 paper.tex
typo

@@
 the adoption of the library, we also limit to a bare minimum the set of external
-depencies. As such, NumPy and SciPy -- the two most common dependencies in the
+dependencies. As such, NumPy and SciPy -- the two most common dependencies in the
 Python scientific ecosystem -- are actually the only two required dependencies
@@
 
-To ease code reuse, object-oriented techniques and design patterns are used all
-throughout the codebase. In particular, the \textit{duck typing} design pattern
-is widely exploited to simplify implementation and skip the introduction of
-superfluous classes. This technique indeed allows both for extensibility and
+To ease code reuse, object-oriented techniques and design patterns are also used
+all throughout the codebase. In particular, the \textit{duck typing} design
+pattern is widely exploited to simplify implementation and skip the introduction
+of superfluous classes. This technique indeed allows both for extensibility and
 flexibility: user-defined estimators that follow our API and conventions should

### 2013-06-15 paper.tex
parallelism

@@
 possible.  Whenever practicable, algorithms implemented in \sklearn are thus
-directly written in pure Python. For intensive numerical computations, the
-library mostly relies on NumPy vector operations, which allows for the code to
-remain concise and readable while being efficient. For all algorithms that
+directly written in pure Python. For intensive numerical computations though,
+the library mostly relies on NumPy vector operations, which allows for the code
+to remain concise and readable while being efficient. For all algorithms that
 cannot be easily and efficiently expressed as NumPy operations, we rely on
@@
 models, some graph-based clustering algorithms or decision trees. To facilitate
-the adoption of the library, we also limit to a bare minimum the set of external
-dependencies. As such, NumPy and SciPy -- the two most common dependencies in the
-Python scientific ecosystem -- are actually the only two required dependencies
-for the library to fully work. Some modules may rely on other external
-libraries, but then we tend to prefer to directly ship them within the codebase
-rather than adding new dependencies. In particular, the codebase ships modified
-versions of the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
-\citep{chang2011libsvm,fan2008}, both written in C{}\verb!++! and wrapped using
-Cython modules.
+the adoption of the library, we also try to limit to a bare minimum the set of
+external dependencies. As such, NumPy and SciPy -- the two most common
+dependencies in the Python scientific ecosystem -- are actually the only two
+required dependencies for the library to fully work. Some modules may rely on
+other external libraries, but in such cases we tend to prefer to directly ship
+them within the codebase rather than adding new dependencies. In particular, the
+codebase includes modified versions of the \textsf{LIBSVM} and
+\textsf{LIBLINEAR} libraries \citep{chang2011libsvm,fan2008}, both written in
+C{}\verb!++! and wrapped using Cython modules.
 
@@
 
-At this point in time, support for multi-core programming in Python,
-at least in the CPython reference implementation targeted by scikit-learn,
-is suboptimal.
-The interpreter supports multithreading,
-but due to the ``Global Interpreter Lock'' (GIL),
-no two threads in a process may at the same time execute Python code.
-It follows that multithreading is not a viable solution
-even for embarrassingly parallel problems,
-such as fitting and evaluation multiple models in a grid search procedure.
-The workaround is to use Python's multiprocessing module,
-which emulates threads using multiple operating system-level processes,
-but this in turn leads to high memory use and overhead
-because of the copying of training and test sets
-into the various processes' address spaces.
+In many cases, machine learning algorithms can be decomposed into independent
+smaller tasks that may each be solved in parallel, thereby reducing the overall
+computing times. Some algorithms in \sklearn are implemented as such, and
+exploit multi-core architectures.
+
+At this point in time, support for multi-core programming in Python -- at least
+in the CPython implementation targeted by scikit-learn -- is however suboptimal.
+The interpreter supports multithreading, but due to the \textit{global
+interpreter lock}, no two threads in a process may execute Python code at the
+same time. It follows that multithreading is not a viable solution for
+parallelizing parts of an algorithm, even for embarrassingly parallel problems,
+such as fitting and evaluating multiple models in a grid search procedure. The
+workaround that we use is to rely on the Python multiprocessing module, which
+emulates threads using multiple operating system-level processes. However this
+solution is not perfect as it may in turn lead to a large memory overhead since
+objects (e.g., input datasets) need to be copied within each process own memory
+space. In some cases, it may unfortunately even happen that the overhead becomes
+so significant that fully exploiting a multi-core architecture becomes
+infeasible due memory constraints. As such, a proper implementation of parallel
+machine learning algorithms remain a challenging task for the project.
 

### 2013-06-16 paper.tex
Move duck typing and trim / merge "Cython for inner loops".
@glouppe: I moved the duck typing paragraph to the meta-estimator
subsection, where I think the concept could introduced nicely (e.g.,
by saying that duck typing allows to use any estimator as base
estimator).

@larsmans: I trimmed and moved the part "Cython for inner loops".
BTW, SGD uses loss functions implemented in Cython but they are passed
as argument from the pure Python side. Yet, the performance seems
reasonable. I wonder if we could use a similar approach for kernel-based
methods...

@@
 
+\textcolor{red}{Point out that any external estimator can be used as base
+estimator thanks to duck typing}\\
+
+To ease code reuse, object-oriented techniques and design patterns are also used
+all throughout the codebase. In particular, the \textit{duck typing} design
+pattern is widely exploited to simplify implementation and skip the introduction
+of superfluous classes. This technique indeed allows both for extensibility and
+flexibility: user-defined estimators that follow our API and conventions should
+be reusable (e.g., in a pipeline) as any built-in estimators, without them
+actually inheriting from our base classes.
+
 \subsection{Pipelines and feature unions}
@@
 \section{Implementation}
+\label{sec:implementation}
 
-\label{sec:implementation}
+\subsection{Cython for performance}
 
@@
 
-% scikit-learn is primarily implemented in Python and Cython
-% ,
-% a language that extends Python with static typing
-% and a compiler that produces C extension modules
-% for the Python runtime system.
-% In addition, it includes (modified versions of)
-% the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
-% used for training support vector machines
-% and logistic regression models \citep{chang2011libsvm, fan2008}.
-% These libraries are both written in C{}\verb!++!
-% and wrapped using Cython modules.
-
-% As much as possible is implemented in ``pure'' Python
-% with NumPy for numerical work,
-% to avoid the compile-edit-debug cycle needed for Cython
-% and to keep the codebase readable for as large an audience as possible.
-% However, not all algorithms are easily and efficiently expressed
-% as NumPy vector operations
-% (among other factors, because they often have linear space requirements).
-% Algorithms that fall into this category have to be implemented in Cython
-% to obtain decent performance and scalability.
-% Examples include $k$-means, stochastic gradient descent for linear models
-% and some graph-based clustering algorithms.
-
-To ease code reuse, object-oriented techniques and design patterns are also used
-all throughout the codebase. In particular, the \textit{duck typing} design
-pattern is widely exploited to simplify implementation and skip the introduction
-of superfluous classes. This technique indeed allows both for extensibility and
-flexibility: user-defined estimators that follow our API and conventions should
-be reusable (e.g., in a pipeline) as any built-in estimators, without them
-actually inheriting from our base classes.
-
-% While object-oriented techniques and design patterns
-% are used throughout scikit-learn,
-% they mostly serve to facilitate code reuse and are not considered a project goal.
-% Where possible, the Python principle of ``duck typing'' is exploited
-% to simplify the implementation and skip the introduction of superfluous classes.
-% This allows for extensibility and flexibility at the same time:
-% user-defined estimators that follow scikit-learn conventions
-% should be usable in pipelines and other composite objects
-% without them actually inheriting from scikit-learn base classes.
+A problem arising from the use of Cython, however, is that passing pure Python
+functions as argument, while possible, degrades performance significantly, due
+to the runtime checks needed for each function invocation. This especially
+problematic, for example for kernel methods, which would benefit from accepting
+a kernel function as argument. A potential solution to this problem would be for
+\sklearn to provide a Cython-level API, thereby alllowing the user to pass
+Cython-implemented kernel functions.
 
@@
 machine learning algorithms remain a challenging task for the project.
-
-\subsection{Cython for inner loops}
-
-A problem arising from the use of Cython
-concerns the mismatch between on the one hand Python functions,
-which are first-class runtime objects endowed with methods and metadata
-that identify them as ``callable'',
-that dynamically check their number and types of arguments,
-and that can return any Python type of object;
-and compiled C or Cython functions on the other hand,
-which exhibit static typing.
-Where it may sometimes be convenient for estimators
-to take functions as arguments
-(e.g., in the case of kernel machine learning,
-where one might like to use a custom kernel),
-but such usage is not at present possible and may never be.
-Passing C or Cython-compiled functions
-would require an API for the kernel method
-to be implemented at the C or Cython level,
-only to be used from compiled code.
-Passing arbitrary Python functions would be possible,
-but performance can be expected to degrade significantly
-because of the runtime checks needed for each invocation of the kernel function.
-(The current solution in scikit-learn is to let the user pass
-not a kernel function,
-but a precomputed Gram matrix for an entire set of objects.
-The linear, polynomial and RBF kernels are treated as special cases in
-the \textsf{LIBSVM}-based SVM learner,
-where they are implemented in C{}\verb!++!.)
 

### 2013-06-16 paper.tex
remove subsections in Implementation

@@
 
-\subsection{Cython for performance}
+% \subsection{Cython for performance}
 
@@
 easily maintainable by external contributors, while being as efficient as
-possible.  Whenever practicable, algorithms implemented in \sklearn are thus
+possible.
+
+Whenever practicable, algorithms implemented in \sklearn are thus
 directly written in pure Python. For intensive numerical computations though,
 the library mostly relies on NumPy vector operations, which allows for the code
-to remain concise and readable while being efficient. For all algorithms that
-cannot be easily and efficiently expressed as NumPy operations, we rely on
-Cython \citep{behnel2011cython}  for writing the critical parts and achieving
-decent performance and scalability. Cython is a compiled programming language
-that extends Python with static typing. It produces efficient C extension
-modules that are directly importable from the Python runtime system. Examples of
+to remain concise, readable and efficient. For all algorithms that cannot be
+easily and efficiently expressed as NumPy operations, we rely on Cython
+\citep{behnel2011cython}  for writing the critical parts and achieving decent
+performance and scalability. Cython is a compiled programming language that
+extends Python with static typing. It produces efficient C extension modules
+that are directly importable from the Python runtime system. Examples of
 algorithms written in Cython include stochastic gradient descent for linear
-models, some graph-based clustering algorithms or decision trees. To facilitate
-the adoption of the library, we also try to limit to a bare minimum the set of
-external dependencies. As such, NumPy and SciPy -- the two most common
-dependencies in the Python scientific ecosystem -- are actually the only two
-required dependencies for the library to fully work. Some modules may rely on
-other external libraries, but in such cases we tend to prefer to directly ship
-them within the codebase rather than adding new dependencies. In particular, the
-codebase includes modified versions of the \textsf{LIBSVM} and
+models, some graph-based clustering algorithms or decision trees.  A problem
+arising from the use of Cython, however, is that passing pure Python functions
+as argument, while possible, degrades performance significantly due to the
+runtime checks needed for each function invocation. This is especially
+problematic for example for kernel methods, which would benefit from accepting a
+kernel function as argument. A potential solution to this problem would be for
+\sklearn to provide a Cython-level API, thereby alllowing the user to pass
+Cython-implemented kernel functions.
+
+To facilitate the adoption of the library, we also try to limit to a bare
+minimum the set of external dependencies. As such, NumPy and SciPy -- the two
+most common dependencies in the Python scientific ecosystem -- are actually the
+only two required dependencies for the library to fully work. Some modules may
+rely on other external libraries, but in such cases we tend to prefer to
+directly ship them within the codebase rather than adding new dependencies. In
+particular, the codebase includes modified versions of the \textsf{LIBSVM} and
 \textsf{LIBLINEAR} libraries \citep{chang2011libsvm,fan2008}, both written in
@@
 
-A problem arising from the use of Cython, however, is that passing pure Python
-functions as argument, while possible, degrades performance significantly, due
-to the runtime checks needed for each function invocation. This especially
-problematic, for example for kernel methods, which would benefit from accepting
-a kernel function as argument. A potential solution to this problem would be for
-\sklearn to provide a Cython-level API, thereby alllowing the user to pass
-Cython-implemented kernel functions.
-
-\subsection{Parallelism}
-\label{sec:parallel}
-
-In many cases, machine learning algorithms can be decomposed into independent
+Finally, some algorithms in \sklearn also support multi-core architectures.
+Machine learning algorithms can indeed often be decomposed into independent
 smaller tasks that may each be solved in parallel, thereby reducing the overall
-computing times. Some algorithms in \sklearn are implemented as such, and
-exploit multi-core architectures.
-
-At this point in time, support for multi-core programming in Python -- at least
-in the CPython implementation targeted by scikit-learn -- is however suboptimal.
-The interpreter supports multithreading, but due to the \textit{global
-interpreter lock}, no two threads in a process may execute Python code at the
-same time. It follows that multithreading is not a viable solution for
-parallelizing parts of an algorithm, even for embarrassingly parallel problems,
-such as fitting and evaluating multiple models in a grid search procedure. The
-workaround that we use is to rely on the Python multiprocessing module, which
-emulates threads using multiple operating system-level processes. However this
-solution is not perfect as it may in turn lead to a large memory overhead since
-objects (e.g., input datasets) need to be copied within each process own memory
-space. In some cases, it may unfortunately even happen that the overhead becomes
-so significant that fully exploiting a multi-core architecture becomes
-infeasible due memory constraints. As such, a proper implementation of parallel
-machine learning algorithms remain a challenging task for the project.
+computing times. At this point in time, support for multi-core programming in
+Python -- at least in the CPython implementation targeted by scikit-learn -- is
+however suboptimal. The interpreter supports multithreading, but due to the
+\textit{global interpreter lock}, no two threads in a process may execute Python
+code at the same time. It follows that multithreading is not a viable solution
+for parallelizing parts of an algorithm, even for embarrassingly parallel
+problems, such as fitting and evaluating multiple models in a grid search
+procedure. The workaround that we use is to rely on the Python multiprocessing
+module, which emulates threads using multiple operating system- level processes.
+However this solution is not perfect as it may in turn lead to a large memory
+overhead since objects (e.g., input datasets) need to be copied within each
+process own memory space. In some cases, it may unfortunately even happen that
+the overhead becomes so significant that fully exploiting a multi- core
+architecture becomes infeasible due memory constraints. As such, a proper
+implementation of parallel machine learning algorithms remain a challenging task
+for the project.
 
@@
 At a lower-level, a potential point of improvement is parallel processing,
-which, as described in Section~\ref{sec:parallel}, is only implemented at a
+which, as described in Section~\ref{sec:implementation}, is only implemented at a
 rather coarse-grained level. Cython built-in support for OpenMP
@@
 grained multicore support in \sklearn.
+
+\subsection*{Acknowledgements}
+
+\textcolor{red}{Mention BSD license, INRIA and stuff.}
 

### 2013-06-16 paper.tex
wording

@@
 rely on other external libraries, but in such cases we tend to prefer to
-directly ship them within the codebase rather than adding new dependencies. In
-particular, the codebase includes modified versions of the \textsf{LIBSVM} and
-\textsf{LIBLINEAR} libraries \citep{chang2011libsvm,fan2008}, both written in
-C{}\verb!++! and wrapped using Cython modules.
+directly ship them within the codebase rather than adding new dependencies for
+the user to install. In particular, the codebase includes modified versions of
+the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
+\citep{chang2011libsvm,fan2008}, both written in C{}\verb!++! and wrapped using
+Cython modules.
 
-Finally, some algorithms in \sklearn also support multi-core architectures.
+Finally, some estimators in \sklearn also support multi-core architectures.
 Machine learning algorithms can indeed often be decomposed into independent

### 2013-06-16 paper.tex
spellings

@@
 and scientific packages centered around the NumPy and SciPy libraries. NumPy
-\citep{vanderwalt2011} augments Python with a numeric array datatype and proper
+\citep{vanderwalt2011} augments Python with a numeric array data type and proper
 array computing primitives while SciPy \citep{varoquaux2013scipy} extends it
@@
        Learning algorithms are the only objects to be of custom classes.
-       Datasets and hyperparameter names and values are represented as standard
+       Datasets and hyper-parameter names and values are represented as standard
        Python, NumPy or SciPy classes rather than new ones whenever feasible.
@@
        Some learning algorithms are also naturally viewed
-       as meta-algorithms parameterized on other algorithms.
+       as meta-algorithms parametrized on other algorithms.
        Whenever possible, those algorithms are implemented and composed from
@@
 extends Python with static typing. It produces efficient C extension modules
-that are directly importable from the Python runtime system. Examples of
+that are directly importable from the Python run-time system. Examples of
 algorithms written in Cython include stochastic gradient descent for linear
@@
 as argument, while possible, degrades performance significantly due to the
-runtime checks needed for each function invocation. This is especially
+run-time checks needed for each function invocation. This is especially
 problematic for example for kernel methods, which would benefit from accepting a
 kernel function as argument. A potential solution to this problem would be for
-\sklearn to provide a Cython-level API, thereby alllowing the user to pass
+\sklearn to provide a Cython-level API, thereby allowing the user to pass
 Cython-implemented kernel functions.
@@
 process own memory space. In some cases, it may unfortunately even happen that
-the overhead becomes so significant that fully exploiting a multi- core
+the overhead becomes so significant that fully exploiting a multi-core
 architecture becomes infeasible due memory constraints. As such, a proper

### 2013-06-16 paper.tex
fix imports in code sample and explain feature selection method

@@
 computing both linear PCA and kernel PCA components on \texttt{X\_train}
-(through a \texttt{FeatureUnion}), selecting the 100 best of those features and
-then feeding them to an $\ell_2$-regularized logistic regression model:
+(through a \texttt{FeatureUnion}),
+selecting the 100 best of those features according to an ANOVA test
+and then feeding them to an $\ell_2$-regularized logistic regression model:
 \begin{pythoncode}
-from sklearn.pipeline import Pipeline
+from sklearn.pipeline import FeatureUnion, Pipeline
 from sklearn.decomposition import PCA, KernelPCA
 from sklearn.feature_selection import SelectKBest
+from sklearn.linear_model import LogisticRegression
 

### 2013-06-16 paper.tex
section on extending scikit-learn
Emphasis is on avoiding frameworky code.
Feel free to excise the witty quote if you think it's not appropriate;
I just had to share it.

@@
 
-To ease code reuse, object-oriented techniques and design patterns are also used
-all throughout the codebase. In particular, the \textit{duck typing} design
-pattern is widely exploited to simplify implementation and skip the introduction
-of superfluous classes. This technique indeed allows both for extensibility and
-flexibility: user-defined estimators that follow our API and conventions should
-be reusable (e.g., in a pipeline) as any built-in estimators, without them
-actually inheriting from our base classes.
-
 \subsection{Pipelines and feature unions}
@@
 \texttt{score} methods to the best estimator, thus mimicking its behavior.
+
+\subsection{Extending \sklearn}
+
+Any object that conforms to a small superset of the API
+lined out in Section~\ref{sec:core-api},
+(in particular, that implements methods \texttt{get\_params}
+and \texttt{set\_params} for setting hyperparameters;
+refer to the documentation at \url{http://scikit-learn.org} for details)
+can be used as a \sklearn estimator
+and be ``plugged into'' a pipeline, grid search or other meta-estimator
+To define such objects, a user can either derive a new class
+from \texttt{sklearn.base.BaseEstimator},
+or define the entire class from scratch,
+without even an \texttt{import sklearn}
+in the relevant program module.
+The Python principle of ``duck typing''
+is employed to make sure that even instances of a class
+that does not derive from any \sklearn class can be a first-class estimator.
+
+In other places of the library,
+in particular in the feature extraction and vectorization code
+for text documents,
+the toolkit is designed to be extensible.
+Here, the estimators provide ``hooks'' for user-defined code:
+objects or functions that provide a specific API
+can be given as arguments at vectorizer construction time.
+\sklearn code then calls into this code
+to obtain the desired behavior.
+Again, such user code can be kept agnostic
+of the \sklearn class hierarchy,
+operating on standard Python objects such as strings instead.
+
+The design principle here is that user code
+should not be tied to \sklearn,
+which is a \textit{library}, not a \textit{framework}.
+This principle avoids a variant of the well-known problem
+with object-oriented design,
+which is that users wanting a ``banana'' get
+``a gorilla holding the banana and the entire jungle''
+(Joe~Armstrong, interviewed by \citealp{seibel2009coders}:213):
+a program that is not reusable with another machine learning toolkit
+(or other solution to the problem at hand),
+because it is intimately tied to \sklearn.
 

### 2013-06-17 paper.tex
wording

@@
 
-% \subsection{Cython for performance}
+Our implementation guidelines emphasize on writing efficient but readable code.
+In particular, we focus on making the codebase easily maintainable and
+understandable in order to favour external contributions. Whenever practicable,
+algorithms implemented in \sklearn are thus directly written in pure Python. For
+intensive numerical computations though, the library mostly relies on NumPy
+vector operations, which allows for the code to remain concise, readable and
+efficient. For all algorithms that cannot be easily and efficiently expressed as
+NumPy operations, we rely on Cython \citep{behnel2011cython}  for writing the
+critical parts and achieving decent performance and scalability. Cython is a
+compiled programming language that extends Python with static typing. It
+produces efficient C extension modules that are directly importable from the
+Python run-time system. Examples of algorithms written in Cython include
+stochastic gradient descent for linear models, some graph-based clustering
+algorithms or decision trees.  A problem arising from the use of Cython,
+however, is that passing pure Python functions as argument, while possible,
+degrades performance significantly due to the run-time checks needed for each
+function invocation. This is especially problematic for example for kernel
+methods, which would benefit from accepting a kernel function as argument. A
+potential solution to this problem would be for \sklearn to provide a Cython-
+level API, thereby allowing the user to pass Cython-implemented kernel
+functions.
 
-Our implementation guidelines emphasize on writing short and readable code,
-easily maintainable by external contributors, while being as efficient as
-possible.
-
-Whenever practicable, algorithms implemented in \sklearn are thus
-directly written in pure Python. For intensive numerical computations though,
-the library mostly relies on NumPy vector operations, which allows for the code
-to remain concise, readable and efficient. For all algorithms that cannot be
-easily and efficiently expressed as NumPy operations, we rely on Cython
-\citep{behnel2011cython}  for writing the critical parts and achieving decent
-performance and scalability. Cython is a compiled programming language that
-extends Python with static typing. It produces efficient C extension modules
-that are directly importable from the Python run-time system. Examples of
-algorithms written in Cython include stochastic gradient descent for linear
-models, some graph-based clustering algorithms or decision trees.  A problem
-arising from the use of Cython, however, is that passing pure Python functions
-as argument, while possible, degrades performance significantly due to the
-run-time checks needed for each function invocation. This is especially
-problematic for example for kernel methods, which would benefit from accepting a
-kernel function as argument. A potential solution to this problem would be for
-\sklearn to provide a Cython-level API, thereby allowing the user to pass
-Cython-implemented kernel functions.
-
-To facilitate the adoption of the library, we also try to limit to a bare
-minimum the set of external dependencies. As such, NumPy and SciPy -- the two
-most common dependencies in the Python scientific ecosystem -- are actually the
-only two required dependencies for the library to fully work. Some modules may
-rely on other external libraries, but in such cases we tend to prefer to
-directly ship them within the codebase rather than adding new dependencies for
-the user to install. In particular, the codebase includes modified versions of
-the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
+To facilitate the adoption and the distribution of the library, we also try to
+limit to a bare minimum the set of external dependencies. As such, NumPy and
+SciPy -- the two most common dependencies in the Python scientific ecosystem --
+are actually the only two required dependencies for the library to fully work.
+Some modules may rely on other external libraries, but in such cases we tend to
+prefer to directly ship them within the codebase rather than adding new
+dependencies for the user to install. In particular, the codebase includes
+modified versions of the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
 \citep{chang2011libsvm,fan2008}, both written in C{}\verb!++! and wrapped using

### 2013-06-17 paper.tex
acknowledgements

@@
 
-\textcolor{red}{Mention BSD license, INRIA and stuff.}
+The project is distributed under the simplified BSD license, encouraging
+its use in both academic and commercial settings.
+
+The authors and contributors acknowledge active support from INRIA. Past and
+present sponsors of the project also include Google -- for funding code
+scholarships (\textit{Google Summer of Code}) --, the Python Software Fundation and
+Tinyclues -- for funding coding sprints.
 

### 2013-06-17 paper.tex
move license to intro
extending scikit-learn

@@
 purpose high-level language. Among other things, it includes classical learning
-algorithms, model selection tools and preprocessing procedures.
+algorithms, model selection tools and preprocessing procedures. The project is
+distributed under the simplified BSD license, encouraging its use in both
+academic and commercial settings.
 
@@
 
-\textcolor{red}{TODO}
-
-\textcolor{red}{Point out that any external estimator can be used as base
-estimator thanks to duck typing}\\
+\textcolor{red}{TODO: Point out that any external estimator can be used as base
+estimator thanks to duck typing}
 
@@
 
-Any object that conforms to a small superset of the API
-lined out in Section~\ref{sec:core-api},
-(in particular, that implements methods \texttt{get\_params}
-and \texttt{set\_params} for setting hyperparameters;
-refer to the documentation at \url{http://scikit-learn.org} for details)
-can be used as a \sklearn estimator
-and be ``plugged into'' a pipeline, grid search or other meta-estimator
-To define such objects, a user can either derive a new class
-from \texttt{sklearn.base.BaseEstimator},
-or define the entire class from scratch,
-without even an \texttt{import sklearn}
-in the relevant program module.
-The Python principle of ``duck typing''
-is employed to make sure that even instances of a class
-that does not derive from any \sklearn class can be a first-class estimator.
+To ease code reuse, simplify implementation and skip the introduction of
+superfluous classes, the principle of \textit{duck typing} is exploited all
+throughout the codebase. This technique allows both for extensibility and
+flexibility: as long as an estimator follows our API and conventions (as
+outlined in Section~\ref{sec:core-api}), then it should be as reusable as any
+built-in estimator (e.g., it can plugged within pipelines or grid search
+estimators). That is, estimators are specified only by their interface, and not
+by inheritance. In particular, this means that external developers can define
+their own estimators without being forced to inherit from any internal class of
+\sklearn, and rest assured that they can be used as first-class estimators.
 
-In other places of the library,
-in particular in the feature extraction and vectorization code
-for text documents,
-the toolkit is designed to be extensible.
-Here, the estimators provide ``hooks'' for user-defined code:
-objects or functions that provide a specific API
-can be given as arguments at vectorizer construction time.
-\sklearn code then calls into this code
-to obtain the desired behavior.
-Again, such user code can be kept agnostic
-of the \sklearn class hierarchy,
-operating on standard Python objects such as strings instead.
+In other places of the library, in particular in the feature extraction and
+vectorization code for text documents, the toolkit is also designed to be
+extensible. Here, estimators provide hooks for user-defined code: objects or
+functions that provide a specific API can be given as arguments at vectorizer
+construction time. The library then calls into this code to obtain the desired
+behavior. Again, such external user code can be kept agnostic of the \sklearn
+class hierarchy and operates, for example, on standard Python objects such as
+strings instead.
 
-The design principle here is that user code
-should not be tied to \sklearn,
-which is a \textit{library}, not a \textit{framework}.
-This principle avoids a variant of the well-known problem
-with object-oriented design,
-which is that users wanting a ``banana'' get
-``a gorilla holding the banana and the entire jungle''
-(Joe~Armstrong, interviewed by \citealp{seibel2009coders}:213):
-a program that is not reusable with another machine learning toolkit
-(or other solution to the problem at hand),
-because it is intimately tied to \sklearn.
+Our rule of thumb is that user code should not be tied to \sklearn -- which is a
+\textit{library}, and not a \textit{framework}. This principle indeed avoids a
+well-known problem with object-oriented design, which is that users wanting a
+``banana'' should not get ``a gorilla holding the banana and the entire jungle''
+(Joe~Armstrong, interviewed by \citealp{seibel2009coders}). That is, programs
+using \sklearn should be easily reusable with other machine learning toolkits or
+in other contexts, because they don't need to be intimately tied to it.
 
@@
 
-The project is distributed under the simplified BSD license, encouraging
-its use in both academic and commercial settings.
-
 The authors and contributors acknowledge active support from INRIA. Past and

### 2013-06-17 paper.tex
wording
typo

@@
 advanced API mechanisms built from the core interface. Section
-\ref{sec:implementation} describes implementation trade-offs and discusses
-limitations of the technologies we rely on. Section \ref{sec:comparison}
-then compares \emph{scikit-learn} to other major projects in terms of API.
-Finally, conclusions and directions of future works are included in section
+\ref{sec:implementation} describes implementation principles and discusses
+limitations of the technologies we rely on. Section \ref{sec:comparison} then
+compares \emph{scikit-learn} to other major projects in terms of API. Finally,
+conclusions and directions of future works are included in section
 \ref{sec:conclusions}.
@@
 methods, which would benefit from accepting a kernel function as argument. A
-potential solution to this problem would be for \sklearn to provide a Cython-
-level API, thereby allowing the user to pass Cython-implemented kernel
+potential solution to this problem would be for \sklearn to provide a Cython-level
+API, thereby allowing the user to pass Cython-implemented kernel
 functions.

### 2013-06-17 paper.tex
shorter title
TODO still just over one line, must use smaller title font
Capitalize first letter (as we do in the JMLR paper)
remove \textit in subtitles
title

@@
 % working title, please change
-\title{Python as a machine learning language:
-       experiences from the scikit-learn project}
+\title{Scikit-learn: API design for machine learning}
 
@@
 \begin{abstract}
-The \sklearn project is an increasingly popular open source machine
-learning library designed to be simple and efficient, accessible to non-experts,
-and reusable in various contexts. In this paper, we present and discuss our
-design choices for the application programming interface (API) of the project.
-In particular, we describe the simple and elegant interface shared by all
-learning and processing units in the library and then discuss its advantages in
-terms of composition and reusability. The paper also comments on implementation
-details specific to the Python ecosystem and analyzes obstacles faced by users
-and developers of the library.
+The \sklearn project is an increasingly popular open source machine learning
+library written in Python and designed to be simple and efficient, accessible to
+non-experts, and reusable in various contexts. In this paper, we present and
+discuss our design choices for the application programming interface (API) of
+the project. In particular, we describe the simple and elegant interface shared
+by all learning and processing units in the library and then discuss its
+advantages in terms of composition and reusability. The paper also comments on
+implementation details specific to the Python ecosystem and analyzes obstacles
+faced by users and developers of the library.
 \end{abstract}
@@
 
-\subsection{Extending \sklearn}
+\subsection{Extending scikit-learn}
 

### 2013-06-17 paper.tex
external estimators covered, remove todo
partial application and meta-estimators

@@
 behavior of the estimator.
+Initializing and fitting an estimator object
+proceeds in a way that is similar to partial function application:
+an estimator can be considered a function
+that maps hyperparameter settings to an actual learning algorithm.
+The advantage of this API
+over one where the object initializer would perform the learning
+will become especially apparent in Section~\ref{sec:advanced-api}.
 
@@
 learned coefficients of a linear model).
+The return value of \texttt{fit} is always the object it was called on,
+which now serves as a model of its input
+and can be used to perform predictions or transformations of input data.
 
@@
 
-Finally, let us also note that the \texttt{fit} method always returns the estimator
-it was called on, hence making method chaining possible.  Using this technique,
-we can thus rewrite the \texttt{StandardScaler} example above in a single line:
+Additionally, by leveraging that fact that \texttt{fit}
+always returns the estimator it was called on,
+the \texttt{StandardScaler} example above
+can be rewritten in a single line using a ``method chaining'' style:
 \begin{pythoncode}
@@
 \end{pythoncode}
-In addition, \texttt{fit(X\_train).transform(X\_train)} can further
+Finally, \texttt{fit(X\_train).transform(X\_train)} can further
 be rewritten as \texttt{fit\_transform(X\_train)}.  Combining the \texttt{fit}
@@
 allows to share computations, which often makes a huge difference in practice.
-In the same spirit, \sklearn sometimes provides the
-\texttt{fit\_predict} method for combining the \texttt{fit} and \texttt{predict}
-methods. This is especially useful for retrieving cluster predictions with
-respect to the training data \texttt{X\_train} in clustering algorithms.
+In the same spirit, clustering algorithms provide a
+\texttt{fit\_predict} method
+that is equivalent to \texttt{fit} followed by \texttt{predict},
+and thus returns the cluster labels assigned to the training data samples.
 
@@
 
-\textcolor{red}{TODO: Point out that any external estimator can be used as base
-estimator thanks to duck typing}
+Some machine learning algorithms are most naturally expressed
+as modification or extensions of other, simpler algorithms.
+Examples include ensemble estimators
+and various multiclass and multilabel classification schemes.
+Such algorithms are implemented as so-called ``meta-estimators'',
+that take an estimator as input and use that to perform learning and prediction.
+
+As an example, a logistic regression classifier
+by default uses a one-vs.-rest (aka one-vs.-all) scheme internally
+for performing multiclass classification.
+If multiclass classification based on error-correcting codes
+\citep{dietterich1995solving} is desired instead,
+this can achieved with an \texttt{OutputCodeClassifier}
+that wraps a \texttt{LogisticRegression}:
+\begin{pythoncode}
+from sklearn.linear_model import LogisticRegression
+from sklearn.multiclass import OutputCodeClassifier
+
+ecoc_lr = OutputCodeClassifier(LogisticRegression(penalty="l1"))
+\end{pythoncode}
+
+% TODO: show a grid search on this estimator
 

### 2013-06-17 paper.tex
wordings

@@
 behavior of the estimator.
-Initializing and fitting an estimator object
-proceeds in a way that is similar to partial function application:
-an estimator can be considered a function
-that maps hyperparameter settings to an actual learning algorithm.
-The advantage of this API
-over one where the object initializer would perform the learning
-will become especially apparent in Section~\ref{sec:advanced-api}.
+Initializing and fitting an estimator object proceeds in a way that is similar
+to partial function application: an estimator can be considered as a function
+that maps hyper-parameter settings to an actual learning algorithm. The
+advantage of this API over one where the object initializer would perform the
+learning will become especially apparent in Section~\ref{sec:advanced-api}.
 
@@
 
-Additionally, by leveraging that fact that \texttt{fit}
-always returns the estimator it was called on,
-the \texttt{StandardScaler} example above
-can be rewritten in a single line using a ``method chaining'' style:
+Additionally, by leveraging the fact that \texttt{fit} always returns the
+estimator it was called on, the \texttt{StandardScaler} example above can be
+rewritten in a single line using method chaining:
 \begin{pythoncode}

### 2013-06-17 paper.tex
move parallel processing from implementation to future work
Also trimmed it quite a bit. Left out the memory overhead argument
because it isn't entirely fair; this is a by-product of parallel
processing in general, as well as our implementation. It is aggravated
by the environment, not caused by it.

@@
 
-Finally, some estimators in \sklearn also support multi-core architectures.
-Machine learning algorithms can indeed often be decomposed into independent
-smaller tasks that may each be solved in parallel, thereby reducing the overall
-computing times. At this point in time, support for multi-core programming in
-Python -- at least in the CPython implementation targeted by scikit-learn -- is
-however suboptimal. The interpreter supports multithreading, but due to the
-\textit{global interpreter lock}, no two threads in a process may execute Python
-code at the same time. It follows that multithreading is not a viable solution
-for parallelizing parts of an algorithm, even for embarrassingly parallel
-problems, such as fitting and evaluating multiple models in a grid search
-procedure. The workaround that we use is to rely on the Python multiprocessing
-module, which emulates threads using multiple operating system- level processes.
-However this solution is not perfect as it may in turn lead to a large memory
-overhead since objects (e.g., input datasets) need to be copied within each
-process own memory space. In some cases, it may unfortunately even happen that
-the overhead becomes so significant that fully exploiting a multi-core
-architecture becomes infeasible due memory constraints. As such, a proper
-implementation of parallel machine learning algorithms remain a challenging task
-for the project.
-
 \section{Comparison with other packages}
@@
 
-\section{Conclusions and future works}
+\section{Conclusions and future work}
 \label{sec:conclusions}
@@
 
-At a lower-level, a potential point of improvement is parallel processing,
-which, as described in Section~\ref{sec:implementation}, is only implemented at a
-rather coarse-grained level. Cython built-in support for OpenMP
-\citep{dagum1998openmp} remains a viable candidate technology for more fine-
-grained multicore support in \sklearn.
+At a lower-level, parallel processing is a potential point of improvement.
+Some estimators in \sklearn are already able to leverage multi-core processors,
+but only in a coarse-grained fashion.
+At present, parallel processing is difficult to accomplish in the Python environment;
+\sklearn targets the main implementation, CPython,
+which cannot execute Python code on multiple CPUs simultaneously.
+It follows that any parallel task decomposition must either be done
+inside Cython modules,
+or at a level high enough to warrant the overhead
+of creating multiple OS-level processes,
+and the ensuing inter-process communication.
+Parallel grid search is an example of the latter approach
+which has already been implemented.
+Recent versions of Cython include support for the OpenMP standard
+\citep{dagum1998openmp},
+which is a viable candidate technology
+for more fine-grained multicore support in \sklearn.
 

### 2013-06-18 paper.tex
Cosmit.

@@
 
-As the above snippet shows, \texttt{Pipeline} and \texttt{FeatureUnion} can be
-combined to create complex objects. In particular, since both are themselves
-estimators, they can be used to form nested workflows.
+As the above example shows, \texttt{Pipeline} and \texttt{FeatureUnion} can be
+combined to create complex and nested workflows.
 

### 2013-06-18 paper.tex
improve partial application stuff

@@
 
-The constructor of an estimator object takes as input a set of named constant
-hyper-parameters. At initialization, all given parameters are attached to the
-estimator instance as public attributes and serve to determine the estimator
-behavior (e.g., the $C$ constant in SVMs). Parameters set as public attributes
-are also used for introspection purposes, e.g., in model selection routines. To
-ease their use, default hyper-parameter values are provided for all built-in
-estimators. These default values are set to be relevant in most common
-situations in order to make estimators as efficient as possible
-\textit{out-of-box} for non-experts. Finally, the constructor of an estimator
-never performs any actual learning (this is the role of the \texttt{fit} method,
-see below).  Its only goal is to set the parameter values and freeze the
-behavior of the estimator.
-Initializing and fitting an estimator object proceeds in a way that is similar
-to partial function application: an estimator can be considered as a function
-that maps hyper-parameter settings to an actual learning algorithm. The
-advantage of this API over one where the object initializer would perform the
-learning will become especially apparent in Section~\ref{sec:advanced-api}.
+Initializing of estimator objects and actual learning are strictly separated,
+in a way that is similar to partial function application:
+an estimator is initialized from a set of hyper-parameter settings
+(e.g., regularization parameters, random seeds)
+and can be considered a function
+that maps these settings to actual learning algorithms.
+The object initializer does not see any actual data.
+All it does is attach the given parameters to the object.
+Hyper-parameter settings are public attributes
+for the sake of convenient model inspection.
 
-The estimator interface defines a \texttt{fit} method. This method is called
+Actual learning is performed by the \texttt{fit} method. This method is called
 with training data (e.g., supplied as two arrays \texttt{X\_train} and
 \texttt{y\_train} in supervised learning estimators). Its task is to run a
-learning algorithm and to determine model- specific parameters from the training
+learning algorithm and to determine model-specific parameters from the training
 data and set these as attributes on the estimator object. As a convention, the
@@
 is suffixed with a trailing underscore (e.g., \texttt{coef\_} for the
-learned coefficients of a linear model).
-The return value of \texttt{fit} is always the object it was called on,
-which now serves as a model of its input
-and can be used to perform predictions or transformations of input data.
+learned coefficients of a linear model),
+again to facilitate model inspection.
+In the partial application view,
+\texttt{fit} is a function from data to a model of that data,
+the return value always being the object it was called on,
+which can be used to perform predictions or transformations of input data.
+The choice to let a single object serve dual purpose as estimator and model
+simplifies the \sklearn class hierarchy
+compared to an approach were estimators and models are separated.
 
-As an example, let us consider a supervised learning task using logistic regression.
+As an example to illustrate the initialize-\texttt{fit} sequence,
+let us consider a supervised learning task using logistic regression.
 Given the API defined above, solving this problem is as simple as:
@@
 thus set to the default values. Upon calling \texttt{fit}, the model is then
-learned from the training arrays \texttt{X\_train} and \texttt{y\_train}. Since
-all estimators share the same interface, using a different learning algorithm is
-as simple as replacing the constructor. For example, to build a Random Forest on
+learned from the training arrays \texttt{X\_train} and \texttt{y\_train}.
+(In this example, \texttt{fit} is called solely for its side effects.)
+Since all estimators share the same interface, using a different learning algorithm is
+as simple as replacing the constructor (the class name).
+For example, to build a random forest on
 the same training data, one would simply have to replace

### 2013-06-18 paper.tex
copyedit

@@
 
-The \textit{predictor} interface defines a \texttt{predict}
-method taking as input some new data \texttt{X\_test} and producing as output
+The \textit{predictor} interface extends the notion of an estimator
+by adding a \texttt{predict}
+method that takes an array \texttt{X\_test} and produces
 predictions for \texttt{X\_test}, based on the learned parameters of the
@@
 error in the case of regression. Not however that, from an API point of view,
-the only guideline is that the \texttt{score} method returns a numerical value
-such that the higher the better. As such, some unsupervised learning estimators
-also exposes a \texttt{score} function, computing for instance the likelihood of
-the given data.
+the only guideline is that the \texttt{score} method returns a value
+that quantifies the quality of its predictions (the higher, the better).
+Unsupervised learning estimators may also exposes a \texttt{score} function
+to compute, for instance, the likelihood of the given data under its model.
 

### 2013-06-18 paper.tex
put some stuff back in that I deleted by mistake

@@
 Hyper-parameter settings are public attributes
-for the sake of convenient model inspection.
+for the sake of convenient model inspection,
+which is especially important in model selection settings.
+For ease of use, default hyper-parameter values are provided
+for all built-in estimators.
+These default values are set to be relevant in most common
+situations in order to make estimators as effective as possible
+\textit{out-of-box} for non-experts.
 

### 2013-06-19 paper.tex
wordings

@@
 
-Initializing of estimator objects and actual learning are strictly separated,
+Estimator initialization and actual learning are strictly separated,
 in a way that is similar to partial function application:
-an estimator is initialized from a set of hyper-parameter settings
-(e.g., regularization parameters, random seeds)
-and can be considered a function
-that maps these settings to actual learning algorithms.
-The object initializer does not see any actual data.
+an estimator is initialized from a set of named constant hyper-parameter values
+(e.g., the $C$ constant in SVMs)
+and can be considered as a function
+that maps these values to actual learning algorithms.
+The constructor of an estimator does not see any actual data, nor does it perform any actual learning.
 All it does is attach the given parameters to the object.
-Hyper-parameter settings are public attributes
-for the sake of convenient model inspection,
+For the sake of convenient model inspection, hyper-parameters are set as public attributes,
 which is especially important in model selection settings.
-For ease of use, default hyper-parameter values are provided
+For ease of use, default hyper-parameter values are also provided
 for all built-in estimators.
@@
 In the partial application view,
-\texttt{fit} is a function from data to a model of that data,
-the return value always being the object it was called on,
-which can be used to perform predictions or transformations of input data.
+\texttt{fit} is a function from data to a model of that data.
+As such, it always returns the estimator object it was called on,
+which now serves as a model of its input and can be used to perform predictions or transformations of input data.
 The choice to let a single object serve dual purpose as estimator and model
@@
 \end{pythoncode}
-Finally, \texttt{fit(X\_train).transform(X\_train)} can further
+In this particular example, \texttt{fit(X\_train).transform(X\_train)} can also further
 be rewritten as \texttt{fit\_transform(X\_train)}.  Combining the \texttt{fit}
-and \texttt{transform} methods into a single method \texttt{fit\_transform}
-allows to share computations, which often makes a huge difference in practice.
+and \texttt{transform} methods into a single \texttt{fit\_transform} method
+indeed allows to share computations, which often makes a huge difference in practice.
 In the same spirit, clustering algorithms provide a

### 2013-06-19 paper.tex
meta-estimators

@@
 
-Some machine learning algorithms are most naturally expressed
-as modification or extensions of other, simpler algorithms.
-Examples include ensemble estimators
-and various multiclass and multilabel classification schemes.
-Such algorithms are implemented as so-called ``meta-estimators'',
-that take an estimator as input and use that to perform learning and prediction.
+In machine learning, some algorithms are expressed as meta-algorithms built
+and parametrized on simpler algorithms. Examples include ensemble methods which
+build and combine several simpler models (e.g., decision trees), or multiclass
+and multilabel classification schemes which can be used to turn a binary
+classifier into a multiclass or multilabel classifier. In \sklearn, such algorithms are
+implemented as \textit{meta-estimators}. They take as input an existing base
+estimator and use it internally for learning and making predictions.
 
 As an example, a logistic regression classifier
-by default uses a one-vs.-rest (aka one-vs.-all) scheme internally
+ uses by default a one-vs.-rest  scheme
 for performing multiclass classification.
@@
 \citep{dietterich1995solving} is desired instead,
-this can achieved with an \texttt{OutputCodeClassifier}
-that wraps a \texttt{LogisticRegression}:
+then this can achieved  using an \texttt{OutputCodeClassifier} meta-estimator
+wrapping a \texttt{LogisticRegression} estimator:
 \begin{pythoncode}

### 2013-06-19 paper.tex
copyedit FeatureUnion description
Select only 10 features, which is more likely to actually work
when PCA and KernelPCA get to determine their numbers of components
automatically.

@@
 A \texttt{FeatureUnion} object is a transformer combining multiple transformers
-into a single one that joins their outputs. From an API point of view, a
+into a single one that concatenates their outputs,
+in the sense that a union of transformers that map inputs having $d$ features
+to $d'$ and $d''$ features, respectively,
+is a transformer that maps its $d$ input features to $d' + d''$ features
+by calling its underlying transformers in parallel.
+From an API point of view, a
 \texttt{FeatureUnion} takes as input a list of transformers. Calling
 \texttt{fit} on the union is the same as calling \texttt{fit} independently on
-each of the transformers. For transforming data, the base transformers are
-applied in parallel and their outputs are concatenated end-to-end into a larger
-output.
+each of the transformers.
 
@@
 (through a \texttt{FeatureUnion}),
-selecting the 100 best of those features according to an ANOVA test
+selecting the 10 best of those features according to an ANOVA test
 and then feeding them to an $\ell_2$-regularized logistic regression model:
@@
 Pipeline([("feat_union", fu),
-          ("feat_sel", SelectKBest(k=100)),
+          ("feat_sel", SelectKBest(k=10)),
           ("log_reg", LogisticRegression(penalty="l2"))
@@
 \end{pythoncode}
-
 As the above example shows, \texttt{Pipeline} and \texttt{FeatureUnion} can be

### 2013-06-19 paper.tex
rewrite meta-estimator part, no longer a subsection

@@
 
-Building on the core interface introduced in the previous section, we now
-present in this section advanced API mechanisms for building meta-estimators,
-composing complex estimators and then selecting models.
-
-\subsection{Meta-estimators}
-
-In machine learning, some algorithms are expressed as meta-algorithms built
-and parametrized on simpler algorithms. Examples include ensemble methods which
+Some machine learning algorithms are expressed naturally
+as meta-algorithms parametrized on simpler algorithms.
+Examples include ensemble methods which
 build and combine several simpler models (e.g., decision trees), or multiclass
@@
 estimator and use it internally for learning and making predictions.
+All meta-estimators implement the regular estimator interface.
 
@@
 \end{pythoncode}
+
+Since using meta-estimators requires the construction of nested objects,
+the decision to implement a meta-estimator
+rather than fold the behavior it implements
+into existing estimators classes
+is always based on a trade-off between genericity and ease of use.
+Relating to the example just given,
+all \sklearn classifiers are designed to do multiclass classification
+and the use of the \texttt{multiclass} module
+is only necessary in advanced use cases.
 

### 2013-06-19 paper.tex
rm Cython problems

@@
 stochastic gradient descent for linear models, some graph-based clustering
-algorithms or decision trees.  A problem arising from the use of Cython,
-however, is that passing pure Python functions as argument, while possible,
-degrades performance significantly due to the run-time checks needed for each
-function invocation. This is especially problematic for example for kernel
-methods, which would benefit from accepting a kernel function as argument. A
-potential solution to this problem would be for \sklearn to provide a Cython-level
-API, thereby allowing the user to pass Cython-implemented kernel
-functions.
+algorithms and decision trees.
 

### 2013-06-19 paper.tex
rewrite requirements section

@@
 
-To facilitate the adoption and the distribution of the library, we also try to
-limit to a bare minimum the set of external dependencies. As such, NumPy and
-SciPy -- the two most common dependencies in the Python scientific ecosystem --
-are actually the only two required dependencies for the library to fully work.
-Some modules may rely on other external libraries, but in such cases we tend to
-prefer to directly ship them within the codebase rather than adding new
-dependencies for the user to install. In particular, the codebase includes
-modified versions of the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
+To facilitate the adoption and the distribution of \sklearn,
+the set of external dependencies is kept to a bare minimum:
+only Python, NumPy and SciPy are required for a functioning installation.
+Binary distributions of these are available for the major platforms.
+Visualisation functionality depends on Matplotlib \citep{hunter2007matplotlib}
+and/or Graphviz \citep{gansner2000},
+but neither is required to do machine learning.
+When feasible, external libraries are integrated into the codebase.
+In particular, the codebase includes modified versions
+of \textsf{LIBSVM} and \textsf{LIBLINEAR}
 \citep{chang2011libsvm,fan2008}, both written in C{}\verb!++! and wrapped using

### 2013-06-19 paper.tex
requirements: shorten sentence
condense "Extending" section

@@
 To ease code reuse, simplify implementation and skip the introduction of
-superfluous classes, the principle of \textit{duck typing} is exploited all
+superfluous classes, the principle of \textit{duck typing} is exploited
 throughout the codebase. This technique allows both for extensibility and
-flexibility: as long as an estimator follows our API and conventions (as
-outlined in Section~\ref{sec:core-api}), then it should be as reusable as any
-built-in estimator (e.g., it can plugged within pipelines or grid search
-estimators). That is, estimators are specified only by their interface, and not
-by inheritance. In particular, this means that external developers can define
-their own estimators without being forced to inherit from any internal class of
-\sklearn, and rest assured that they can be used as first-class estimators.
+flexibility: as long as an estimator follows the API and conventions
+outlined in Section~\ref{sec:core-api}, then it can be used in lieu of a
+built-in estimator (e.g., it can plugged into pipelines or grid search).
+That is, estimators are defined only by interface, not by inheritance,
+and external developers are not forced to inherit from any \sklearn class.
 
@@
 behavior. Again, such external user code can be kept agnostic of the \sklearn
-class hierarchy and operates, for example, on standard Python objects such as
-strings instead.
+class hierarchy.
+Communication proceeds by passing object of standard Python/NumPy types.
 
@@
 When feasible, external libraries are integrated into the codebase.
-In particular, the codebase includes modified versions
-of \textsf{LIBSVM} and \textsf{LIBLINEAR}
+In particular, modified versions of \textsf{LIBSVM} and \textsf{LIBLINEAR}
 \citep{chang2011libsvm,fan2008}, both written in C{}\verb!++! and wrapped using

### 2013-06-19 paper.tex
"at the time" -> "at present" and "many" -> "some" in missing algos
break long sentence about FeatureUnion and move some stuff around

@@
 A \texttt{FeatureUnion} object is a transformer combining multiple transformers
-into a single one that concatenates their outputs,
-in the sense that a union of transformers that map inputs having $d$ features
+into a single one that concatenates their outputs.
+I.e., a union of transformers that map inputs having $d$ features
 to $d'$ and $d''$ features, respectively,
@@
 
+\texttt{Pipeline} and \texttt{FeatureUnion} can be
+combined to create complex and nested workflows.
 As an example, the following snippet illustrates how to create a complex worflow
@@
 \end{pythoncode}
-As the above example shows, \texttt{Pipeline} and \texttt{FeatureUnion} can be
-combined to create complex and nested workflows.
 
@@
 Although it is coming close to a 1.0 release, \sklearn is by no means finished.
-At the time, it does not support many classical machine learning tasks. Among
+At present, it does not support some classical machine learning tasks. Among
 others, missing features include neural networks, ensemble meta-estimators for

### 2013-06-19 paper.tex
missing verb (thanks Gilles)
Compress bibliography to a single page
If necessary, a sectionbib argument to natbib saves an additional
one or two lines.
page number for gorilla quote

@@
 ``banana'' should not get ``a gorilla holding the banana and the entire jungle''
-(Joe~Armstrong, interviewed by \citealp{seibel2009coders}). That is, programs
+(J.~Armstrong, cited by \citealp[p.~213]{seibel2009coders}).  That is, programs
 using \sklearn should be easily reusable with other machine learning toolkits or
@@
 In particular, modified versions of \textsf{LIBSVM} and \textsf{LIBLINEAR}
-\citep{chang2011libsvm,fan2008}, both written in C{}\verb!++! and wrapped using
-Cython modules.
+\citep{chang2011libsvm,fan2008}, both written in C{}\verb!++!,
+and wrapped using Cython modules, are shipped with \sklearn.
 
@@
 
-\bibliographystyle{plainnat}
+\bibliographystyle{abbrvnat}
 \DeclareRobustCommand{\VAN}[3]{#3}

### 2013-06-19 paper.tex
trim use of em/en dashes
I've never really liked these, as they take the flow out of sentences,
even more than parentheses. They also aren't necessary when commas suffice.

@@
 and Bootstrap. The score function used by default is the \texttt{score} method
-provided by the estimator. However, the library provides a variety of
-score functions that the user can choose instead, including -- among many others
--- accuracy, AUC, $F_1$-score for classification; $R^2$ score or mean squared
-error for regression and the Rand Index for clustering.
+provided by the estimator. However, the library provides a large variety of
+score functions that the user can choose instead,
+including accuracy, AUC, $F_1$-score for classification;
+$R^2$ score or mean squared error for regression
+and the Rand index for clustering.
 
@@
 
-Our rule of thumb is that user code should not be tied to \sklearn -- which is a
+Our rule of thumb is that user code should not be tied to \sklearn---which is a
 \textit{library}, and not a \textit{framework}. This principle indeed avoids a
@@
 The authors and contributors acknowledge active support from INRIA. Past and
-present sponsors of the project also include Google -- for funding code
-scholarships (\textit{Google Summer of Code}) --, the Python Software Fundation and
-Tinyclues -- for funding coding sprints.
+present sponsors of the project also include Google for funding code
+scholarships (\textit{Google Summer of Code}),
+the Python Software Fundation and Tinyclues for funding coding sprints.
 

### 2013-06-19 paper.tex
move the verb back at the beginning of the sentence

@@
 When feasible, external libraries are integrated into the codebase.
-In particular, modified versions of \textsf{LIBSVM} and \textsf{LIBLINEAR}
+In particular, \sklearn includes modified versions of \textsf{LIBSVM} and \textsf{LIBLINEAR}
 \citep{chang2011libsvm,fan2008}, both written in C{}\verb!++!,
-and wrapped using Cython modules, are shipped with \sklearn.
+and wrapped using Cython modules.
 

### 2013-06-19 paper.tex
grammar, typo, compress a sentence
Left the verb in this time.

--- 2013-06-19 paper.tex
@@
 
-Our implementation guidelines emphasize on writing efficient but readable code.
+Our implementation guidelines emphasize writing efficient but readable code.
 In particular, we focus on making the codebase easily maintainable and
@@
 While the text vectorization part of \sklearn
-also uses irerables to some extent,
-the emphasis is still largely on (mini-)batch processing,
+also uses iterables to some extent,
+the emphasis is still on (mini-)batch processing,
 even in the truly stateless, O(1) memory vectorizers
@@
 reduces various forms of overhead
-and allows users to make use of the abundance of memory in modern machines.
+and makes effective use of the abundance of memory in modern computers.
 No attempt is made to hide this batch-oriented processing from the user,

### 2013-06-19 paper.tex
put back meta-estimators sub-section
In lack of something better, I think it is best at the time to put back the meta-estimators section. Yet, this subsection still requires some work for the message to be clear and explicit.

@@
 
+Building on the core interface introduced in the previous section, we now
+present in this section advanced API mechanisms for building meta-estimators,
+composing complex estimators and then selecting models. We also discuss design
+choices allowing for an easy use and extension of \sklearn.
+
+\subsection{Meta-estimators}
+
 Some machine learning algorithms are expressed naturally

### 2013-06-19 paper.tex
PCA pipeline example runs across page boundaries; make it float
There are more cases where this happens, but this is the longest
snippet.

@@
 combined to create complex and nested workflows.
-As an example, the following snippet illustrates how to create a complex worflow
-computing both linear PCA and kernel PCA components on \texttt{X\_train}
+Figure~\ref{union+pipeline} illustrates how to create a complex estimator
+that computes both linear PCA and kernel PCA features on \texttt{X\_train}
 (through a \texttt{FeatureUnion}),
-selecting the 10 best of those features according to an ANOVA test
-and then feeding them to an $\ell_2$-regularized logistic regression model:
+selects the 10 best features in the combination according to an ANOVA test
+and feeds those to an $\ell_2$-regularized logistic regression model.
+
+\begin{figure}[h]
 \begin{pythoncode}
@@
 \end{pythoncode}
+\caption{Combined use of \texttt{Pipeline} and \texttt{FeatureUnion}.}
+\label{union+pipeline}
+\end{figure}
 

### 2013-06-19 paper.tex
floating figure for k-means example too
XXX there's an asymmetry between the examples now...

@@
 \end{pythoncode}
+
 Some unsupervised learning
-estimators may however also implement that interface, as illustrated below:
+estimators may however also implement the \texttt{predict} interface.
+The code in Figure~\ref{kmeans-predict} fits a $k$-means model with $k=10$
+on training data \texttt{X\_train},
+then uses the model's \texttt{predict} method
+to obtain cluster labels for unseen data \texttt{X\_test}.
+
+\begin{figure}[h]
 \begin{pythoncode}
@@
 \end{pythoncode}
-In the above example, a \texttt{KMeans} estimator is first fit on the training
-data \texttt{X\_train}. Then the \texttt{predict} method is called on the test
-data \texttt{X\_test} in order to retrieve predicted cluster labels.
-The \texttt{predict} method in \sklearn thus offers a unified way
-to compute predictions with respect to unseen data.
+\caption{Finding closest centroids for unseen data after $k$-means clustering.
+         \texttt{clust\_pred} will hold integer indices of centroids.}
+\label{kmeans-predict}
+\end{figure}
 

### 2013-06-19 paper.tex
emphasize that StandardScaler is stateful
User questions boiling down to "why not just scale train and test
separately" pop up every now and then.
superfluous comma

@@
 Of course, in practice, it is important to apply the same preprocessing to the
-test data \texttt{X\_test}. This can easily be done as follows:
+test data \texttt{X\_test}.
+Since a \texttt{StandardScaler} stores the mean and standard deviation
+that it computed for the training set,
+applying it to an unseen test set \texttt{X\_test}
+maps that into the appropriate region of feature space:
 \begin{pythoncode}
@@
 In particular, \sklearn includes modified versions of \textsf{LIBSVM} and \textsf{LIBLINEAR}
-\citep{chang2011libsvm,fan2008}, both written in C{}\verb!++!,
+\citep{chang2011libsvm,fan2008}, both written in C{}\verb!++!
 and wrapped using Cython modules.

### 2013-06-19 paper.tex
clustering example figure
XXX the float actually makes it take more space, even though the font
is smaller and even without the caption.
Is there a smarter way to keep it from crossing the page boundary?

@@
 Some unsupervised learning
-estimators may however also implement the \texttt{predict} interface.
+estimators may also implement the \texttt{predict} interface.
 The code in Figure~\ref{kmeans-predict} fits a $k$-means model with $k=10$
@@
 then uses the model's \texttt{predict} method
-to obtain cluster labels for unseen data \texttt{X\_test}.
+to obtain cluster labels (integer indices) for unseen data \texttt{X\_test}.
 
@@
 \end{pythoncode}
-\caption{Finding closest centroids for unseen data after $k$-means clustering.
-         \texttt{clust\_pred} will hold integer indices of centroids.}
+\caption{Finding closest centroids for unseen data after $k$-means clustering.}
 \label{kmeans-predict}

### 2013-06-19 paper.tex
*every* transformer has fit_transform
Also "huge difference" -> "it depends" (but it's always better).

@@
 \end{pythoncode}
-In this particular example, \texttt{fit(X\_train).transform(X\_train)} can also further
-be rewritten as \texttt{fit\_transform(X\_train)}.  Combining the \texttt{fit}
-and \texttt{transform} methods into a single \texttt{fit\_transform} method
-indeed allows to share computations, which often makes a huge difference in practice.
+
+Furthermore, every estimator allows \texttt{fit(X\_train).transform(X\_train)}
+to be written as \texttt{fit\_transform(X\_train)}.
+The combined \texttt{fit\_transform} method prevents repeated computations.
+Depending on the transformer,
+this may either skip only an input validation step,
+or in fact use a more efficient algorithm for the transformation.
 In the same spirit, clustering algorithms provide a

### 2013-06-19 paper.tex
emphasize importance of separate initialization in ECOC example
XXX @mblondel, I'm not really familiar with this algorithm;
maybe it's easier to use an OvR naive Bayes instead?

@@
 If multiclass classification based on error-correcting codes
-\citep{dietterich1995solving} is desired instead,
+is desired instead,
 then this can achieved  using an \texttt{OutputCodeClassifier} meta-estimator
@@
 \end{pythoncode}
+The \texttt{OutputCodeClassifier} object
+\textit{clones} the LR estimator multiple times for learning,
+resulting in a set of estimator objects with the same settings
+that each perform a binary classification
+(see \citealt{dietterich1995solving} for the algorithm).
+The snippet thus exemplifies the importance
+of separating object instantiation and actual learning.
 

### 2013-06-19 paper.tex
copyedit/elaborate SVM grid search example
Rearranged the param_grid to make it more suggestive of two separate
grids for two kernels.  (I didn't even know this was possible! :)

@@
 
-As an example, the snippet below illustrates how to optimize the
-hyper-parameters of an SVM classifier. It maximizes the $F_1$ score of the
-classifier through a 10-fold cross-validation scheme.
+As an example, the snippet below illustrates how to find the
+hyper-parameter settings of an SVM classifier (SVC)
+that maximize its $F_1$ score,
+through 10-fold cross-validation on the training set.
 \begin{pythoncode}
+from sklearn.grid_search import GridSearchCV
 from sklearn.svm import SVC
-from sklearn.grid_search import GridSearchCV
 
 param_grid = [
-  {"C": [1, 10, 100, 1000], "kernel": ["linear"]},
-  {"C": [1, 10, 100, 1000], "gamma": [0.001, 0.0001],
-   "kernel": ["rbf"]},
+  {"kernel": ["linear"], "C": [1, 10, 100, 1000]},
+  {"kernel": ["rbf"], "C": [1, 10, 100, 1000],
+                      "gamma": [0.001, 0.0001]}
 ]
@@
 considered. This is because it only makes sense to search for \texttt{gamma}
-when \texttt{kernel} is set to \texttt{"rbf"}.  Additionally, we see that
+when \texttt{kernel} is set to \texttt{"rbf"}.
+An SVM with a linear kernel will accept a $\gamma$ parameter,
+but ignore it, so the grid search would waste time
+trying out effectively equivalent parameter settings.
+Additionally, we see that
 \texttt{GridSearchCV} has a \texttt{predict} method, like any other classifier.
@@
 delegate the \texttt{predict}, \texttt{predict\_proba}, \texttt{transform} and
-\texttt{score} methods to the best estimator, thus mimicking its behavior.
+\texttt{score} methods to the best estimator
+(optionally after re-fitting it on the whole training set).
 

### 2013-06-19 paper.tex
nested pipelines example (maybe subsection?)

@@
 
+% XXX do we need a subsection header here?
+% "Recursive composition of estimators"?
+Figure~\ref{nested-gridsearch} shows how different meta-estimators interact.
+This snippet builds a composite estimator in several steps.
+First, a linear SVM classifier is combined in a pipeline with
+a feature selector based on $\chi^2$ test statistics.
+This is then fed as input to a \texttt{OneVsRestClassifier}
+to build a $K$-way multiclass estimator
+consisting of $K$ binary classifiers,
+after determining the value of $K$ at \texttt{fit} time
+from the training labels.
+Each of these binary classifiers performs feature selection
+independently of all the others.
+This composite classifier is then combined
+with a \textsf{tf--idf} feature extractor using yet another pipeline.
+Because the result is again an estimator,
+we can perform grid search to find optimal values for the hyper-parameters
+of each step of the recursive pipeline.
+
+The \texttt{param\_grid} in Figure~\ref{nested-gridsearch}
+uses a \sklearn-specific notation
+to address the individual components of the nested structure.
+This notation uses double underscores to separate components,
+so that \texttt{tfidf\_\_sublinear\_tf}
+addresses the \texttt{sublinear\_tf} setting
+of the \textsf{tf--idf} feature extractor
+(when turned on, term frequencies are replaced by their logarithms).
+Note the \texttt{estimator} component,
+which address the base estimator (the first pipeline)
+inside the OvR meta-estimator.
+This is required because the wrapper may itself have settings
+that might conflict with the estimator that it wraps;
+this is not required for pipelines,
+because those have no settings of their own.
+
+\begin{figure}[t]
+\begin{pythoncode}
+from sklearn.grid_search import GridSearchCV
+from sklearn.feature_extraction.text import TfidfVectorizer
+from sklearn.feature_selection import SelectKBest, chi2
+from sklearn.multiclass import OneVsRestClassifier
+from sklearn.pipeline import Pipeline
+from sklearn.svm import LinearSVC
+
+svm_chi2 = Pipeline([("chi2", SelectKBest(chi2)),
+                     ("svm", LinearSVC())])
+ovr_svm = OneVsRestClassifier(svm_chi2)
+
+doc_clf = Pipeline([("tfidf", TfidfVectorizer()),
+                    ("ovr", ovr_svm)])
+
+param_grid = {
+  "tfidf__sublinear_tf": [True, False],
+  "ovr__estimator__chi2__k": [200, 500, 1000],
+  "ovr__estimator__svm_C": [1, 10, 100]}
+
+clf = GridSearchCV(doc_clf, param_grid, cv=10)
+\end{pythoncode}
+\caption{Parameter search with nested pipelines for text classification.}
+\label{nested-gridsearch}
+\end{figure}
+
 \subsection{Extending scikit-learn}

### 2013-06-19 paper.tex
trim implementation section a bit
We don't just use NumPy for "intensive" processing but whenever it
leads to clearer code.

--- 2013-06-19 paper.tex
@@
 understandable in order to favour external contributions. Whenever practicable,
-algorithms implemented in \sklearn are thus directly written in pure Python. For
-intensive numerical computations though, the library mostly relies on NumPy
-vector operations, which allows for the code to remain concise, readable and
-efficient. For all algorithms that cannot be easily and efficiently expressed as
-NumPy operations, we rely on Cython \citep{behnel2011cython}  for writing the
-critical parts and achieving decent performance and scalability. Cython is a
+algorithms implemented in \sklearn are written in Python,
+using Numpy vector operations for numerical work.
+This allows for the code to remain concise, readable and
+efficient. For critical algorithms that cannot be easily and efficiently
+expressed as NumPy operations, we rely on Cython \citep{behnel2011cython}
+to achieve decent performance and scalability. Cython is a
 compiled programming language that extends Python with static typing. It

### 2013-06-19 paper.tex
rm Julia comparison, replace R citation by preferred one from R project

@@
 
-% XXX Is the Julia comparison fair?
-% The goal of that language seems to be a bit different:
-% to provide a compiler to implement numeric algorithms without vectorized idioms,
-% maybe more like Cython than Python...
 In comparison to specialized languages for numeric and statistical programming
-such as \textsc{matlab}, R \citep{trancon2012r} and Julia \citep{bezanson2012julia},
+such as \textsc{matlab} and R \citep{r},
 the use of Python as a host language has the distinct advantage
-that it is a mature, \textit{general purpose} language:
+that it is a \textit{general purpose} language:
 it has strong language and standard library support for such tasks as

### 2013-06-19 paper.tex
minor trimming and copyedit

@@
 provided by the estimator. However, the library provides a large variety of
-score functions that the user can choose instead,
-including accuracy, AUC, $F_1$-score for classification;
-$R^2$ score or mean squared error for regression
+score functions that the user can choose from,
+including accuracy and $F_1$-score for classification;
+$R^2$ score or mean squared error for regression;
 and the Rand index for clustering.
@@
 
-To facilitate the adoption and the distribution of \sklearn,
+To facilitate the installation and thus adoption of \sklearn,
 the set of external dependencies is kept to a bare minimum:
@@
 will have to perform ``on the side''.
-In addition, third-party libraries are available for virtually any task
-that modern computer programs routinely perform.
+In addition, third-party libraries are available for many other routine tasks.
 
@@
 Although its stated goal is not machine learning,
-it does exemplify a style of API design for scalable processing of ``big data''
-that takes a different approach than the comparable models in \sklearn.
+it does exemplify API design for scalable processing of ``big data''
+that takes a different approach than comparable models in \sklearn.
 Gensim's way of dealing with large datasets is to use algorithms
@@
 the emphasis is still on (mini-)batch processing,
-even in the truly stateless, O(1) memory vectorizers
+even in the stateless, O(1) memory vectorizers
 that implement the hashing trick of \citet{weinberger2009}.
@@
 and makes effective use of the abundance of memory in modern computers.
-No attempt is made to hide this batch-oriented processing from the user,
+We make no attempt to hide this batch-oriented processing from the user,
 allowing control over the amount of memory actually dedicated

### 2013-06-19 paper.tex
omit all the import statements; down to 13 pages
After all, this is not tutorial material.

@@
 let us consider a supervised learning task using logistic regression.
-Given the API defined above, solving this problem is as simple as:
+Given the API defined above, solving this problem is as simple as:\footnote{
+  \texttt{import} statement are omitted for brevity
+  throughout this paper.}
 \begin{pythoncode}
-from sklearn.linear_model import LogisticRegression
-
 clf = LogisticRegression(penalty="l1")
@@
 \begin{pythoncode}
-from sklearn.cluster import KMeans
-
 km = KMeans(n_clusters=10)
@@
 \begin{pythoncode}
-from sklearn.preprocessing import StandardScaler
-
 scaler = StandardScaler()
@@
 \begin{pythoncode}
-from sklearn.linear_model import LogisticRegression
-from sklearn.multiclass import OutputCodeClassifier
-
 ecoc_lr = OutputCodeClassifier(LogisticRegression(penalty="l1"))
@@
 \begin{pythoncode}
-from sklearn.pipeline import FeatureUnion, Pipeline
-from sklearn.decomposition import PCA, KernelPCA
-from sklearn.feature_selection import SelectKBest
-from sklearn.linear_model import LogisticRegression
-
 fu = FeatureUnion([("pca", PCA()),
@@
 \begin{pythoncode}
-from sklearn.grid_search import GridSearchCV
-from sklearn.svm import SVC
-
 param_grid = [

### 2013-06-19 paper.tex
replace ECOC with OvO; easier to explain, -1 reference

@@
 for performing multiclass classification.
-If multiclass classification based on error-correcting codes
-is desired instead,
-then this can achieved  using an \texttt{OutputCodeClassifier} meta-estimator
+If a one-vs.-one scheme is desired instead,
+this can achieved using a \texttt{OneVsOneClassifier} meta-estimator
 wrapping a \texttt{LogisticRegression} estimator:
 \begin{pythoncode}
-ecoc_lr = OutputCodeClassifier(LogisticRegression(penalty="l1"))
+ovo_lr = OneVsOneClassifier(LogisticRegression(penalty="l1"))
 \end{pythoncode}
-The \texttt{OutputCodeClassifier} object
+The \texttt{OneVsOneClassifier} object
 \textit{clones} the LR estimator multiple times for learning,
-resulting in a set of estimator objects with the same settings
-that each perform a binary classification
-(see \citealt{dietterich1995solving} for the algorithm).
-The snippet thus exemplifies the importance
+resulting in a set of $\frac{K(K-1)}{2}$ estimator objects
+for $K$-way classification,
+all with the same settings,
+that each perform a binary classification and vote to make the final decision.
+The snippet exemplifies the importance
 of separating object instantiation and actual learning.

### 2013-06-19 paper.tex
copyedit acks

@@
 The authors and contributors acknowledge active support from INRIA. Past and
-present sponsors of the project also include Google for funding code
-scholarships (\textit{Google Summer of Code}),
+present sponsors of the project also include Google for funding
+scholarships through its Summer of Code program,
 the Python Software Fundation and Tinyclues for funding coding sprints.

### 2013-06-19 paper.tex
copyedit, trim
Three substantive changes:
* the implementation section no longer discusses Python limitations;
* transformers have fit_transform, not all estimators;
* vectorizers don't only handle text documents.

@@
 core developers, mostly researchers from various fields (e.g., machine learning,
-natural language processing, or neuroscience). The project also greatly benefits
+natural language processing, or neuroscience). The project also benefits
 from many occasional contributors proposing small code or documentation fixes or
@@
 specific quality guidelines, such as style consistency and unit-test coverage.
-In addition, complete documentation and examples are required for all features.
-Finally, additions to the code base must pass code review by at least two
-developers not involved in the implementation of the proposed addition.
+Complete documentation and examples are required for all features,
+and major modifications must pass code review by at least two
+developers not involved in the implementation of the proposed change.
 
-Within the past few years, the project has proved valuable to an ever increasing
+Over the past few years, the project has proved valuable to an ever increasing
 user base. Its popularity can be gaged from various indicators such as hundreds
@@
 challenges (e.g., Kaggle), or significant statistics derived from our
-repositories and mailing lists.  At the time of writing, the project is watched
-by 1309 people and forked 667 times on GitHub, and the mailing-list receives
-more than 300 mails per month. The version control logs for the development
-version
+repositories and mailing list.  At the time of writing, the project is watched
+by 1309 people and forked 667 times on GitHub; the mailing list receives
+more than 300 mails per month; and version control logs
 % ddaa494c116e3c16bf032003c5cccbed851733d2
-also count 183 unique contributors to the project.
+show 183 unique contributors to the codebase.
 
@@
 advanced API mechanisms built from the core interface. Section
-\ref{sec:implementation} describes implementation principles and discusses
-limitations of the technologies we rely on. Section \ref{sec:comparison} then
+\ref{sec:implementation} briefly describes the implementation.
+Section \ref{sec:comparison} then
 compares \emph{scikit-learn} to other major projects in terms of API. Finally,
@@
 
-As an example to illustrate the initialize-\texttt{fit} sequence,
+To illustrate the initialize-\texttt{fit} sequence,
 let us consider a supervised learning task using logistic regression.
@@
 Transformers also include a variety of learning algorithms, such as
-dimension reduction (PCA, manifold learning techniques), kernel
-approximation techniques, and any other approach that can be used to
-derive a modified representation of the data.
+dimension reduction (PCA, manifold learning), kernel approximation,
+and other mappings from one feature space to another.
 
@@
 
-Furthermore, every estimator allows \texttt{fit(X\_train).transform(X\_train)}
+Furthermore, every transformer allows \texttt{fit(X\_train).transform(X\_train)}
 to be written as \texttt{fit\_transform(X\_train)}.
@@
 Depending on the transformer,
-this may either skip only an input validation step,
+it may skip only an input validation step,
 or in fact use a more efficient algorithm for the transformation.
@@
 that is equivalent to \texttt{fit} followed by \texttt{predict},
-and thus returns the cluster labels assigned to the training data samples.
+returning cluster labels assigned to the training samples.
 
@@
 for performing multiclass classification.
-If a one-vs.-one scheme is desired instead,
-this can achieved using a \texttt{OneVsOneClassifier} meta-estimator
-wrapping a \texttt{LogisticRegression} estimator:
+A different scheme can be achieved
+by a meta-estimator wrapping a logistic regression estimator:
 \begin{pythoncode}
@@
 
-\texttt{Pipeline} objects are used for chaining multiple estimators into a
+\texttt{Pipeline} objects chain multiple estimators into a
 single one. This is useful since a machine learning workflow typically involves
-a fixed sequence of steps in processing data (e.g., feature extraction,
-normalization, feature selection, learning and making predictions). A sequence
-of $N$ such steps can be combined into a composite object called a
-\texttt{Pipeline}. The first $N-1$ steps must be transformers and the last can
+a fixed sequence of processing steps (e.g., feature extraction,
+dimensionality reduction, learning and making predictions). A sequence
+of $N$ such steps can be combined into a \texttt{Pipeline}
+if the first $N-1$ steps are transformers; the last can
 be either a predictor or a transformer. Conceptually, fitting a pipeline to a
@@
 
-A \texttt{FeatureUnion} object is a transformer combining multiple transformers
+\texttt{FeatureUnion} combines multiple transformers
 into a single one that concatenates their outputs.
@@
 by calling its underlying transformers in parallel.
-From an API point of view, a
+In terms of the API, a
 \texttt{FeatureUnion} takes as input a list of transformers. Calling
@@
 
-As an example, the snippet below illustrates how to find the
-hyper-parameter settings of an SVM classifier (SVC)
-that maximize its $F_1$ score,
+The snippet below illustrates how to find
+hyper-parameter settings for an SVM classifier (SVC)
+that maximize $F_1$ score
 through 10-fold cross-validation on the training set.
@@
 
-In other places of the library, in particular in the feature extraction and
-vectorization code for text documents, the toolkit is also designed to be
+In other places of the library, in particular in the vectorization code
+for unstructured inputs, the toolkit is also designed to be
 extensible. Here, estimators provide hooks for user-defined code: objects or
 functions that provide a specific API can be given as arguments at vectorizer
-construction time. The library then calls into this code to obtain the desired
-behavior. Again, such external user code can be kept agnostic of the \sklearn
+construction time. The library then calls into this code.
+Again, such external user code can be kept agnostic of the \sklearn
 class hierarchy.

### 2013-06-19 paper.tex
future work: serialization
Also, 1.0 being near actually implies problems needing a solution.

@@
 
-Although it is coming close to a 1.0 release, \sklearn is by no means finished.
-At present, it does not support some classical machine learning tasks. Among
-others, missing features include neural networks, ensemble meta-estimators for
-bagging or subsampling strategies, or missing value completion algorithms.
-Topics like structured prediction or reinforcement learning are however
+A \sklearn~1.0 release is nearing,
+but several problems must still be resolved.
+At present, the library does not support some classical machine learning tasks,
+including neural networks, ensemble meta-estimators for
+bagging or subsampling strategies and missing value completion algorithms.
+Topics like structured prediction or reinforcement learning are
 considered off-topic for the project, since they would require quite different
-data representations and API.
+data representations and APIs.
 
@@
 
+Finally, a long-term solution for model persistence is missing.
+Currently, Python's \texttt{pickle} module is recommended for serialization,
+but this only offers a file format,
+not a way of preserving compatibility between versions.
+Also, it has security problems because its deserializer
+may execute arbitrary Python code,
+so models from untrusted sources cannot be safely ``unpickled''.
+
 \subsection*{Acknowledgements}

### 2013-06-20 paper.tex
remove figures (no snippet is broken on 2 pages)

@@
 
-Some unsupervised learning
-estimators may also implement the \texttt{predict} interface.
-The code in Figure~\ref{kmeans-predict} fits a $k$-means model with $k=10$
-on training data \texttt{X\_train},
-then uses the model's \texttt{predict} method
-to obtain cluster labels (integer indices) for unseen data \texttt{X\_test}.
-
-\begin{figure}[h]
+Some unsupervised learning estimators may also implement the \texttt{predict}
+interface. The code in the snippet below fits a $k$-means model with $k=10$ on
+training data \texttt{X\_train}, and then uses the  \texttt{predict} method to
+obtain cluster labels (integer indices) for unseen data \texttt{X\_test}.
 \begin{pythoncode}
@@
 \end{pythoncode}
-\caption{Finding closest centroids for unseen data after $k$-means clustering.}
-\label{kmeans-predict}
-\end{figure}
 
@@
 combined to create complex and nested workflows.
-Figure~\ref{union+pipeline} illustrates how to create a complex estimator
+The following snippet illustrates how to create a complex estimator
 that computes both linear PCA and kernel PCA features on \texttt{X\_train}
@@
 selects the 10 best features in the combination according to an ANOVA test
-and feeds those to an $\ell_2$-regularized logistic regression model.
-
-\begin{figure}[h]
+and feeds those to an $\ell_2$-regularized logistic regression model:
 \begin{pythoncode}
@@
 \end{pythoncode}
-\caption{Combined use of \texttt{Pipeline} and \texttt{FeatureUnion}.}
-\label{union+pipeline}
-\end{figure}
 

### 2013-06-20 paper.tex
wordings (abstract + the scikit-learn project)

@@
 \begin{abstract}
-The \sklearn project is an increasingly popular open source machine learning
-library written in Python and designed to be simple and efficient, accessible to
+The \sklearn project is an increasingly popular machine learning
+library. Written in Python, it is designed to be simple and efficient, accessible to
 non-experts, and reusable in various contexts. In this paper, we present and
@@
 \citep{vanderwalt2011} augments Python with a numeric array data type and proper
-array computing primitives while SciPy \citep{varoquaux2013scipy} extends it
+array computing primitives, while SciPy \citep{varoquaux2013scipy} extends it
 further by making common numerical operations available, either by implementing
@@
 benchmarked against several competitors. In this paper, we instead focus on an
-in-depth discussion and analysis of the design choices used in building the
+in-depth discussion and analysis of the design choices made for building the
 library. We first present in section \ref{sec:core-api} the central application

### 2013-06-20 paper.tex
wordings (2., 2.1, 2.2)

@@
   \item[Consistency.]
-       All objects (basic or composed) share a consistent interface composed of
+       All objects (basic or composite) share a consistent interface composed of
        a limited set of methods. This interface is documented in a consistent
@@
        This keeps \sklearn easy to use and easy to combine with other libraries.
-       % XXX do we need to elaborate more on this, or do we expect everyone
-       % to be an experienced Java/adapter pattern hater?
   \item[Composition.]
@@
 Weka \citep{hall2009weka}, they bring the prime advantage of allowing us to rely
-% Lars: "orders of magnitude faster" is an implementation detail
-on all NumPy and SciPy vector operations while allowing us at the same time to
-keep the code short and readable.  This design choice has also been motivated by
+on all NumPy and SciPy vector operations while keeping
+the code short and readable.  This design choice has also been motivated by
 the fact that, given their pervasiveness in many other scientific Python
@@
 arrays and SciPy sparse matrices.
-% Gilles: the phrase was too long
-% hence helping them to familiarize with the project.
 From a practical point of view, these formats also provide a bunch of
@@
 data and set these as attributes on the estimator object. As a convention, the
-parameters learned by an estimator are exposed as public attributes whose name
-is suffixed with a trailing underscore (e.g., \texttt{coef\_} for the
+parameters learned by an estimator are exposed as public attributes whose names
+are suffixed with a trailing underscore (e.g., \texttt{coef\_} for the
 learned coefficients of a linear model),
@@
 Given the API defined above, solving this problem is as simple as:\footnote{
-  \texttt{import} statement are omitted for brevity
+  \texttt{import} statements are omitted for brevity
   throughout this paper.}

### 2013-06-20 paper.tex
wordings (2.3, 2.4)

@@
 Of course, in practice, it is important to apply the same preprocessing to the
-test data \texttt{X\_test}.
-Since a \texttt{StandardScaler} stores the mean and standard deviation
-that it computed for the training set,
-applying it to an unseen test set \texttt{X\_test}
-maps that into the appropriate region of feature space:
+test data \texttt{X\_test}. Since a \texttt{StandardScaler} estimator stores the
+mean and standard deviation that it computed for the training set, transforming
+an unseen test set \texttt{X\_test} maps it into the appropriate region of
+feature space:
 \begin{pythoncode}

### 2013-06-20 paper.tex
wordings (3.1, 3.2)

@@
 Building on the core interface introduced in the previous section, we now
-present in this section advanced API mechanisms for building meta-estimators,
+present advanced API mechanisms for building meta-estimators,
 composing complex estimators and then selecting models. We also discuss design
@@
 \end{pythoncode}
-The \texttt{OneVsOneClassifier} object
-\textit{clones} the LR estimator multiple times for learning,
+For learning, the \texttt{OneVsOneClassifier} object
+\textit{clones} the logistic regression estimator multiple times,
 resulting in a set of $\frac{K(K-1)}{2}$ estimator objects
 for $K$-way classification,
-all with the same settings,
-that each perform a binary classification and vote to make the final decision.
+all with the same settings.
+For predictions, all estimators perform a binary classification and then vote to make the final decision.
 The snippet exemplifies the importance
@@
 
-Since using meta-estimators requires the construction of nested objects,
+Since meta-estimators requires the construction of nested objects,
 the decision to implement a meta-estimator
@@
 
-\texttt{FeatureUnion} combines multiple transformers
-into a single one that concatenates their outputs.
-I.e., a union of transformers that map inputs having $d$ features
-to $d'$ and $d''$ features, respectively,
-is a transformer that maps its $d$ input features to $d' + d''$ features
-by calling its underlying transformers in parallel.
-In terms of the API, a
-\texttt{FeatureUnion} takes as input a list of transformers. Calling
-\texttt{fit} on the union is the same as calling \texttt{fit} independently on
-each of the transformers.
+\texttt{FeatureUnion} objects combine multiple transformers into a single one
+that concatenates their outputs. For example, a union of two transformers that
+each maps inputs having $d$ features to $d'$ and $d''$ features respectively is
+a transformer that maps its $d$ input features to $d' + d''$ features. In terms
+of API, a \texttt{FeatureUnion} takes as input a list of transformers.
+Calling \texttt{fit} on the union is the same as calling \texttt{fit}
+independently on each of the transformers and then joining their outputs.
 

### 2013-06-20 paper.tex
wordings (3.3, 3.4)

@@
 when \texttt{kernel} is set to \texttt{"rbf"}.
-An SVM with a linear kernel will accept a $\gamma$ parameter,
-but ignore it, so the grid search would waste time
-trying out effectively equivalent parameter settings.
+An SVM with a linear kernel could accept a $\gamma$ parameter,
+but would ignore it and thus waste computing time trying out effectively
+equivalent settings if a single hyper-parameter grid was used insteead.
 Additionally, we see that
@@
 functions that provide a specific API can be given as arguments at vectorizer
-construction time. The library then calls into this code.
+construction time. The library then calls into this code to obtain the desired behavior.
 Again, such external user code can be kept agnostic of the \sklearn

### 2013-06-20 paper.tex
the description of FeatureUnion.transform is a definition, not an example
+ a small grammar fix

@@
 
-Since meta-estimators requires the construction of nested objects,
+Since meta-estimators require the construction of nested objects,
 the decision to implement a meta-estimator
@@
 \texttt{FeatureUnion} objects combine multiple transformers into a single one
-that concatenates their outputs. For example, a union of two transformers that
-each maps inputs having $d$ features to $d'$ and $d''$ features respectively is
+that concatenates their outputs. A union of two transformers that
+map inputs having $d$ features to $d'$ and $d''$ features respectively is
 a transformer that maps its $d$ input features to $d' + d''$ features. In terms

### 2013-06-21 paper.tex
generalize FU definition
fixes last comments

@@
 benchmarked against several competitors. In this paper, we instead focus on an
-in-depth discussion and analysis of the design choices made for building the
+in-depth discussion and analysis of the design choices made when building the
 library. We first present in section \ref{sec:core-api} the central application
@@
 data and set these as attributes on the estimator object. As a convention, the
-parameters learned by an estimator are exposed as public attributes whose names
-are suffixed with a trailing underscore (e.g., \texttt{coef\_} for the
+parameters learned by an estimator are exposed as public attributes with names
+suffixed with a trailing underscore (e.g., \texttt{coef\_} for the
 learned coefficients of a linear model),
@@
 map inputs having $d$ features to $d'$ and $d''$ features respectively is
-a transformer that maps its $d$ input features to $d' + d''$ features. In terms
-of API, a \texttt{FeatureUnion} takes as input a list of transformers.
+a transformer that maps its $d$ input features to $d' + d''$ features.
+This generalizes in the obvious way to more than two transformers.
+In terms of API, a \texttt{FeatureUnion} takes as input a list of transformers.
 Calling \texttt{fit} on the union is the same as calling \texttt{fit}
@@
 when \texttt{kernel} is set to \texttt{"rbf"}.
-An SVM with a linear kernel could accept a $\gamma$ parameter,
-but would ignore it and thus waste computing time trying out effectively
-equivalent settings if a single hyper-parameter grid was used insteead.
+An SVM with a linear kernel accepts a $\gamma$ parameter,
+but actually ignores it. Computing time would thus be wasted, trying out effectively
+equivalent settings, if a single hyper-parameter grid was used instead.
 Additionally, we see that

### 2013-06-21 paper.tex
fixes
Put imports back.

@@
 
-To illustrate the initialize-\texttt{fit} sequence,
+To illustrate the initialize-fit sequence,
 let us consider a supervised learning task using logistic regression.
-Given the API defined above, solving this problem is as simple as:\footnote{
-  \texttt{import} statements are omitted for brevity
-  throughout this paper.}
+Given the API defined above, solving this problem is as simple as the following
+example.
+\newpage
 \begin{pythoncode}
+from sklearn.linear_model import LogisticRegression
+
 clf = LogisticRegression(penalty="l1")
@@
 obtain cluster labels (integer indices) for unseen data \texttt{X\_test}.
+\newpage
 \begin{pythoncode}
+from sklearn.cluster import KMeans
+
 km = KMeans(n_clusters=10)
@@
 \begin{pythoncode}
+from sklearn.preprocessing import StandardScaler
+
 scaler = StandardScaler()
@@
 \begin{pythoncode}
+from sklearn.multiclass import OneVsOneClassifier
+
 ovo_lr = OneVsOneClassifier(LogisticRegression(penalty="l1"))
@@
 selects the 10 best features in the combination according to an ANOVA test
-and feeds those to an $\ell_2$-regularized logistic regression model:
+and feeds those to an $\ell_2$-regularized logistic regression model.
+\newpage
 \begin{pythoncode}
+from sklearn.pipeline import FeatureUnion, Pipeline
+from sklearn.decomposition import PCA, KernelPCA
+from sklearn.feature_selection import SelectKBest
+
 fu = FeatureUnion([("pca", PCA()),
@@
 \begin{pythoncode}
+from sklearn.grid_search import GridSearchCV
+from sklearn.svm import SVC
+
 param_grid = [
@@
 functions that provide a specific API can be given as arguments at vectorizer
-construction time. The library then calls into this code to obtain the desired behavior.
+construction time. The library then calls into this code, communicating with it by passing objects of standard Python/Numpy types.
 Again, such external user code can be kept agnostic of the \sklearn
 class hierarchy.
-Communication proceeds by passing object of standard Python/NumPy types.
 

### 2013-06-22 paper.tex
Title :)
acknowledge my fundings
nested objects place a burden *on the user*

@@
 
-% working title, please change
-\title{Scikit-learn: API design for machine learning}
+\title{ API design for machine learning software: experiences from the
+scikit-learn project}
 
@@
 
-Since meta-estimators require the construction of nested objects,
+Since meta-estimators require users to construct nested objects,
 the decision to implement a meta-estimator
@@
 
+Lars Buitinck is a ... Gilles Louppe is a research fellow of the Fonds de la
+Recherche Scientifique-FNRS (Belgium) and acknowledges its financial support.
+Mathieu Blondel is a ...
+
 \bibliographystyle{abbrvnat}

### 2013-06-22 paper.tex
grid search doesn't work for clustering, so Rand index is irrelevant

@@
 score functions that the user can choose from,
-including accuracy and $F_1$-score for classification;
-$R^2$ score or mean squared error for regression;
-and the Rand index for clustering.
+including accuracy, AUC, $F_1$ score (classification),
+$R^2$ score and mean squared error (regression).
 

### 2013-06-22 paper.tex
trim Gensim comparison a bit

@@
 Gensim's way of dealing with large datasets is to use algorithms
-that have $O(1)$ space complexity and can be updated online,
-called \textit{document streaming} by \citeauthor{rehurek2010gensim}.
+that have $O(1)$ space complexity and can be updated online.
 The API is designed around the Python concept of an \textit{iterable}

### 2013-06-22 paper.tex
minor copyedit

@@
 further by making common numerical operations available, either by implementing
-those in Python/NumPy or by wrapping existing C/C{}\verb!++!/Fortran
+these in Python/NumPy or by wrapping existing C/C{}\verb!++!/Fortran
 implementations. In a similar vein, \sklearn extends Python with a set of
@@
        as meta-algorithms parametrized on other algorithms.
-       Whenever possible, those algorithms are implemented and composed from
+       Whenever possible, such algorithms are implemented and composed from
        existing building blocks.
@@
 multidimensional arrays for dense data and as SciPy sparse matrices for sparse
-data. While those may seem rather unsophisticated data representations when
-compared to more object-oriented constructs, such as those used by
+data. While these may seem rather unsophisticated data representations when
+compared to more object-oriented constructs, such as the ones used by
 Weka \citep{hall2009weka}, they bring the prime advantage of allowing us to rely
@@
 features) or feature extraction techniques (e.g., vectorization of text
-documents) also implement the same interface. Even those estimators that are
+documents) also implement the same interface. Even processing steps that are
 entirely stateless and therefore do not require the \texttt{fit} method to
@@
 into existing estimators classes
-is always based on a trade-off between genericity and ease of use.
+is always based on a trade-off between generality and ease of use.
 Relating to the example just given,
@@
 also take a cross-validation scheme and a score function.  The library provides
-various such cross-validation schemes, including K-Fold (default), Leave-One-Out
-and Bootstrap. The score function used by default is the \texttt{score} method
+various such cross-validation schemes, including $k$-fold (default),
+leave-one-out and bootstrapping.
+The score function used by default is the \texttt{score} method
 provided by the estimator. However, the library provides a large variety of
@@
 superfluous classes, the principle of \textit{duck typing} is exploited
-throughout the codebase. This technique allows both for extensibility and
+throughout the codebase. This technique allows for both extensibility and
 flexibility: as long as an estimator follows the API and conventions
 outlined in Section~\ref{sec:core-api}, then it can be used in lieu of a
-built-in estimator (e.g., it can plugged into pipelines or grid search).
+built-in estimator (e.g., it can be plugged into pipelines or grid search).
 That is, estimators are defined only by interface, not by inheritance,
@@
 ``banana'' should not get ``a gorilla holding the banana and the entire jungle''
-(J.~Armstrong, cited by \citealp[p.~213]{seibel2009coders}).  That is, programs
-using \sklearn should be easily reusable with other machine learning toolkits or
-in other contexts, because they don't need to be intimately tied to it.
+(J.~Armstrong, cited by \citealp[p.~213]{seibel2009coders}).
+That is, programs using \sklearn should not be intimately tied to it,
+so that their code can be reused with other toolkits or in other contexts.
 

### 2013-06-22 paper.tex
rename fu to union in example
For some reason, fu looks obscene to me.

@@
 
-fu = FeatureUnion([("pca", PCA()),
+union = FeatureUnion([("pca", PCA()),
                       ("kpca", KernelPCA(kernel="rbf"))])
 
-Pipeline([("feat_union", fu),
+Pipeline([("feat_union", union),
           ("feat_sel", SelectKBest(k=10)),

### 2013-06-22 paper.tex
model selection: emphasize generalization performance

@@
 As introduced in Section~\ref{sec:estimators}, hyper-parameters set in the
-constructor of an estimator determine its behavior. In particular, depending on
-the problem, some hyper-parameters produce estimators that are better than
-others. The problem of \textit{model selection} is therefore to find, within
+constructor of an estimator
+determine the behavior of its learning algorithm
+and hence the performance of the resulting model on unseen data.
+The problem of \textit{model selection} is therefore to find, within
 some hyper-parameter space, the best combination of hyper-parameters, with
@@
 tree, too small values of the \texttt{max\_depth} parameter will make the tree
-underfit while too large values will make the tree overfit the training data.
-The problem is therefore to find the best trade-off in order to maximize
-the performance of the tree on unseen data.
+underfit, while too large values will make it overfit.
 

### 2013-06-22 paper.tex

@@
 be used to combine typical machine learning workflows into a single object which
-is itself an estimator. Thanks to duck typing, composed
-estimators can be used wherever usual estimators can be used. In particular,
+is itself an estimator, and can be used wherever usual estimators can be used.
+In particular,
 they integrate within model selection routines, allowing one to optimize at once
@@
 To ease code reuse, simplify implementation and skip the introduction of
-superfluous classes, the principle of \textit{duck typing} is exploited
-throughout the codebase. This technique allows for both extensibility and
+superfluous classes, the Python principle of \textit{duck typing} is exploited
+throughout the codebase. This means that estimators are defined by interface,
+not by inheritance, where the interface is entirely implicit
+as far as the programming language is concerned.
+Duck typing allows both for extensibility and
 flexibility: as long as an estimator follows the API and conventions
 outlined in Section~\ref{sec:core-api}, then it can be used in lieu of a
-built-in estimator (e.g., it can be plugged into pipelines or grid search).
-That is, estimators are defined only by interface, not by inheritance,
+built-in estimator (e.g., it can be plugged into pipelines or grid search)
 and external developers are not forced to inherit from any \sklearn class.

### 2013-06-23 paper.tex
conclusion; mention nolearn
Fixes #19.

@@
 
-\textcolor{red}{TODO: include some overall conclusions (e.g., reasons of the success sklearn)}
+We have discussed the scikit-learn API
+and the way it maps machine learning concepts and tasks
+onto objects and operations in the Python programming language.
+We motivated choices and compromises made in its design.
+We argue that the API is powerful, usable and extensible,
+as witnessed by the appearance of third-party packages
+that follow the scikit-learn conventions and tie into the toolkit.\footnote{
+  E.g.\ nolearn, \url{http://pythonhosted.org/nolearn};
+  this mention does not imply endorsement of this package.}
+While part of the scikit-learn API is necessarily Python-specific,
+core concepts may be applicable to
+machine learning applications and toolkits
+written in other (dynamic) programming languages.
 

### 2013-06-24 paper.tex
Institutions.
Not that bad...

@@
 
-\author{Lars~Buitinck~\inst{1,2} \and
-        Gilles~Louppe~\inst{1,2} \and
-        Mathieu~Blondel~\inst{1,2} \and
-        Fabian~Pedregosa~\inst{2} \and
-        Andreas~Mueller~\inst{2} \and
-        Olivier~Grisel~\inst{2} \and
-        Vlad~Niculae~\inst{2} \and
-        Peter~Prettenhofer~\inst{2} \and
-        Alexandre~Gramfort~\inst{2} \and
-        Jacques~Grobler~\inst{2} \and
-        Robert~Layton~\inst{2} \and
-        Jake~Vanderplas~\inst{2} \and
+\author{Lars~Buitinck~\inst{1} \and
+        Gilles~Louppe~\inst{2} \and
+        Mathieu~Blondel~\inst{3} \and
+        Fabian~Pedregosa~\inst{4} \and
+        Andreas~Mueller~\inst{5} \and
+        Olivier~Grisel~\inst{6} \and
+        Vlad~Niculae~\inst{7} \and
+        Peter~Prettenhofer~\inst{8} \and
+        Alexandre~Gramfort~\inst{9} \and
+        Jacques~Grobler~\inst{4} \and
+        Robert~Layton~\inst{10} \and
+        Jake~Vanderplas~\inst{11} \and
         Arnaud~Joly~\inst{2} \and
-        Brian Holt~\inst{2} \and
-        Gaël~Varoquaux~\inst{2}}
+        Brian Holt~\inst{12} \and
+        Gaël~Varoquaux~\inst{4}}
 
-\institute{Main authors of the present paper \and
-           Core contributors to the project}
+\institute{University of Amsterdam \and
+           University of Liège \and
+           Kobe University \and
+           Parietal, INRIA Saclay \and
+           University of Bonn \and
+           Nuxeo \and 
+           University of Bucharest \and
+           Bauhaus-Universitat \and
+           Telecom ParisTech \and
+           University of Ballarat \and
+           University of Washington \and
+           Samsung Electronics Research Institute}
 
@@
 
-Lars Buitinck is a ... Gilles Louppe is a research fellow of the Fonds de la
-Recherche Scientifique-FNRS (Belgium) and acknowledges its financial support.
-Mathieu Blondel is a ...
-
 \bibliographystyle{abbrvnat}

### 2013-06-24 paper.tex
Remove manual page breaks.
affiliations (mine and @pprett's)

@@
 
-\institute{University of Amsterdam \and
+\institute{Informatics Institute, University of Amsterdam \and
            University of Liège \and
@@
            University of Bucharest \and
-           Bauhaus-Universitat \and
+           Ciuvo GmbH \and
            Telecom ParisTech \and
@@
 example.
-\newpage
 \begin{pythoncode}
@@
 obtain cluster labels (integer indices) for unseen data \texttt{X\_test}.
-\newpage
 \begin{pythoncode}
@@
 and feeds those to an $\ell_2$-regularized logistic regression model.
-\newpage
 \begin{pythoncode}

### 2013-06-24 paper.tex
@ogrisel's affiliation

--- 2013-06-24 paper.tex
@@
            University of Bonn \and
-           Nuxeo \and 
+           Independent consultant \and
            University of Bucharest \and
-           Bauhaus-Universitat \and
+           Ciuvo GmbH \and
            Telecom ParisTech \and

### 2013-06-24 paper.tex
update my affiliation

--- 2013-06-24 paper.tex
@@
         Peter~Prettenhofer~\inst{8} \and
-        Alexandre~Gramfort~\inst{9} \and
+        Alexandre~Gramfort~\inst{4,9} \and
         Jacques~Grobler~\inst{4} \and
@@
            University of Bonn \and
-           Nuxeo \and 
+           Independent consultant \and
            University of Bucharest \and
-           Bauhaus-Universitat \and
-           Telecom ParisTech \and
+           Ciuvo GmbH \and
+           Institut Mines-Telecom, Telecom ParisTech, CNRS LTCI \and
            University of Ballarat \and

### 2013-06-24 paper.tex
typo in @jaquesgrobler's name

@@
         Alexandre~Gramfort~\inst{4,9} \and
-        Jacques~Grobler~\inst{4} \and
+        Jaques~Grobler~\inst{4} \and
         Robert~Layton~\inst{10} \and

### 2013-06-24 paper.tex
Various small improvements

@@
 by 1309 people and forked 667 times on GitHub; the mailing list receives
-more than 300 mails per month; and version control logs
+more than 300 mails per month; version control logs
 % ddaa494c116e3c16bf032003c5cccbed851733d2
-show 183 unique contributors to the codebase.
+show 183 unique contributors to the codebase; the online documentation website
+receives 37k unique visitors and 295k page views per month.
 
@@
 
-\texttt{Pipeline} objects chain multiple estimators into a
-single one. This is useful since a machine learning workflow typically involves
-a fixed sequence of processing steps (e.g., feature extraction,
-dimensionality reduction, learning and making predictions). A sequence
-of $N$ such steps can be combined into a \texttt{Pipeline}
-if the first $N-1$ steps are transformers; the last can
-be either a predictor or a transformer. Conceptually, fitting a pipeline to a
-training set amounts to the following recursive procedure: i) when only one step
-remains, call its \texttt{fit} method; ii) otherwise, \texttt{fit} the first
-step, use it to \texttt{transform} the training set and \texttt{fit} the rest of
-the pipeline with the transformed data. The pipeline exposes all the methods the
-last estimator in the pipe exposes. That is, if the last estimator is a
-predictor, the pipeline can itself be used as a predictor. If the last estimator
-is a transformer, then the pipeline is itself a transformer.
+\texttt{Pipeline} objects chain multiple estimators into a single one. This is
+useful since a machine learning workflow typically involves a fixed sequence of
+processing steps (e.g., feature extraction, dimensionality reduction, learning
+and making predictions). A sequence of $N$ such steps can be combined into a
+\texttt{Pipeline} if the first $N-1$ steps are transformers; the last can be
+either a predictor, a transformer or both. Conceptually, fitting a pipeline to
+a training set amounts to the following recursive procedure: i) when only one
+step remains, call its \texttt{fit} method; ii) otherwise, \texttt{fit} the
+first step, use it to \texttt{transform} the training set and \texttt{fit} the
+rest of the pipeline with the transformed data. The pipeline exposes all the
+methods the last estimator in the pipe exposes. That is, if the last estimator
+is a predictor, the pipeline can itself be used as a predictor. If the last
+estimator is a transformer, then the pipeline is itself a transformer.
 
@@
 various such cross-validation schemes, including $k$-fold (default),
-leave-one-out and bootstrapping.
+leave-one-out and iterated shuffled $k$-folds with optional sub-sampling.
 The score function used by default is the \texttt{score} method
@@
 bagging or subsampling strategies and missing value completion algorithms.
-Topics like structured prediction or reinforcement learning are
-considered off-topic for the project, since they would require quite different
+Tasks like structured prediction or reinforcement learning are
+considered off-scope for the project, since they would require quite different
 data representations and APIs.

### 2013-06-24 paper.tex

--- 2013-06-24 paper.tex
@@
 Weka \citep{hall2009weka}, they bring the prime advantage of allowing us to rely
-on all NumPy and SciPy vector operations while keeping
+on efficient NumPy and SciPy vectorized operations while keeping
 the code short and readable.  This design choice has also been motivated by
@@
 thus set to the default values. Upon calling \texttt{fit}, the model is then
-learned from the training arrays \texttt{X\_train} and \texttt{y\_train}.
-(In this example, \texttt{fit} is called solely for its side effects.)
+learned from the training arrays \texttt{X\_train} and \texttt{y\_train},
+and stored within the object for later use.
+%% (In this example, \texttt{fit} is called solely for its side effects.)
 Since all estimators share the same interface, using a different learning algorithm is

### 2013-06-24 paper.tex
rework the conclusion and separate out future work

--- 2013-06-24 paper.tex
@@
 
-\section{Conclusions and future work}
-\label{sec:conclusions}
-
-We have discussed the scikit-learn API
-and the way it maps machine learning concepts and tasks
-onto objects and operations in the Python programming language.
-We motivated choices and compromises made in its design.
-We argue that the API is powerful, usable and extensible,
-as witnessed by the appearance of third-party packages
-that follow the scikit-learn conventions and tie into the toolkit.\footnote{
-  E.g.\ nolearn, \url{http://pythonhosted.org/nolearn};
-  this mention does not imply endorsement of this package.}
-While part of the scikit-learn API is necessarily Python-specific,
-core concepts may be applicable to
-machine learning applications and toolkits
-written in other (dynamic) programming languages.
-
+\section{Future Work}
 A \sklearn~1.0 release is nearing,
@@
 
+\section{Conclusions and future work}
+\label{sec:conclusions}
+
+We have discussed the scikit-learn API
+and the way it maps machine learning concepts and tasks
+onto objects and operations in the Python programming language.
+We have shown how a consistent API across the package makes scikit-learn
+very \textbf{usable} in practice: experimenting with different learning
+algorithm is as simple as substituting a new class definition.
+Through composition interfaces such as Pipelines, Feature Unions,
+and meta-estimators, these simple building blocks lead to an API which is
+\textbf{powerful}, and can accomplish a wide variety of learning tasks
+within a small amount of easy-to-read code.
+Through duck-typing, the consistent API leads to a framework that is
+easily \textbf{extensible}, and allows user-defined estimators to be
+incorporated into the scikit-learn workflow without any explicit object
+inheritance.
+
+The simplicity, power, and extensibility of the scikit-learn API is evidenced
+by the large and growing user-base, the use of scikit-learn to solve real
+problems across a wide array of fields,
+as well as the appearance of third-party packages
+that follow the scikit-learn conventions and tie into the toolkit.\footnote{
+  E.g.\ nolearn, \url{http://pythonhosted.org/nolearn};
+  this mention does not imply endorsement of this package.}
+While part of the scikit-learn API is necessarily Python-specific,
+core concepts may be applicable to
+machine learning applications and toolkits
+written in other (dynamic) programming languages.
+
 \subsection*{Acknowledgements}

### 2013-06-24 paper.tex
Fix latex comment broken by text wrapping and add missing 'and'
move 1.0 release mention to the end

@@
 \section{Future Work}
-A \sklearn~1.0 release is nearing,
-but several problems must still be resolved.
+There are several problems with the current scikit-learn interface which
+will be resolved in the future.
 At present, the library does not support some classical machine learning tasks,
@@
 so models from untrusted sources cannot be safely ``unpickled''.
+
+These API issues will be addressed in the future in preparation for
+the 1.0 release of scikit-learn.
 

### 2013-06-24 paper.tex
COSMIT Fixed my name

--- 2013-06-24 paper.tex
@@
         Fabian~Pedregosa~\inst{4} \and
-        Andreas~Mueller~\inst{5} \and
+        Andreas~C.~M\"uller~\inst{5} \and
         Olivier~Grisel~\inst{6} \and

### 2013-06-24 paper.tex
rename sections; make forward referencing correct

--- 2013-06-24 paper.tex
@@
 Section \ref{sec:comparison} then
-compares \emph{scikit-learn} to other major projects in terms of API. Finally,
-conclusions and directions of future works are included in section
-\ref{sec:conclusions}.
+compares \emph{scikit-learn} to other major projects in terms of API.
+Section \ref{sec:future_work} outlines some of the work to be done before
+the scikit-learn 1.0 release.  Finally, we summarize and conclude in
+section \ref{sec:conclusions}.
 
@@
 \section{Future Work}
+\label{sec:future_work}
 There are several problems with the current scikit-learn interface which
@@
 
-\section{Conclusions and future work}
+\section{Conclusion}
 \label{sec:conclusions}

### 2013-06-24 paper.tex

@@
 repositories and mailing list.  At the time of writing, the project is watched
-by 1309 people and forked 667 times on GitHub; the mailing list receives
-more than 300 mails per month; and version control logs
+by 1309 people and forked 667 times on GitHub; the mailing list receives more
+than 300 mails per month; version control logs
 % ddaa494c116e3c16bf032003c5cccbed851733d2
-show 183 unique contributors to the codebase.
+show 183 unique contributors to the codebase and the online documentation
+website receives 37k unique visitors and 295k page views per month.
 

### 2013-06-25 paper.tex
Fix indentation.
Packages using the scikit-learn conventions.

@@
 
+While part of the scikit-learn API is necessarily Python-specific,
+core concepts may be applicable to
+machine learning applications and toolkits
+written in other (dynamic) programming languages.
 The simplicity, power, and extensibility of the scikit-learn API is evidenced
@@
 as well as the appearance of third-party packages
-that follow the scikit-learn conventions and tie into the toolkit.\footnote{
-  E.g.\ nolearn, \url{http://pythonhosted.org/nolearn};
-  this mention does not imply endorsement of this package.}
-While part of the scikit-learn API is necessarily Python-specific,
-core concepts may be applicable to
-machine learning applications and toolkits
-written in other (dynamic) programming languages.
+that follow the scikit-learn conventions. Examples of such packages include
+\textit{astroML}\footnote{\url{http://astroml.org/}}, a package providing
+machine learning tools for astronomers and
+\textit{wiseRF}\footnote{\url{http://wise.io}}, a commercial random forest
+implementation. The source code of 
+the recently-proposed sparse multiclass
+algorithm \citep{mblondel-mlj2013}, released as
+part of the 
+\textit{lightning}\footnote{\url{https://github.com/mblondel/lightning}}
+package, also follows the scikit-learn conventions. To maximize ease of use, we
+encourage more researchers to follow the scikit-learn conventions when
+releasing their software in the future.
 

### 2013-06-25 paper.tex
add astroML reference; fix typos; update user statistics

@@
 core developers, mostly researchers from various fields (e.g., machine learning,
-natural language processing, or neuroscience). The project also benefits
+natural language processing, neuroscience, astrophysics, etc.). The project also benefits
 from many occasional contributors proposing small code or documentation fixes or
@@
 Over the past few years, the project has proved valuable to an ever increasing
-user base. Its popularity can be gaged from various indicators such as hundreds
+user base. Its popularity can be gauged from various indicators such as the hundreds
 of citations in scientific publications, successes in various machine learning
@@
 repositories and mailing list.  At the time of writing, the project is watched
-by 1309 people and forked 667 times on GitHub; the mailing list receives more
+by 1365 people and forked 693 times on GitHub; the mailing list receives more
 than 300 mails per month; version control logs
@@
 show 183 unique contributors to the codebase and the online documentation
-website receives 37k unique visitors and 295k page views per month.
+website receives 37,000 unique visitors and 295,000 page views per month.
 
@@
 that follow the scikit-learn conventions. Examples of such packages include
-\textit{astroML}\footnote{\url{http://astroml.org/}}, a package providing
+\textit{astroML}\footnote{\url{http://astroml.org/}}
+\citep{vanderplas2012astroML}, a package providing
 machine learning tools for astronomers and

### 2013-06-25 paper.tex
Smaller font size for bibliography.
Cosmit.
Add project URL and reset footnote counter.
Small improvements

@@
 
+\setcounter{footnote}{0}
+
 \section{The scikit-learn project}
 
-The \sklearn project \citep{pedregosa2011} is an open source machine learning
+The \sklearn project\footnote{\url{http://scikit-learn.org}}
+\citep{pedregosa2011} is an open source machine learning
 library for the Python programming language. The ambition of the project is to
@@
 purpose high-level language. Among other things, it includes classical learning
-algorithms, model selection tools and preprocessing procedures. The project is
+algorithms; model evalutation, selection and evaluation tools; and
+preprocessing procedures. The project is
 distributed under the simplified BSD license, encouraging its use in both
@@
 such samples.  As such, matrices of numerical values constitute a natural and
-widely used representation in machine learning. In this view, each row in the
-matrix then corresponds to a sample of the dataset and each column to one of the
-variables of the problem.
+a widely used representation in machine learning: each row in the
+matrix corresponds to one sample of the dataset and each column to one
+variable of the problem.
 
@@
 computes the coefficient of determination between \texttt{y\_test} and
-\texttt{predict(X\_test)} in the case of classification, or the mean squared
-error in the case of regression. Not however that, from an API point of view,
+\texttt{predict(X\_test)} in regression, or the accuracy
+in classification. Not however that, from an API point of view,
 the only guideline is that the \texttt{score} method returns a value
@@
 Although its stated goal is not machine learning,
-it does exemplify API design for scalable processing of ``big data''
+it exemplifies API design for scalable processing of ``big data''
 that takes a different approach than comparable models in \sklearn.
@@
 
-\section{Future Work}
+\section{Future work}
 \label{sec:future_work}
@@
 will be resolved in the future.
-At present, the library does not support some classical machine learning tasks,
+At present, the library does not support some classical machine learning
+algorithms,
 including neural networks, ensemble meta-estimators for
@@
 
+{\small
 \bibliographystyle{abbrvnat}
@@
 \bibliography{paper}
+}
 

### 2013-06-25 paper.tex
Shorten sentence

--- 2013-06-25 paper.tex
@@
 \texttt{predict(X\_test)} in regression, or the accuracy
-in classification. Not however that, from an API point of view,
+in classification. From an API point of view,
 the only guideline is that the \texttt{score} method returns a value

### 2013-06-25 paper.tex
minor rewords

--- 2013-06-25 paper.tex
@@
 purpose high-level language. Among other things, it includes classical learning
-algorithms; model evalutation, selection and evaluation tools; and
+algorithms, model evalutation and selection tools, as well as
 preprocessing procedures. The project is
@@
 it exemplifies API design for scalable processing of ``big data''
-that takes a different approach than comparable models in \sklearn.
+with an approach different from the comparable models in \sklearn.
 Gensim's way of dealing with large datasets is to use algorithms

### 2013-06-25 paper.tex
Cosmit.

@@
 it exemplifies API design for scalable processing of ``big data''
-with an approach different from the comparable models in \sklearn.
+which takes a different approach from the comparable models in \sklearn.
 Gensim's way of dealing with large datasets is to use algorithms

### 2013-06-25 paper.tex
typo
scikit-learn -> \sklearn + copyedit conclusion

@@
 purpose high-level language. Among other things, it includes classical learning
-algorithms, model evalutation and selection tools, as well as
+algorithms, model evaluation and selection tools, as well as
 preprocessing procedures. The project is
@@
 programming interface (API) and then describe, in section \ref{sec:advanced-api},
-advanced API mechanisms built from the core interface. Section
-\ref{sec:implementation} briefly describes the implementation.
+advanced API mechanisms built from the core interface.
+Section \ref{sec:implementation} briefly describes the implementation.
 Section \ref{sec:comparison} then
-compares \emph{scikit-learn} to other major projects in terms of API.
+compares \sklearn to other major projects in terms of API.
 Section \ref{sec:future_work} outlines some of the work to be done before
-the scikit-learn 1.0 release.  Finally, we summarize and conclude in
-section \ref{sec:conclusions}.
+the \sklearn 1.0 release.
+Finally, we summarize and conclude in section \ref{sec:conclusions}.
 
@@
 \label{sec:future_work}
-There are several problems with the current scikit-learn interface which
+There are several problems with the current \sklearn interface which
 will be resolved in the future.
@@
 These API issues will be addressed in the future in preparation for
-the 1.0 release of scikit-learn.
+the 1.0 release of \sklearn.
 
@@
 
-We have discussed the scikit-learn API
+We have discussed the \sklearn API
 and the way it maps machine learning concepts and tasks
 onto objects and operations in the Python programming language.
-We have shown how a consistent API across the package makes scikit-learn
+We have shown how a consistent API across the package makes \sklearn
 very \textbf{usable} in practice: experimenting with different learning
@@
 easily \textbf{extensible}, and allows user-defined estimators to be
-incorporated into the scikit-learn workflow without any explicit object
+incorporated into the \sklearn workflow without any explicit object
 inheritance.
 
-While part of the scikit-learn API is necessarily Python-specific,
+While part of the \sklearn API is necessarily Python-specific,
 core concepts may be applicable to
@@
 written in other (dynamic) programming languages.
-The simplicity, power, and extensibility of the scikit-learn API is evidenced
-by the large and growing user-base, the use of scikit-learn to solve real
+The power, and extensibility of the \sklearn API is evidenced
+by the large and growing user-base, its use to solve real
 problems across a wide array of fields,
 as well as the appearance of third-party packages
-that follow the scikit-learn conventions. Examples of such packages include
+that follow the \sklearn conventions. Examples of such packages include
 \textit{astroML}\footnote{\url{http://astroml.org/}}
@@
 the recently-proposed sparse multiclass
-algorithm \citep{mblondel-mlj2013}, released as
+algorithm of \citet{mblondel-mlj2013}, released as
 part of the
 \textit{lightning}\footnote{\url{https://github.com/mblondel/lightning}}
-package, also follows the scikit-learn conventions. To maximize ease of use, we
-encourage more researchers to follow the scikit-learn conventions when
-releasing their software in the future.
+package, also follows the \sklearn conventions.
+To maximize ease of use, we encourage researchers
+to follow these conventions when releasing machine learning software.
 

### 2013-06-25 paper.tex
comparison to VW, SofiaML, Weka (CLI tools)
Comparison section needs heavy editing.

@@
 
-\textcolor{red}{TODO: compare with weka, R, vowpal rabbit, ... ?}\\
 \textcolor{red}{TODO: emphasize scikit-learn's place in the overall scipy ecosystem, and its role in spurring the development of more basic tools.}\\
-\textcolor{red}{TODO: some packages start to adopt our own API.}
 
@@
 (both open source and proprietary) with which \sklearn competes.
+
+Many machine learning packages exist, with new ones still often appearing,
+that are intended to be used as command-line tools
+(and sometimes do not offer any type of API).
+Examples include Weka \citep{hall2009weka}, SofiaML \citep{sculley2010web}
+and Vowpal~Wabbit \citep{langford2009sparse}.
+While these packages have the advantage
+that their users are not tied to a particular programming language,
+the users will find that they still need to do programming,
+and will do so in a variety of languages.
+We argue that this makes it harder for researchers and developers
+to exchange software and collaborate on software.
 
@@
 Another package that deserves mention is the Gensim topic modeling toolkit
-\citep{rehurek2010gensim}.
-Although its stated goal is not machine learning,
-it exemplifies API design for scalable processing of ``big data''
-which takes a different approach from the comparable models in \sklearn.
+\citep{rehurek2010gensim},
+which exemplifies API design for scalable processing of ``big data''.
 Gensim's way of dealing with large datasets is to use algorithms

### 2013-06-26 paper.tex
Explain what is a scikit and mention scikit-image.
Fixes #25

@@
 these in Python/NumPy or by wrapping existing C/C{}\verb!++!/Fortran
-implementations. In a similar vein, \sklearn extends Python with a set of
-composable machine learning operations.
+implementations. The \textit{scikits} were created to complement SciPy with
+domain-specific toolkits. Currently, the two most popular and feature-complete
+ones are by far \sklearn and 
+\textit{scikit-image}\footnote{\url{http://scikit-image.org}}, an
+image-processing library.
 

### 2013-06-26 paper.tex
Clarify gensim paragraph.

@@
 \textit{generators}).
-While the text vectorization part of \sklearn
-also uses iterables to some extent,
-the emphasis is still on (mini-)batch processing,
+While the text vectorizers part of \sklearn
+also use iterables to some extent,
+they still produce entire sparse matrices, intended to be used for batch or
+mini-batch learning. This is the case
 even in the stateless, O(1) memory vectorizers
@@
 reduces various forms of overhead
-and makes effective use of the abundance of memory in modern computers.
-We make no attempt to hide this batch-oriented processing from the user,
+and allows effective use of the vectorized operations provided by NumPy and 
+SciPy.  We make no attempt to hide this batch-oriented processing from the user,
 allowing control over the amount of memory actually dedicated

### 2013-06-26 paper.tex
Improve GUI / command line comparison.

@@
 
-\section{Comparison with other packages}
-
+\section{Related software}
 \label{sec:comparison}
-
-\textcolor{red}{TODO: emphasize scikit-learn's place in the overall scipy ecosystem, and its role in spurring the development of more basic tools.}\\
 
@@
 (both open source and proprietary) with which \sklearn competes.
-
-Many machine learning packages exist, with new ones still often appearing,
-that are intended to be used as command-line tools
+Although it can also be used through its Java API, Weka \citep{hall2009weka} is
+popularly used through its graphical user interface (GUI). We argue that a GUI
+lacks flexibility, makes experiments difficult to reproduce and has a high
+maintenance cost from a development point of view. For these reasons, we make no
+attempt to provide such a GUI in \sklearn.
+Other existing machine learning packages 
+such as SofiaML \citep{sculley2010web}
+and Vowpal~Wabbit \citep{langford2009sparse}
+are intended to be used as command-line tools
 (and sometimes do not offer any type of API).
-Examples include Weka \citep{hall2009weka}, SofiaML \citep{sculley2010web}
-and Vowpal~Wabbit \citep{langford2009sparse}.
 While these packages have the advantage
 that their users are not tied to a particular programming language,
-the users will find that they still need to do programming,
+the users will find that they still need programming to process inputs/outputs,
 and will do so in a variety of languages.
 We argue that this makes it harder for researchers and developers
-to exchange software and collaborate on software.
+to exchange and collaborate on software.
 

### 2013-06-26 paper.tex
Cosmit in Matlab/R paragraph.

@@
 it has strong language and standard library support for such tasks as
-string handling/text processing, interprocess communication, networking
-and many of the other tasks that many machine learning programs
-(whether academic or commercial)
-will have to perform ``on the side''.
-In addition, third-party libraries are available for many other routine tasks.
+string/text processing, interprocess communication, networking
+and many of the other tasks that machine learning programs
+(whether academic or commercial) routinely need to perform.
+However, \textsc{matlab} and R remain the preferred language of many
+researchers. As such, much scientific software is still open-sourced primarily
+in these two languages.
 

### 2013-06-26 paper.tex
Comparison w/ Matlab/R: mention NumPy/SciPy

@@
 such as \textsc{matlab} and R \citep{r},
-the use of Python as a host language has the distinct advantage
-that it is a \textit{general purpose} language:
-it has strong language and standard library support for such tasks as
+Python has the distinct advantage
+that it is a \textit{general purpose} language, while still offering similar 
+advantages to \textsc{matlab} and R through the use of NumPy and SciPy.
+Python has strong language and standard library support for such tasks as
 string/text processing, interprocess communication, networking

### 2013-06-26 paper.tex
Cosmit.
Use project URL for sofia-ml and VW.

@@
 Other existing machine learning packages 
-such as SofiaML \citep{sculley2010web}
-and Vowpal~Wabbit \citep{langford2009sparse}
+such as SofiaML\footnote{\url{http://code.google.com/p/sofia-ml/}}
+and Vowpal~Wabbit\footnote{\url{http://hunch.net/~vw/}}
 are intended to be used as command-line tools
@@
 package, also follows the \sklearn conventions.
-To maximize ease of use, we encourage researchers
-to follow these conventions when releasing machine learning software.
+To maximize ease of use, we encourage more researchers
+to follow these conventions when releasing their software.
 

### 2013-06-26 paper.tex
Cosmit.

@@
 domain-specific toolkits. Currently, the two most popular and feature-complete
-ones are by far \sklearn and 
-\textit{scikit-image}\footnote{\url{http://scikit-image.org}}, an
-image-processing library.
+ones are by far \sklearn for machine learning and 
+\textit{scikit-image}\footnote{\url{http://scikit-image.org}} for image
+processing.
 

### 2013-06-26 paper.tex
inputs -> input, outputs -> output
I prefer to use these as uncountable nouns; did that where it looked
better than the countable usage.

@@
 performance on some input data. In supervised learning estimators, this method
-takes as inputs arrays \texttt{X\_test} and \texttt{y\_test} and typically
+takes as input arrays \texttt{X\_test} and \texttt{y\_test} and typically
 computes the coefficient of determination between \texttt{y\_test} and
@@
 algorithms are all provided as transformers within the library.  In our example,
-if one wanted to preprocess the inputs \texttt{X\_train} before fitting the
+if one wanted to preprocess the input \texttt{X\_train} before fitting the
 logistic regression estimator, e.g., by standardizing the feature values, one
@@
 that concatenates their outputs. A union of two transformers that
-map inputs having $d$ features to $d'$ and $d''$ features respectively is
+map input having $d$ features to $d'$ and $d''$ features respectively is
 a transformer that maps its $d$ input features to $d' + d''$ features.
@@
 In \sklearn, model selection is supported in two distinct meta-estimators,
-\texttt{GridSearchCV} and \texttt{RandomizedSearchCV}.  They take as inputs an
+\texttt{GridSearchCV} and \texttt{RandomizedSearchCV}.  They take as input an
 estimator (basic or composite), whose hyper-parameters must be optimized, and a
@@
 In other places of the library, in particular in the vectorization code
-for unstructured inputs, the toolkit is also designed to be
+for unstructured input, the toolkit is also designed to be
 extensible. Here, estimators provide hooks for user-defined code: objects or
@@
 that their users are not tied to a particular programming language,
-the users will find that they still need programming to process inputs/outputs,
+the users will find that they still need programming to process input/output,
 and will do so in a variety of languages.

### 2013-06-26 paper.tex
copyedit Matlab comparison

@@
 Python has the distinct advantage
-that it is a \textit{general purpose} language, while still offering similar 
-advantages to \textsc{matlab} and R through the use of NumPy and SciPy.
+that it is a \textit{general purpose} language,
+while NumPy and SciPy extend it functionality
+similar to that offered by \textsc{matlab} and R.
 Python has strong language and standard library support for such tasks as

### 2013-06-26 paper.tex
spell check using aspell

@@
 functions that provide a specific API can be given as arguments at vectorizer
-construction time. The library then calls into this code, communicating with it by passing objects of standard Python/Numpy types.
+construction time. The library then calls into this code, communicating with it by passing objects of standard Python/NumPy types.
 Again, such external user code can be kept agnostic of the \sklearn
@@
 In particular, we focus on making the codebase easily maintainable and
-understandable in order to favour external contributions. Whenever practicable,
+understandable in order to favor external contributions. Whenever practicable,
 algorithms implemented in \sklearn are written in Python,
-using Numpy vector operations for numerical work.
+using NumPy vector operations for numerical work.
 This allows for the code to remain concise, readable and
@@
 Binary distributions of these are available for the major platforms.
-Visualisation functionality depends on Matplotlib \citep{hunter2007matplotlib}
+Visualization functionality depends on Matplotlib \citep{hunter2007matplotlib}
 and/or Graphviz \citep{gansner2000},
@@
 At a lower-level, parallel processing is a potential point of improvement.
-Some estimators in \sklearn are already able to leverage multi-core processors,
+Some estimators in \sklearn are already able to leverage multicore processors,
 but only in a coarse-grained fashion.
@@
 
-\subsection*{Acknowledgements}
+\subsection*{Acknowledgments}
 
@@
 scholarships through its Summer of Code program,
-the Python Software Fundation and Tinyclues for funding coding sprints.
+the Python Software Foundation and Tinyclues for funding coding sprints.
 

### 2013-06-26 paper.tex
forgot "with"

@@
 that it is a \textit{general purpose} language,
-while NumPy and SciPy extend it functionality
+while NumPy and SciPy extend it with functionality
 similar to that offered by \textsc{matlab} and R.

### 2013-06-26 paper.tex
Fix VW URL.

@@
 such as SofiaML\footnote{\url{http://code.google.com/p/sofia-ml/}}
-and Vowpal~Wabbit\footnote{\url{http://hunch.net/~vw/}}
+and Vowpal~Wabbit\footnote{\url{http://hunch.net/$\sim$vw/}}
 are intended to be used as command-line tools

### 2013-06-26 paper.tex
Some formulation changes

@@
        Whenever an operation requires a user-defined parameter,
-       a default and appropriate value is defined by the library.
+       an appropriate default value is defined by the library.
        The default value should cause the operation to be performed
@@
 arrays and SciPy sparse matrices.
-From a practical point of view, these formats also provide a bunch of
+From a practical point of view, these formats also provide a collection of
 data loading and conversion tools which make them very easy to use in many
@@
 unsupervised learning algorithms (e.g., for classification, regression or
-clustering) are offered as objects implementing that interface. Machine
+clustering) are offered as objects implementing this interface. Machine
 learning tasks like feature extraction, feature selection or dimensionality
@@
 the same training data, one would simply have to replace
-\texttt{LogisticRegression(penalty="l1")} in the snippet above by \\
+\texttt{LogisticRegression(penalty="l1")} in the snippet above by
 \texttt{RandomForestClassifier()}.
@@
 features) or feature extraction techniques (e.g., vectorization of text
-documents) also implement the same interface. Even processing steps that are
+documents) also implement the \textit{estimator} interface. Even processing steps that are
 entirely stateless and therefore do not require the \texttt{fit} method to
@@
 present advanced API mechanisms for building meta-estimators,
-composing complex estimators and then selecting models. We also discuss design
+composing complex estimators and selecting models. We also discuss design
 choices allowing for an easy use and extension of \sklearn.
@@
 be used to combine typical machine learning workflows into a single object which
-is itself an estimator, and can be used wherever usual estimators can be used.
+is itself an estimator, and can be employed wherever usual estimators can be used.
 In particular,
@@
 constructor of an estimator
-determine the behavior of its learning algorithm
+determine the behavior of the learning algorithm
 and hence the performance of the resulting model on unseen data.
@@
 some hyper-parameter space, the best combination of hyper-parameters, with
-respect to some user-chosen criterion. For example, in the case of a decision
+respect to some user-specified criterion. For example, in the case of a decision
 tree, too small values of the \texttt{max\_depth} parameter will make the tree
@@
 expressed as NumPy operations, we rely on Cython \citep{behnel2011cython}
-to achieve decent performance and scalability. Cython is a
+to achieve competative performance and scalability. Cython is a
 compiled programming language that extends Python with static typing. It

### 2013-06-27 paper.tex
Typo.

@@
 expressed as NumPy operations, we rely on Cython \citep{behnel2011cython}
-to achieve competative performance and scalability. Cython is a
+to achieve competitive performance and scalability. Cython is a
 compiled programming language that extends Python with static typing. It

### 2013-06-27 paper.tex
Cosmit.

@@
 such samples.  As such, matrices of numerical values constitute a natural and
-a widely used representation in machine learning: each row in the
+widely used representation in machine learning: each row in the
 matrix corresponds to one sample of the dataset and each column to one

### 2013-06-27 paper.tex
Grammar

@@
 to tie a library to a suboptimal implementation strategy. As such, batch
-processing enable fast implementations in lower-level languages (where memory
+processing enables fast implementations in lower-level languages (where memory
 hierarchy effects and the possibility of internal parallelization come into

### 2013-06-27 paper.tex
Confidence.
Cosmit.

@@
 Along with the \texttt{predict} method, predictors may also implement methods
-for quantifying the certainty of a prediction. In the case of estimators
+for quantifying the confidence of a prediction. In the case of estimators
 implementing linear models, the \texttt{decision\_function} method can be used
@@
 Additionally, we see that
-\texttt{GridSearchCV} has a \texttt{predict} method, like any other classifier.
+\texttt{GridSearchCV} has a \texttt{predict} method, just like any other classifier.
 This is because \texttt{GridSearchCV} and \texttt{RandomizedSearchCV} in fact

### 2013-06-27 paper.tex
Cosmit.

@@
 extensible. Here, estimators provide hooks for user-defined code: objects or
-functions that provide a specific API can be given as arguments at vectorizer
+functions that follow a specific API can be given as arguments at vectorizer
 construction time. The library then calls into this code, communicating with it by passing objects of standard Python/NumPy types.

### 2013-06-27 paper.tex
Make subsection for the general principles.
Should hopefully address @amueller's concern.
Improving the tone in the comparisons section and the future work section

@@
 fitting models, a \textit{predictor} interface for making predictions and a
-\textit{transformer} interface for converting data. As much as possible, our
+\textit{transformer} interface for converting data. In this section, we describe
+these three interfaces, after reviewing our general principles and data
+representation choices.
+
+\subsection{General principles}
+
+As much as possible, our
 design choices have been guided so as to avoid the proliferation of framework
@@
 features) or feature extraction techniques (e.g., vectorization of text
-documents) also implement the \textit{estimator} interface. Even processing steps that are
-entirely stateless and therefore do not require the \texttt{fit} method to
-perform any useful work implement the estimator interface. As we will illustrate
+documents) also implement the \textit{estimator} interface. Even stateless
+processing steps, that do not require the \texttt{fit} method to
+perform useful work, implement the estimator interface. As we will illustrate
 in the next sections, this design pattern is indeed of prime importance for
@@
 Although it can also be used through its Java API, Weka \citep{hall2009weka} is
-popularly used through its graphical user interface (GUI). We argue that a GUI
-lacks flexibility, makes experiments difficult to reproduce and has a high
-maintenance cost from a development point of view. For these reasons, we make no
-attempt to provide such a GUI in \sklearn.
+popularly used through its graphical user interface (GUI). The target audience
+of the \sklearn is capable of programming, and therefore we focus on developing
+a usable and consistent API, rather than expend effort into creating a GUI. In
+addition, running machine learning through a GUI makes reproducibility difficult
+(although Weka has developed a GUI for managing such workflows).
+
 Other existing machine learning packages 
@@
 and will do so in a variety of languages.
-We argue that this makes it harder for researchers and developers
-to exchange and collaborate on software.
+The focus of \sklearn to incorporate the entire package into a single package still
+allows for the use of external tools to process input/output, however gives the added
+benefit of being able to run these processes within the same development environment.
+In addition, this makes it easier for researchers and developers
+to exchange and collaborate on software, as dependencies and setup are kept to a
+minimum.
 
@@
 string/text processing, interprocess communication, networking
-and many of the other tasks that machine learning programs
+and many of the other auxiliary tasks that machine learning programs
 (whether academic or commercial) routinely need to perform.
-However, \textsc{matlab} and R remain the preferred language of many
-researchers. As such, much scientific software is still open-sourced primarily
-in these two languages.
+While support for many of these tasks is improving in languages such as
+\textsc{matlab} and R, they still lag behind Python in their general purpose
+applicability.
+In many applications of machine learning these tasks, such as data access,
+data preprocessing and reporting, can be a more significant task than applying 
+the actual learning algorithm.
 
@@
 
-\section{Future work}
+\section{Future directions}
 \label{sec:future_work}
-There are several problems with the current \sklearn interface which
-will be resolved in the future.
+There are several directions that the \sklearn project
+aims to focus on in future development.
 At present, the library does not support some classical machine learning

### 2013-06-27 paper.tex
Improving the tone in the comparisons section and the future work section

@@
 features) or feature extraction techniques (e.g., vectorization of text
-documents) also implement the \textit{estimator} interface. Even stateless
-processing steps, that do not require the \texttt{fit} method to
-perform useful work, implement the estimator interface. As we will illustrate
+documents) also implement the \textit{estimator} interface. Even processing steps that are
+entirely stateless and therefore do not require the \texttt{fit} method to
+perform any useful work implement the estimator interface. As we will illustrate
 in the next sections, this design pattern is indeed of prime importance for
@@
 addition, running machine learning through a GUI makes reproducibility difficult
-(although Weka has developed a GUI for managing such workflows).
+(although WEKA has developed a GUI for managing such workflows).
 
@@
 While support for many of these tasks is improving in languages such as
-\textsc{matlab} and R, they still lag behind Python in their general purpose
+\textsc{matlab} and R, they still languish behind python in their general purpose
 applicability.
@@
 data preprocessing and reporting, can be a more significant task than applying 
-the actual learning algorithm.
+the machine learning algorithm.
+%However, \textsc{matlab} and R remain the preferred language of many
+%researchers. As such, much scientific software is still open-sourced primarily
+%in these two languages.
 
@@
 \label{sec:future_work}
-There are several directions that the \sklearn project
-aims to focus on in future development.
+There are several directions that the \sklearn contributors aim to focus on in future development.
 At present, the library does not support some classical machine learning

### 2013-06-27 paper.tex
principles: swap feasible and possible
We use standard data types whenever they suffice.

We make a more thorough tradeoff when deciding to implement things as
meta-estimators.

--- 2013-06-27 paper.tex
@@
        Datasets and hyper-parameter names and values are represented as standard
-       Python, NumPy or SciPy classes rather than new ones whenever feasible.
+       Python, NumPy or SciPy classes rather than new ones whenever possible.
        This keeps \sklearn easy to use and easy to combine with other libraries.
@@
        as meta-algorithms parametrized on other algorithms.
-       Whenever possible, such algorithms are implemented and composed from
+       Whenever feasible, such algorithms are implemented and composed from
        existing building blocks.

### 2013-06-27 paper.tex
trim grid search/randomized search section
The description of randomized search was far too long for a section
that is supposed to discuss APIs, not implementation.

@amueller, @mblondel, please check if it's still correct.

@@
 estimator (basic or composite), whose hyper-parameters must be optimized, and a
-hyper-parameter grid, which defines the values to search. Optionally, they can
-also take a cross-validation scheme and a score function.  The library provides
+set of hyperparameter settings to search through.
+This set is represented as a mapping of parameter names
+to a set of discrete choices in the case of grid search,
+which exhaustively enumerates the ``grid'' (cartesian product)
+of complete parameter choices.
+Randomized search is a smarter algorithm
+that avoids the combinatorial explosion in grid search
+by sampling a fixed number of times from its parameter distributions
+(see \citealt{bergstra2012}).
+The rest of this section focuses mostly on grid search for simplicity.
+
+Optionally, the model selection algorithms
+also take a cross-validation scheme and a score function.  \sklearn provides
 various such cross-validation schemes, including $k$-fold (default),
@@
 
-\texttt{GridSearchCV} and \texttt{RandomizedSearchCV} differ in the way they
-search the hyper-parameter grid.  \texttt{GridSearchCV} exhaustively searches
-for all hyper-parameter combinations from the grid. This scheme can be expensive
-when there are many possible hyper-parameter combinations.
-\texttt{RandomizedSearchCV}, on the other hand, only searches for a limited
-number of combinations, randomly sampled from the grid. Such scheme has been
-shown to work well in practice while considerably reducing the computational
-cost \citep{bergstra2012}. For each hyper-parameter combination and each train /
-validation split generated by the cross-validation scheme, \texttt{GridSearchCV}
-and \texttt{RandomizedSearchCV} fit the base estimator on the training set and
+For each hyper-parameter combination and each train/validation split
+generated by the cross-validation scheme, \texttt{GridSearchCV}
+and \texttt{RandomizedSearchCV} fit their base estimator on the training set and
 evaluate its performance on the validation set.  In the end, the best performing

### 2013-06-27 paper.tex
minor copyedit to grid search example

@@
 Additionally, we see that
-\texttt{GridSearchCV} has a \texttt{predict} method, just like any other classifier.
-This is because \texttt{GridSearchCV} and \texttt{RandomizedSearchCV} in fact
-delegate the \texttt{predict}, \texttt{predict\_proba}, \texttt{transform} and
+\texttt{GridSearchCV} has a \texttt{predict} method, just like any other classifier:
+it delegates the \texttt{predict}, \texttt{predict\_proba}, \texttt{transform} and
 \texttt{score} methods to the best estimator

### 2013-06-27 paper.tex
trim introduction
It ran from page 1, bottom, to page 3, top; now stops at page 2, bottom.

@@
 reusable in various scientific areas. The project is not a novel domain-specific
-language, but a library that provides machine learning idioms to a general-
-purpose high-level language. Among other things, it includes classical learning
+language, but a library that provides machine learning idioms to a
+general-purpose high-level language. Among other things, it includes classical learning
 algorithms, model evaluation and selection tools, as well as
@@
 and scientific packages centered around the NumPy and SciPy libraries. NumPy
-\citep{vanderwalt2011} augments Python with a numeric array data type and proper
-array computing primitives, while SciPy \citep{varoquaux2013scipy} extends it
-further by making common numerical operations available, either by implementing
+\citep{vanderwalt2011} augments Python with a contiguous numeric array datatype
+and fast array computing primitives,
+while SciPy \citep{varoquaux2013scipy} extends it
+further with common numerical operations, either by implementing
 these in Python/NumPy or by wrapping existing C/C{}\verb!++!/Fortran
@@
 Started in 2007, \sklearn is developed by an international team of a dozen of
-core developers, mostly researchers from various fields (e.g., machine learning,
-natural language processing, neuroscience, astrophysics, etc.). The project also benefits
-from many occasional contributors proposing small code or documentation fixes or
+core developers, mostly researchers from various fields (e.g.,
+computer science, neuroscience, astrophysics). The project also benefits
+from many occasional contributors proposing small bugfixes or
 improvements. Development proceeds on GitHub\footnote{\url{https://github.com/scikit-learn}},
@@
 specific quality guidelines, such as style consistency and unit-test coverage.
-Complete documentation and examples are required for all features,
-and major modifications must pass code review by at least two
+Documentation and examples are required for all features,
+and major changes must pass code review by at least two
 developers not involved in the implementation of the proposed change.
 
-Over the past few years, the project has proved valuable to an ever increasing
-user base. Its popularity can be gauged from various indicators such as the hundreds
+\sklearn's popularity can be gauged from various indicators such as the hundreds
 of citations in scientific publications, successes in various machine learning
-challenges (e.g., Kaggle), or significant statistics derived from our
+challenges (e.g., Kaggle), and statistics derived from our
 repositories and mailing list.  At the time of writing, the project is watched
@@
 show 183 unique contributors to the codebase and the online documentation
-website receives 37,000 unique visitors and 295,000 page views per month.
+receives 37,000 unique visitors and 295,000 pageviews per month.
 
 In a previous paper \citep{pedregosa2011}, \sklearn was briefly presented and
-benchmarked against several competitors. In this paper, we instead focus on an
-in-depth discussion and analysis of the design choices made when building the
-library. We first present in section \ref{sec:core-api} the central application
-programming interface (API) and then describe, in section \ref{sec:advanced-api},
-advanced API mechanisms built from the core interface.
+benchmarked against several competitors. In this paper, we instead present an
+in-depth analysis of design choices made when building the
+library. We first present in section~\ref{sec:core-api} the central application
+programming interface (API) and then describe, in section~\ref{sec:advanced-api},
+advanced API mechanisms built on the core interface.
 Section \ref{sec:implementation} briefly describes the implementation.
@@
 Section \ref{sec:future_work} outlines some of the work to be done before
-the \sklearn 1.0 release.
+a \sklearn 1.0 release.
 Finally, we summarize and conclude in section \ref{sec:conclusions}.

### 2013-06-27 paper.tex
actually, the focus isn't that strongly on grid search
copyedit model selection further
Replaced iterated k-fold with subsampling by stratified k-fold;
we're discussing APIs, not listing all the fancy algorithms that we have.

@@
 (see \citealt{bergstra2012}).
-The rest of this section focuses mostly on grid search for simplicity.
 
@@
 various such cross-validation schemes, including $k$-fold (default),
-leave-one-out and iterated shuffled $k$-folds with optional sub-sampling.
-The score function used by default is the \texttt{score} method
-provided by the estimator. However, the library provides a large variety of
-score functions that the user can choose from,
-including accuracy, AUC, $F_1$ score (classification),
-$R^2$ score and mean squared error (regression).
+stratified $k$-fold and leave-one-out.
+The score function used by default is the estimator's \texttt{score} method,
+but the library provides a variety of
+alternatives that the user can choose from,
+including accuracy, AUC and $F_1$ score for classification,
+$R^2$ score and mean squared error for regression.
 
@@
 when \texttt{kernel} is set to \texttt{"rbf"}.
-An SVM with a linear kernel accepts a $\gamma$ parameter,
-but actually ignores it. Computing time would thus be wasted, trying out effectively
-equivalent settings, if a single hyper-parameter grid was used instead.
+An SVM with a linear kernel accepts a $\gamma$ parameter, but ignores it,
+so using a single parameter grid would waste computing time
+trying out effectively equivalent settings.
 Additionally, we see that

### 2013-06-27 paper.tex
Non-proliferation of classes: less intimidating phrasing
s/fold/integrate/ in meta-estimators (@amueller was confused :)

@@
        Datasets and hyper-parameter names and values are represented as standard
-       Python, NumPy or SciPy classes rather than new ones whenever possible.
+       Python, NumPy or SciPy objects whenever possible.
        This keeps \sklearn easy to use and easy to combine with other libraries.
@@
 the decision to implement a meta-estimator
-rather than fold the behavior it implements
+rather than integrate the behavior it implements
 into existing estimators classes

### 2013-06-27 paper.tex
emphasize that preprocessing may involve learning

@@
 processing steps (e.g., feature extraction, dimensionality reduction, learning
-and making predictions). A sequence of $N$ such steps can be combined into a
+and making predictions), many of which perform some kind of learning.
+A sequence of $N$ such steps can be combined into a
 \texttt{Pipeline} if the first $N-1$ steps are transformers; the last can be

### 2013-06-27 paper.tex
trim to get standardscaler example before the page break

@@
 
-Along with the \texttt{predict} method, predictors may also implement methods
-for quantifying the confidence of a prediction. In the case of estimators
-implementing linear models, the \texttt{decision\_function} method can be used
-to measure the distance of the samples from the discriminating hyperplane. Some
+Apart from \texttt{predict}, predictors may also implement methods
+for quantifying the confidence of a prediction. In the case of
+linear models, the \texttt{decision\_function} method returns
+the distance of the samples to the separating hyperplane. Some
 predictors also provide a \texttt{predict\_proba} method which returns
@@
 algorithms are all provided as transformers within the library.  In our example,
-if one wanted to preprocess the input \texttt{X\_train} before fitting the
-logistic regression estimator, e.g., by standardizing the feature values, one
-could thus simply write:
+to standardize the input \texttt{X\_train} to zero mean and unit variance
+before fitting the logistic regression estimator,
+one would write:
 \begin{pythoncode}

### 2013-06-27 paper.tex
rename first section Introduction
"The scikit-learn project" also starts off the first sentence,
and the first sentence of the abstract.
split long paragraph
Cosmit.

@@
 
-\section{The scikit-learn project}
+\section{Introduction}
 
@@
 \texttt{Pipeline} if the first $N-1$ steps are transformers; the last can be
-either a predictor, a transformer or both. Conceptually, fitting a pipeline to
+either a predictor, a transformer or both.
+
+Conceptually, fitting a pipeline to
 a training set amounts to the following recursive procedure: i) when only one
@@
 bagging or subsampling strategies and missing value completion algorithms.
-Tasks like structured prediction or reinforcement learning are
-considered off-scope for the project, since they would require quite different
+However, tasks like structured prediction or reinforcement learning are
+considered off-topic for the project, since they would require quite different
 data representations and APIs.

### 2013-06-27 paper.tex
Precise used classes used in numpy, scipy and standard python for datasets, hyperparameter and values

--- 2013-06-27 paper.tex
@@
        Learning algorithms are the only objects to be of custom classes.
-       Datasets and hyper-parameter names and values are represented as standard
-       Python, NumPy or SciPy objects whenever possible.
+       Datasets are represented as standard Python list, NumPy array or
+       Scipy sparse matrices.
+       Hyper-parameter names and values are represented as standard
+       Python string or number whenever possible.
        This keeps \sklearn easy to use and easy to combine with other libraries.

### 2013-06-27 paper.tex
Emphasize that we consider a dataset a pair of matrix

@@
 between input variables $X_1, ..., X_p$, called features, and
-some output variable $Y$. A sample is then defined as a tuple of values $([x_1,
-..., x_p]^\mathrm{T}, y)$ of these variables and a dataset as a collection of
-such samples.  As such, matrices of numerical values constitute a natural and
-widely used representation in machine learning: each row in the
-matrix corresponds to one sample of the dataset and each column to one
-variable of the problem.
+some output variable $Y$. A sample is then defined as a pair of input/output
+values $([x_1,..., x_p]^\mathrm{T}, y)$. of these variables. A widely used
+representation of a dataset, a collection of
+such samples, is a set of two matrix of numerical values: one for the input
+values and one for the output values. Each row of these matrix
+corresponds to one sample of the dataset and each column to one variable of
+the problem.
 

### 2013-06-27 paper.tex
Code indentation
Explain max_depth parameter

@@
 some hyper-parameter space, the best combination of hyper-parameters, with
-respect to some user-specified criterion. For example, in the case of a decision
-tree, too small values of the \texttt{max\_depth} parameter will make the tree
-underfit, while too large values will make it overfit.
+respect to some user-specified criterion. For example, a decision
+tree with too small values of the maximal tree depth
+parameter  will lead to underfitting, while too large
+values will make it overfit.
 

### 2013-06-27 paper.tex
Typo punctuation
Consistency

@@
 some output variable $Y$. A sample is then defined as a pair of input/output
-values $([x_1,..., x_p]^\mathrm{T}, y)$. of these variables. A widely used
+values $([x_1,..., x_p]^\mathrm{T}, y)$ of these variables. A widely used
 representation of a dataset, a collection of
-such samples, is a set of two matrix of numerical values: one for the input
+such samples, is a pair of matrix with numerical values: one for the input
 values and one for the output values. Each row of these matrix

### 2013-06-27 paper.tex
s/choices/combinations/ at @mblondel's request

@@
        Learning algorithms are the only objects to be of custom classes.
-       Datasets are represented as standard Python list, NumPy array or
-       Scipy sparse matrices.
-       Hyper-parameter names and values are represented as standard
-       Python string or number whenever possible.
+       Datasets and hyper-parameter names and values are represented as standard
+       Python, NumPy or SciPy objects whenever possible.
        This keeps \sklearn easy to use and easy to combine with other libraries.
@@
 between input variables $X_1, ..., X_p$, called features, and
-some output variable $Y$. A sample is then defined as a pair of input/output
-values $([x_1,..., x_p]^\mathrm{T}, y)$ of these variables. A widely used
-representation of a dataset, a collection of
-such samples, is a pair of matrix with numerical values: one for the input
-values and one for the output values. Each row of these matrix
-corresponds to one sample of the dataset and each column to one variable of
-the problem.
+some output variable $Y$. A sample is then defined as a tuple of values $([x_1,
+..., x_p]^\mathrm{T}, y)$ of these variables and a dataset as a collection of
+such samples.  As such, matrices of numerical values constitute a natural and
+widely used representation in machine learning: each row in the
+matrix corresponds to one sample of the dataset and each column to one
+variable of the problem.
 
@@
 some hyper-parameter space, the best combination of hyper-parameters, with
-respect to some user-specified criterion. For example, a decision
-tree with too small values of the maximal tree depth
-parameter  will lead to underfitting, while too large
-values will make it overfit.
+respect to some user-specified criterion. For example, in the case of a decision
+tree, too small values of the \texttt{max\_depth} parameter will make the tree
+underfit, while too large values will make it overfit.
 
@@
 which exhaustively enumerates the ``grid'' (cartesian product)
-of complete parameter choices.
+of complete parameter combinations.
 Randomized search is a smarter algorithm

### 2013-06-27 paper.tex
trim introduction back to a page and a bit

@@
 domain-specific toolkits. Currently, the two most popular and feature-complete
-ones are by far \sklearn for machine learning and 
+ones are by far \sklearn and
 \textit{scikit-image}\footnote{\url{http://scikit-image.org}} for image
@@
 
-In a previous paper \citep{pedregosa2011}, \sklearn was briefly presented and
-benchmarked against several competitors. In this paper, we instead present an
-in-depth analysis of design choices made when building the
-library. We first present in section~\ref{sec:core-api} the central application
+\citet{pedregosa2011} briefly presented \sklearn and
+benchmarked it against several competitors.
+In this paper, we instead present an
+in-depth analysis of design choices made when building the library,
+detailing how we organized and operationalized
+common machine learning concepts.
+We first present in section~\ref{sec:core-api} the central application
 programming interface (API) and then describe, in section~\ref{sec:advanced-api},
@@
 a \sklearn 1.0 release.
-Finally, we summarize and conclude in section \ref{sec:conclusions}.
+We summarize and conclude in section \ref{sec:conclusions}.
 

### 2013-06-27 paper.tex
clarify that sklearn is not an image processing library

@@
 ones are by far \sklearn and
-\textit{scikit-image}\footnote{\url{http://scikit-image.org}} for image
-processing.
+\textit{scikit-image},\footnote{\url{http://scikit-image.org}}
+which does image processing.
 
@@
        Learning algorithms are the only objects to be of custom classes.
-       Datasets and hyper-parameter names and values are represented as standard
-       Python, NumPy or SciPy objects whenever possible.
+       Datasets are represented as NumPy arrays or SciPy sparse matrices.
+       Hyper-parameter names and values are represented as standard
+       Python strings or numbers whenever possible.
        This keeps \sklearn easy to use and easy to combine with other libraries.
@@
 between input variables $X_1, ..., X_p$, called features, and
-some output variable $Y$. A sample is then defined as a tuple of values $([x_1,
-..., x_p]^\mathrm{T}, y)$ of these variables and a dataset as a collection of
-such samples.  As such, matrices of numerical values constitute a natural and
-widely used representation in machine learning: each row in the
-matrix corresponds to one sample of the dataset and each column to one
-variable of the problem.
+some output variable $Y$. A sample is then defined as a pair of
+values $([x_1,..., x_p]^\mathrm{T}, y)$ of these variables. A widely used
+representation of a dataset, a collection of
+such samples, is a pair of matrix with numerical values: one for the input
+values and one for the output values. Each row of these matrix
+corresponds to one sample of the dataset and each column to one variable of
+the problem.
 
@@
 some hyper-parameter space, the best combination of hyper-parameters, with
-respect to some user-specified criterion. For example, in the case of a decision
-tree, too small values of the \texttt{max\_depth} parameter will make the tree
-underfit, while too large values will make it overfit.
+respect to some user-specified criterion. For example, a decision
+tree with too small a value for the maximal tree depth
+parameter will tend to underfit, while too large a value will make it overfit.
 

### 2013-06-27 paper.tex
copyedit/trim: get scaler example back on its page

@@
 In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
-setting the \texttt{penalty} hyper-parameter to \texttt{"l1"}, hence enabling
-$\ell_1$ regularization. Other hyper-parameters (such as \texttt{C}, which
-defines the strength of the regularization) are not explicitly defined and are
-thus set to the default values. Upon calling \texttt{fit}, the model is then
+setting the \texttt{penalty} hyper-parameter to \texttt{"l1"} for
+$\ell_1$ regularization. Other hyper-parameters (such as \texttt{C},
+the strength of the regularization) are not explicitly given and
+thus set to the default values. Upon calling \texttt{fit}, a model is
 learned from the training arrays \texttt{X\_train} and \texttt{y\_train},
@@
 Since all estimators share the same interface, using a different learning algorithm is
-as simple as replacing the constructor (the class name).
-For example, to build a random forest on
-the same training data, one would simply have to replace
+as simple as replacing the constructor (the class name);
+to build a random forest on
+the same data, one would simply replace
 \texttt{LogisticRegression(penalty="l1")} in the snippet above by
@@
 \texttt{predict(X\_test)} in regression, or the accuracy
-in classification. From an API point of view,
-the only guideline is that the \texttt{score} method returns a value
+in classification.
+The only requirement is that the \texttt{score} method return a value
 that quantifies the quality of its predictions (the higher, the better).

### 2013-06-27 paper.tex
copyedit
* Introduce supervised learning, classification and regression
* Explain rbf
* Scikit-learn is a library, not a framework

@@
 
-\title{ API design for machine learning software: experiences from the
+\title{API design for machine learning software: experiences from the
 scikit-learn project}
@@
 
-In itself, the project is a programming library, i.e. a collection of classes
+In itself, the project is a programming library, i.e.\ a collection of classes
 and functions that users import into Python programs. Using \sklearn therefore
@@
 advanced API mechanisms built on the core interface.
-Section \ref{sec:implementation} briefly describes the implementation.
-Section \ref{sec:comparison} then
-compares \sklearn to other major projects in terms of API.
-Section \ref{sec:future_work} outlines some of the work to be done before
+Section~\ref{sec:implementation} briefly describes the implementation.
+Section~\ref{sec:comparison} then
+compares \sklearn to other major projects in terms of API\@.
+Section~\ref{sec:future_work} outlines some of the work to be done before
 a \sklearn 1.0 release.
-We summarize and conclude in section \ref{sec:conclusions}.
+We summarize and conclude in section~\ref{sec:conclusions}.
 
@@
 In most machine learning tasks, data is modeled as a set of variables.  For
-example, in a classification or regression task, the goal is to find a mapping
-between input variables $X_1, ..., X_p$, called features, and
-some output variable $Y$. A sample is then defined as a pair of
-values $([x_1,..., x_p]^\mathrm{T}, y)$ of these variables. A widely used
+example, in a supervised learning task, the goal is to find a mapping
+from input variables $X_1, \ldots X_p$, called features, to
+some output variables $Y$. A sample is then defined as a pair of
+values $([x_1, \ldots, x_p]^\mathrm{T}, y)$ of these variables. A widely used
 representation of a dataset, a collection of
-such samples, is a pair of matrix with numerical values: one for the input
-values and one for the output values. Each row of these matrix
+such samples, is a pair of matrices with numerical values: one for the input
+values and one for the output values. Each row of these matrices
 corresponds to one sample of the dataset and each column to one variable of
@@
 that quantifies the quality of its predictions (the higher, the better).
-Unsupervised learning estimators may also exposes a \texttt{score} function
+Unsupervised learning estimators may also expose a \texttt{score} function
 to compute, for instance, the likelihood of the given data under its model.
@@
 considered. This is because it only makes sense to search for \texttt{gamma}
-when \texttt{kernel} is set to \texttt{"rbf"}.
-An SVM with a linear kernel accepts a $\gamma$ parameter, but ignores it,
+when \texttt{kernel} is set to a radial basis function kernel \texttt{"rbf"}.
+An SVM with a linear kernel accepts a \texttt{gamma} parameter, but ignores it,
 so using a single parameter grid would waste computing time
@@
 popularly used through its graphical user interface (GUI). The target audience
-of the \sklearn is capable of programming, and therefore we focus on developing
-a usable and consistent API, rather than expend effort into creating a GUI. In
+of \sklearn is capable of programming, and therefore we focus on developing
+a usable and consistent API, rather than expend effort into creating a GUI\@. In
 addition, running machine learning through a GUI makes reproducibility difficult
@@
 and will do so in a variety of languages.
-The focus of \sklearn to incorporate the entire package into a single package still
+The focus of \sklearn to incorporate the entire process into a single package still
 allows for the use of external tools to process input/output, however gives the added
@@
 within a small amount of easy-to-read code.
-Through duck-typing, the consistent API leads to a framework that is
+Through duck-typing, the consistent API leads to a library that is
 easily \textbf{extensible}, and allows user-defined estimators to be
@@
 
-The authors and contributors acknowledge active support from INRIA. Past and
+The authors and contributors acknowledge active support from INRIA\@. Past and
 present sponsors of the project also include Google for funding

### 2013-06-27 paper.tex
Introduce supervised learning, classification and regression

--- 2013-06-27 paper.tex
@@
 In most machine learning tasks, data is modeled as a set of variables.  For
-example, in a classification or regression task, the goal is to find a mapping
+example in supervised learning task, the goal is to find a mapping
 between input variables $X_1, ..., X_p$, called features, and
-some output variable $Y$. A sample is then defined as a pair of
+some discrete as in classification task or continous as in regression task
+output variables $Y$. A sample is then defined as a pair of
 values $([x_1,..., x_p]^\mathrm{T}, y)$ of these variables. A widely used
 representation of a dataset, a collection of
-such samples, is a pair of matrix with numerical values: one for the input
-values and one for the output values. Each row of these matrix
+such samples, is a pair of matrices with numerical values: one for the input
+values and one for the output values. Each row of these matrices
 corresponds to one sample of the dataset and each column to one variable of

### 2013-06-27 paper.tex
Typo

@@
 which now serves as a model of its input and can be used to perform predictions or transformations of input data.
-The choice to let a single object serve dual purpose as estimator and model
+The choice to let a single object serves dual purpose as estimator and model
 simplifies the \sklearn class hierarchy

### 2013-06-27 paper.tex
R2 score => coefficient of determination

@@
 including accuracy, AUC and $F_1$ score for classification,
-$R^2$ score and mean squared error for regression.
+coefficient of determination and mean squared error for regression.
 

### 2013-06-27 paper.tex
Write in plain letter what is rbf

--- 2013-06-27 paper.tex
@@
 In most machine learning tasks, data is modeled as a set of variables.  For
-example in supervised learning task, the goal is to find a mapping
-between input variables $X_1, ..., X_p$, called features, and
-some discrete as in classification task or continous as in regression task
-output variables $Y$. A sample is then defined as a pair of
-values $([x_1,..., x_p]^\mathrm{T}, y)$ of these variables. A widely used
+example, in a classification or regression task, the goal is to find a mapping
+between input variables $X_1, \ldots X_p$, called features, and
+some output variable $Y$. A sample is then defined as a pair of
+values $([x_1, \ldots, x_p]^\mathrm{T}, y)$ of these variables. A widely used
 representation of a dataset, a collection of
-such samples, is a pair of matrices with numerical values: one for the input
-values and one for the output values. Each row of these matrices
+such samples, is a pair of matrix with numerical values: one for the input
+values and one for the output values. Each row of these matrix
 corresponds to one sample of the dataset and each column to one variable of
@@
 of the \sklearn is capable of programming, and therefore we focus on developing
-a usable and consistent API, rather than expend effort into creating a GUI. In
+a usable and consistent API, rather than expend effort into creating a GUI\@. In
 addition, running machine learning through a GUI makes reproducibility difficult

### 2013-06-27 paper.tex

@@
 which now serves as a model of its input and can be used to perform predictions or transformations of input data.
-The choice to let a single object serves dual purpose as estimator and model
+The choice to let a single object serve dual purpose as estimator and model
 simplifies the \sklearn class hierarchy

### 2013-06-28 paper.tex
the scikit-learn -> scikit-learn

--- 2013-06-27 paper.tex
@@
 that quantifies the quality of its predictions (the higher, the better).
-Unsupervised learning estimators may also expose a \texttt{score} function
+Unsupervised learning estimators may also exposes a \texttt{score} function
 to compute, for instance, the likelihood of the given data under its model.
@@
 including accuracy, AUC and $F_1$ score for classification,
-coefficient of determination and mean squared error for regression.
+$R^2$ score and mean squared error for regression.
 
@@
 considered. This is because it only makes sense to search for \texttt{gamma}
-when \texttt{kernel} is set to a radial basis function kernel \texttt{"rbf"}.
-An SVM with a linear kernel accepts a \texttt{gamma} parameter, but ignores it,
+when \texttt{kernel} is set to \texttt{"rbf"}.
+An SVM with a linear kernel accepts a $\gamma$ parameter, but ignores it,
 so using a single parameter grid would waste computing time

### 2013-06-28 paper.tex
copyedit/trim double parameter grid part, again

@@
 In most machine learning tasks, data is modeled as a set of variables.  For
-example, in a classification or regression task, the goal is to find a mapping
-between input variables $X_1, \ldots X_p$, called features, and
-some output variable $Y$. A sample is then defined as a pair of
+example, in a supervised learning task, the goal is to find a mapping
+from input variables $X_1, \ldots X_p$, called features, to
+some output variables $Y$. A sample is then defined as a pair of
 values $([x_1, \ldots, x_p]^\mathrm{T}, y)$ of these variables. A widely used
 representation of a dataset, a collection of
-such samples, is a pair of matrix with numerical values: one for the input
-values and one for the output values. Each row of these matrix
+such samples, is a pair of matrices with numerical values: one for the input
+values and one for the output values. Each row of these matrices
 corresponds to one sample of the dataset and each column to one variable of
@@
 \end{pythoncode}
-Note that, in the above example, two distinct hyper-parameter grids are
-considered. This is because it only makes sense to search for \texttt{gamma}
-when \texttt{kernel} is set to \texttt{"rbf"}.
-An SVM with a linear kernel accepts a $\gamma$ parameter, but ignores it,
+In this example, two distinct hyper-parameter grids are
+considered for the linear and radial basis function (RBF) kernels;
+an SVM with a linear kernel accepts a $\gamma$ parameter, but ignores it,
 so using a single parameter grid would waste computing time

### 2013-06-28 paper.tex
copyedit Comparison section

@@
 and will do so in a variety of languages.
-The focus of \sklearn to incorporate the entire process into a single package still
-allows for the use of external tools to process input/output, however gives the added
-benefit of being able to run these processes within the same development environment.
-In addition, this makes it easier for researchers and developers
+By contrast, \sklearn allows users to implement that entire workflow
+in a single program, written in a single language,
+and developed in a single working environment.
+This also makes it easier for researchers and developers
 to exchange and collaborate on software, as dependencies and setup are kept to a
@@
 
-In comparison to specialized languages for numeric and statistical programming
-such as \textsc{matlab} and R \citep{r},
-Python has the distinct advantage
+Similar benefits hold in the case of specialized languages
+for numeric and statistical programming
+such as \textsc{matlab} and R \citep{r}.
+In comparison to these, though, Python has the distinct advantage
 that it is a \textit{general purpose} language,
@@
 
-Another package that deserves mention is the Gensim topic modeling toolkit
+Within the realm of Python,
+a package that deserves mention is the Gensim topic modeling toolkit
 \citep{rehurek2010gensim},
-which exemplifies API design for scalable processing of ``big data''.
+which exemplifies a different style of API design
+geared toward scalable processing of ``big data''.
 Gensim's way of dealing with large datasets is to use algorithms

### 2013-06-28 paper.tex
my institute and Gilles' funding

@@
 
-\institute{Informatics Institute, University of Amsterdam \and
+\institute{ILPS, Informatics Institute, University of Amsterdam \and
            University of Liège \and
@@
 
+Gilles~Louppe is a research fellow of the Belgian
+Fonds~de~la~Recherche~Scientifique (FNRS)
+and acknowledges its financial support.
+
 {\small

### 2013-06-28 paper.tex
missing comma spotted by @vene

@@
 by sampling a fixed number of times from its parameter distributions
-(see \citealt{bergstra2012}).
+(see \citealp{bergstra2012}).
 

### 2013-06-28 paper.tex
off-topic -> out of scope (@ogrisel + @amueller)

@@
 However, tasks like structured prediction or reinforcement learning are
-considered off-topic for the project, since they would require quite different
-data representations and APIs.
+considered out of scope for the project,
+since they would require quite different data representations and APIs.
 

### 2013-06-28 paper.tex
@arjoly's acks
abstract and intro: separate project and library

@@
 \begin{abstract}
-The \sklearn project is an increasingly popular machine learning
+\sklearn is an increasingly popular machine learning
 library. Written in Python, it is designed to be simple and efficient, accessible to
@@
 The \sklearn project\footnote{\url{http://scikit-learn.org}}
-\citep{pedregosa2011} is an open source machine learning
+\citep{pedregosa2011} provides an open source machine learning
 library for the Python programming language. The ambition of the project is to
@@
 algorithms, model evaluation and selection tools, as well as
-preprocessing procedures. The project is
+preprocessing procedures. The library is
 distributed under the simplified BSD license, encouraging its use in both
@@
 
-In itself, the project is a programming library, i.e.\ a collection of classes
+\sklearn is a library, i.e.\ a collection of classes
 and functions that users import into Python programs. Using \sklearn therefore
@@
 
-From its core, the library has been designed to tie in with the set of numeric
+The library has been designed to tie in with the set of numeric
 and scientific packages centered around the NumPy and SciPy libraries. NumPy
@@
 
-Gilles~Louppe is a research fellow of the Belgian
-Fonds~de~la~Recherche~Scientifique (FNRS)
-and acknowledges its financial support.
+Gilles~Louppe and Arnaud~Joly are research fellows of the Belgian
+Fonds de la Recherche Scientifique (FNRS)
+and acknowledge its financial support.
 

### 2013-07-19 paper.tex
fix typo spotted by reviewer

@@
 simplifies the \sklearn class hierarchy
-compared to an approach were estimators and models are separated.
+compared to an approach where estimators and models are separated.
 

### 2013-07-30 paper.tex
explain why estimators serve as models
Third reviewer wanted more extensive comparison with alternative
approaches.

--- 2013-06-28 paper.tex
@@
 \texttt{fit} is a function from data to a model of that data.
-As such, it always returns the estimator object it was called on,
+It always returns the estimator object it was called on,
 which now serves as a model of its input and can be used to perform predictions or transformations of input data.
+
 The choice to let a single object serve dual purpose as estimator and model
-simplifies the \sklearn class hierarchy
-compared to an approach were estimators and models are separated.
+simplifies the \sklearn API
+compared to an approach where estimators and models are separated,
+forming two parallel class hierarchies.
+While from a mathematical point of view,
+models are more similar to one another than algorithms,
+we argue that this is less true in practical machine learning.
+As an example, consider the case of linear models,
+described mathematically by a weight matrix $w$ and an intercept vector $b$,
+suggesting a class \texttt{LinearModel}
+with such a matrix and vector as attributes.
+However, to use such a model in practice,
+one first needs to know whether the task at hand is
+regression, $y = w^\top x + b$,
+or classification, $y = \arg \max_k[w_{:,k}^\top x + b_k]$,
+calling for two classes.
+Then, the various linear probabilistic classifiers must be distuingished:
+logistic regression, with $P(y \mid x) = \sigma(w^\top x + b)$,
+multinomial naïve Bayes, with $\log P(y \mid x) = w^\top x + b$,
+and the winnow probability estimator \citep{zhang2002}, with
+$P(y \mid x) = \frac{1}{2} (1 + \mathrm{median} \{-1, w^\top x + b, 1 \})$,
+would all need either distinct classes,
+or a common class that remembers the training regime.
+Furthermore, separating estimators and models
+would make online/out-of-core learning
+(not further discussed in the present paper) more complicated.
 

### 2013-08-05 paper.tex
Rewording to better introduce scikits
Places the footnote with the word, not the comma after it

--- 2013-07-19 paper.tex
@@
 these in Python/NumPy or by wrapping existing C/C{}\verb!++!/Fortran
-implementations. The \textit{scikits} were created to complement SciPy with
+implementations. Building upon this stack, a series of libraries called
+\textit{scikits} were created, to complement SciPy with
 domain-specific toolkits. Currently, the two most popular and feature-complete
 ones are by far \sklearn and
-\textit{scikit-image},\footnote{\url{http://scikit-image.org}}
+\textit{scikit-image}\footnote{\url{http://scikit-image.org}},
 which does image processing.

### 2013-08-05 paper.tex
Typo

@@
 ones are by far \sklearn and
-\textit{scikit-image}\footnote{\url{http://scikit-image.org}},
+\textit{scikit-image},\footnote{\url{http://scikit-image.org}}
 which does image processing.
 
-Started in 2007, \sklearn is developed by an international team of a dozen of
+Started in 2007, \sklearn is developed by an international team of over a dozen
 core developers, mostly researchers from various fields (e.g.,

### 2013-08-05 paper.tex
Typo

--- 2013-08-05 paper.tex
@@
 
-Started in 2007, \sklearn is developed by an international team of a dozen of
+Started in 2007, \sklearn is developed by an international team of over a dozen of
 core developers, mostly researchers from various fields (e.g.,

### 2013-08-05 paper.tex
Fixed the grammar
Formalised the wording

--- 2013-08-05 paper.tex
@@
 ones are by far \sklearn and
-\textit{scikit-image},\footnote{\url{http://scikit-image.org}}
+\textit{scikit-image}\footnote{\url{http://scikit-image.org}},
 which does image processing.
@@
 compares \sklearn to other major projects in terms of API\@.
-Section~\ref{sec:future_work} outlines some of the work to be done before
+Section~\ref{sec:future_work} outlines some of the objectives and goals for
 a \sklearn 1.0 release.

### 2013-08-05 paper.tex
As it was, this sentence gave little information

--- 2013-08-05 paper.tex
@@
 compares \sklearn to other major projects in terms of API\@.
-Section~\ref{sec:future_work} outlines some of the work to be done before
+Section~\ref{sec:future_work} outlines some of the objectives for
 a \sklearn 1.0 release.
-We summarize and conclude in section~\ref{sec:conclusions}.
+We conclude by summarizing the major points of this paper in
+section~\ref{sec:conclusions}.
 

### 2013-08-05 paper.tex
As it was, this sentence gave little information
Grammar

--- 2013-08-05 paper.tex
@@
 a \sklearn 1.0 release.
-We summarize and conclude in section~\ref{sec:conclusions}.
+We conclude by summarizing the major points of this paper in
+section~\ref{sec:conclusions}.
 
@@
   \item[Non-proliferation of classes.]
-       Learning algorithms are the only objects to be of custom classes.
+       Learning algorithms are the only objects to be represented using custom classes.
        Datasets are represented as NumPy arrays or SciPy sparse matrices.

### 2013-08-05 paper.tex
Grammar
I don't think any parameter set works for 'most' cases

--- 2013-08-05 paper.tex
@@
 for all built-in estimators.
-These default values are set to be relevant in most common
+These default values are set to be relevant in many common
 situations in order to make estimators as effective as possible

### 2013-08-05 paper.tex
Comma not needed

--- 2013-08-05 paper.tex
@@
 Actual learning is performed by the \texttt{fit} method. This method is called
-with training data (e.g., supplied as two arrays \texttt{X\_train} and
+with training data (e.g. supplied as two arrays \texttt{X\_train} and
 \texttt{y\_train} in supervised learning estimators). Its task is to run a

### 2013-08-05 paper.tex
Comma not needed

@@
 parameters learned by an estimator are exposed as public attributes with names
-suffixed with a trailing underscore (e.g., \texttt{coef\_} for the
+suffixed with a trailing underscore (e.g. \texttt{coef\_} for the
 learned coefficients of a linear model),

### 2013-08-05 paper.tex
Slightly better wording
Another e.g. comma

@@
 composing complex estimators and selecting models. We also discuss design
-choices allowing for an easy use and extension of \sklearn.
+choices allowing for the easy usage and extension of \sklearn.
 
@@
 Examples include ensemble methods which
-build and combine several simpler models (e.g., decision trees), or multiclass
+build and combine several simpler models (e.g. decision trees), or multiclass
 and multilabel classification schemes which can be used to turn a binary

### 2013-08-05 paper.tex
Rewording

--- 2013-08-05 paper.tex
@@
 ones are by far \sklearn and
-\textit{scikit-image}\footnote{\url{http://scikit-image.org}},
+\textit{scikit-image},\footnote{\url{http://scikit-image.org}}
 which does image processing.
@@
 compares \sklearn to other major projects in terms of API\@.
-Section~\ref{sec:future_work} outlines some of the objectives and goals for
+Section~\ref{sec:future_work} outlines some of the objectives for
 a \sklearn 1.0 release.
@@
 Actual learning is performed by the \texttt{fit} method. This method is called
-with training data (e.g. supplied as two arrays \texttt{X\_train} and
+with training data (e.g., supplied as two arrays \texttt{X\_train} and
 \texttt{y\_train} in supervised learning estimators). Its task is to run a
@@
 composing complex estimators and selecting models. We also discuss design
-choices allowing for an easy use and extension of \sklearn.
+choices allowing for easy usage and extension of \sklearn.
 
@@
 In particular,
-they integrate within model selection routines, allowing one to optimize at once
-every step of a complex workflow. Composition of estimators can be done in two
+they integrate within model selection routines, allowing one to optimize
+every step of a complex workflow at once. Composition of estimators can be done in two
 ways: either sequentially through \texttt{Pipeline} objects or in a parallel

### 2013-08-05 paper.tex
Wording

--- 2013-08-05 paper.tex
@@
 and/or Graphviz \citep{gansner2000},
-but neither is required to do machine learning.
+however neither is required to do machine learning.
 When feasible, external libraries are integrated into the codebase.

### 2013-08-05 paper.tex
'to do machine learning' was a bit casual

@@
 and/or Graphviz \citep{gansner2000},
-however neither is required to do machine learning.
+however neither is required to install or use \sklearn.
 When feasible, external libraries are integrated into the codebase.

### 2013-08-05 paper.tex
Wording

--- 2013-08-05 paper.tex
@@
 is itself an estimator, and can be employed wherever usual estimators can be used.
-In particular,
-they integrate within model selection routines, allowing one to optimize
-every step of a complex workflow at once. Composition of estimators can be done in two
-ways: either sequentially through \texttt{Pipeline} objects or in a parallel
+In particular, \sklearn's model selection routines
+can be applied to composite estimators, allowing global optimization
+of all parameters in a complex workflow.
+Composition of estimators can be done in two
+ways: either sequentially through \texttt{Pipeline} objects, or in a parallel
 fashion through \texttt{FeatureUnion} objects.
@@
 and/or Graphviz \citep{gansner2000},
-but neither is required to do machine learning.
+but neither is required to perform machine learning or prediction.
 When feasible, external libraries are integrated into the codebase.
@@
 geared toward scalable processing of ``big data''.
-Gensim's way of dealing with large datasets is to use algorithms
+Gensim's method of dealing with large datasets is to use algorithms
 that have $O(1)$ space complexity and can be updated online.

### 2013-08-05 paper.tex
rm pleonasm

@@
 a \sklearn 1.0 release.
-We conclude by summarizing the major points of this paper in
-section~\ref{sec:conclusions}.
+We summarize and conclude in section~\ref{sec:conclusions}.
 
@@
   \item[Non-proliferation of classes.]
-       Learning algorithms are the only objects to be represented using custom classes.
+       Learning algorithms are the only objects to be of custom classes.
        Datasets are represented as NumPy arrays or SciPy sparse matrices.
@@
 for all built-in estimators.
-These default values are set to be relevant in many common
+These default values are set to be relevant in most common
 situations in order to make estimators as effective as possible
@@
 composing complex estimators and selecting models. We also discuss design
-choices allowing for easy usage and extension of \sklearn.
+choices allowing for an easy use and extension of \sklearn.
 
@@
 is itself an estimator, and can be employed wherever usual estimators can be used.
-In particular, \sklearn's model selection routines
-can be applied to composite estimators, allowing global optimization
-of all parameters in a complex workflow.
-Composition of estimators can be done in two
-ways: either sequentially through \texttt{Pipeline} objects, or in a parallel
+In particular,
+they integrate within model selection routines, allowing one to optimize at once
+every step of a complex workflow. Composition of estimators can be done in two
+ways: either sequentially through \texttt{Pipeline} objects or in a parallel
 fashion through \texttt{FeatureUnion} objects.
@@
 and/or Graphviz \citep{gansner2000},
-but neither is required to perform machine learning or prediction.
+but neither is required to do machine learning.
 When feasible, external libraries are integrated into the codebase.
@@
 geared toward scalable processing of ``big data''.
-Gensim's method of dealing with large datasets is to use algorithms
+Gensim's way of dealing with large datasets is to use algorithms
 that have $O(1)$ space complexity and can be updated online.

### 2013-08-05 paper.tex
missing comma

@@
 \citep{vanderplas2012astroML}, a package providing
-machine learning tools for astronomers and
+machine learning tools for astronomers, and
 \textit{wiseRF}\footnote{\url{http://wise.io}}, a commercial random forest

### 2013-08-05 paper.tex
trim to get scaling example back on one page

@@
 Apart from \texttt{predict}, predictors may also implement methods
-for quantifying the confidence of a prediction. In the case of
+that quantify the confidence of predictions. In the case of
 linear models, the \texttt{decision\_function} method returns
-the distance of the samples to the separating hyperplane. Some
+the distance of samples to the separating hyperplane. Some
 predictors also provide a \texttt{predict\_proba} method which returns
-class probabilities for the input samples \texttt{X\_test}.
+class probabilities.
 
-Finally, predictors should also provide a \texttt{score} function to assess the
-performance on some input data. In supervised learning estimators, this method
+Finally, predictors must provide a \texttt{score} function to assess their
+performance on a batch of input data. In supervised estimators, this method
 takes as input arrays \texttt{X\_test} and \texttt{y\_test} and typically
@@
 that quantifies the quality of its predictions (the higher, the better).
-Unsupervised learning estimators may also expose a \texttt{score} function
+An unsupervised estimator may also expose a \texttt{score} function
 to compute, for instance, the likelihood of the given data under its model.

### 2013-08-09 paper.tex
Summarize discussion of #34

--- 2013-08-05 paper.tex
@@
 these in Python/NumPy or by wrapping existing C/C{}\verb!++!/Fortran
-implementations. Building upon this stack, a series of libraries called
-\textit{scikits} were created, to complement SciPy with
+implementations. The \textit{scikits} were created to complement SciPy with
 domain-specific toolkits. Currently, the two most popular and feature-complete
@@
 
-Started in 2007, \sklearn is developed by an international team of over a dozen
+Started in 2007, \sklearn is developed by an international team of a dozen of
 core developers, mostly researchers from various fields (e.g.,
@@
 compares \sklearn to other major projects in terms of API\@.
-Section~\ref{sec:future_work} outlines some of the objectives for
+Section~\ref{sec:future_work} outlines some of the work to be done before
 a \sklearn 1.0 release.
-We conclude by summarizing the major points of this paper in
-section~\ref{sec:conclusions}.
+We summarize and conclude in section~\ref{sec:conclusions}.
 
@@
 \texttt{fit} is a function from data to a model of that data.
-As such, it always returns the estimator object it was called on,
+It always returns the estimator object it was called on,
 which now serves as a model of its input and can be used to perform predictions or transformations of input data.
-The choice to let a single object serve dual purpose as estimator and model
-simplifies the \sklearn class hierarchy
-compared to an approach where estimators and models are separated.
+
+From the start, the choice to let a single object serve dual purpose as
+estimator and model has mostly been driven by usability and technical
+considerations. From the user point of view, having two coupled instances (i.e.,
+an estimator object, used as a factory, and a model object, produced by the
+estimator) indeed decreases the ease of use and is also more likely to
+unecessarily confuse newcomers. From the developer point of view, decoupling
+estimators from models also creates parallel class hierarchies and increases the
+overall maintenance complexity of the project. For these practical reasons, we
+believe that decoupling estimators from models is not worth the effort. The good
+reason for decoupling however, would be that it makes it possible to ship a
+model in a new environment without having to deal with potentially complex
+software dependencies. Such a feature could however still be implemented in
+\sklearn by making estimators able to export a fitted model to an agnostic model
+description such as PMML.
 

### 2013-08-09 paper.tex
Typo

@@
 \begin{pythoncode}
-from sklearn.linear_model import LogisticRegression
+from clean.linear_model import LogisticRegression
 
@@
 In \sklearn, classical learning algorithms are not the only objects to be
-implemented as estimators. For example, preprocessing routines (e.g., scaling of
+implemented as estimators. For examplee, preprocessing routines (e.g., scaling of
 features) or feature extraction techniques (e.g., vectorization of text

### 2013-08-09 paper.tex
Added reference to PMML

@@
 estimator) indeed decreases the ease of use and is also more likely to
-unecessarily confuse newcomers. From the developer point of view, decoupling
+unnecessarily confuse newcomers. From the developer point of view, decoupling
 estimators from models also creates parallel class hierarchies and increases the
@@
 \sklearn by making estimators able to export a fitted model to an agnostic model
-description such as PMML.
+description such as PMML~\citep{guazzelli2009pmml}.
 

### 2013-08-09 paper.tex
Address larsmans comment

@@
 overall maintenance complexity of the project. For these practical reasons, we
-believe that decoupling estimators from models is not worth the effort. The good
+believe that decoupling estimators from models is not worth the effort. A good
 reason for decoupling however, would be that it makes it possible to ship a
@@
 software dependencies. Such a feature could however still be implemented in
-\sklearn by making estimators able to export a fitted model to an agnostic model
-description such as PMML~\citep{guazzelli2009pmml}.
+\sklearn by making estimators able to export a fitted model, using the
+information from its public attributes, to an agnostic model description such as
+PMML~\citep{guazzelli2009pmml}.
 

### 2013-08-09 paper.tex
Revert 7fed4161ba317fedc63b79ee3f34ce946246f5fa (geh?!)

@@
 \begin{pythoncode}
-from clean.linear_model import LogisticRegression
+from sklearn.linear_model import LogisticRegression
 
@@
 In \sklearn, classical learning algorithms are not the only objects to be
-implemented as estimators. For examplee, preprocessing routines (e.g., scaling of
+implemented as estimators. For example, preprocessing routines (e.g., scaling of
 features) or feature extraction techniques (e.g., vectorization of text

### 2013-08-09 paper.tex

@@
 these in Python/NumPy or by wrapping existing C/C{}\verb!++!/Fortran
-implementations. The \textit{scikits} were created to complement SciPy with
+implementations. Building upon this stack, a series of libraries called
+\textit{scikits} were created, to complement SciPy with
 domain-specific toolkits. Currently, the two most popular and feature-complete
@@
 
-Started in 2007, \sklearn is developed by an international team of a dozen of
+Started in 2007, \sklearn is developed by an international team of over a dozen
 core developers, mostly researchers from various fields (e.g.,
@@
 is itself an estimator, and can be employed wherever usual estimators can be used.
-In particular,
-they integrate within model selection routines, allowing one to optimize at once
-every step of a complex workflow. Composition of estimators can be done in two
-ways: either sequentially through \texttt{Pipeline} objects or in a parallel
+In particular, \sklearn's model selection routines
+can be applied to composite estimators, allowing global optimization
+of all parameters in a complex workflow.
+Composition of estimators can be done in two
+ways: either sequentially through \texttt{Pipeline} objects, or in a parallel
 fashion through \texttt{FeatureUnion} objects.

### 2013-08-09 paper.tex
Feature extraction are implemented as transformers as well
Cite Orange

@@
 data \texttt{X\_test} and yields as output a transformed version of
-\texttt{X\_test}. Preprocessing, feature selection and dimensionality reduction
+\texttt{X\_test}. Preprocessing, feature selection, feature extraction and dimensionality reduction
 algorithms are all provided as transformers within the library.  In our example,
@@
 
-Recent years have witnessed a rising interest in machine learning and data mining
-with applications in many fields.
-With this rise comes a host of machine learning packages
-(both open source and proprietary) with which \sklearn competes.
-Although it can also be used through its Java API, Weka \citep{hall2009weka} is
-popularly used through its graphical user interface (GUI). The target audience
-of \sklearn is capable of programming, and therefore we focus on developing
-a usable and consistent API, rather than expend effort into creating a GUI\@. In
-addition, running machine learning through a GUI makes reproducibility difficult
-(although Weka has developed a GUI for managing such workflows).
+Recent years have witnessed a rising interest in machine learning and data
+mining with applications in many fields. With this rise comes a host of machine
+learning packages (both open source and proprietary) with which \sklearn
+competes. Some of those, including Weka~\citep{hall2009weka} or
+Orange~\citep{Demsar2004}, offer and focus on a graphical user interface (GUI)
+which allows novices to easily use and apply machine learning algorithms. By
+contrast, the target audience of \sklearn is capable of programming, and
+therefore we focus on developing a usable and consistent API, rather than expend
+effort into creating a GUI\@. In addition, while GUIs are useful tools, they
+sometimes make reproducibility difficult in the case of complex workflows
+(although those packages usually have developed a GUI for managing complex
+tasks).
 

### 2013-08-09 paper.tex
Weka and Orange offer an API

@@
 competes. Some of those, including Weka~\citep{hall2009weka} or
-Orange~\citep{Demsar2004}, offer and focus on a graphical user interface (GUI)
-which allows novices to easily use and apply machine learning algorithms. By
+Orange~\citep{Demsar2004}, offer APIs but actually focus on the use of a graphical user interface (GUI)
+which allows novices to easily apply machine learning algorithms. By
 contrast, the target audience of \sklearn is capable of programming, and

### 2013-08-13 paper.tex
Cite sofia-ml

--- 2013-08-09 paper.tex
@@
 Other existing machine learning packages
-such as SofiaML\footnote{\url{http://code.google.com/p/sofia-ml/}}
-and Vowpal~Wabbit\footnote{\url{http://hunch.net/$\sim$vw/}}
+such as SofiaML\footnote{\url{https://code.google.com/p/sofia-ml}}~\citep{sculley2009large}
+and Vowpal~Wabbit\footnote{\url{http://hunch.net/$\sim$vw}}
 are intended to be used as command-line tools
@@
 that follow the \sklearn conventions. Examples of such packages include
-\textit{astroML}\footnote{\url{http://astroml.org/}}
+\textit{astroML}\footnote{\url{http://astroml.org}}
 \citep{vanderplas2012astroML}, a package providing

### 2013-08-20 paper.tex
twocolumn format and decoupled from LLNCS for CiSE
Also changed some affiliations and upped the sales pitch.

@@
-\documentclass{llncs}
+\documentclass[10pt,twocolumn]{article}
 
+\usepackage{authblk}
 \usepackage[utf8]{inputenc}
 \usepackage{amsmath}
-%\usepackage{hyperref}
+\usepackage{hyperref}
 \usepackage[authoryear,round]{natbib}
@@
 
-\pagestyle{headings}
+%\pagestyle{headings}
 
@@
 
-\author{Lars~Buitinck~\inst{1} \and
-        Gilles~Louppe~\inst{2} \and
-        Mathieu~Blondel~\inst{3} \and
-        Fabian~Pedregosa~\inst{4} \and
-        Andreas~C.~M\"uller~\inst{5} \and
-        Olivier~Grisel~\inst{6} \and
-        Vlad~Niculae~\inst{7} \and
-        Peter~Prettenhofer~\inst{8} \and
-        Alexandre~Gramfort~\inst{4,9} \and
-        Jaques~Grobler~\inst{4} \and
-        Robert~Layton~\inst{10} \and
-        Jake~Vanderplas~\inst{11} \and
-        Arnaud~Joly~\inst{2} \and
-        Brian Holt~\inst{12} \and
-        Gaël~Varoquaux~\inst{4}}
+\author[1]{Lars~Buitinck}
+\author[2]{Gilles~Louppe}
+\author[3]{Mathieu~Blondel}
+\author[4]{Fabian~Pedregosa}
+\author[5]{Andreas~C.~Müller}
+\author[4]{Olivier~Grisel}
+\author[6]{Vlad~Niculae}
+\author[7]{Peter~Prettenhofer}
+\author[4,8]{Alexandre~Gramfort}
+\author[4]{Jaques~Grobler}
+\author[9]{Robert~Layton}
+\author[10]{Jake~Vanderplas}
+\author[2]{Arnaud~Joly}
+\author[11]{Brian Holt}
+\author[4]{Gaël~Varoquaux}
 
-\institute{ILPS, Informatics Institute, University of Amsterdam \and
-           University of Liège \and
-           Kobe University \and
-           Parietal, INRIA Saclay \and
-           University of Bonn \and
-           Independent consultant \and
-           University of Bucharest \and
-           Ciuvo GmbH \and
-           Institut Mines-Telecom, Telecom ParisTech, CNRS LTCI \and
-           University of Ballarat \and
-           University of Washington \and
-           Samsung Electronics Research Institute}
+\affil[1]{Netherlands eScience Center}
+\affil[2]{University of Liège}
+\affil[3]{Kobe University}
+\affil[4]{Parietal, INRIA Saclay}
+\affil[5]{University of Bonn}
+\affil[6]{University of Bucharest}
+\affil[7]{Ciuvo GmbH}
+\affil[8]{Institut Mines-Telecom, Telecom ParisTech, CNRS LTCI}
+\affil[9]{University of Ballarat}
+\affil[10]{University of Washington}
+\affil[11]{Samsung Electronics Research Institute}
 
@@
 \begin{abstract}
-\sklearn is an increasingly popular machine learning
-library. Written in Python, it is designed to be simple and efficient, accessible to
-non-experts, and reusable in various contexts. In this paper, we present and
+\sklearn is an increasingly popular machine learning library
+for the Python programming language.
+It is designed to be simple and efficient,
+accessible to non-experts, and reusable in various contexts:
+it has a user base in neuroscience, astronomy, computational linguistics,
+as well as industry.
+
+In this paper, we present and
 discuss our design choices for the application programming interface (API) of
@@
 implementation details specific to the Python ecosystem and analyzes obstacles
-faced by users and developers of the library.
+faced by users and developers of the library.\footnote{
+  An earlier version of this paper was presented
+  at the ECML/PKDD Workshop on Languages for Data Mining and Machine Learning,
+  September 2013, Prague.}
 \end{abstract}
 
-\setcounter{footnote}{0}
+%\setcounter{footnote}{0}
 

### 2013-09-09 paper.tex
standard article format

@@
-\documentclass[10pt,twocolumn]{article}
+\documentclass{article}
 
-\usepackage{authblk}
 \usepackage[utf8]{inputenc}
 \usepackage{amsmath}
-\usepackage{hyperref}
+\usepackage{authblk}
+%\usepackage{hyperref}
+\usepackage{minted}
 \usepackage[authoryear,round]{natbib}
@@
 \usepackage{xspace}
-\usepackage{minted}
+\usepackage{url}
 
@@
 
-%\pagestyle{headings}
+\pagestyle{headings}
 
@@
 
+\date{}
+
 % Dutch name sorting hack as per http://tex.stackexchange.com/a/40750/2806
@@
 
-\begin{abstract}
-\sklearn is an increasingly popular machine learning library
-for the Python programming language.
-It is designed to be simple and efficient,
-accessible to non-experts, and reusable in various contexts:
-it has a user base in neuroscience, astronomy, computational linguistics,
-as well as industry.
-
-In this paper, we present and
+\sklearn is an increasingly popular machine learning
+library. Written in Python, it is designed to be simple and efficient, accessible to
+non-experts, and reusable in various contexts. In this paper, we present and
 discuss our design choices for the application programming interface (API) of
@@
 implementation details specific to the Python ecosystem and analyzes obstacles
-faced by users and developers of the library.\footnote{
-  An earlier version of this paper was presented
-  at the ECML/PKDD Workshop on Languages for Data Mining and Machine Learning,
-  September 2013, Prague.}
-\end{abstract}
-
-%\setcounter{footnote}{0}
+faced by users and developers of the library.
 

### 2013-09-09 paper.tex
twocolumn (examples need to be fixed)
correct the largest code snippets by putting them in figures

@@
-\documentclass{article}
+\documentclass[twocolumn]{article}
 
@@
 \usepackage{authblk}
+\usepackage{dblfloatfix}
 %\usepackage{hyperref}
@@
 combined to create complex and nested workflows.
-The following snippet illustrates how to create a complex estimator
+Figure~\ref{pca-union} illustrates how to create a complex estimator
 that computes both linear PCA and kernel PCA features on \texttt{X\_train}
@@
 and feeds those to an $\ell_2$-regularized logistic regression model.
+
+\begin{figure*}[b]
 \begin{pythoncode}
@@
 \end{pythoncode}
+\caption{A union of linear PCA and kernel PCA for feature extraction.}
+\label{pca-union}
+\end{figure*}
 
@@
 
-The snippet below illustrates how to find
-hyper-parameter settings for an SVM classifier (SVC)
-that maximize $F_1$ score
-through 10-fold cross-validation on the training set.
+\begin{figure*}[b]
 \begin{pythoncode}
@@
 \end{pythoncode}
-In this example, two distinct hyper-parameter grids are
+\caption{Optimizing the hyper-parameters of an SVM for $F_1$ score.}
+\label{svm-gridsearch}
+\end{figure*}
+
+The code in Figure~\ref{svm-gridsearch} illustrates how to find
+hyper-parameter settings for an SVM classifier (SVC)
+that maximize $F_1$ score
+through 10-fold cross-validation on the training set.
+Two distinct hyper-parameter grids are
 considered for the linear and radial basis function (RBF) kernels;

### 2013-09-09 paper.tex
trim/copyedit

@@
 
-\sklearn is an increasingly popular machine learning
-library. Written in Python, it is designed to be simple and efficient, accessible to
-non-experts, and reusable in various contexts. In this paper, we present and
-discuss our design choices for the application programming interface (API) of
-the project. In particular, we describe the simple and elegant interface shared
-by all learning and processing units in the library and then discuss its
-advantages in terms of composition and reusability. The paper also comments on
-implementation details specific to the Python ecosystem and analyzes obstacles
-faced by users and developers of the library.
+\sklearn is a machine learning library for the Python language.\footnote{\url{http://scikit-learn.org}}
+It is designed to be simple and efficient, accessible to
+non-experts, and reusable in various contexts. In this article, we present and
+discuss the uniform application programming interface shared
+by all learning and processing units in the library,
+and the way it facilitates composition and reusability. We also comment on
+implementation details specific to the Python ecosystem
+and analyze potential obstacles faced by users and developers of the library.
 
@@
 
-The \sklearn project\footnote{\url{http://scikit-learn.org}}
-\citep{pedregosa2011} provides an open source machine learning
+The \sklearn project \citep{pedregosa2011} provides an open source machine learning
 library for the Python programming language. The ambition of the project is to
@@
 
-In \sklearn, we chose a representation of data that is as close as
-possible to the matrix representation: datasets are encoded as NumPy
-multidimensional arrays for dense data and as SciPy sparse matrices for sparse
-data. While these may seem rather unsophisticated data representations when
+In \sklearn, the data representation is kept as close as possible
+to the matrix representation: datasets are encoded as
+two-dimensional NumPy arrays or SciPy sparse matrices.
+While these may seem rather unsophisticated data representations when
 compared to more object-oriented constructs, such as the ones used by
@@
 
-For efficiency reasons, the public interface is oriented towards processing
-batches of samples rather than single samples per API call. While classification
-and regression algorithms can indeed make predictions for single samples,
-\sklearn objects are not optimized for this use case. (The few online learning
-algorithms implemented are intended to take mini-batches.) Batch processing makes
+The public interface is oriented towards processing a batch of samples,
+rather than a single sample, in each API call. While classification
+and regression algorithms can make predictions for single samples,
+\sklearn objects are not optimized for this use case. (The online learning
+algorithms in the library are intended to take mini-batches.) Batch processing makes
 optimal use of NumPy and SciPy by preventing the overhead inherent to Python
@@
 detail that leaks into the API, we argue that APIs should be designed so as not
-to tie a library to a suboptimal implementation strategy. As such, batch
-processing enables fast implementations in lower-level languages (where memory
+to tie a library to a suboptimal implementation strategy.
+Batch processing also enables fast implementations in lower-level languages, where memory
 hierarchy effects and the possibility of internal parallelization come into
-play).
+play.
 
@@
 an estimator object, used as a factory, and a model object, produced by the
-estimator) indeed decreases the ease of use and is also more likely to
-unnecessarily confuse newcomers. From the developer point of view, decoupling
+estimator) decreases the ease of use and is also more likely to confuse newcomers.
+From the developer point of view, decoupling
 estimators from models also creates parallel class hierarchies and increases the
@@
 Transformers also include a variety of learning algorithms, such as
-dimension reduction (PCA, manifold learning), kernel approximation,
+dimensionality reduction (PCA, manifold learning), kernel approximation,
 and other mappings from one feature space to another.
 
-Additionally, by leveraging the fact that \texttt{fit} always returns the
-estimator it was called on, the \texttt{StandardScaler} example above can be
-rewritten in a single line using method chaining:
-\begin{pythoncode}
-X_train = StandardScaler().fit(X_train).transform(X_train)
-\end{pythoncode}
-
-Furthermore, every transformer allows \texttt{fit(X\_train).transform(X\_train)}
-to be written as \texttt{fit\_transform(X\_train)}.
+Every transformer allows \texttt{fit(X).transform(X)}
+to be written as \texttt{fit\_transform(X)}.
 The combined \texttt{fit\_transform} method prevents repeated computations.
@@
 also take a cross-validation scheme and a score function.  \sklearn provides
-various such cross-validation schemes, including $k$-fold (default),
+various such schemes, including $k$-fold (default),
 stratified $k$-fold and leave-one-out.
@@
 functions that follow a specific API can be given as arguments at vectorizer
-construction time. The library then calls into this code, communicating with it by passing objects of standard Python/NumPy types.
-Again, such external user code can be kept agnostic of the \sklearn
-class hierarchy.
+construction time. The library then calls into this code,
+communicating with it by passing objects of standard Python/NumPy types,
+and keeping it agnostic of the \sklearn class hierarchy.
 
 Our rule of thumb is that user code should not be tied to \sklearn---which is a
-\textit{library}, and not a \textit{framework}. This principle indeed avoids a
-well-known problem with object-oriented design, which is that users wanting a
+\textit{library}, and not a framework. This principle avoids a
+common problem with object-oriented design, which is that users wanting a
 ``banana'' should not get ``a gorilla holding the banana and the entire jungle''
 (J.~Armstrong, cited by \citealp[p.~213]{seibel2009coders}).
-That is, programs using \sklearn should not be intimately tied to it,
+Programs using \sklearn should not be intimately tied to it,
 so that their code can be reused with other toolkits or in other contexts.
@@
 Our implementation guidelines emphasize writing efficient but readable code.
-In particular, we focus on making the codebase easily maintainable and
-understandable in order to favor external contributions. Whenever practicable,
+In particular, we focus on making the codebase maintainable and
+understandable in order to favor external contributions. Whenever practical,
 algorithms implemented in \sklearn are written in Python,
 using NumPy vector operations for numerical work.
-This allows for the code to remain concise, readable and
-efficient. For critical algorithms that cannot be easily and efficiently
+This allows the code to remain concise, readable and efficient.
+For critical algorithms that cannot be easily and efficiently
 expressed as NumPy operations, we rely on Cython \citep{behnel2011cython}
@@
 compiled programming language that extends Python with static typing. It
-produces efficient C extension modules that are directly importable from the
-Python run-time system. Examples of algorithms written in Cython include
-stochastic gradient descent for linear models, some graph-based clustering
-algorithms and decision trees.
+produces efficient C extension modules that are importable from the
+Python run-time system. Examples of Cython code in \sklearn are
+stochastic gradient descent for linear models, some clustering
+algorithms, and decision trees.
 
-To facilitate the installation and thus adoption of \sklearn,
+To facilitate the installation of \sklearn,
 the set of external dependencies is kept to a bare minimum:
@@
 Binary distributions of these are available for the major platforms.
-Visualization functionality depends on Matplotlib \citep{hunter2007matplotlib}
-and/or Graphviz \citep{gansner2000},
+Visualization functionality uses on Matplotlib \citep{hunter2007matplotlib}
+and Graphviz \citep{gansner2000},
 but neither is required to perform machine learning or prediction.
@@
 \citep{chang2011libsvm,fan2008}, both written in C{}\verb!++!
-and wrapped using Cython modules.
+and wrapped using Cython.
 
@@
 
-Recent years have witnessed a rising interest in machine learning and data
-mining with applications in many fields. With this rise comes a host of machine
-learning packages (both open source and proprietary) with which \sklearn
-competes. Some of those, including Weka~\citep{hall2009weka} or
+\sklearn competes with a many machine learning packages,
+both open source and proprietary.
+Some of those, including Weka~\citep{hall2009weka} or
 Orange~\citep{Demsar2004}, offer APIs but actually focus on the use of a graphical user interface (GUI)
@@
 Other existing machine learning packages
-such as SofiaML\footnote{\url{https://code.google.com/p/sofia-ml}}~\citep{sculley2009large}
-and Vowpal~Wabbit\footnote{\url{http://hunch.net/$\sim$vw}}
+such as SofiaML\footnote{\url{https://code.google.com/p/sofia-ml}}
+and Vowpal~Wabbit\footnote{\url{http://hunch.net/~vw}}
 are intended to be used as command-line tools
-(and sometimes do not offer any type of API).
+(and sometimes do not offer any kind of API).
 While these packages have the advantage
 that their users are not tied to a particular programming language,
-the users will find that they still need programming to process input/output,
+users will find that they still need programming to process input/output,
 and will do so in a variety of languages.
@@
 applicability.
-In many applications of machine learning these tasks, such as data access,
-data preprocessing and reporting, can be a more significant task than applying
+In many applications of machine learning, these tasks, such as data access,
+data preprocessing and reporting, can take more work than applying
 the actual learning algorithm.
@@
 Gensim's method of dealing with large datasets is to use algorithms
-that have $O(1)$ space complexity and can be updated online.
-The API is designed around the Python concept of an \textit{iterable}
-(supported in the language by a restricted form of co-routines called
-\textit{generators}).
-While the text vectorizers part of \sklearn
-also use iterables to some extent,
-they still produce entire sparse matrices, intended to be used for batch or
-mini-batch learning. This is the case
-even in the stateless, O(1) memory vectorizers
-that implement the hashing trick of \citet{weinberger2009}.
+that have small space complexity and can be updated online.
+The API is designed around the Python concept of an ``iterable''
+(supported in the language by a restricted form of co-routines called generators).
+While the text vectorizers in \sklearn
+also leverage iterables to some extent,
+they still produce entire sparse matrices,
+intended to be used for batch or mini-batch learning.
 This way of processing, as argued earlier in Section~\ref{sec:arrays},

### 2013-09-10 paper.tex
Gensim's algos actually don't always take O(1) space
(And when they do, it's commonly the second pass through the data
that does.)
trim FeatureUnion example

@@
 that concatenates their outputs. A union of two transformers that
-map input having $d$ features to $d'$ and $d''$ features respectively is
-a transformer that maps its $d$ input features to $d' + d''$ features.
+produce outputs with (resp.) $d_1$ and $d_2$ features, is
+a transformer that produces outputs with $d_1 + d_2$ features.
 This generalizes in the obvious way to more than two transformers.
-In terms of API, a \texttt{FeatureUnion} takes as input a list of transformers.
 Calling \texttt{fit} on the union is the same as calling \texttt{fit}
@@
 (through a \texttt{FeatureUnion}),
-selects the 10 best features in the combination according to an ANOVA test
 and feeds those to an $\ell_2$-regularized logistic regression model.
@@
 from sklearn.decomposition import PCA, KernelPCA
-from sklearn.feature_selection import SelectKBest
 
@@
 Pipeline([("feat_union", union),
-          ("feat_sel", SelectKBest(k=10)),
           ("log_reg", LogisticRegression(penalty="l2"))
-         ]).fit(X_train, y_train).predict(X_test)
+         ]).fit(X_train, y_train)
 \end{pythoncode}

### 2013-09-10 paper.tex
remove OvO example, it's too long

@@
 Examples include ensemble methods which
-build and combine several simpler models (e.g., decision trees), or multiclass
-and multilabel classification schemes which can be used to turn a binary
-classifier into a multiclass or multilabel classifier. In \sklearn, such algorithms are
-implemented as \textit{meta-estimators}. They take as input an existing base
-estimator and use it internally for learning and making predictions.
+build and combine several simpler models (boosting, bagging, random forests),
+or multiclass classification schemes which combine multiple binary classifiers.
+In \sklearn, such algorithms are implemented as \textit{meta-estimators}.
+They take as input an existing estimator
+and use it internally for learning and making predictions.
 All meta-estimators implement the regular estimator interface.
-
-As an example, a logistic regression classifier
-uses by default a one-vs.-rest scheme
-for performing multiclass classification.
-A different scheme can be achieved
-by a meta-estimator wrapping a logistic regression estimator:
-\begin{pythoncode}
-from sklearn.multiclass import OneVsOneClassifier
-
-ovo_lr = OneVsOneClassifier(LogisticRegression(penalty="l1"))
-\end{pythoncode}
-For learning, the \texttt{OneVsOneClassifier} object
-\textit{clones} the logistic regression estimator multiple times,
-resulting in a set of $\frac{K(K-1)}{2}$ estimator objects
-for $K$-way classification,
-all with the same settings.
-For predictions, all estimators perform a binary classification and then vote to make the final decision.
-The snippet exemplifies the importance
-of separating object instantiation and actual learning.
+% XXX bagging example here
 
@@
 is always based on a trade-off between generality and ease of use.
-Relating to the example just given,
-all \sklearn classifiers are designed to do multiclass classification
-and the use of the \texttt{multiclass} module
-is only necessary in advanced use cases.
-
-% TODO: show a grid search on this estimator
 

### 2013-09-10 paper.tex
more copyediting/trimming

@@
 language, but a library that provides machine learning idioms to a
-general-purpose high-level language. Among other things, it includes classical learning
-algorithms, model evaluation and selection tools, as well as
-preprocessing procedures. The library is
-distributed under the simplified BSD license, encouraging its use in both
-academic and commercial settings.
+general-purpose high-level language. It includes classical learning algorithms,
+model evaluation and selection tools, as well as preprocessing procedures.
+The library is distributed under the simplified BSD license,
+encouraging its use in both academia and industry.
 
 \sklearn is a library, i.e.\ a collection of classes
-and functions that users import into Python programs. Using \sklearn therefore
-requires basic Python programming knowledge. No command-line interface, let
-alone a graphical user interface, is offered for non-programmer users. Instead,
+and functions that users import into Python programs.
+No command-line interface, let alone a graphical user interface,
+is offered for non-programmer users. Instead,
 interactive use is made possible by the Python interactive interpreter, and its
-enhanced replacement IPython \citep{perez2007ipython}, which offer a
-\textsc{matlab}-like working environment specifically designed for scientific
-use.
+enhanced replacement IPython \citep{perez2007ipython}, which offers a
+\textsc{matlab}-like working environment especially suited for scientific use.
 
@@
 
-Started in 2007, \sklearn is developed by an international team of over a dozen
+Since 2007, \sklearn has been developed by an international team of more than a dozen
 core developers, mostly researchers from various fields (e.g.,
@@
 benchmarked it against several competitors.
-In this paper, we instead present an
+In this paper, we present an
 in-depth analysis of design choices made when building the library,
@@
 
-All objects within \sklearn share a uniform common basic API consisting of three
+All objects in \sklearn share a uniform basic API consisting of three
 complementary interfaces: an \textit{estimator} interface for building and
@@
 
-As much as possible, our
-design choices have been guided so as to avoid the proliferation of framework
-code. We try to adopt simple conventions and to limit to a minimum the number of
-methods an object must implement. The API is designed to adhere to the following
-broad principles:
+As much as possible, \sklearn has been designed in such a way
+as to avoid the proliferation of framework code.
+We adopt simple conventions and limit the number of methods
+an object must implement to a minimum.
+The API is designed to adhere to the following principles:
 
@@
 arrays and SciPy sparse matrices.
-From a practical point of view, these formats also provide a collection of
+These formats also provide a collection of
 data loading and conversion tools which make them very easy to use in many
-contexts. Moreover, for tasks where the inputs are text files or semi-structured
+contexts. For tasks where the inputs are text files or semi-structured
 objects, we provide \textit{vectorizer} objects that efficiently convert such
@@
 clustering) are offered as objects implementing this interface. Machine
-learning tasks like feature extraction, feature selection or dimensionality
-reduction are also provided as estimators.
+learning tasks like feature extraction and selection are also provided as estimators.
 
@@
 in a way that is similar to partial function application:
-an estimator is initialized from a set of named constant hyper-parameter values
+an estimator is initialized from a set of named hyper-parameter values
 (e.g., the $C$ constant in SVMs)
-and can be considered as a function
+and can be considered a function
 that maps these values to actual learning algorithms.
-The constructor of an estimator does not see any actual data, nor does it perform any actual learning.
+The constructor does not see any actual data.
 All it does is attach the given parameters to the object.
-For the sake of convenient model inspection, hyper-parameters are set as public attributes,
+For the sake of model inspection, hyper-parameters are set as public attributes,
 which is especially important in model selection settings.
-For ease of use, default hyper-parameter values are also provided
-for all built-in estimators.
-These default values are set to be relevant in many common
-situations in order to make estimators as effective as possible
-\textit{out-of-box} for non-experts.
+Default hyper-parameter values are provided for all built-in estimators.
+These defaults are set to be relevant in many common
+situations in order to make estimators effective \textit{out-of-box}.
 
@@
 learned from the training arrays \texttt{X\_train} and \texttt{y\_train},
-and stored within the object for later use.
+and stored on the object for later use.
 %% (In this example, \texttt{fit} is called solely for its side effects.)
@@
 \texttt{X\_test}. Preprocessing, feature selection, feature extraction and dimensionality reduction
-algorithms are all provided as transformers within the library.  In our example,
+algorithms are all provided as transformers.  In our example,
 to standardize the input \texttt{X\_train} to zero mean and unit variance
@@
 In \sklearn, such algorithms are implemented as \textit{meta-estimators}.
-They take as input an existing estimator
-and use it internally for learning and making predictions.
+They take as input an estimator which is used
+internally for learning and making predictions.
 All meta-estimators implement the regular estimator interface.
@@
 rather than integrate the behavior it implements
-into existing estimators classes
+into existing estimator classes
 is always based on a trade-off between generality and ease of use.
@@
 model on average is retained and exposed as the public attribute
-\texttt{best\_estimator\_}.
+\texttt{best\_estimator\_} (optionally re-fit on the whole training set),
+but \texttt{GridSearchCV} also has a \texttt{predict} method,
+which delegates to the best model's \texttt{predict},
+so the CV object is itself an estimator and a predictor.
 
@@
 clf.fit(X_train, y_train)
-y_pred = clf.predict(X_test)
 \end{pythoncode}
@@
 
-The code in Figure~\ref{svm-gridsearch} illustrates how to find
+The code in Figure~\ref{svm-gridsearch} shows how to find
 hyper-parameter settings for an SVM classifier (SVC)
@@
 trying out effectively equivalent settings.
-Additionally, we see that
-\texttt{GridSearchCV} has a \texttt{predict} method, just like any other classifier:
-it delegates the \texttt{predict}, \texttt{predict\_proba}, \texttt{transform} and
-\texttt{score} methods to the best estimator
-(optionally after re-fitting it on the whole training set).
 
@@
 
-To ease code reuse, simplify implementation and skip the introduction of
-superfluous classes, the Python principle of \textit{duck typing} is exploited
+The Python principle of \textit{duck typing} is exploited
 throughout the codebase. This means that estimators are defined by interface,
@@
 outlined in Section~\ref{sec:core-api}, then it can be used in lieu of a
-built-in estimator (e.g., it can be plugged into pipelines or grid search)
-and external developers are not forced to inherit from any \sklearn class.
+built-in estimator (e.g., in pipelines and grid search)
+and developers are not forced to inherit from any \sklearn class.
 
@@
 construction time. The library then calls into this code,
-communicating with it by passing objects of standard Python/NumPy types,
+communicating with it using objects of standard Python/NumPy types,
 and keeping it agnostic of the \sklearn class hierarchy.
@@
 Our rule of thumb is that user code should not be tied to \sklearn---which is a
-\textit{library}, and not a framework. This principle avoids a
-common problem with object-oriented design, which is that users wanting a
-``banana'' should not get ``a gorilla holding the banana and the entire jungle''
+\textit{library}, not a framework. We try to avoid a
+common problem with object-oriented design: that users wanting a
+``banana'' instead get ``a gorilla holding the banana and the entire jungle''
 (J.~Armstrong, cited by \citealp[p.~213]{seibel2009coders}).
 Programs using \sklearn should not be intimately tied to it,
-so that their code can be reused with other toolkits or in other contexts.
+so that their code can be reused or in other contexts.
 
@@
 Binary distributions of these are available for the major platforms.
-Visualization functionality uses on Matplotlib \citep{hunter2007matplotlib}
+Visualization functionality uses Matplotlib \citep{hunter2007matplotlib}
 and Graphviz \citep{gansner2000},
@@
 
-Other existing machine learning packages
+Other machine learning packages,
 such as SofiaML\footnote{\url{https://code.google.com/p/sofia-ml}}
-and Vowpal~Wabbit\footnote{\url{http://hunch.net/~vw}}
-are intended to be used as command-line tools
-(and sometimes do not offer any kind of API).
-While these packages have the advantage
-that their users are not tied to a particular programming language,
-users will find that they still need programming to process input/output,
+and Vowpal~Wabbit,\footnote{\url{http://hunch.net/~vw}}
+are often designed as command-line tools
+and may not offer any kind of API\@.
+While this has the advantage that users are not tied
+to a particular programming language,
+they will find that they still need programming to process input/output,
 and will do so in a variety of languages.
 By contrast, \sklearn allows users to implement that entire workflow
-in a single program, written in a single language,
-and developed in a single working environment.
+in a single program, written in a single language.
 This also makes it easier for researchers and developers
@@
 algorithms,
-including neural networks, ensemble meta-estimators for
-bagging or subsampling strategies and missing value completion algorithms.
+including neural networks and subsampling strategies.
 However, tasks like structured prediction or reinforcement learning are
@@
 but only in a coarse-grained fashion.
-At present, parallel processing is difficult to accomplish in the Python environment;
+Parallel processing is difficult to accomplish in the Python environment;
 \sklearn targets the main implementation, CPython,
@@
 We have shown how a consistent API across the package makes \sklearn
-very \textbf{usable} in practice: experimenting with different learning
-algorithm is as simple as substituting a new class definition.
-Through composition interfaces such as Pipelines, Feature Unions,
-and meta-estimators, these simple building blocks lead to an API which is
-\textbf{powerful}, and can accomplish a wide variety of learning tasks
-within a small amount of easy-to-read code.
-Through duck-typing, the consistent API leads to a library that is
-easily \textbf{extensible}, and allows user-defined estimators to be
-incorporated into the \sklearn workflow without any explicit object
-inheritance.
+easy to use: experimenting with different learning
+algorithms can be as simple as substituting a class name.
+Through composition interfaces such as pipelines and meta-estimators,
+simple building blocks lead to an API which is powerful,
+and can accomplish a wide variety of learning tasks
+within a small amount of code.
+Through duck-typing, the consistent API leads to an extensible library
+where custom algorithm can be incorporated into the workflow without inheritance.
 
@@
 written in other (dynamic) programming languages.
-The power, and extensibility of the \sklearn API is evidenced
+The power and extensibility of the \sklearn API is evidenced
 by the large and growing user-base, its use to solve real
@@
 as well as the appearance of third-party packages
-that follow the \sklearn conventions. Examples of such packages include
+that follow \sklearn conventions. Examples of such packages include
 \textit{astroML}\footnote{\url{http://astroml.org}}
-\citep{vanderplas2012astroML}, a package providing
-machine learning tools for astronomers, and
-\textit{wiseRF}\footnote{\url{http://wise.io}}, a commercial random forest
-implementation. The source code of
-the recently-proposed sparse multiclass
-algorithm of \citet{mblondel-mlj2013}, released as
-part of the
+\citep{vanderplas2012astroML}, a machine learning toolkit for astronomy,
+and \textit{wiseRF}\footnote{\url{http://wise.io}}, a commercial random forest
+implementation. The source code of the recently-proposed sparse multiclass
+algorithm of \citet{mblondel-mlj2013}, released as part of the
 \textit{lightning}\footnote{\url{https://github.com/mblondel/lightning}}
 package, also follows the \sklearn conventions.
-To maximize ease of use, we encourage more researchers
-to follow these conventions when releasing their software.
+We encourage researchers to follow these conventions when releasing software.
 

### 2013-09-10 paper.tex
OpenMP is probably not the way
rewrite the data representation section
Focus on X first, since there's no y in clustering/dim.red.
More readable math.

@@
 \usepackage{amsmath}
+\usepackage{amsfonts}
 \usepackage{authblk}
@@
 
-In most machine learning tasks, data is modeled as a set of variables.  For
-example, in a supervised learning task, the goal is to find a mapping
-from input variables $X_1, \ldots X_p$, called features, to
-some output variables $Y$. A sample is then defined as a pair of
-values $([x_1, \ldots, x_p]^\mathrm{T}, y)$ of these variables. A widely used
-representation of a dataset, a collection of
-such samples, is a pair of matrices with numerical values: one for the input
-values and one for the output values. Each row of these matrices
-corresponds to one sample of the dataset and each column to one variable of
-the problem.
+Machine learning revolves around data,
+so good datastructures are paramount to designing software for it.
+In most tasks, data is modeled by a set of $d$ numerical variables,
+so that a single \textit{sample} (instance, observation, event)
+is a vector $x \in \mathbb{R}^d$.
+A collection of such samples is naturally regarded
+as the rows in a matrix $X$ of size $n \times d$.
+In the common case of supervised learning (classification, regression),
+we have an additional vector $y$ of length $n$ at training time
+and want an algorithm to produce such a $y$ for new data.
 
-In \sklearn, the data representation is kept as close as possible
-to the matrix representation: datasets are encoded as
-two-dimensional NumPy arrays or SciPy sparse matrices.
-While these may seem rather unsophisticated data representations when
+\sklearn's data representation is kept as close as possible
+to this matrix formulation: datasets are encoded as
+two-dimensional NumPy arrays or SciPy sparse matrices,
+while target vectors are ``flat'' arrays of numbers or strings.
+While these may seem rather unsophisticated when
 compared to more object-oriented constructs, such as the ones used by
-Weka \citep{hall2009weka}, they bring the prime advantage of allowing us to rely
-on efficient NumPy and SciPy vectorized operations while keeping
-the code short and readable.  This design choice has also been motivated by
-the fact that, given their pervasiveness in many other scientific Python
-packages, many scientific users of Python are already familiar with NumPy dense
-arrays and SciPy sparse matrices.
-These formats also provide a collection of
-data loading and conversion tools which make them very easy to use in many
-contexts. For tasks where the inputs are text files or semi-structured
+Weka \citep{hall2009weka},
+they allow us to rely on efficient NumPy and SciPy vector operations
+while keeping the code ``close to the textbook.''
+Given the pervasiveness of NumPy and SciPy in many other scientific Python
+packages, many scientific users of Python
+will already be familiar with these data structures,
+and a collection of available data loading and conversion tools
+facilitate interoperability.
+For tasks where the inputs are text files or semi-structured
 objects, we provide \textit{vectorizer} objects that efficiently convert such
@@
 which has already been implemented.
-Recent versions of Cython include support for the OpenMP standard
-\citep{dagum1998openmp},
-which is a viable candidate technology
-for more fine-grained multicore support in \sklearn.
+Experiments are underway to implement more fine-grained multicore support.
 

### 2013-09-11 paper.tex
rewrite the introduction
In the CiSE format, it's below or next to the abstract, which it was
repeating. Also upped the sales pitch.

@@
 
-The \sklearn project \citep{pedregosa2011} provides an open source machine learning
-library for the Python programming language. The ambition of the project is to
-provide efficient and well-established machine learning tools within a
-programming environment that is accessible to non-machine learning experts and
-reusable in various scientific areas. The project is not a novel domain-specific
-language, but a library that provides machine learning idioms to a
-general-purpose high-level language. It includes classical learning algorithms,
-model evaluation and selection tools, as well as preprocessing procedures.
-The library is distributed under the simplified BSD license,
-encouraging its use in both academia and industry.
+Many science and engineering disciplines are trying to make sense
+of ever-increasing volumes of information.
+Machine learning (ML) promises to alleviate this problem
+by automating the process of finding patterns in the rising sea of data,
+but ML tools can be hard to use
+and in particular, hard to combine
+when multiple algorithms are needed for a task.
 
-\sklearn is a library, i.e.\ a collection of classes
-and functions that users import into Python programs.
-No command-line interface, let alone a graphical user interface,
-is offered for non-programmer users. Instead,
-interactive use is made possible by the Python interactive interpreter, and its
-enhanced replacement IPython \citep{perez2007ipython}, which offers a
-\textsc{matlab}-like working environment especially suited for scientific use.
+The \sklearn project provides a comprehensive suite of machine learning tools,
+packaged as a library for the Python programming language.
+It extends this general-purpose language with machine learning operations:
+learning algorithms, preprocessing tools, model selection procedures
+and a composition mechanism to produce complex ML workflows.
+The ambition of the project is to provide efficient implementations
+of well-established ML algorithms within a
+programming environment that is accessible to non-experts and
+reusable in various scientific areas.
+The library is distributed under the simplified BSD license
+to encourage its use in both academia and industry.
 
-The library has been designed to tie in with the set of numeric
+\sklearn is designed to tie in with the set of numeric
 and scientific packages centered around the NumPy and SciPy libraries. NumPy
@@
 further with common numerical operations, either by implementing
-these in Python/NumPy or by wrapping existing C/C{}\verb!++!/Fortran
-implementations. Building upon this stack, a series of libraries called
-\textit{scikits} were created, to complement SciPy with
-domain-specific toolkits. Currently, the two most popular and feature-complete
-ones are by far \sklearn and
-\textit{scikit-image},\footnote{\url{http://scikit-image.org}}
-which does image processing.
+these in Python/NumPy or by wrapping existing C/C{}\verb!++!/Fortran code.
+The name ``scikit'' derives from (domain-specific) ``SciPy toolkit''
+and is shared with \textit{scikit-image}.\footnote{\url{http://scikit-image.org}}
+IPython \citep{perez2007ipython} and Matplotlib \citep{hunter2007matplotlib}
+complement SciPy to provide a \textsc{matlab}-esque working environment
+suited for scientific use.
 
-Since 2007, \sklearn has been developed by an international team of more than a dozen
-core developers, mostly researchers from various fields (e.g.,
-computer science, neuroscience, astrophysics). The project also benefits
-from many occasional contributors proposing small bugfixes or
+Since 2007, \sklearn has been developed by more than a dozen
+core developers, mostly researchers in various fields
+(a.o.\ computer science, neuroscience, astrophysics). It also benefits
+from occasional contributors proposing small bugfixes or
 improvements. Development proceeds on GitHub\footnote{\url{https://github.com/scikit-learn}},

### 2013-09-11 paper.tex
-PMML, +public attributes
Focus on what we have instead of what we could implement.

@@
 
-From the start, the choice to let a single object serve dual purpose as
-estimator and model has mostly been driven by usability and technical
-considerations. From the user point of view, having two coupled instances (i.e.,
-an estimator object, used as a factory, and a model object, produced by the
-estimator) decreases the ease of use and is also more likely to confuse newcomers.
+The choice to let a single object serve dual purpose as
+estimator and model has been driven by usability and technical considerations.
+Having two coupled instances
+(an estimator object used as a factory, and a model object produced by the estimator)
+makes a library harder to learn and use.
 From the developer point of view, decoupling
-estimators from models also creates parallel class hierarchies and increases the
-overall maintenance complexity of the project. For these practical reasons, we
-believe that decoupling estimators from models is not worth the effort. A good
-reason for decoupling however, would be that it makes it possible to ship a
-model in a new environment without having to deal with potentially complex
-software dependencies. Such a feature could however still be implemented in
-\sklearn by making estimators able to export a fitted model, using the
-information from its public attributes, to an agnostic model description such as
-PMML~\citep{guazzelli2009pmml}.
+estimators from models would create parallel class hierarchies and increases the
+overall maintenance burden.
+A good reason for decoupling would be to make it possible to ship a model
+to a new environment where the full library cannot be installed.
+However, our inspectable setup where model parameters are documented,
+public attributes, and prediction formulas follow standard textbooks,
+goes a long way in solving this problem.
 

### 2013-09-11 paper.tex
some more trimming

@@
 To illustrate the initialize-fit sequence,
-let us consider a supervised learning task using logistic regression.
+let us consider a supervised learning task using logistic regression (LR).
 Given the API defined above, solving this problem is as simple as the following
@@
 from sklearn.linear_model import LogisticRegression
-
 clf = LogisticRegression(penalty="l1")
@@
 to emphasize that \texttt{predict} generalizes to new data). In the case of
-supervised learning estimators, this method typically returns the predicted
-labels or values computed by the model.  Continuing with the previous example,
-predicted labels for \texttt{X\_test} can be obtained using the following
-snippet:
+supervised learning estimators, this method returns the predicted
+labels or values computed by the model:
 \begin{pythoncode}
@@
 Some unsupervised learning estimators may also implement the \texttt{predict}
-interface. The code in the snippet below fits a $k$-means model with $k=10$ on
+interface. The code below fits a $k$-means model with $k=10$ on
 training data \texttt{X\_train}, and then uses the  \texttt{predict} method to
-obtain cluster labels (integer indices) for unseen data \texttt{X\_test}.
+obtain cluster labels for unseen data \texttt{X\_test}.
 \begin{pythoncode}
 from sklearn.cluster import KMeans
-
 km = KMeans(n_clusters=10)
@@
 
-Finally, predictors must provide a \texttt{score} function to assess their
-performance on a batch of input data. In supervised estimators, this method
+Finally, supervised predictors provide a \texttt{score} function to assess their
+performance on a batch of input data. This method
 takes as input arrays \texttt{X\_test} and \texttt{y\_test} and typically
@@
 An unsupervised estimator may also expose a \texttt{score} function
-to compute, for instance, the likelihood of the given data under its model.
+to compute, for instance, data likelihood under its model.
 
@@
 to standardize the input \texttt{X\_train} to zero mean and unit variance
-before fitting the logistic regression estimator,
-one would write:
+before fitting the LR estimator, one would write:
 \begin{pythoncode}
 from sklearn.preprocessing import StandardScaler
-
 scaler = StandardScaler()
@@
 to be written as \texttt{fit\_transform(X)}.
-The combined \texttt{fit\_transform} method prevents repeated computations.
-Depending on the transformer,
-it may skip only an input validation step,
-or in fact use a more efficient algorithm for the transformation.
-In the same spirit, clustering algorithms provide a
-\texttt{fit\_predict} method
+This combined operation prevents repeated computations,
+either by skipping only an input validation step,
+or by using a more efficient algorithm for the transformation.
+In the same spirit, clustering algorithms provide a \texttt{fit\_predict}
 that is equivalent to \texttt{fit} followed by \texttt{predict},
-returning cluster labels assigned to the training samples.
+returning cluster labels for the training samples.
 
@@
 
-Some machine learning algorithms are expressed naturally
-as meta-algorithms parametrized on simpler algorithms.
+Some learning algorithms are expressed naturally
+as meta-algorithms parametrized on simpler ones.
 Examples include ensemble methods which
@@
 the decision to implement a meta-estimator
-rather than integrate the behavior it implements
-into existing estimator classes
+rather than integrate its functionality
+into existing classes
 is always based on a trade-off between generality and ease of use.
@@
 
-In \sklearn, model selection is supported in two distinct meta-estimators,
-\texttt{GridSearchCV} and \texttt{RandomizedSearchCV}.  They take as input an
+In \sklearn, model selection is supported by the meta-estimators
+\texttt{GridSearchCV} and \texttt{RandomizedSearchCV}.  These take as input an
 estimator (basic or composite), whose hyper-parameters must be optimized, and a
@@
 by sampling a fixed number of times from its parameter distributions
-(see \citealp{bergstra2012}).
+\citep{bergstra2012}.
 
@@
 ]
-
 clf = GridSearchCV(SVC(), param_grid, scoring="f1", cv=10)

### 2013-09-11 paper.tex
figures at top of page
Makes the footnotes come out better.

@@
 
-\begin{figure*}[b]
+\begin{figure*}[t]
 \begin{pythoncode}
@@
                       ("kpca", KernelPCA(kernel="rbf"))])
-
 Pipeline([("feat_union", union),

### 2013-09-11 paper.tex
cite Pylearn2; rewrite Gensim comparison and quickly mention partial_fit

@@
 
-Within the realm of Python,
-a package that deserves mention is the Gensim topic modeling toolkit
-\citep{rehurek2010gensim},
-which exemplifies a different style of API design
-geared toward scalable processing of ``big data''.
+Within the realm of Python, two packages deserve mention.
+The Pylearn2 library \citep{goodfellow2013pylearn2}
+is similar to \sklearn in that aims to facilitate reproducible results
+and model inspection.
+However, it is designed for the audience of machine learning researchers,
+rather than practitioners in other fields.
+This leads to a modular toolkit,
+where each element of an algorithm can be tweaked or replaced as needed.
+This provides more control for the user,
+at the cost of more math to be mastered.
+
+Finally, the Gensim topic modeling toolkit \citep{rehurek2010gensim}
+exemplifies an API geared toward ``big data'' processing.
 Gensim's method of dealing with large datasets is to use algorithms
@@
 (supported in the language by a restricted form of co-routines called generators).
-While the text vectorizers in \sklearn
-also leverage iterables to some extent,
-they still produce entire sparse matrices,
-intended to be used for batch or mini-batch learning.
-This way of processing, as argued earlier in Section~\ref{sec:arrays},
-reduces various forms of overhead
-and allows effective use of the vectorized operations provided by NumPy and
-SciPy.  We make no attempt to hide this batch-oriented processing from the user,
-allowing control over the amount of memory actually dedicated
-to \sklearn algorithms.
+Some \sklearn estimators, notably Naïve Bayes and linear model SGD,\footnote{
+  Least-squares, logistic regression and SVM models.}
+have a \texttt{partial\_fit} method to learn from a stream of samples
+that is divided into mini-batches (chunks).
+This way of processing, as argued in Section~\ref{sec:arrays},
+allows for more efficient implementation.
+We make no attempt to hide this batch-oriented processing from the user,
+allowing control over the amount of memory dedicated to \sklearn algorithms.
 

### 2013-09-27 paper.tex
tightly packed bibliography
some minor trimming

@@
 The \sklearn project provides a comprehensive suite of machine learning tools,
-packaged as a library for the Python programming language.
+packaged as a Python library.
 It extends this general-purpose language with machine learning operations:
@@
 The ambition of the project is to provide efficient implementations
-of well-established ML algorithms within a
+of well-established algorithms within a
 programming environment that is accessible to non-experts and
@@
 \bibliographystyle{abbrvnat}
+\setlength{\bibsep}{1mm}
 \DeclareRobustCommand{\VAN}[3]{#3}

### 2013-09-27 paper.tex
stress early on that user-defined estimators are first-class citizens
Someone at ECML wisely remarked that this is a requirement for a truly
compositional library.

@@
   \item[Composition.]
-       Many machine learning tasks are expressible
-       as sequences or combinations of transformations to data.
-       Some learning algorithms are also naturally viewed
-       as meta-algorithms parametrized on other algorithms.
-       Whenever feasible, such algorithms are implemented and composed from
-       existing building blocks.
+       Learning algorithms can be combined into processing pipelines.
+       Ensemble methods and other meta-algorithms
+       are implemented as adapter classes.
+       User-defined algorithms that follow the right conventions
+       can be used as building blocks in the resulting composite algorithms.
   \item[Sensible defaults.]

### 2013-09-27 paper.tex
numerical citations to save space

@@
 \usepackage{minted}
-\usepackage[authoryear,round]{natbib}
+\usepackage[numbers,square]{natbib}
 \usepackage{xcolor}
@@
 ``banana'' instead get ``a gorilla holding the banana and the entire jungle''
-(J.~Armstrong, cited by \citealp[p.~213]{seibel2009coders}).
+(J.~Armstrong, cited in \citep[p.~213]{seibel2009coders}).
 Programs using \sklearn should not be intimately tied to it,
@@
 \textit{astroML}\footnote{\url{http://astroml.org}}
-\citep{vanderplas2012astroML}, a machine learning toolkit for astronomy,
-and \textit{wiseRF}\footnote{\url{http://wise.io}}, a commercial random forest
-implementation. The source code of the recently-proposed sparse multiclass
-algorithm of \citet{mblondel-mlj2013}, released as part of the
-\textit{lightning}\footnote{\url{https://github.com/mblondel/lightning}}
-package, also follows the \sklearn conventions.
+\citep{vanderplas2012astroML}, a machine learning toolkit for astronomy;
+\textit{wiseRF}\footnote{\url{http://wise.io}}, a commercial random forest implementation;
+and \textit{lightning},\footnote{\url{https://github.com/mblondel/lightning}}
+the source code of a recently-proposed sparse multiclass
+algorithm \citep{mblondel-mlj2013}.
 We encourage researchers to follow these conventions when releasing software.

### 2013-09-28 paper.tex
cite Oliphant's CiSE article for NumPy and SciPy

@@
 \sklearn is designed to tie in with the set of numeric
-and scientific packages centered around the NumPy and SciPy libraries. NumPy
-\citep{vanderwalt2011} augments Python with a contiguous numeric array datatype
+and scientific packages centered around the NumPy and SciPy libraries
+\citep{oliphant2007python, vanderwalt2011}.
+NumPy augments Python with a contiguous numeric array datatype
 and fast array computing primitives,
-while SciPy \citep{varoquaux2013scipy} extends it
-further with common numerical operations, either by implementing
-these in Python/NumPy or by wrapping existing C/C{}\verb!++!/Fortran code.
-The name ``scikit'' derives from (domain-specific) ``SciPy toolkit''
+while SciPy extends it further with common numerical operations,
+either by implementing these in Python/NumPy
+or by wrapping existing C/C{}\verb!++!/Fortran code.
+The name ``scikit'' derives from ``SciPy toolkit''
 and is shared with \textit{scikit-image}.\footnote{\url{http://scikit-image.org}}
@@
 to this matrix formulation: datasets are encoded as
-two-dimensional NumPy arrays or SciPy sparse matrices,
+two-dimensional NumPy arrays or SciPy sparse matrices \citep{vanderwalt2011},
 while target vectors are ``flat'' arrays of numbers or strings.

### 2013-10-18 paper.tex
Define some accronyms.
Also, SGD is not a linear model but an optimization method.

@@
 an estimator is initialized from a set of named hyper-parameter values
-(e.g., the $C$ constant in SVMs)
+(e.g., the $C$ constant in Support Vector Machines (SVM))
 and can be considered a function
@@
 \end{pythoncode}
-\caption{Optimizing the hyper-parameters of an SVM for $F_1$ score.}
+\caption{Optimizing the hyper-parameters of a Support Vector Machine for $F_1$
+score.}
 \label{svm-gridsearch}
@@
 (supported in the language by a restricted form of co-routines called generators).
-Some \sklearn estimators, notably Naïve Bayes and linear model SGD,\footnote{
-  Least-squares, logistic regression and SVM models.}
+Some \sklearn estimators, notably Naïve Bayes and Stochastic Gradient Descent
+(for learning linear models such as
+logistic regression and SVM),
 have a \texttt{partial\_fit} method to learn from a stream of samples

### 2013-10-18 paper.tex
Use smaller font size for too large snippet.
Another solution is manual line breaking, see:
http://tex.stackexchange.com/questions/10239/breaking-long-lines-using-latex-with-minted-and-pygments
Copyedit intro.

@@
 
+{\bf Abstract}
 \sklearn is a machine learning library for the Python language.\footnote{\url{http://scikit-learn.org}}
@@
 
-The \sklearn project provides a comprehensive suite of machine learning tools,
-packaged as a Python library.
-It extends this general-purpose language with machine learning operations:
-learning algorithms, preprocessing tools, model selection procedures
+The \sklearn library provides a comprehensive suite of machine learning tools
+for Python.
+It extends this general-purpose programming language with machine learning
+operations: learning algorithms, preprocessing tools, model selection procedures
 and a composition mechanism to produce complex ML workflows.
-The ambition of the project is to provide efficient implementations
+The ambition of the library is to provide efficient implementations
 of well-established algorithms within a
@@
 IPython \citep{perez2007ipython} and Matplotlib \citep{hunter2007matplotlib}
-complement SciPy to provide a \textsc{matlab}-esque working environment
+complement SciPy to provide a \textsc{matlab}-like working environment
 suited for scientific use.
@@
 core developers, mostly researchers in various fields
-(a.o.\ computer science, neuroscience, astrophysics). It also benefits
+(among others, computer science, neuroscience, astrophysics). It also benefits
 from occasional contributors proposing small bugfixes or
@@
 Documentation and examples are required for all features,
-and major changes must pass code review by at least two
-developers not involved in the implementation of the proposed change.
+and major changes must pass code review by at least two other
+developers.
 
@@
 before fitting the LR estimator, one would write:
+\begin{small}
 \begin{pythoncode}
@@
 \end{pythoncode}
+\end{small}
 Of course, in practice, it is important to apply the same preprocessing to the

### 2013-10-18 paper.tex
Copy edit core API.

@@
 an estimator is initialized from a set of named hyper-parameter values
-(e.g., the $C$ constant in Support Vector Machines (SVM))
+(e.g., the $C$ constant in Support Vector Machines, SVM)
 and can be considered a function
@@
 It always returns the estimator object it was called on,
-which now serves as a model of its input and can be used to perform predictions or transformations of input data.
+which now serves as a model and can be used to perform predictions
+or transformations of new data.
 
@@
 To illustrate the initialize-fit sequence,
-let us consider a supervised learning task using logistic regression (LR).
+let us consider a supervised learning task using logistic regression.
 Given the API defined above, solving this problem is as simple as the following
 example.
+\begin{footnotesize}
 \begin{pythoncode}
@@
 \end{pythoncode}
+\end{footnotesize}
 In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
@@
 Since all estimators share the same interface, using a different learning algorithm is
-as simple as replacing the constructor (the class name);
-to build a random forest on
+as simple as replacing the constructor (the class name).
+To build a random forest on
 the same data, one would simply replace
@@
 to standardize the input \texttt{X\_train} to zero mean and unit variance
-before fitting the LR estimator, one would write:
+before fitting the logistic regression estimator, one would write:
 \begin{small}

### 2013-10-18 paper.tex
Move model selection snippet.
Copy edit advanced API.

@@
 and hence the performance of the resulting model on unseen data.
+For example, a decision
+tree with too small a value for the maximal tree depth
+parameter will tend to underfit, while too large a value will make it overfit.
 The problem of \textit{model selection} is therefore to find, within
 some hyper-parameter space, the best combination of hyper-parameters, with
-respect to some user-specified criterion. For example, a decision
-tree with too small a value for the maximal tree depth
-parameter will tend to underfit, while too large a value will make it overfit.
+respect to some user-specified criterion.
 
@@
 of complete parameter combinations.
-Randomized search is a smarter algorithm
+Randomized search is a more efficient algorithm
 that avoids the combinatorial explosion in grid search
@@
 so the CV object is itself an estimator and a predictor.
-
-\begin{figure*}[t]
-\begin{pythoncode}
-from sklearn.grid_search import GridSearchCV
-from sklearn.svm import SVC
-
-param_grid = [
-  {"kernel": ["linear"], "C": [1, 10, 100, 1000]},
-  {"kernel": ["rbf"], "C": [1, 10, 100, 1000],
-   "gamma": [0.001, 0.0001]}
-]
-clf = GridSearchCV(SVC(), param_grid, scoring="f1", cv=10)
-clf.fit(X_train, y_train)
-\end{pythoncode}
-\caption{Optimizing the hyper-parameters of a Support Vector Machine for $F_1$
-score.}
-\label{svm-gridsearch}
-\end{figure*}
 
@@
 so that their code can be reused or in other contexts.
+
+\begin{figure*}[t]
+\begin{pythoncode}
+from sklearn.grid_search import GridSearchCV
+from sklearn.svm import SVC
+
+param_grid = [
+  {"kernel": ["linear"], "C": [1, 10, 100, 1000]},
+  {"kernel": ["rbf"], "C": [1, 10, 100, 1000],
+   "gamma": [0.001, 0.0001]}
+]
+clf = GridSearchCV(SVC(), param_grid, scoring="f1", cv=10)
+clf.fit(X_train, y_train)
+\end{pythoncode}
+\caption{Optimizing the hyper-parameters of a Support Vector Machine for $F_1$
+score.}
+\label{svm-gridsearch}
+\end{figure*}
 

### 2013-10-18 paper.tex
Move snippets again.
Copy edit implementation and related work.

@@
 
+\begin{figure*}[t]
+\begin{pythoncode}
+from sklearn.pipeline import FeatureUnion, Pipeline
+from sklearn.decomposition import PCA, KernelPCA
+
+union = FeatureUnion([("pca", PCA()),
+                      ("kpca", KernelPCA(kernel="rbf"))])
+Pipeline([("feat_union", union),
+          ("log_reg", LogisticRegression(penalty="l2"))
+         ]).fit(X_train, y_train)
+\end{pythoncode}
+\caption{A union of linear PCA and kernel PCA for feature extraction.}
+\label{pca-union}
+\end{figure*}
+
 A distinguishing feature of the \sklearn API is its ability to
@@
 
+\subsection{Model selection}
+
 \begin{figure*}[t]
 \begin{pythoncode}
-from sklearn.pipeline import FeatureUnion, Pipeline
-from sklearn.decomposition import PCA, KernelPCA
+from sklearn.grid_search import GridSearchCV
+from sklearn.svm import SVC
 
-union = FeatureUnion([("pca", PCA()),
-                      ("kpca", KernelPCA(kernel="rbf"))])
-Pipeline([("feat_union", union),
-          ("log_reg", LogisticRegression(penalty="l2"))
-         ]).fit(X_train, y_train)
+param_grid = [
+  {"kernel": ["linear"], "C": [1, 10, 100, 1000]},
+  {"kernel": ["rbf"], "C": [1, 10, 100, 1000],
+   "gamma": [0.001, 0.0001]}
+]
+clf = GridSearchCV(SVC(), param_grid, scoring="f1", cv=10)
+clf.fit(X_train, y_train)
 \end{pythoncode}
-\caption{A union of linear PCA and kernel PCA for feature extraction.}
-\label{pca-union}
+\caption{Optimizing the hyper-parameters of a Support Vector Machine for $F_1$
+score.}
+\label{svm-gridsearch}
 \end{figure*}
-
-\subsection{Model selection}
 
@@
 
-\begin{figure*}[t]
-\begin{pythoncode}
-from sklearn.grid_search import GridSearchCV
-from sklearn.svm import SVC
-
-param_grid = [
-  {"kernel": ["linear"], "C": [1, 10, 100, 1000]},
-  {"kernel": ["rbf"], "C": [1, 10, 100, 1000],
-   "gamma": [0.001, 0.0001]}
-]
-clf = GridSearchCV(SVC(), param_grid, scoring="f1", cv=10)
-clf.fit(X_train, y_train)
-\end{pythoncode}
-\caption{Optimizing the hyper-parameters of a Support Vector Machine for $F_1$
-score.}
-\label{svm-gridsearch}
-\end{figure*}
-
 \section{Implementation}
@@
 Python run-time system. Examples of Cython code in \sklearn are
-stochastic gradient descent for linear models, some clustering
+Stochastic Gradient Descent for linear models, some clustering
 algorithms, and decision trees.
@@
 
-\sklearn competes with a many machine learning packages,
+\sklearn competes with many machine learning packages,
 both open source and proprietary.
@@
 The Pylearn2 library \citep{goodfellow2013pylearn2}
-is similar to \sklearn in that aims to facilitate reproducible results
+is similar to \sklearn in that it aims to facilitate reproducible results
 and model inspection.
@@
 This provides more control for the user,
-at the cost of more math to be mastered.
+at the cost of more advanced concepts to be mastered.
 
@@
 (supported in the language by a restricted form of co-routines called generators).
-Some \sklearn estimators, notably Naïve Bayes and Stochastic Gradient Descent
-(for learning linear models such as
-logistic regression and SVM),
+Some \sklearn estimators, notably Naïve Bayes and Stochastic Gradient Descent,
 have a \texttt{partial\_fit} method to learn from a stream of samples

### 2013-10-18 paper.tex
change back to my old affiliation
... after discussion with my new boss, who said work performed at UvA
should be credited to UvA. That saves a footnote :)
drop the SVM acronym altogether
The only thing we have to explain now is SVC, which was previously
unexpanded.

@@
 
-\affil[1]{Netherlands eScience Center}
+\affil[1]{ISLA, University of Amsterdam}
 \affil[2]{University of Liège}
@@
 an estimator is initialized from a set of named hyper-parameter values
-(e.g., the $C$ constant in Support Vector Machines, SVM)
+(e.g., the $C$ constant in support vector machines)
 and can be considered a function
@@
 The code in Figure~\ref{svm-gridsearch} shows how to find
-hyper-parameter settings for an SVM classifier (SVC)
+hyper-parameter settings for an support vector classifier (SVC)
 that maximize $F_1$ score
@@
 considered for the linear and radial basis function (RBF) kernels;
-an SVM with a linear kernel accepts a $\gamma$ parameter, but ignores it,
+an SVC with a linear kernel accepts a $\gamma$ parameter, but ignores it,
 so using a single parameter grid would waste computing time

### 2013-10-18 paper.tex
"stochastic gradient descent" in lowercase, not a brand name
minibatch SGD -> minibatch k-means
Shows that clustering benefits from partial_fit too.

@@
 Python run-time system. Examples of Cython code in \sklearn are
-Stochastic Gradient Descent for linear models, some clustering
+stochastic gradient descent for linear models, some clustering
 algorithms, and decision trees.
@@
 (supported in the language by a restricted form of co-routines called generators).
-Some \sklearn estimators, notably Naïve Bayes and Stochastic Gradient Descent,
+Some \sklearn estimators, among others the implementations of naïve Bayes
+and minibatch $k$-means \citep{sculley2010web},
 have a \texttt{partial\_fit} method to learn from a stream of samples

### 2013-10-18 paper.tex
copyedit intro again
I don't think they really do abstracts in CiSE.

@@
 
-{\bf Abstract}
 \sklearn is a machine learning library for the Python language.\footnote{\url{http://scikit-learn.org}}
@@
 Since 2007, \sklearn has been developed by more than a dozen
-core developers, mostly researchers in various fields
-(among others, computer science, neuroscience, astrophysics). It also benefits
+core developers, mostly researchers in fields ranging from
+neuroscience to astrophysics. It also benefits
 from occasional contributors proposing small bugfixes or

### 2013-10-19 paper.tex
Nitpicks in sections 1 and 2

@@
 as to avoid the proliferation of framework code.
-We adopt simple conventions and limit the number of methods
-an object must implement to a minimum.
+We adopt simple conventions and limit to a minimum the number of methods
+an object must implement.
 The API is designed to adhere to the following principles:
@@
 to a new environment where the full library cannot be installed.
-However, our inspectable setup where model parameters are documented,
-public attributes, and prediction formulas follow standard textbooks,
+However, our inspectable setup where model parameters are documented
+public attributes and prediction formulas follow standard textbooks,
 goes a long way in solving this problem.

### 2013-10-19 paper.tex
Typos

@@
 The code in Figure~\ref{svm-gridsearch} shows how to find
-hyper-parameter settings for an support vector classifier (SVC)
-that maximize $F_1$ score
+hyper-parameter settings for a support vector classifier (SVC)
+that maximizes $F_1$ score
 through 10-fold cross-validation on the training set.
@@
 
-Our rule of thumb is that user code should not be tied to \sklearn---which is a
+Our rule of thumb is that user code should not be tied to \sklearn -- which is a
 \textit{library}, not a framework. We try to avoid a
@@
 Programs using \sklearn should not be intimately tied to it,
-so that their code can be reused or in other contexts.
+so that their code can be reused in other contexts.
 

### 2013-10-19 paper.tex
Don't cite mini-batch k-means
Indeed, we do not cite any of the other mentioned algorithms. I don't think we should make an exception for this one.

@@
 Some \sklearn estimators, among others the implementations of naïve Bayes
-and minibatch $k$-means \citep{sculley2010web},
+and minibatch $k$-means,
 have a \texttt{partial\_fit} method to learn from a stream of samples

### 2013-10-19 paper.tex
Nitpick

@@
 (supported in the language by a restricted form of co-routines called generators).
-Some \sklearn estimators, among others the implementations of naïve Bayes
+Some \sklearn estimators, among which the implementations of naïve Bayes
 and minibatch $k$-means,

### 2013-10-21 paper.tex
some more copy-editing/trimming

@@
        manner for all objects.
-  \item[Inspection.]
-       Constructor parameters and parameter values determined by learning
-       algorithms are stored and exposed as public attributes.
+  \item[Inspectability.]
+       Parameters, whether set by the user or learned from data,
+       are stored and exposed as public attributes.
   \item[Non-proliferation of classes.]
@@
        Datasets are represented as NumPy arrays or SciPy sparse matrices.
-       Hyper-parameter names and values are represented as standard
-       Python strings or numbers whenever possible.
+       User-set parameter names and values are represented as standard
+       Python strings or numbers, not custom objects.
        This keeps \sklearn easy to use and easy to combine with other libraries.
-  \item[Composition.]
+  \item[Composability.]
        Learning algorithms can be combined into processing pipelines.
@@
        Whenever an operation requires a user-defined parameter,
-       an appropriate default value is defined by the library.
-       The default value should cause the operation to be performed
-       in a sensible way (giving a baseline solution for the task at hand).
+       a sensible default value is defined by the library
+       to give a baseline solution for the task at hand.
 \end{description}
@@
 \begin{pythoncode}
-from sklearn.linear_model import LogisticRegression
+from sklearn.linear_model import \
+    LogisticRegression
 clf = LogisticRegression(penalty="l1")
@@
 perform useful work, implement the estimator interface. As we will illustrate
-in the next sections, this design pattern is indeed of prime importance for
+in the next sections, this design pattern is of prime importance for
 consistency, composition and model selection reasons.
@@
 \end{small}
-Of course, in practice, it is important to apply the same preprocessing to the
-test data \texttt{X\_test}. Since a \texttt{StandardScaler} estimator stores the
-mean and standard deviation that it computed for the training set, transforming
-an unseen test set \texttt{X\_test} maps it into the appropriate region of
-feature space:
-\begin{pythoncode}
-X_test = scaler.transform(X_test)
-\end{pythoncode}
+To obtain good results, it is necessary to apply the exact same preprocessing to the
+test data \texttt{X\_test}.
+For this reason, a \texttt{StandardScaler} stores the mean and standard deviation
+that it computed for the training set
+and will use these when asked to \texttt{transform}
+an unseen set of samples \texttt{X\_test}:
+
 Transformers also include a variety of learning algorithms, such as
-dimensionality reduction (PCA, manifold learning), kernel approximation,
+dimensionality reduction, manifold learning, kernel approximation,
 and other mappings from one feature space to another.
@@
 
-As introduced in Section~\ref{sec:estimators}, hyper-parameters set in the
-constructor of an estimator
+Hyper-parameters values set in the constructor of an estimator
+(Section~\ref{sec:estimators})
 determine the behavior of the learning algorithm
-and hence the performance of the resulting model on unseen data.
-For example, a decision
-tree with too small a value for the maximal tree depth
-parameter will tend to underfit, while too large a value will make it overfit.
-The problem of \textit{model selection} is therefore to find, within
-some hyper-parameter space, the best combination of hyper-parameters, with
-respect to some user-specified criterion.
+and the performance of the resulting model on unseen data.
+The problem of \textit{model selection} is to find
+the optimal combination of hyper-parameters, with
+respect to some user-specified criterion. For example, a decision
+tree with too small a maximal tree depth parameter will fit its training set badly,
+while too large a value will make it overfit.
 
@@
 \texttt{GridSearchCV} and \texttt{RandomizedSearchCV}.  These take as input an
-estimator (basic or composite), whose hyper-parameters must be optimized, and a
-set of hyperparameter settings to search through.
+estimator, basic or composite,
+and a set of hyperparameter settings to search through.
 This set is represented as a mapping of parameter names
@@
 
-The code in Figure~\ref{svm-gridsearch} shows how to find
-hyper-parameter settings for a support vector classifier (SVC)
-that maximizes $F_1$ score
-through 10-fold cross-validation on the training set.
+Figure~\ref{svm-gridsearch} shows code that finds
+settings for an SVM classifier (SVC)
+that maximize $F_1$ score
+through 10-fold cross-validation.
 Two distinct hyper-parameter grids are
-considered for the linear and radial basis function (RBF) kernels;
-an SVC with a linear kernel accepts a $\gamma$ parameter, but ignores it,
-so using a single parameter grid would waste computing time
+considered for the linear and radial basis function (RBF) kernels:
+an SVM with a linear kernel accepts a $\gamma$ parameter, but ignores it,
+so using a single grid would waste computing time
 trying out effectively equivalent settings.
@@
 When feasible, external libraries are integrated into the codebase.
-In particular, \sklearn includes modified versions of \textsf{LIBSVM} and \textsf{LIBLINEAR}
-\citep{chang2011libsvm,fan2008}, both written in C{}\verb!++!
-and wrapped using Cython.
+In particular, \sklearn includes modified versions of the C{}\verb!++! libraries
+\textsf{LIBSVM} and \textsf{LIBLINEAR} \citep{chang2011libsvm,fan2008}.
 
@@
 and Vowpal~Wabbit,\footnote{\url{http://hunch.net/~vw}}
-are often designed as command-line tools
-and may not offer any kind of API\@.
+are often designed as command-line tools without any kind of API\@.
 While this has the advantage that users are not tied
@@
 By contrast, \sklearn allows users to implement that entire workflow
-in a single program, written in a single language.
-This also makes it easier for researchers and developers
+in a single language.
+This makes it easier for researchers and developers
 to exchange and collaborate on software, as dependencies and setup are kept to a
@@
 In comparison to these, though, Python has the distinct advantage
-that it is a \textit{general purpose} language,
+that it is a \textit{general-purpose} language,
 while NumPy and SciPy extend it with functionality
@@
 While support for many of these tasks is improving in languages such as
-\textsc{matlab} and R, they still lag behind Python in their general purpose
-applicability.
-In many applications of machine learning, these tasks, such as data access,
-data preprocessing and reporting, can take more work than applying
+\textsc{matlab} and R, they still lag behind Python in this regard.
+In many applications of machine learning, data access,
+preprocessing, and reporting can take more work than applying
 the actual learning algorithm.
@@
 (supported in the language by a restricted form of co-routines called generators).
-Some \sklearn estimators, among which the implementations of naïve Bayes
+Some \sklearn estimators, including the implementations of naïve Bayes
 and minibatch $k$-means,
@@
 This way of processing, as argued in Section~\ref{sec:arrays},
-allows for more efficient implementation.
+allows for efficient implementation.
 We make no attempt to hide this batch-oriented processing from the user,
@@
 which cannot execute Python code on multiple CPUs simultaneously.
-It follows that any parallel task decomposition must either be done
+It follows that parallel task decomposition must either be done
 inside Cython modules,
 or at a level high enough to warrant the overhead
-of creating multiple OS-level processes,
-and the ensuing inter-process communication.
+of inter-process communication.
 Parallel grid search is an example of the latter approach

### 2013-11-08 paper.tex
change title per discussion with @mblondel, @glouppe

@@
 
-\title{API design for machine learning software: experiences from the
-scikit-learn project}
+\title{API design for machine learning: experiences from scikit-learn}
 
@@
 \begin{pythoncode}
-from sklearn.linear_model import \
-    LogisticRegression
+from sklearn.linear_model import LogisticRegression
 clf = LogisticRegression(penalty="l1")

### 2014-02-12 paper.tex
remove some of the bragging

@@
 developers.
-
-\sklearn's popularity can be gauged from various indicators such as the hundreds
-of citations in scientific publications, successes in various machine learning
-challenges (e.g., Kaggle), and statistics derived from our
-repositories and mailing list.  At the time of writing, the project is watched
-by 1365 people and forked 693 times on GitHub; the mailing list receives more
-than 300 mails per month; version control logs
-% ddaa494c116e3c16bf032003c5cccbed851733d2
-show 183 unique contributors to the codebase and the online documentation
-receives 37,000 unique visitors and 295,000 pageviews per month.
 

### 2014-02-12 paper.tex
change my affiliation (again)
stress composition and model selection in the introduction

@@
 
-\affil[1]{ISLA, University of Amsterdam}
+\affil[1]{ISLA, University of Amsterdam; Netherlands eScience Center}
 \affil[2]{University of Liège}
@@
 common machine learning concepts.
-We first present in section~\ref{sec:core-api} the central application
-programming interface (API) and then describe, in section~\ref{sec:advanced-api},
-advanced API mechanisms built on the core interface.
+We first present in section~\ref{sec:core-api}
+the central concepts of our application programming interface (API).
+In section~\ref{sec:advanced-api}, we discuss
+advanced API mechanisms built on the core interface,
+focusing on composition of operations
+and on the important problem of model selection.
 Section~\ref{sec:implementation} briefly describes the implementation.

### 2014-02-12 paper.tex
remove scikit-image ref

@@
 or by wrapping existing C/C{}\verb!++!/Fortran code.
-The name ``scikit'' derives from ``SciPy toolkit''
-and is shared with \textit{scikit-image}.\footnote{\url{http://scikit-image.org}}
 IPython \citep{perez2007ipython} and Matplotlib \citep{hunter2007matplotlib}

### 2014-02-12 paper.tex
example of fit_transform's efficiency

@@
 or by using a more efficient algorithm for the transformation.
+For example, dimensionality reduction
+using truncated singular value decomposition (SVD)
+computes $U_k \Sigma_k V_k^\mathsf{T} \approx X$ in its \texttt{fit} phase;
+its transform phase then computes $X_k = XV$,
+but the combined \texttt{fit\_transform}
+can compute the same result more economically as $X_k = U\Sigma^\mathsf{T}$.
 In the same spirit, clustering algorithms provide a \texttt{fit\_predict}

### 2014-02-12 paper.tex
rewrite abstract to stress the general API rather than the lib
+ cite self

@@
 
-\sklearn is a machine learning library for the Python language.\footnote{\url{http://scikit-learn.org}}
-It is designed to be simple and efficient, accessible to
-non-experts, and reusable in various contexts. In this article, we present and
-discuss the uniform application programming interface shared
-by all learning and processing units in the library,
-and the way it facilitates composition and reusability. We also comment on
-implementation details specific to the Python ecosystem
-and analyze potential obstacles faced by users and developers of the library.
+We describe a general, high-level application programming interface (API)
+for machine learning and data mining.
+This API, originally conceived for the Python library \sklearn
+but now adopted by several machine learning packages,
+is designed to be domain-independent and accessible
+to workers in the wide variety of scientific and engineering fields
+that employ machine learning.
+We present and discuss how a small set of primitive concepts
+facilititates composition of operations into complex workflows.\footnote{
+  An earlier version of this paper was presented at the ECML~PKDD
+  Workshop on Languages for Machine Learning \citep{buitinck2013ecml}.
+}
 

### 2014-02-18 paper.tex
less advertising/implementation stuff

@@
 or by wrapping existing C/C{}\verb!++!/Fortran code.
-IPython \citep{perez2007ipython} and Matplotlib \citep{hunter2007matplotlib}
-complement SciPy to provide a \textsc{matlab}-like working environment
-suited for scientific use.
-
-Since 2007, \sklearn has been developed by more than a dozen
-core developers, mostly researchers in fields ranging from
-neuroscience to astrophysics. It also benefits
-from occasional contributors proposing small bugfixes or
-improvements. Development proceeds on GitHub\footnote{\url{https://github.com/scikit-learn}},
-a platform which greatly facilitates this kind of
-collaboration. Because of the large number of developers, emphasis is
-put on keeping the project maintainable. In particular, code must follow
-specific quality guidelines, such as style consistency and unit-test coverage.
-Documentation and examples are required for all features,
-and major changes must pass code review by at least two other
-developers.
 
@@
 Binary distributions of these are available for the major platforms.
-Visualization functionality uses Matplotlib \citep{hunter2007matplotlib}
-and Graphviz \citep{gansner2000},
-but neither is required to perform machine learning or prediction.
-When feasible, external libraries are integrated into the codebase.
-In particular, \sklearn includes modified versions of the C{}\verb!++! libraries
-\textsf{LIBSVM} and \textsf{LIBLINEAR} \citep{chang2011libsvm,fan2008}.
 

### 2014-02-18 paper.tex
"generalize" introduction

@@
 
-The \sklearn library provides a comprehensive suite of machine learning tools
-for Python.
-It extends this general-purpose programming language with machine learning
-operations: learning algorithms, preprocessing tools, model selection procedures
-and a composition mechanism to produce complex ML workflows.
-The ambition of the library is to provide efficient implementations
-of well-established algorithms within a
-programming environment that is accessible to non-experts and
-reusable in various scientific areas.
-The library is distributed under the simplified BSD license
-to encourage its use in both academia and industry.
+In this paper, we present a high-level application programming interface (API)
+for machine learning tasks.
+The API presented here is primarily that of \sklearn \citep{pedregosa2011},
+an open~source Python library
+that provides a comprehensive suite of machine learning tools,
+extending the ``ecosystem'' of scientific Python tools
+centered around NumPy and SciPy \citep{oliphant2007python, vanderwalt2011}.
+\sklearn's API principles have been adopted by other packages including
+\textit{astroML} \citep{vanderplas2012astroML},
+a machine learning toolkit for astronomy;
+\textit{wiseRF}\footnote{\url{http://wise.io}},
+a commercial random~forest implementation;
+and \textit{lightning},\footnote{\url{https://github.com/mblondel/lightning}}
+the source code of a recently-proposed sparse multiclass
+algorithm \citep{mblondel-mlj2013}.
+% XXX PyStruct, seqlearn?
+While code examples in this article are in Python and use \sklearn,
+the principles exposed are general
+and should apply in any modern programming language.
 
-\sklearn is designed to tie in with the set of numeric
-and scientific packages centered around the NumPy and SciPy libraries
-\citep{oliphant2007python, vanderwalt2011}.
-NumPy augments Python with a contiguous numeric array datatype
-and fast array computing primitives,
-while SciPy extends it further with common numerical operations,
-either by implementing these in Python/NumPy
-or by wrapping existing C/C{}\verb!++!/Fortran code.
-
-\citet{pedregosa2011} briefly presented \sklearn and
-benchmarked it against several competitors.
 In this paper, we present an
